### 2025-05-21 advances/informe1.qmd
Update

---
title: "Preferences for market justice: the role of intergenerational social mobility and beliefs in meritocracy in Chile"
format:
  docx:
    number-sections: true
  html: 
    code-fold: true
    html-math-method: katex
    number-sections: true
  pdf: 
    geometry: margin=2cm
    template-partials: 
      - title.tex
    keep-tex: true
    include-in-header:
      text: |
        \usepackage[noblocks]{authblk}
        \renewcommand*{\Authsep}{, }
        \renewcommand*{\Authand}{, }
        \renewcommand*{\Authands}{, }
        \renewcommand\Affilfont{\small}
    number-sections: true
editor: source
bibliography: "../input/bib/tesis-magister.bib"
csl: "../input/bib/apa6.csl"
author:
  - name: Andreas Laffert
    orcid: 0000-0002-9008-2454
    email: alaffertt@estudiante.uc.cl 
    affiliations:
      - ref: 1
affiliations: 
  - id: 1
    name: Instituto de Sociología, Pontificia Universidad Católica de Chile
citeproc: true
link-citations: true
linestretch: 1.15       
mainfont: Times New Roman
fontsize: 12pt          
colorlinks: true
fig-height: 4
fig-width: 7.5
---

```{r setup, include=FALSE}
 knitr::opts_chunk$set(echo=FALSE, warning = FALSE,message = FALSE, cache = TRUE,out.width = '85%',fig.pos= "H"
                       # , fig.align = 'center'
                       )
 # knitr::opts_knit$set(base.url = "../") #relative path for .html output file
 # knitr::opts_knit$set(root.dir = "../") #relative path for chunks within .rmd files
 options(scipen=999)
 options(kableExtra.auto_format = FALSE)
 rm(list=ls())
 options(knitr.kable.NA = '')
 options(knitr.graphics.error = FALSE)
 Sys.setlocale("LC_ALL", "ES_ES.UTF-8")
```

```{r}
# This document was last modified at {{< modified >}}
# 
# and it was last rendered at {{< now >}}

```

# Introduction

Since the early 1980s, many countries have experienced a widespread retreatment of universal welfare programs and have adopted a trend toward privatization and commoditization of various public goods, welfare policies and social services [@gingrich_making_2011; @streeck_how_2016; @salamon_marketization_1993]. Neoliberal reforms reshaped the architecture of welfare-state institutions across Western democracies and Latin America, extending market criteria into areas of social reproduction that were traditionally governed by the state. This shift diminished the role of public provision of core social services [@arrizabalo_milagro_1995; @busemeyer_welfare_2020]. Echoing Polanyi's [-@polanyi_great_1975] insight that markets constitute a distinct moral order, the institutional diffusion of market rules has fostered a corresponding moral economy---a constellation of norms and values about fair allocation embedded in, and reinforced by, social institutions [@mau_moral_2014; @svallforsMoralEconomyClass2006a]. Within this moral economy, a growing body of research has been addressed the extent to which, and the mechanisms through which, citizens deem it just that access to social services such as health care, pensions, and education be determined by market-based criteria, a phenomenon labelled market justice preferences [@busemeyer_skills_2014; @immergut_it_2020; @castillo_socialization_2024; @koos_moral_2019; @lindh_bringing_2023; @lindh_public_2015]. Understanding these preferences is crucial because they can legitimize economic inequality by framing it as the natural outcome of fair competition, driven by individual responsibility and limited state intervention [@svallfors_political_2007].

Existing literature consistently demonstrates that market justice preferences are shaped by both the economic and institutional contexts of countries and individuals' positions within social stratification. Recent studies indicate that in countries with greater public provision of social services or more extensive welfare states, market justice preferences tend to be lower [@busemeyer_skills_2014; @immergut_it_2020], whereas in contexts with higher levels of privatization, preferences for market-based criteria are more pronounced [@castillo_socialization_2024; @lindh_public_2015]. In such contexts, market justice preferences tend to rise as individuals "ascend" the social structure, with those in more privileged positions in terms of class, education, and income holding stronger preferences for market-based solutions compared to those in more disadvantaged or at-risk positions [@busemeyer_skills_2014; @immergut_it_2020; @lindh_public_2015; @svallfors_political_2007; @kluegel_legitimation_1999]. Empirically, it has been consistently demonstrated that those in socioeconomically advantaged positions endorse the idea that those with higher incomes should be able to pay more for better social services in the domains of education [@lee_fairness_2023], healthcare [@vondemknesebeck_are_2016; @immergut_it_2020] and old age pensions [@castillo_socialization_2024; @otero_power_2024].

Although it is clear that one's social position influences market preferences, the question of how upward or downward mobility within the social structure affects these preferences remains unanswered. This is far from trivial, as many individuals in Western democracies and Latin America have experienced varying degrees of social mobility [@lopez-roldan_comparative_2021; @breen_social_2004]. The social origins and destinations of individuals exert different effects on a range of attitudes [@gugushvili_trends_2014], while movement between these positions exposes individuals to distinct processes and mechanisms that can profoundly influence their perspectives, shaping how they perceive the world and evaluate what they consider fair [@gugushvili_trends_2014]). Some of the current research on attitudes toward inequality has concluded that the type of social mobility---upward or downward---that individuals experience, whether measured objectively or subjectively, has differentiated effects on support for redistribution [@alesina_intergenerational_2018; @benabou_social_2001; @jaime-castillo_social_2019; @schmidt_experience_2011], system legitimacy, and meritocratic beliefs [@day_movin_2017; @mijs_belief_2022], as well as on tolerance for economic inequality [@shariff_income_2016; @gugushvili_intergenerational_2016c], attributions about poverty and wealth [@bucca_merit_2016; @gugushvili_intergenerational_2016], and support for welfare policies and market economies [@gugushvili_subjective_2017; @wen_does_2021]. Thus, building on these contributions, expanding the analysis of the effects of social mobility on market justice preferences could contribute to this emerging literature and shed light on the mechanisms behind the justification of inequalities in access to social services.

Market justice preferences are shaped not only by objective socioeconomic factors but also by perceptions and beliefs about meritocracy. Meritocracy asserts that inequality is an inherent feature of societies but can be legitimized through principles such as effort and talent [@young_rise_1958; @davis_principles_2001]. Previous studies have shown that individuals with stronger meritocratic beliefs tend to perceive less inequality, attributing economic differences to individual achievements [@mijs_paradox_2019; @wilson_role_2003]. Additionally, these beliefs justify greater inequality, as they are associated with attitudes that uphold the legitimization of social differences [@batruch_belief_2023]. In unequal societies where the distribution of goods and services is predominantly governed by market logic, such beliefs can play a critical role in the acceptance of social inequalities. Recent research by @castillo_socialization_2024 demonstrates that students in Chile who believe that effort and talent are rewarded in their country hold stronger preferences for market justice in access to healthcare, pensions, and education. By incorporating perceptions of meritocracy into this study, we can better understand how preferences toward market justice are shaped not only by an individual's structural social position but also by beliefs about the justification of these inequalities. Moreover, these beliefs are often solidified early in life, reinforced by institutions that promote values such as effort and individual skills as means to upward social mobility [@castillo_socialization_2024; @reynolds_perceptions_2014].

Beyond the isolated effects of social mobility and beliefs in meritocracy on market justice preferences, these factors interact to shape these preferences in a complex way. One proposed mechanism in the literature regarding the process of social mobility that leads to differences in distributive justice preferences is the psychological mechanism of self-serving bias in causal attribution [@schmidt_experience_2011; @gugushvili_intergenerational_2016c]. Self-serving bias suggests that people attribute failures, such as downward mobility, to factors beyond their control, while explaining successes, such as upward mobility, by pointing to their own merits and effort [@miller_selfserving_1975]. Individuals who experience upward mobility tend to view their position as a result of their own effort and skills, making them more likely to support the idea that individuals are responsible for their own success or failure. Research shows that upward mobility is associated with lower redistribution preferences [@schmidt_experience_2011; @gugushvili_intergenerational_2016c] and stronger legitimacy of income inequality [@shariff_income_2016]. In contrast, individuals who experience downward mobility tend to attribute their failure to external factors, such as inequality, and are more supportive of redistribution and the idea that individuals should not be blamed for their economic failure [@schmidt_experience_2011; @gugushvili_intergenerational_2016c]. Taken together, I suggest that meritocratic beliefs may mirror this self-serving bias mechanism, as they reinforce the tendency to justify one's social status as the result of personal merits, which is closely related to the attribution bias.

Focusing in Chile, this article aims to achieve two main objectives: first, to analyze the extent to which intergenerational social mobility affects market justice preferences on healthcare, pensions and education, and second, to examine how meritocratic beliefs may moderate this relationship. The central argument is that upward social mobility will have a positive effect on market justice preferences, while downward mobility will have a negative effect. Additionally, meritocratic beliefs will moderate this relationship by reflecting a self-serving bias mechanism of causal attribution.

The Chilean case is particularly intriguing for studying market justice preferences. Despite being one of the most prosperous nations in Latin America, it has one of the highest levels of economic inequality in the region. The poorest 50% of Chileans capture only 10% of the total income and hold negative wealth, while the richest 1% receives nearly 27% of the income and controls 49.6% of the wealth [@chancel_world_2022]. This economic inequality exists alongside notable social mobility, with significant upward mobility in recent decades [@torche_intergenerational_2014; @espinoza_estratificacion_2013]. Chile's social policy regime is characterized by a welfare model that relies heavily on private provision in healthcare, pensions, and education systems, often segmented by individuals' ability to pay and heavily dependent on state subsidies [@boccardo_30_2020].

Against this backdrop, the main research question is: To what extent does intergenerational social mobility affect market justice preferences, and how do meritocratic beliefs explain this relationship in Chile? This study aims to contribute to the literature by providing evidence from a Latin American developing country, highlighting how changes in individuals' socioeconomic position and meritocratic beliefs influence market justice preferences in critical areas like healthcare, pensions, and education.

# Data, variables and methods

## Data

This study draws on data from the Chilean Longitudinal Social Survey (ELSOC) of the Center for Social Conflict and Cohesion Studies (COES). The survey is an annual panel collected from 2016 to 2023, comprising two independent samples (original and refreshment), featuring permanent and rotating questionnaire modules. It evaluates how individuals think, feel, and behave regarding conflict and social cohesion in Chile. The sampling design is complex, probabilistic, clustered, multistage, and stratified by city size, targeting men and women aged 18--75 who habitually reside in private dwellings in urban areas spanning 40 cities (92 municipalities, 13 regions). This analysis focuses on 2023, the most recent wave, which includes 1,737 respondents in the original sample and 989 in the refreshment sample. Further details on sampling, attrition, and weighting can be found at https://coes.cl/encuesta-panel/, and the dataset is publicly available at https://dataverse.harvard.edu/dataverse/elsoc.

## Variables

*Outcome variables*

**Market justice preferences**: The outcome variables in this study are market justice preferences. This construct is operationalized through three variables that address the degree of justification regarding whether access to social services in healthcare, pensions, and education should be income conditional. Specifically, the justification of inequality in healthcare is assessed by the question: "Is it fair in Chile that people with higher incomes can access better healthcare than people with lower incomes?" The same question is asked for pensions and education. In all cases, respondents indicate their level of agreement on a five-point Likert scale ranging from 1 ("strongly disagree") to 5 ("strongly agree"). Although these items theoretically reflect the same underlying concept of market justice [@castillo_socialization_2024; @lindh_public_2015], they are analyzed separately because healthcare, pensions, and education may prompt distinct response patterns [@lee_fairness_2023; @busemeyer_skills_2014; @immergut_it_2020].

*Independent variables*

**Social mobility**: Intergenerational social mobility is treated as an exposure indicating whether respondents occupy a different class position from their fathers. Following Breen and Ermisch's [-@breen_effects_2024] framework for estimating causal effects of mobility, the variable is constructed in two stages.

1.  Class assignment. Using the International Socio-Economic Index of Occupational Status (ISEI) derived from three-digit ISCO-08 codes, both paternal (origin) and respondent (destination) occupations are grouped into three strata:

    -   Upper (codes 100--299: managers, directors, professionals)
    -   Middle (codes 300--499: skilled white‑collar workers)
    -   Lower (codes 500 and above: manual, semi‑skilled, or unskilled workers)

2.  Propensity-score estimation. To model mobility as a treatment---i.e., a transition from origin *j* to destination *k*---I estimate multinomial logit propensity scores using covariates that influence an individual's likelihood of moving between strata: (a) father's educational level, (b) presence of both parents at age 15, (c) nationality, (d) age, (e) sex, and (f) ethnicity.

These propensity scores are subsequently employed to adjust for selection into mobility when assessing its effect on market justice preferences.

**Meritocracy**: Meritocratic perception is operationalized through two components: one addressing effort and another focusing on talent [@young_rise_1958]. The item used to gauge effort is: "In Chile, people are rewarded for their efforts," while the item for talent is: "In Chile, people are rewarded for their intelligence and skills." In both cases, respondents indicate their level of agreement on a five-point Likert scale, ranging from 1 ("strongly disagree") to 5 ("strongly agree").

*Controls*

Sociodemographic and attitudinal variables are included to control for potential composition effects in the population. In terms of sociodemographic characteristics, I incorporate per capita household income quantile, educational level (1=Primary or below, 2=Secondary, 3=Technical, 4=University or above), age (in years), and sex (1=Male, 2=Female), which have been shown to significantly influence market justice preferences [@castillo_socialization_2024; @lindh_public_2015]. Regarding attitudinal variables, I include political identification (1=Left, 2=Center, 3=Right, 4=No identification) and subjective social status (ranging from 1 to 10) because they may confound the relationship between market justice preferences and social mobility and meritocracy.

## Methods

Following Breen and Ermisch's [-@breen_effects_2024] strategy for estimating the causal effect of social mobility on market justice preferences, I employ linear regression models with inverse-probability weights (IPW) for mobility conditional on class of origin. This approach allows estimation of the average causal effect of moving from an origin class *j* to a destination class *k* by comparing individuals who actually reach *k* with their counterfactual outcome had they instead moved to an alternative destination *k′* [@breen_effects_2024]. The resulting estimand is the average treatment effect on the treated (ATT).

All the analyses will be conducted using R software.

# References

::: {#refs}
:::