### 2025-05-27 advances/informe2.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
]{article}

\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
  \setmainfont[]{Times New Roman}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=2cm]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage[noblocks]{authblk}
\renewcommand*{\Authsep}{, }
\renewcommand*{\Authand}{, }
\renewcommand*{\Authands}{, }
\renewcommand\Affilfont{\small}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Preferences for mercantilization: the role of intergenerational social mobility and beliefs in meritocracy in Chile},
  pdfauthor={Andreas Laffert Tamayo},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Preferences for mercantilization: the role of intergenerational
social mobility and beliefs in meritocracy in Chile}

  \author{Andreas Laffert Tamayo}
            \affil{%
                  Instituto de Sociología, Pontificia Universidad
                  Católica de Chile
              }

\date{}
\begin{document}
\maketitle

\setstretch{1.15}
\section{Introduction}\label{introduction}

What is the legitimate extent of market inequality in the eyes of the
public? Since the early 1980s, many countries have experienced a
widespread retreat from universal welfare programs and a shift toward
the privatization and commodification of public goods, welfare policies,
and social services (\citeproc{ref-gingrich_making_2011}{Gingrich,
2011}; \citeproc{ref-streeck_how_2016}{Streeck, 2016}). In Latin
America, as elsewhere, neoliberal reforms reshaped welfare-state
institutions by extending market logic into domains of social
reproduction that were traditionally governed by the state
(\citeproc{ref-arrizabalo_milagro_1995}{Arrizabalo, 1995};
\citeproc{ref-ferre_welfare_2023}{Ferre, 2023}). This transformation
reduced the role of public provision and increased the presence of
private actors in core social services
(\citeproc{ref-harvey_breve_2015}{Harvey, 2015}). Echoing Polanyi's
(\citeproc{ref-polanyi_great_1975}{1975}) insight that markets
constitute a distinct moral order, the institutional diffusion of market
rules has fostered a corresponding moral economy: a constellation of
norms and values concerning fair allocation, embedded in institutions
and shaping individual subjectivities
(\citeproc{ref-mau_inequality_2015}{Mau, 2015};
\citeproc{ref-svallforsMoralEconomyClass2006a}{Svallfors, 2006}). Within
this framework, a growing body of research examines the extent to which,
and the mechanisms by which, citizens consider it fair that the
allocation of services like health care, pensions, and education be
governed by market-based criteria---a phenomenon known as \emph{market
justice preferences} (\citeproc{ref-busemeyer_skills_2014}{Busemeyer,
2014}; \citeproc{ref-castillo_socialization_2024}{Castillo et al.,
2024}; \citeproc{ref-immergut_it_2020}{Immergut \& Schneider, 2020};
\citeproc{ref-koos_moral_2019}{Koos \& Sachweh, 2019};
\citeproc{ref-lindh_public_2015}{Lindh, 2015};
\citeproc{ref-lindh_bringing_2023}{Lindh \& McCall, 2023}).
Understanding these preferences is crucial, as they contribute to
legitimizing economic inequality by framing it as the fair result of
individual responsibility and limited state intervention
(\citeproc{ref-mau_inequality_2015}{Mau, 2015}).

Existing literature shows that market justice preferences are shaped by
both the economic and institutional context and individuals' positions
within social stratification. Grounded in the notion that economic
institutions influence people's normative attitudes
(\citeproc{ref-jacksonActorsInstitutions2010}{Jackson, 2010}), studies
find that countries with stronger public provision or more expansive
welfare states exhibit lower levels of market justice preferences
(\citeproc{ref-busemeyer_skills_2014}{Busemeyer, 2014};
\citeproc{ref-immergut_it_2020}{Immergut \& Schneider, 2020}), while
more privatized contexts show stronger support for market-based criteria
(\citeproc{ref-castillo_socialization_2024}{Castillo et al., 2024};
\citeproc{ref-lindh_public_2015}{Lindh, 2015}). In such contexts, market
justice preferences tend to rise as individuals ``ascend'' the social
structure, with those in more privileged positions in terms of class,
education, and income holding stronger preferences for market-based
solutions compared to those in more disadvantaged or at-risk positions
(\citeproc{ref-castillo_socialization_2024}{Castillo et al., 2024};
\citeproc{ref-immergut_it_2020}{Immergut \& Schneider, 2020};
\citeproc{ref-lee_fairness_2023}{Lee \& Stacey, 2023};
\citeproc{ref-lindh_public_2015}{Lindh, 2015};
\citeproc{ref-otero_power_2024}{Otero \& Mendoza, 2024};
\citeproc{ref-svallfors_political_2007}{Svallfors, 2007};
\citeproc{ref-vondemknesebeck_are_2016}{Von Dem Knesebeck et al.,
2016}).

Market justice preferences are shaped not only by objective
socioeconomic conditions but also by popular beliefs about inequality.
Among these, meritocracy is a key normative principle underpinning
market-based distributive preferences
(\citeproc{ref-mau_inequality_2015}{Mau, 2015}). It frames inequality as
inevitable but justifiable through effort and talent
(\citeproc{ref-davis_principles_2001}{Davis \& Moore, 2001};
\citeproc{ref-young_rise_1958}{Young, 1958}). Studies show that
individuals with stronger meritocratic beliefs tend to perceive less
inequality and legitimize it by attributing economic differences to
personal achievement (\citeproc{ref-batruch_belief_2023}{Batruch et al.,
2023}; \citeproc{ref-mijs_paradox_2019}{Mijs, 2019};
\citeproc{ref-wilson_role_2003}{Wilson, 2003}). In highly unequal
societies where access to services is largely governed by market logic,
such beliefs play a critical role in normalizing inequality. Recent
evidence from Chile shows that students who believe effort and talent
are rewarded in their country express stronger preferences for
market-based access to healthcare, pensions, and education
(\citeproc{ref-castillo_socialization_2024}{Castillo et al., 2024}).

Although it is clear that one's social position influences market
justice preferences, the question of how upward or downward mobility
within the social structure affects these preferences remains
unanswered. This question is far from trivial, especially in Latin
America, where many have experienced various forms of mobility amid high
economic inequality and deep welfare privatization
(\citeproc{ref-ferre_welfare_2023}{Ferre, 2023};
\citeproc{ref-lopez-roldan_comparative_2021}{López-Roldán \& Fachelli,
2021}; \citeproc{ref-torche_intergenerational_2014}{Torche, 2014}).
Social origins and destinations affect attitudes toward inequality in
distinct ways (\citeproc{ref-day_movin_2017}{Day \& Fiske, 2017};
\citeproc{ref-gugushvili_intergenerational_2016}{Gugushvili, 2016b},
\citeproc{ref-gugushvili_subjective_2017}{2017};
\citeproc{ref-jaime-castillo_social_2019}{Jaime-Castillo \&
Marqués-Perales, 2019}; \citeproc{ref-mijs_belief_2022}{Mijs et al.,
2022}; \citeproc{ref-wen_does_2021}{Wen \& Witteveen, 2021}), while
movement between these positions exposes individuals to different
experiences and mechanisms that shape their views on what is fair
(\citeproc{ref-gugushvili_trends_2014}{Gugushvili, 2014};
\citeproc{ref-mau_inequality_2015}{Mau, 2015}). Building on this
research, examining the effects of social mobility on market justice
preferences can help to illuminate how inequalities in access to social
services are justified among individuals who have experienced, or not,
changes in their social standing, and what are the normative mechanisms
that guide this justification (\citeproc{ref-mau_inequality_2015}{Mau,
2015}).

Beyond their isolated effects, social mobility and meritocratic beliefs
interact in complex ways to shape market justice preferences. Among
others, a key mechanism proposed in the literature to explain how
mobility influences distributive justice preferences is the
psychological process of self-serving bias in causal attribution
(\citeproc{ref-gugushvili_intergenerational_2016c}{Gugushvili, 2016a};
\citeproc{ref-schmidt_experience_2011}{Schmidt, 2011}). This bias
suggests that individuals attribute failures---such as downward
mobility---to external factors, while crediting successes---such as
upward mobility---to their own merit and effort
(\citeproc{ref-miller_selfserving_1975}{Miller \& Ross, 1975}). Those
who experience upward mobility tend to view their social position as
earned, making them more likely to believe that individuals are
responsible for their own success or failure. Research shows that upward
mobility is associated with weaker preferences for redistribution
(\citeproc{ref-alesina_intergenerational_2018}{Alesina et al., 2018};
\citeproc{ref-gugushvili_intergenerational_2016c}{Gugushvili, 2016a};
\citeproc{ref-jaime-castillo_social_2019}{Jaime-Castillo \&
Marqués-Perales, 2019}; \citeproc{ref-schmidt_experience_2011}{Schmidt,
2011}) and stronger legitimacy of income inequality
(\citeproc{ref-shariff_income_2016}{Shariff et al., 2016}). In contrast,
individuals who experience downward mobility tend to blame structural
factors like inequality and are more supportive of redistribution while
rejecting merit-based explanations
(\citeproc{ref-gugushvili_trends_2014}{Gugushvili, 2014}). Taken
together, I argue that meritocratic beliefs may reinforce this
self-serving attribution mechanism by legitimizing one's social status
as the outcome of personal merit, closely tied to attribution bias.

Against this background, this article pursues two main objectives:
first, to analyze the extent to which intergenerational social mobility
influences market justice preferences regarding healthcare, pensions,
and education; and second, to examine how meritocratic beliefs may
moderate this relationship. Building on a theoretical framework that
emphasizes how neoliberal transformations---particularly through the
privatization and commodification of key areas of social
reproduction---have profoundly reshaped processes of subject formation
(\citeproc{ref-mau_inequality_2015}{Mau, 2015}), the central argument is
that upward mobility increases support for market justice preferences,
while downward mobility decreases it. Moreover, meritocratic beliefs are
expected to moderate this relationship by reflecting a self-serving
attribution bias, whereby individuals justify their social position in
terms of personal merit.

This study focuses on Chile, a particularly intriguing case for
examining market justice preferences. Despite significant economic
growth and poverty reduction, Chile has some of the highest levels of
inequality in Latin America and among OECD countries
(\citeproc{ref-chancel_world_2022}{Chancel et al., 2022};
\citeproc{ref-flores_top_2020}{Flores et al., 2020}). This inequality
coexists with short-range upward mobility among lower-class segments
moving into middle strata, though strong barriers remain to reaching
higher positions (\citeproc{ref-espinoza_estratificacion_2013}{Espinoza
et al., 2013}; \citeproc{ref-torche_intergenerational_2014}{Torche,
2014}). What makes Chile especially salient is that much of this
inequality is rooted in deep neoliberal reforms that institutionalized
the privatization and commodification of key social sectors
(\citeproc{ref-madariaga_three_2020}{Madariaga, 2020}). Introduced
during the dictatorship (1973--1989) and expanded in democracy, these
reforms enabled the unprecedented emergence of markets in health,
pensions, and education, with provision segmented by individuals'
ability to pay and supported by public subsidies
(\citeproc{ref-boccardo_30_2020}{Boccardo, 2020}). In parallel---and
despite waves of protest against inequality and commodification from
2006 to 2019 (\citeproc{ref-somma_no_2021}{Somma et al.,
2021})---Chilean subjectivities have been increasingly shaped by
neoliberal discourses and market logics, influencing their attitudes
toward inequality and welfare distribution
(\citeproc{ref-araujo_desafios_2012}{Araujo \& Martuccelli, 2012};
\citeproc{ref-canalesceron_sujeto_2021}{Canales Cerón et al., 2021}).

In this context, the questions that guide this research are as follows:

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\tightlist
\item
  To what extent does intergenerational social mobility influence market
  justice preferences regarding healthcare, pensions, and education in
  Chile?
\item
  How do meritocratic beliefs condition or moderate this relationship in
  the Chilean context?
\end{enumerate}

To address these questions, this study draws on large-scale,
representative survey data collected in 2018 from the urban Chilean
population aged 18 to 75 (n = 2,726). The next section outlines the
theoretical framework linking market justice preferences, social
mobility, and meritocratic beliefs, and proposes a set of hypotheses.
This is followed by a description of the data, variables, and analytical
strategy. The final sections present the empirical findings, offer an
interpretation of the results, and conclude with a discussion of their
implications.

\section{Theoretical and empirical
background}\label{theoretical-and-empirical-background}

\subsection{Market justice
preferences}\label{market-justice-preferences}

Although redistribution typically refers to the state's capacity to
reallocate resources from the advantaged to the vulnerable, market
institutions also play a central role in distributing socially valuable
goods and rewards (Koos \& Sachweh, 2019; Lindh \& McCall, 2023). As
Polanyi (1957) observed, economic integration in capitalist societies is
primarily organized through market exchange, governed by a
self-regulating price system embedded in institutional frameworks. These
institutions are not mere aggregates of individual behavior, but social
realities endowed with rules, mechanisms, and normative meanings that
shape everyday thinking (Koos \& Sachweh, 2019; Immergut, 1998). In this
sense, the economic order is mirrored in a moral economy: collectively
shared norms and beliefs about justice in distribution, embedded and
reinforced through institutions (Mau, 2015; Svallfors, 2006). While most
research from this perspective has focused on welfare institutions,
recent scholarship has brought the market back into focus as a site of
distributive justice beliefs and institutional responses to inequality
(Lindh \& McCall, 2023). In many countries, privatization and
commodification have expanded market logic into core areas of social
reproduction, such as healthcare, education, pensions, and social
security, deepening inequality in access to these services (Gingrich,
2011; Ferre, 2023). Yet public support for market-based welfare
provision has grown, even in traditional welfare states, particularly
among higher-income groups who view private alternatives as more
efficient or higher in quality (Busemeyer \& Iversen, 2020). This shift
calls for broader inquiry into how institutions like markets structure
access to resources and legitimize inequalities (Lindh \& McCall, 2023;
Mau, 2015).

The legitimacy of market-based inequalities is closely tied to beliefs
about distributive justice grounded in market principles. Beyond
stratification structures and socioeconomic conditions, the study of
inequality also examines the beliefs individuals hold about the causes
of social inequality, the normative principles that underpin those
beliefs, the factors that shape them, and their consequences for
attitudes and behavior (Kluegel \& Smith, 1981). In this regard,
empirical research on distributive justice focuses on individuals'
conceptions of how goods and rewards \emph{should} be distributed in
society (Jasso et al., año). This line of inquiry allows for examining
the extent to which economic inequality is perceived as just, or in a
certain sense, legitimate (Castillo, 2011). One prominent strand
investigates the legitimacy of wage inequality, particularly salary
differences across occupations (Jasso, 1978; Castillo, 2011; Wegener,
1999). Another examines how individuals justify market-generated
inequalities in access to core social services such as healthcare,
education, and pensions. Here, legitimacy stems from the belief that
access to these goods should follow market-based criteria (Castillo et
al., 2024; Lindh, 2015). In such views, these services are treated as
legitimate commodities; goods that can be traded, priced, and evaluated
through market logic (Busemeyer \& Iversen, 2020).

Market justice preferences refer to normative beliefs that legitimize
the idea that access to core social services---such as healthcare,
education, or pensions---should be determined by market-based criteria.
Following Janmaat's (2013, p.~359) distinction, these preferences fall
under the category of ``beliefs,'' understood as normative ideas about
what inequality should look like, as opposed to ``perceptions,'' which
refer to subjective evaluations of existing inequality. Market justice
preferences reflect the view that access to these services should depend
on individuals' ability to pay, thus justifying inequalities generated
by market mechanisms (Kluegel \& Smith, 1999; Lindh, 2015). The concept
draws on Lane's (1968) classic contrast between market justice and
political justice: the former is grounded in the idea of earned
deserts---where rewards reflect effort, productivity, and skill---while
the latter prioritizes need and equality, typically expressed in welfare
state policies. Lane argued that markets and states differ in purpose
(efficiency vs.~need), logic (individual vs.~collective), and fairness
criteria (merit vs.~equality). Market justice assumes that markets are
neutral, self-regulating systems in which fair procedures yield outcomes
proportional to merit. Inequality, from this perspective, is not only
expected but legitimate---as long as it arises from fair competition. In
this way, market justice offers a moral lens through which individuals
can view the commodification and stratified access to social services as
fair and justified (Kluegel \& Smith, 1999; Lindh, 2015).

The empirical study of market justice preferences has relied on diverse
strategies to operationalize how individuals evaluate inequalities
produced by market allocation. One common approach assesses whether
people consider it fair that access to key social services---such as
healthcare, education, or pensions---depends on income. This strategy
builds on foundational work by Kluegel and Smith (1981), who explored
the normative justifications behind support for economic inequality.
More recent studies have expanded this logic beyond income to include
market-mediated access to welfare goods. For instance, Immergut and
Schneider (2020) and von dem Knesebeck et al.~(2016) examine whether
respondents believe it is just that those with higher incomes receive
better healthcare. Similarly, Lee and Stacey (2023) apply this approach
to educational opportunities. These studies typically use targeted
survey items that ask respondents to evaluate specific distributive
scenarios, allowing for comparative analyses across countries or welfare
regimes. Complementing these efforts, more recent research has
introduced composite indicators designed to capture broader orientations
toward market-based allocation. For example, Castillo et al.~(2024)
propose a single-item measure that summarizes individual support for
income-based access across multiple domains---health, education, and
pensions---within the Chilean context. Together, these instruments aim
to capture to what extent individuals perceive market-generated
inequality as justified.

Comparative empirical research has identified several individual-level
factors that influence support for market justice. Individuals in more
advantaged socioeconomic positions---those with higher income,
education, and occupational status---are consistently more likely to
endorse market-based distributive principles (Svallfors, 2007; Koos \&
Sachweh, 2019; Lindh, 2015). For example, Lindh (2015) finds that
individuals from the service class are more likely to support
market-based access to healthcare and education than skilled and
unskilled workers across 17 relatively affluent countries. In a
comparative analysis, Svallfors (2007) observes that this expected class
pattern appears clearly only in Sweden, where support for private
education and healthcare varies systematically by class. Busemeyer
(2014) similarly shows that support for private education is stronger
among high-income groups, while Immergut and Schneider (2020) and von
dem Knesebeck et al.~(2016) report comparable findings for healthcare,
suggesting that wealthier individuals perceive private provision as a
means of maintaining relative advantage. In Chile, Otero and Mendoza
(2024) show that individuals with higher income and university education
express stronger support for market allocation in healthcare, education,
and pensions. Beliefs about inequality and political orientation also
matter. In Chile, Castillo et al.~(2024) show that individuals who have
strong meritocratic beliefs are more likely to support market-based
distribution, while Lee and Stacy (2023) in Australia suggest that these
preferences are also greater as people lean toward economic
conservatism. In this sense, market justice preferences are shaped by
the interplay between structural position and normative reasoning.

However, individual characteristics alone do not fully explain variation
in support for market justice. A growing body of cross-national research
highlights the importance of institutional arrangements in shaping these
preferences. For instance, Immergut and Schneider (2020) find that in
countries with higher public spending on healthcare, individuals are
less likely to view income-based access as fair. Similarly, Busemeyer
(2014) shows that increased public investment in education is associated
with lower support for privatized provision. Conversely, Lindh (2015)
finds that in countries with more market-oriented welfare systems,
support for market-based distribution tends to be higher, suggesting
that individual attitudes often align with institutional outputs. These
findings are consistent with neo-institutionalist and policy feedback
theories (Campbell, 2012), which argue that institutions do more than
redistribute resources---they also shape the normative categories
through which individuals assess who is deserving of support (Immergut,
1990). In this view, dominant values and preferences are both shaped by
and embedded in institutional configurations (Busemeyer, 2014),
reinforcing the notion that institutions are not neutral structures, but
active producers of the moral frameworks that legitimize or challenge
inequality.

\subsection{Social mobility}\label{social-mobility}

The study of social mobility, its drivers and consequences has long been
central to sociology. Classic theorists explored not only movement
across social hierarchies but also its broader implications for class
conflict, norm stability, and institutional change (Breen \& Ermisch,
2024). Sorokin (1927) introduced the concept formally, defining mobility
as the shift of individuals, values, or objects between positions within
a stratification system, and distinguishing between horizontal and
vertical forms. Later work differentiated intergenerational from
intragenerational mobility, as well as absolute mobility---driven by
structural change---from relative mobility, which captures the extent to
which origins constrain destinations (Eyles et al., 2022). In Latin
America, research shows high absolute but low relative mobility (Bucca,
2016): although educational expansion and economic modernization have
enabled some upward movement, status reproduction remains strong,
especially among elites (Torche, 2015; López-Roldán \& Fachelli, 2021).
This reflects deep structural inequality, segmented education systems,
stratified labor markets, and legacies of dependent development that
restrict access to mobility channels and reinforce the intergenerational
transmission of advantage.

Beyond mapping mobility patterns, growing research has examined its
subjective and attitudinal effects. Mobility effects---defined as
outcomes resulting from movement between origin and destination classes
(Breen \& Ermisch, 2024)---have long attracted theoretical interest.
Sorokin's (1959) dissociative hypothesis posits that mobility, whether
upward or downward, may produce psychological strain due to conflicting
norms between class contexts, leading to lower life satisfaction.
Similarly, Lenski (1954) argued that status inconsistency---mismatches
among education, income, and occupation---can undermine well-being.
These ideas underpin extensive empirical work linking intergenerational
mobility to outcomes such as life satisfaction, mental health, and
stress (Hadjar, Gugushvili, Präg \& Gugushvili).

Research on intergenerational social mobility has increasingly examined
its effects on attitudes toward economic inequality. A key area of
inquiry focuses on how upward and downward mobility influence support
for redistribution, though findings remain mixed. Alesina et al.~(2018)
show that individuals with pessimistic expectations about their
mobility---particularly those anticipating downward movement---are more
likely to support generous redistributive policies. Similarly, Ares
(2020) finds that upwardly mobile individuals tend to be less supportive
of state-led redistribution compared to those who have experienced
downward mobility. Comparative studies by Schmidt (2011) and Gugushvili
(2016) likewise report that subjective upward mobility is associated
with weaker preferences for redistribution, while downward mobility
strengthens redistributive support. However, recent causal evidence from
Breen and Ermisch (2024) suggests the opposite: upward mobility may
increase redistributive preferences, while downward mobility may reduce
them.

Beyond redistribution, other studies have explored the impact of
mobility on broader beliefs about inequality. Gugushvili (2016) finds
that upwardly mobile individuals are more likely to adopt
individualistic attributions of poverty and to legitimize income
inequality, particularly in post-socialist societies. Similarly, Bucca
(2016) shows that subjective upward mobility reinforces individualistic
explanations of wealth in seven Latin American countries. At the macro
level, Shariff et al.~(2016) demonstrate that higher levels of national
economic mobility correlate with greater tolerance of inequality. In
contrast, Day and Fiske (2019) find that low perceived mobility
undermines belief in meritocracy and a just world, thereby weakening
system justification. Taken together, this body of research suggests
that mobility shapes attitudes toward inequality and justice through
multiple, and sometimes contradictory, mechanisms.

The literature on the effects of social mobility has proposed various
mechanisms to explain how and why changes in social position may
influence individual outcomes (Freyr \& Rhem, año). One of the most
prominent is the self-interest mechanism, which posits that individuals
who experience upward or downward mobility undergo a shift in their
material interests, thereby altering their perceptions and preferences
(Ares, 2020; Langsæther, Evans \& O'Grady, 2021; Freyr \& Rhem, año).
Closely related to this logic is the Prospect of Upward Mobility (POUM)
hypothesis, which suggests that individuals may oppose redistribution
not because of their current position, but because they anticipate
improving their status in the future (REF). A second line of explanation
draws on the framework of theories of socialization. Within this
tradition, hypotheses such as acculturation, socialization, and status
maximization propose that individuals adjust their attitudes based on
the norms and values of either their class of origin or their
destination, or a combination of both (Jaime-Castillo \&
Marqués-Perales, 2019). However, most of these mechanisms focus
primarily on the indirect effects of origin and destination positions,
rather than on the direct effect of experience of movement itself. In
response to this gap, recent research has highlighted the role of
self-serving bias in causal attribution processes, suggesting that
individuals tend to explain their mobility trajectories in ways that
justify their current position, which in turn shapes their beliefs and
preferences (Schmidt, 2011; Gugushvili, 2016; Molina, Bucca \& Macy,
2019).

The self-serving bias mechanism builds on a value-oriented perspective,
emphasizing that individuals' experiences of social mobility shape their
causal attributions, which in turn influence their beliefs about justice
and distributive preferences (Gugushvili, 2014). Causal attribution
refers to the process through which individuals generate explanations
for their own behavior and outcomes, as well as those of others
(Gugushvili, year). In this view, people's interpretations of economic
inequality depend on whether they believe such disparities reflect
unequal individual contributions. Individuals who adopt an internal
attribution framework tend to see success or failure as rooted in
personal characteristics such as effort, talent, or merit. In contrast,
those who regard inequality as unjust are more likely to adopt an
external attribution model, viewing outcomes as the result of structural
barriers beyond individual control (Kluegel \& Smith, 1999). This
mechanism, often described as intrapersonal causal attribution, reflects
how people explain their own socioeconomic positions---typically
attributing their successes to internal qualities while blaming failures
on external circumstances (Miller \& Ross, 1975). Over time, individuals
may revise their beliefs and attitudes: while early views are shaped by
their social origin, these are later adjusted in light of personal
experiences of mobility and the perceived role of ascribed versus
achieved factors in determining socioeconomic outcomes (Gugushvili,
year).

Empirical research has provided support for the self-serving bias
mechanism in shaping redistributive preferences and attitudes toward the
legitimacy of inequality. For instance, Schmidt (2011) finds that
individuals who experience upward mobility are more likely to interpret
their success as the result of personal effort or merit, and
consequently perceive redistribution as less necessary. Conversely,
individuals who experience downward mobility tend to attribute their
decline to external circumstances---such as structural inequality or
unemployment---and show stronger support for redistribution. In a
comparative analysis across different welfare domains, Gugushvili
demonstrates that upward mobility is associated with lower support for
government spending on housing and pensions, while individuals who
experience downward mobility express lower support for healthcare and
education spending, but favor increased investment in housing and
pensions, reflecting the material nature of these domains. Moreover,
Gugushvili (2016) finds that upward mobility is linked to greater
justification of income inequality, suggesting that improved social
standing reinforces an attributional view in which success is seen as
the result of individual characteristics---thus legitimizing inequality
as a fair outcome.

These findings suggest that causal attribution processes mediate the
relationship between mobility experiences and justice-related attitudes.
Accordingly, this study proposes the following hypothesis:

\(H1\): Experiencing upward (downward) social mobility is positively
(negatively) associated with greater support for market justice in
healthcare, pensions, and education.

\subsection{Meritocracy}\label{meritocracy}

Meritocracy constitutes a central ideological framework for legitimizing
different types of social inequality, for instance through market
justice beliefs. Rooted in the belief that rewards and positions should
be allocated based on individual effort and talent, meritocracy operates
as a normative ideal and a descriptive belief about how society
functions. As initially conceptualized by Michael Young
(\citeproc{ref-young_rise_1962}{\textbf{young\_rise\_1962?}}), the term
was meant to critique a system in which merit-based stratification
becomes a new form of inequality. However, over time, meritocracy has
been widely supported in many societies as a fair and desirable
principle of distribution, particularly within liberal democracies and
market-oriented economies (\citeproc{ref-sandel_tyranny_2020}{Sandel,
2020}; \citeproc{ref-mijs_paradox_2021}{\textbf{mijs\_paradox\_2021?}}).
From a sociological perspective, the belief in meritocracy is more than
a cognitive assessment; it reflects a moral lens through which
individuals interpret inequality. People who believe that success
results from hard work and talent are more likely to view social and
economic disparities as legitimate
(\citeproc{ref-batruch_belief_2023}{Batruch et al., 2023};
\citeproc{ref-castillo_percepcion_2019}{\textbf{castillo\_percepcion\_2019?}}).
Conversely, if they see outcomes as driven by luck, social origin, or
systemic barriers, inequality is more likely to be perceived as unjust.
This distinction becomes crucial in societies with persistent structural
inequality, where public narratives often emphasize personal
responsibility and merit while overlooking entrenched disadvantages.

I adopt a multidimensional perspective on meritocracy, distinguishing
between two key dimensions: effort-based and talent-based perceptions.
This distinction is essential, as it captures different pathways through
which individuals justify inequality. Effort-based meritocracy
emphasizes hard work and perseverance as the basis for social rewards,
aligning closely with cultural narratives of personal responsibility. A
talent-based meritocracy, by contrast, emphasizes intelligence and
innate abilities, which are often perceived as less malleable and more
unequally distributed. Both dimensions have been shown to correlate with
acceptance of inequality, but they may carry distinct implications for
how inequality is justified in specific domains
(\citeproc{ref-castillo_multidimensional_2023}{\textbf{castillo\_multidimensional\_2023?}}).
The relevance of this distinction is supported by recent studies, which
show that individuals respond differently to these dimensions. For
instance, perceptions that effort is rewarded in society are more
strongly associated with positive evaluations of fairness and acceptance
of unequal outcomes (\citeproc{ref-batruch_belief_2023}{Batruch et al.,
2023}). This may be because effort is seen as a controllable and morally
virtuous trait, whereas talent is often perceived as a natural
advantage. Consequently, effort-based meritocracy is likely more potent
in legitimizing inequality, particularly in neoliberal contexts.

These dimensions of meritocracy reflect how respondents perceive
society's distributive logic, regardless of whether they endorse
meritocratic principles. This distinction aligns with recent findings
indicating that individuals distinguish between how merit is perceived
in society and how it should ideally operate, which in turn shapes their
preferences for redistribution and justice
(\citeproc{ref-tejero-peregrina_perceived_2025}{Tejero-Peregrina et al.,
2025}). Meritocratic beliefs serve as symbolic justifications for
unequal outcomes, particularly when access is stratified by income or
social opportunity. Prior studies have shown that individuals who
perceive higher levels of meritocracy tend to express stronger support
for unequal distributions that reflect market outcomes
(\citeproc{ref-castillo_socialization_2024}{Castillo et al., 2024};
\citeproc{ref-castillo_percepcion_2019}{\textbf{castillo\_percepcion\_2019?}}).

In addition to influencing individual attitudes toward inequality,
meritocratic beliefs can contribute to social division and the
stigmatization of disadvantaged groups. Recent research has demonstrated
that exposure to meritocratic narratives can reinforce the belief that
poverty results from individual failings rather than systemic
conditions, reducing support for redistributive measures and increasing
the stigmatization of the poor
(\citeproc{ref-hoyt_mindsets_2023}{\textbf{hoyt\_mindsets\_2023?}}).
This reinforces negative stereotypes and reduces empathy toward
individuals from lower socioeconomic backgrounds. Moreover,
(\citeproc{ref-busemeyer_positive_2021}{\textbf{busemeyer\_positive\_2021?}})
argues that meritocratic narratives can serve as feedback mechanisms
that shape public opinion and well-being by framing individuals'
understanding of welfare outcomes as deserved or undeserved within
existing institutional structures. This psychological mechanism
highlights the normative power of meritocracy in stabilizing unequal
systems by shaping political attitudes and personal perceptions of
success and failure.

Importantly, recent research has explored how meritocratic beliefs
interact with experiences of intergenerational mobility to shape
distributive attitudes. The belief that one's success is earned can lead
upwardly mobile individuals to internalize meritocratic narratives and
justify existing inequalities, reinforcing support for market justice
(Gugushvili, 2016; Molina et al., 2019). Conversely, downwardly mobile
individuals who maintain strong meritocratic beliefs may interpret their
status as a personal failure, reducing their support for redistribution
(Day \& Fiske, 2019). At the macro level, Shariff et al.~(2016) show
that higher perceived mobility increases tolerance for inequality,
suggesting that meritocracy and mobility are mutually reinforcing.

Taken together, this literature supports the idea that meritocratic
beliefs can moderate the relationship between mobility and market
justice preferences. Individuals who experience mobility---especially
upward---may draw on meritocratic narratives to legitimize both their
own status and broader inequalities, thereby strengthening their support
for market-based distribution. Accordingly, the second hypothesis of
this study is:

\(H2\): The positive (negative) relationship between upward (downward)
social mobility and support for market justice in healthcare, pensions,
and education is moderated by meritocratic beliefs; specifically, this
association is stronger (weaker) among individuals with higher
perceptions of meritocracy.

\subsection{The Chilean context}\label{the-chilean-context}

\section{Data, variables and methods}\label{data-variables-and-methods}

\subsection{Data}\label{data}

This study draws on data from the Chilean Longitudinal Social Survey
(ELSOC) of the Center for Social Conflict and Cohesion Studies (COES).
The survey is an annual panel collected from 2016 to 2023, comprising
two independent samples (original and refreshment), featuring permanent
and rotating questionnaire modules. It evaluates how individuals think,
feel, and behave regarding conflict and social cohesion in Chile. The
sampling design is complex, probabilistic, clustered, multistage, and
stratified by city size, targeting men and women aged 18--75 who
habitually reside in private dwellings in urban areas spanning 40 cities
(92 municipalities, 13 regions). This analysis focuses on 2023, the most
recent wave, which includes 1,737 respondents in the original sample and
989 in the refreshment sample. Further details on sampling, attrition,
and weighting can be found at https://coes.cl/encuesta-panel/, and the
dataset is publicly available at
https://dataverse.harvard.edu/dataverse/elsoc.

\subsection{Variables}\label{variables}

\emph{Outcome variables}

\textbf{Market justice preferences}: The outcome variables in this study
are market justice preferences. This construct is operationalized
through three variables that address the degree of justification
regarding whether access to social services in healthcare, pensions, and
education should be income conditional. Specifically, the justification
of inequality in healthcare is assessed by the question: ``Is it fair in
Chile that people with higher incomes can access better healthcare than
people with lower incomes?'' The same question is asked for pensions and
education. In all cases, respondents indicate their level of agreement
on a five-point Likert scale ranging from 1 (``strongly disagree'') to 5
(``strongly agree''). Although these items theoretically reflect the
same underlying concept of market justice
(\citeproc{ref-castillo_socialization_2024}{Castillo et al., 2024};
\citeproc{ref-lindh_public_2015}{Lindh, 2015}), they are analyzed
separately because healthcare, pensions, and education may prompt
distinct response patterns
(\citeproc{ref-busemeyer_skills_2014}{Busemeyer, 2014};
\citeproc{ref-immergut_it_2020}{Immergut \& Schneider, 2020};
\citeproc{ref-lee_fairness_2023}{Lee \& Stacey, 2023}).

\emph{Independent variables}

\textbf{Social mobility}: Intergenerational social mobility is treated
as an exposure indicating whether respondents occupy a different class
position from their fathers. Following Breen and Ermisch's
(\citeproc{ref-breen_effects_2024}{2024}) framework for estimating
causal effects of mobility, the variable is constructed in two stages.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Class assignment. Using the International Socio-Economic Index of
  Occupational Status (ISEI) derived from three-digit ISCO-08 codes,
  both paternal (origin) and respondent (destination) occupations are
  grouped into three strata:

  \begin{itemize}
  \tightlist
  \item
    Upper (codes 100-299: managers, directors, professionals)
  \item
    Middle (codes 300-499: skilled white-collar workers)
  \item
    Lower (codes 500 and above: manual, semi-skilled, or unskilled
    workers)
  \end{itemize}
\item
  Propensity-score estimation. To model mobility as a treatment---i.e.,
  a transition from origin \emph{j} to destination \emph{k}---I estimate
  multinomial logit propensity scores using covariates that influence an
  individual's likelihood of moving between strata: (a) father's
  educational level, (b) presence of both parents at age 15, (c)
  nationality, (d) age, (e) sex, and (f) ethnicity.
\end{enumerate}

These propensity scores are subsequently employed to adjust for
selection into mobility when assessing its effect on market justice
preferences.

\textbf{Meritocracy}: Meritocratic perception is operationalized through
two components: one addressing effort and another focusing on talent
(\citeproc{ref-young_rise_1958}{Young, 1958}). The item used to gauge
effort is: ``In Chile, people are rewarded for their efforts,'' while
the item for talent is: ``In Chile, people are rewarded for their
intelligence and skills.'' In both cases, respondents indicate their
level of agreement on a five-point Likert scale, ranging from 1
(``strongly disagree'') to 5 (``strongly agree'').

\emph{Controls}

Sociodemographic and attitudinal variables are included to control for
potential composition effects in the population. In terms of
sociodemographic characteristics, I incorporate per capita household
income quantile, educational level (1=Primary or below, 2=Secondary,
3=Technical, 4=University or above), age (in years), and sex (1=Male,
2=Female), which have been shown to significantly influence market
justice preferences (\citeproc{ref-castillo_socialization_2024}{Castillo
et al., 2024}; \citeproc{ref-lindh_public_2015}{Lindh, 2015}). Regarding
attitudinal variables, I include political identification (1=Left,
2=Center, 3=Right, 4=No identification) and subjective social status
(ranging from 1 to 10) because they may confound the relationship
between market justice preferences and social mobility and meritocracy.

\subsection{Methods}\label{methods}

Following Breen and Ermisch's (\citeproc{ref-breen_effects_2024}{2024})
strategy for estimating the causal effect of social mobility on market
justice preferences, I employ linear regression models with
inverse-probability weights (IPW) for mobility conditional on class of
origin. This approach allows estimation of the average causal effect of
moving from an origin class \emph{j} to a destination class \emph{k} by
comparing individuals who actually reach \emph{k} with their
counterfactual outcome had they instead moved to an alternative
destination \emph{k′} (\citeproc{ref-breen_effects_2024}{Breen \&
Ermisch, 2024}). The resulting estimand is the average treatment effect
on the treated (ATT).

All the analyses will be conducted using R software.

\section{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-alesina_intergenerational_2018}
Alesina, A., Stantcheva, S., \& Teso, E. (2018). Intergenerational
{Mobility} and {Preferences} for {Redistribution}. \emph{American
Economic Review}, \emph{108}(2), 521--554.
\url{https://doi.org/10.1257/aer.20162015}

\bibitem[\citeproctext]{ref-araujo_desafios_2012}
Araujo, K., \& Martuccelli, D. (2012). \emph{Desaf{í}os comunes: Retrato
de la sociedad chilena y sus individuos} (1a. ed). Santiago: LOM
Ediciones.

\bibitem[\citeproctext]{ref-arrizabalo_milagro_1995}
Arrizabalo, X. (1995). \emph{{Milagro o quimera: la econom{í}a chilena
durante la dictadura}}. Libros de la Catarata.

\bibitem[\citeproctext]{ref-batruch_belief_2023}
Batruch, A., Jetten, J., Van De Werfhorst, H., Darnon, C., \& Butera, F.
(2023). Belief in {School Meritocracy} and the {Legitimization} of
{Social} and {Income Inequality}. \emph{Social Psychological and
Personality Science}, \emph{14}(5), 621--635.
\url{https://doi.org/10.1177/19485506221111017}

\bibitem[\citeproctext]{ref-boccardo_30_2020}
Boccardo, G. (2020). \emph{30 a{ñ}os de privatizaciones en {Chile}: Lo
que la pandemia revel{ó}} (Nodo XXI). Santiago.

\bibitem[\citeproctext]{ref-breen_effects_2024}
Breen, R., \& Ermisch, J. (2024). The {Effects} of {Social Mobility}.
\emph{Sociological Science}, \emph{11}, 467--488.
\url{https://doi.org/10.15195/v11.a17}

\bibitem[\citeproctext]{ref-busemeyer_skills_2014}
Busemeyer, M. R. (2014). \emph{Skills and {Inequality}: {Partisan
Politics} and the {Political Economy} of {Education Reforms} in {Western
Welfare States}}. Cambridge University Press.

\bibitem[\citeproctext]{ref-canalesceron_sujeto_2021}
Canales Cerón, M., Orellana Calderón, V. S., \& Guajardo Mañán, F.
(2021). Sujeto y cotidiano en la era neoliberal: El caso de la
educaci{ó}n chilena. \emph{Revista Mexicana de Ciencias Pol{í}ticas y
Sociales}, \emph{67}(244).
\url{https://doi.org/10.22201/fcpys.2448492xe.2022.244.70386}

\bibitem[\citeproctext]{ref-castillo_socialization_2024}
Castillo, J. C., Salgado, M., Carrasco, K., \& Laffert, A. (2024). The
{Socialization} of {Meritocracy} and {Market Justice Preferences} at
{School}. \emph{Societies}, \emph{14}(11), 214.
\url{https://doi.org/10.3390/soc14110214}

\bibitem[\citeproctext]{ref-chancel_world_2022}
Chancel, L., Piketty, T., Saez, E., \& Zucman, G. (2022). World
inequality report 2022.
https://bibliotecadigital.ccb.org.co/handle/11520/27510.

\bibitem[\citeproctext]{ref-davis_principles_2001}
Davis, K., \& Moore, W. E. (2001). Some {Principles} of
{Stratification}. In \emph{Social {Stratification}, {Class}, {Race}, and
{Gender} in {Sociological Perspective}, {Second Edition}} (2nd ed.).
Routledge.

\bibitem[\citeproctext]{ref-day_movin_2017}
Day, M. V., \& Fiske, S. T. (2017). Movin' on {Up}? {How Perceptions} of
{Social Mobility Affect Our Willingness} to {Defend} the {System}.
\emph{Social Psychological and Personality Science}, \emph{8}(3),
267--274. \url{https://doi.org/10.1177/1948550616678454}

\bibitem[\citeproctext]{ref-espinoza_estratificacion_2013}
Espinoza, V., Barozet, E., \& Méndez, M. L. (2013). {Estratificaci{ó}n y
movilidad social bajo un}.

\bibitem[\citeproctext]{ref-ferre_welfare_2023}
Ferre, J. C. (2023). Welfare regimes in twenty-first-century {Latin
America}. \emph{Journal of International and Comparative Social Policy},
\emph{39}(2), 101--127. \url{https://doi.org/10.1017/ics.2023.16}

\bibitem[\citeproctext]{ref-flores_top_2020}
Flores, I., Sanhueza, C., Atria, J., \& Mayer, R. (2020). Top {Incomes}
in {Chile}: {A Historical Perspective} on {Income Inequality},
1964--2017. \emph{Review of Income and Wealth}, \emph{66}(4), 850--874.
\url{https://doi.org/10.1111/roiw.12441}

\bibitem[\citeproctext]{ref-gingrich_making_2011}
Gingrich, J. R. (2011). \emph{Making {Markets} in the {Welfare State}:
{The Politics} of {Varying Market Reforms}} (1st ed.). Cambridge
University Press. \url{https://doi.org/10.1017/CBO9780511791529}

\bibitem[\citeproctext]{ref-gugushvili_trends_2014}
Gugushvili, A. (2014). Trends, {Covariates} and {Consequences} of
{Intergenerational Social Mobility} in {Post- Socialist Societies}.

\bibitem[\citeproctext]{ref-gugushvili_intergenerational_2016c}
Gugushvili, A. (2016a). Intergenerational objective and subjective
mobility and attitudes towards income differences: Evidence from
transition societies. \emph{Journal of International and Comparative
Social Policy}, \emph{32}(3), 199--219.
\url{https://doi.org/10.1080/21699763.2016.1206482}

\bibitem[\citeproctext]{ref-gugushvili_intergenerational_2016}
Gugushvili, A. (2016b). Intergenerational {Social Mobility} and {Popular
Explanations} of {Poverty}: {A Comparative Perspective}. \emph{Social
Justice Research}, \emph{29}(4), 402--428.
\url{https://doi.org/10.1007/s11211-016-0275-9}

\bibitem[\citeproctext]{ref-gugushvili_subjective_2017}
Gugushvili, A. (2017). Subjective {Intergenerational Mobility} and
{Support} for {Welfare State Programmes}.

\bibitem[\citeproctext]{ref-harvey_breve_2015}
Harvey, D. (2015). \emph{{Breve historia del neoliberalismo}}. Madrid
(Espa{ñ}a): Ediciones Akal.

\bibitem[\citeproctext]{ref-immergut_it_2020}
Immergut, E. M., \& Schneider, S. M. (2020). Is it unfair for the
affluent to be able to purchase {``better''} healthcare? {Existential}
standards and institutional norms in healthcare attitudes across 28
countries. \emph{Social Science \& Medicine}, \emph{267}, 113146.
\url{https://doi.org/10.1016/j.socscimed.2020.113146}

\bibitem[\citeproctext]{ref-jacksonActorsInstitutions2010}
Jackson, G. (2010). Actors and institutions. In G. Morgan, J. L.
Campbell, C. Crouch, O. K. Pedersen, \& R. Whitley (Eds.), \emph{The
{Oxford Handbook} of {Comparative Institutional Analysis}} (1st ed.).
Oxford University Press.
\url{https://doi.org/10.1093/oxfordhb/9780199233762.001.0001}

\bibitem[\citeproctext]{ref-jaime-castillo_social_2019}
Jaime-Castillo, A. M., \& Marqués-Perales, I. (2019). Social mobility
and demand for redistribution in {Europe}: A comparative analysis.
\emph{The British Journal of Sociology}, \emph{70}(1), 138--165.
\url{https://doi.org/10.1111/1468-4446.12363}

\bibitem[\citeproctext]{ref-koos_moral_2019}
Koos, S., \& Sachweh, P. (2019). The moral economies of market
societies: Popular attitudes towards market competition, redistribution
and reciprocity in comparative perspective. \emph{Socio-Economic
Review}, \emph{17}(4), 793--821.
\url{https://doi.org/10.1093/ser/mwx045}

\bibitem[\citeproctext]{ref-lee_fairness_2023}
Lee, J.-S., \& Stacey, M. (2023). Fairness perceptions of educational
inequality: The effects of self-interest and neoliberal orientations.
\emph{The Australian Educational Researcher}.
\url{https://doi.org/10.1007/s13384-023-00636-6}

\bibitem[\citeproctext]{ref-lindh_public_2015}
Lindh, A. (2015). Public {Opinion} against {Markets}? {Attitudes}
towards {Market Distribution} of {Social Services} -- {A Comparison} of
17 {Countries}. \emph{Social Policy \& Administration}, \emph{49}(7),
887--910. \url{https://doi.org/10.1111/spol.12105}

\bibitem[\citeproctext]{ref-lindh_bringing_2023}
Lindh, A., \& McCall, L. (2023). Bringing the market in: An expanded
framework for understanding popular responses to economic inequality.
\emph{Socio-Economic Review}, \emph{21}(2), 1035--1055.
\url{https://doi.org/10.1093/ser/mwac018}

\bibitem[\citeproctext]{ref-lopez-roldan_comparative_2021}
López-Roldán, P., \& Fachelli, S. (Eds.). (2021). \emph{Towards a
{Comparative Analysis} of {Social Inequalities} between {Europe} and
{Latin America}}. Cham: Springer International Publishing.
\url{https://doi.org/10.1007/978-3-030-48442-2}

\bibitem[\citeproctext]{ref-madariaga_three_2020}
Madariaga, A. (2020). The three pillars of neoliberalism: {Chile}'s
economic policy trajectory in comparative perspective.
\emph{Contemporary Politics}, \emph{26}(3), 308--329.
\url{https://doi.org/10.1080/13569775.2020.1735021}

\bibitem[\citeproctext]{ref-mau_inequality_2015}
Mau, S. (2015). \emph{Inequality, {Marketization} and the {Majority
Class}: {Why Did} the {European Middle Classes Accept Neo-Liberalism}?}
Houndmills: Palgrave Macmillan.

\bibitem[\citeproctext]{ref-mijs_paradox_2019}
Mijs, J. (2019). The paradox of inequality: Income inequality and belief
in meritocracy go hand in hand. \emph{Socio-Economic Review},
\emph{19}(1), 7--35. \url{https://doi.org/10.1093/ser/mwy051}

\bibitem[\citeproctext]{ref-mijs_belief_2022}
Mijs, J., Daenekindt, S., de Koster, W., \& van der Waal, J. (2022).
Belief in {Meritocracy Reexamined}: {Scrutinizing} the {Role} of
{Subjective Social Mobility}. \emph{Social Psychology Quarterly},
\emph{85}(2), 131--141. \url{https://doi.org/10.1177/01902725211063818}

\bibitem[\citeproctext]{ref-miller_selfserving_1975}
Miller, D. T., \& Ross, M. (1975). Self-serving biases in the
attribution of causality: {Fact} or fiction? \emph{Psychological
Bulletin}, \emph{82}(2), 213--225.
\url{https://doi.org/10.1037/h0076486}

\bibitem[\citeproctext]{ref-otero_power_2024}
Otero, G., \& Mendoza, M. (2024). The {Power} of {Diversity}: {Class},
{Networks} and {Attitudes Towards Inequality}. \emph{Sociology},
\emph{58}(4), 851--876. \url{https://doi.org/10.1177/00380385231217625}

\bibitem[\citeproctext]{ref-polanyi_great_1975}
Polanyi, K. (1975). \emph{The great transformation} (Repr). New York,
NY: Octagon Books.

\bibitem[\citeproctext]{ref-sandel_tyranny_2020}
Sandel, M. J. (2020). \emph{The tyranny of merit: {What}'s become of the
common good?} (First edition). New York: {Farrar, Straus and Giroux}.

\bibitem[\citeproctext]{ref-schmidt_experience_2011}
Schmidt, A. W. (2011). The experience of social mobility and the
formation of attitudes toward redistribution. In.

\bibitem[\citeproctext]{ref-shariff_income_2016}
Shariff, A. F., Wiwad, D., \& Aknin, L. B. (2016). Income {Mobility
Breeds Tolerance} for {Income Inequality}: {Cross-National} and
{Experimental Evidence}. \emph{Perspectives on Psychological Science},
\emph{11}(3), 373--380. \url{https://doi.org/10.1177/1745691616635596}

\bibitem[\citeproctext]{ref-somma_no_2021}
Somma, N. M., Bargsted, M., Disi Pavlic, R., \& Medel, R. M. (2021). No
water in the oasis: The {Chilean Spring} of 2019--2020. \emph{Social
Movement Studies}, \emph{20}(4), 495--502.
\url{https://doi.org/10.1080/14742837.2020.1727737}

\bibitem[\citeproctext]{ref-streeck_how_2016}
Streeck, W. (2016). \emph{How will capitalism end? Essays on a failing
system}. London: Verso.

\bibitem[\citeproctext]{ref-svallforsMoralEconomyClass2006a}
Svallfors, S. (2006). \emph{The {Moral Economy} of {Class}: {Class} and
{Attitudes} in {Comparative Perspective}}. Stanford University Press.

\bibitem[\citeproctext]{ref-svallfors_political_2007}
Svallfors, S. (Ed.). (2007). \emph{The {Political Sociology} of the
{Welfare State}: {Institutions}, {Social Cleavages}, and {Orientations}}
(1st ed.). Stanford University Press.
\url{https://doi.org/10.2307/j.ctvr0qv0q}

\bibitem[\citeproctext]{ref-tejero-peregrina_perceived_2025}
Tejero-Peregrina, L., Willis, G., Sánchez-Rodríguez, Á., \&
Rodríguez-Bailón, R. (2025). From {Perceived Economic Inequality} to
{Support} for {Redistribution}: {The Role} of {Meritocracy Perception}.
\emph{International Review of Social Psychology}, \emph{38}(1), 4.
\url{https://doi.org/10.5334/irsp.1013}

\bibitem[\citeproctext]{ref-torche_intergenerational_2014}
Torche, F. (2014). Intergenerational {Mobility} and {Inequality}: {The
Latin American Case}. \emph{Annual Review of Sociology}, \emph{40}(1),
619--642. \url{https://doi.org/10.1146/annurev-soc-071811-145521}

\bibitem[\citeproctext]{ref-vondemknesebeck_are_2016}
Von Dem Knesebeck, O., Vonneilich, N., \& Kim, T. J. (2016). Are health
care inequalities unfair? {A} study on public attitudes in 23 countries.
\emph{International Journal for Equity in Health}, \emph{15}(1), 61.
\url{https://doi.org/10.1186/s12939-016-0350-8}

\bibitem[\citeproctext]{ref-wen_does_2021}
Wen, F., \& Witteveen, D. (2021). Does perceived social mobility shape
attitudes toward government and family educational investment?
\emph{Social Science Research}, \emph{98}, 102579.
\url{https://doi.org/10.1016/j.ssresearch.2021.102579}

\bibitem[\citeproctext]{ref-wilson_role_2003}
Wilson, C. (2003). The {Role} of a {Merit Principle} in {Distributive
Justice}. \emph{The Journal of Ethics}, \emph{7}(3), 277--314.
\url{https://doi.org/10.1023/A:1024667228488}

\bibitem[\citeproctext]{ref-young_rise_1958}
Young, M. (1958). \emph{The rise of the meritocracy}. New Brunswick,
N.J., U.S.A: Transaction Publishers.

\end{CSLReferences}

\end{document}