### 2025-06-27 advances/informe3.qmd

---
title: "Preferences for mercantilization: the role of intergenerational social mobility and beliefs in meritocracy in Chile"
format:
  html: 
    code-fold: true
    html-math-method: katex
    number-sections: true
    code-link: true
    title-block-banner: true
    theme:
      - pretty.scss
  pdf: 
    geometry: margin=2cm
    template-partials: 
      - title.tex
    keep-tex: true
    include-in-header:
      text: |
        \usepackage[noblocks]{authblk}
        \renewcommand*{\Authsep}{, }
        \renewcommand*{\Authand}{, }
        \renewcommand*{\Authands}{, }
        \renewcommand\Affilfont{\small}
    number-sections: true
editor: source
bibliography: "../input/bib/tesis-magister.bib"
csl: "../input/bib/apa6.csl"
author:
  - name: Andreas Laffert Tamayo
    orcid: 0000-0002-9008-2454
    email: alaffertt@estudiante.uc.cl 
    affiliations:
      - ref: 1
affiliations: 
  - id: 1
    name: Instituto de Sociología, Pontificia Universidad Católica de Chile
link-citations: true
linestretch: 1.5       
mainfont: Times New Roman
fontsize: 13pt          
fig-height: 4
fig-width: 7.5
---

```{r setup, include=FALSE}
 knitr::opts_chunk$set(echo=FALSE, warning = FALSE,message = FALSE, cache = TRUE,out.width = '85%',fig.pos= "H"
                       # , fig.align = 'center'
                       )
 # knitr::opts_knit$set(base.url = "../") #relative path for .html output file
 # knitr::opts_knit$set(root.dir = "../") #relative path for chunks within .rmd files
 options(scipen=999)
 options(kableExtra.auto_format = FALSE)
 rm(list=ls())
 options(knitr.kable.NA = '')
 options(knitr.graphics.error = FALSE)
 Sys.setlocale("LC_ALL", "ES_ES.UTF-8")
```

```{r}
# This document was last modified at {{< modified >}}
# 
# and it was last rendered at {{< now >}}

```

# Introduction

:::: {.justify}

What is the legitimate extent of market inequality in the eyes of the public? Since the early 1980s, many countries have experienced a widespread retreat from universal welfare programs and a shift toward the privatization and commodification of public goods, welfare policies, and social services [@gingrich_making_2011; @streeck_how_2016]. In Latin America, as elsewhere, neoliberal reforms reshaped welfare-state institutions by extending market logic into domains of social reproduction that were traditionally governed by the state [@ferre_welfare_2023]. This transformation reduced the role of public provision and increased the presence of private actors in core social services [@ferre_welfare_2023]. Echoing Polanyi's [-@polanyi_great_1975] insight that markets constitute a distinct moral order, the institutional diffusion of market rules has fostered a corresponding moral economy: a constellation of norms and values concerning fair allocation, embedded in institutions and shaping individual subjectivities [@mau_inequality_2015; @svallforsMoralEconomyClass2006a]. Within this framework, a growing body of research examines the extent to which, and the mechanisms by which, citizens consider it fair that the allocation of services like health care, pensions, and education be governed by market-based criteria---a phenomenon known as *market justice preferences* [@busemeyer_skills_2014; @immergut_it_2020; @castillo_socialization_2024; @castillo_perceptions_2025; @koos_moral_2019; @lindh_bringing_2023; @lindh_public_2015]. Understanding these preferences is crucial, as they contribute to legitimizing economic inequality by framing it as the fair result of individual responsibility and limited state intervention [@mau_inequality_2015].

Existing literature shows that market justice preferences are shaped by both the economic and institutional context and individuals' positions within social stratification. Grounded in the notion that economic institutions influence people's normative attitudes [@immergut_theoretical_1998], studies find that countries with stronger public provision or more expansive welfare states exhibit lower levels of market justice preferences [@busemeyer_skills_2014; @immergut_it_2020], while more privatized contexts show stronger support for market-based criteria [@castillo_socialization_2024; @lindh_public_2015]. In such contexts, market justice preferences tend to rise as individuals "ascend" the social structure, with those in more privileged positions in terms of class, education, and income holding stronger preferences for market-based solutions compared to those in more disadvantaged or at-risk positions [@lee_fairness_2023; @vondemknesebeck_are_2016; @immergut_it_2020; @castillo_socialization_2024; @otero_power_2024; @lindh_public_2015; @svallfors_political_2007].

Market justice preferences are shaped not only by objective socioeconomic conditions but also by popular beliefs about inequality. Among these, meritocracy is a key normative principle underpinning market-based distributive preferences [@mau_inequality_2015]. Studies show that individuals with stronger meritocratic beliefs tend to perceive less inequality and legitimize it by attributing economic differences to personal achievement [@mijs_paradox_2019; @wilson_role_2003; @batruch_belief_2023]. In highly unequal societies where access to services is largely governed by market logic, such beliefs play a critical role in normalizing inequality. Recent evidence from Chile shows that students who believe effort and talent are rewarded in their country express stronger preferences for market-based access to healthcare, pensions, and education [@castillo_socialization_2024].

Although it is clear that one's social position influences market justice preferences, the question of how upward or downward mobility within the social structure affects these preferences remains unanswered. This question is far from trivial, especially in Latin America, where many have experienced various forms of mobility amid high economic inequality and deep welfare privatization [@lopez-roldan_comparative_2021; @torche_intergenerational_2014; @ferre_welfare_2023]. Social origins and destinations affect attitudes toward inequality in distinct ways [ @jaime-castillo_social_2019; @gugushvili_intergenerational_2016; @day_movin_2017; @gugushvili_subjective_2017; @wen_does_2021; @mijs_belief_2022], while movement between these positions exposes individuals to different experiences and mechanisms that shape their views on what is fair [@gugushvili_trends_2014; @mau_inequality_2015]. Building on this research, examining the effects of social mobility on market justice preferences can help to illuminate how inequalities in access to social services are justified among individuals who have experienced, or not, changes in their social standing, and what are the normative mechanisms that guide this justification [@mau_inequality_2015].

Beyond their isolated effects, social mobility and meritocratic beliefs interact in complex ways to shape market justice preferences. Among others, a key mechanism proposed in the literature to explain how mobility influences distributive justice preferences is the psychological process of self-serving bias in causal attribution [@schmidt_experience_2011; @gugushvili_intergenerational_2016c]. This bias suggests that individuals attribute failures---such as downward mobility---to external factors, while crediting successes---such as upward mobility---to their own merit and effort [@miller_selfserving_1975]. Those who experience upward mobility tend to view their social position as earned, making them more likely to believe that individuals are responsible for their own success or failure. Research shows that upward mobility is associated with weaker preferences for redistribution [@schmidt_experience_2011; @gugushvili_intergenerational_2016c; @jaime-castillo_social_2019; @alesina_intergenerational_2018] and stronger legitimacy of income inequality [@shariff_income_2016]. In contrast, individuals who experience downward mobility tend to blame structural factors like inequality and are more supportive of redistribution while rejecting merit-based explanations [@gugushvili_trends_2014]. Taken together, I argue that meritocratic beliefs may reinforce this self-serving attribution mechanism by legitimizing one's social status as the outcome of personal merit, closely tied to attribution bias.

Against this background, this article pursues two main objectives: first, to analyze the extent to which intergenerational social mobility influences market justice preferences regarding healthcare, pensions, and education; and second, to examine how meritocratic beliefs may moderate this relationship. Building on a theoretical framework that emphasizes how neoliberal transformations---particularly through the privatization and commodification of key areas of social reproduction---have profoundly reshaped processes of subject formation [@mau_inequality_2015], the central argument is that upward mobility increases support for market justice preferences, while downward mobility decreases it. Moreover, meritocratic beliefs are expected to moderate this relationship by reflecting a self-serving attribution bias, whereby individuals justify their social position in terms of personal merit.

This study focuses on Chile, a particularly intriguing case for examining market justice preferences. Despite significant economic growth and poverty reduction, Chile has some of the highest levels of inequality in Latin America and among OECD countries [@chancel_world_2022; @flores_top_2020]. This inequality coexists with short-range upward mobility among lower-class segments moving into middle strata, though strong barriers remain to reaching higher positions [@torche_intergenerational_2014; @espinoza_estratificacion_2013]. What makes Chile especially salient is that much of this inequality is rooted in deep neoliberal reforms that institutionalized the privatization and commodification of key social sectors [@madariaga_three_2020]. Introduced during the dictatorship (1973--1989) and expanded in democracy, these reforms enabled the unprecedented emergence of markets in health, pensions, and education, with provision segmented by individuals' ability to pay and supported by public subsidies [@boccardo_30_2020]. In parallel---and despite waves of protest against inequality and commodification from 2006 to 2019 [@somma_no_2021]---Chilean subjectivities have been increasingly shaped by neoliberal discourses and market logics, influencing their attitudes toward inequality and welfare distribution [@araujo_desafios_2012; @canalesceron_sujeto_2021].

In this context, the questions that guide this research are as follows:

(1) To what extent does intergenerational social mobility influence market justice preferences regarding healthcare, pensions, and education in Chile?
(2) How do meritocratic beliefs condition or moderate this relationship in the Chilean context?

To address these questions, this study draws on large-scale, representative survey data collected in 2018 from the urban Chilean population aged 18 to 75 (n = 2,726). The next section outlines the theoretical framework linking market justice preferences, social mobility, and meritocratic beliefs, and proposes a set of hypotheses. This is followed by a description of the data, variables, and analytical strategy. The final sections present the empirical findings, offer an interpretation of the results, and conclude with a discussion of their implications.

# Theoretical and empirical background

## Market justice preferences

Beyond state’s capacity to reallocate resources from the advantaged to the vulnerable, market institutions also play a central role in distributing socially valuable goods and rewards [@koos_moral_2019; @lindh_bringing_2023]. As Polanyi [-@polanyi_great_1975] observed, economic integration in capitalist societies is primarily organized through market exchange, governed by a self-regulating price system embedded in institutional frameworks. These institutions are not mere aggregates of individual behavior, but social realities endowed with rules, mechanisms, and normative meanings that shape everyday thinking [@koos_moral_2019; @immergut_theoretical_1998]. In this sense, the economic order is mirrored in a moral economy: collectively shared norms and beliefs about justice in distribution, embedded and reinforced through institutions [@mau_inequality_2015; @svallfors_political_2007]. While most research from this perspective has focused on welfare institutions, recent scholarship has brought the market back into focus as a site of distributive justice beliefs and institutional responses to inequality [@lindh_bringing_2023]. In many countries, privatization and commodification have expanded market logic into core areas of social reproduction, such as healthcare, education, pensions, and social security, deepening inequality in access to these services [@gingrich_making_2011; @ferre_welfare_2023]. Yet public support for market-based welfare provision has grown, even in traditional welfare states, particularly among higher-income groups who view private alternatives as more efficient or higher in quality [@busemeyer_welfare_2020]. This shift calls for broader inquiry into how institutions like markets structure access to resources and legitimize inequalities [@satz_por_2019; @mau_inequality_2015].

The legitimacy of market-based inequalities is closely tied to beliefs about distributive justice grounded in market principles. Beyond stratification structures and socioeconomic conditions, the study of inequality also focuses on individuals’ beliefs about the origins of inequality, the normative frameworks that sustain those beliefs, the mechanisms that shape them, and their implications for attitudes and behavior [@kluegel_beliefs_1981]. In this regard, empirical research on distributive justice focuses on individuals’ conceptions of how goods and rewards _should_ be distributed in society [@jasso_distributive_2016]. This line of inquiry allows for examining the extent to which economic inequality is perceived as just, or in a certain sense, legitimate [@castillo_legitimacy_2011]. Most of the research on this field explores the legitimacy of wage inequality, specially salary gaps between jobs at opposite ends of the occupational hierarchy [@castillo_legitimacy_2011; @jasso_gender_1999; @jasso_justice_1978]. Recently, others examines how individuals justify market-generated inequalities in access to core social services such as healthcare, education, pensions or social security. Here, legitimacy stems from the belief that access to these goods should follow market-based criteria [@castillo_socialization_2024;  @lindh_public_2015]. In such views, these services are treated as legitimate commodities; goods that can be traded, priced, and evaluated through market logic [@busemeyer_welfare_2020].

Market justice preferences refer to normative beliefs that legitimize the idea that access to core social services—such as healthcare, education, or pensions—should be determined by market-based criteria. Following Janmaat’s [-@janmaat_subjective_2013, p. 359] distinction, these preferences fall under the category of “beliefs,” understood as normative ideas about what inequality should look like, as opposed to “perceptions,” which refer to subjective evaluations of existing inequality. Market justice preferences reflect the view that access to these services should depend on individuals’ ability to pay, thus justifying inequalities generated by market mechanisms [@kluegel_legitimation_1999;  @lindh_public_2015]. The concept draws on Lane’s [-@lane_market_1986] classic contrast between market justice and political justice: the former is grounded in the idea of earned deserts—where rewards reflect effort, productivity, and skill—while the latter prioritizes need and equality, typically expressed in welfare state policies. Lane argued that markets and states differ in purpose (efficiency vs. need), logic (individual vs. collective), and fairness criteria (merit vs. equality). Market justice assumes that markets are neutral, self-regulating systems in which fair procedures yield outcomes proportional to merit. Inequality, from this perspective, is not only expected but legitimate—as long as it arises from fair competition. In this way, market justice offers a moral lens through which individuals can view the commodification and stratified access to social services as fair and justified [@kluegel_legitimation_1999; @lindh_public_2015].

The empirical study of market justice preferences has employed various strategies to capture how individuals assess inequalities arising from market allocation. A common approach evaluates whether people find it fair that access to essential social services—such as healthcare, education, or pensions—depends on income. This builds on foundational work by Kluegel and Smith [-@kluegel_legitimation_1999], who examined normative justifications for capitalist systems. Subsequent studies have extended this logic to a wider set of welfare goods. For example, Immergut and Schneider [-@immergut_it_2020] and von dem Knesebeck et al. [-@vondemknesebeck_are_2016] explore whether respondents believe it is just that higher-income individuals receive better healthcare. Similarly, Lee and Stacey [-@lee_fairness_2023] and @castillo_changes_2025 apply this framework to education. Complementing these efforts, more recent research has introduced composite indicators to capture broader orientations toward market-based allocation. Lindh [-@lindh_public_2015], for instance, constructs an index averaging support for income-based access to healthcare and education in comparative perspective, while Castillo et al. [-@castillo_socialization_2024] propose a single-item measure covering health, education, and pensions in Chile. These instruments seek to gauge to what extent individuals view market-generated inequalities as legitimate. Taken together, they capture two core dimensions of market distribution: the role of economic resources as a key determinant of outcomes, and the framing of services as tradable commodities that can be bought and sold according to ability to pay [@lindh_public_2015].

Comparative empirical research has identified several individual-level factors that influence support for market justice. Individuals in more advantaged socioeconomic positions—those with higher income, education, and occupational status—are consistently more likely to endorse market-based distributive principles [@koos_moral_2019; @lindh_public_2015; @svallfors_political_2007]. For example, Lindh [-@lindh_public_2015] finds that individuals from the service class are more likely to support market-based access to healthcare and education than skilled and unskilled workers across 17 relatively affluent countries. In a comparative analysis, Svallfors [-@svallfors_political_2007] observes that this expected class pattern appears clearly only in Sweden, where support for private education and healthcare varies systematically by class. Busemeyer [-@busemeyer_skills_2014] similarly shows that support for private education is stronger among high-income groups, while Immergut and Schneider [-@immergut_it_2020] and von dem Knesebeck et al. [-@vondemknesebeck_are_2016] report comparable findings for healthcare, suggesting that wealthier individuals perceive private provision as a means of maintaining relative advantage. In Chile, Otero and Mendoza [-@otero_power_2024] show that individuals with higher income and university education express stronger support for market allocation in healthcare, education, and pensions. Beliefs about inequality and political orientation also matter. Castillo et al. [-@castillo_socialization_2024] and @castillo_perceptions_2025 show that individuals who have strong meritocratic beliefs are more likely to support market-based distribution in Chile, while Lee and Stacy [-@lee_fairness_2023] in Australia suggest that these preferences are also greater as people lean toward economic conservatism. In this sense, market justice preferences are shaped by the interplay between structural position and normative reasoning.

However, individual characteristics alone do not fully explain variation in support for market justice. A growing body of cross-national research highlights the importance of institutional arrangements in shaping these preferences. For instance, Immergut and Schneider [-@immergut_it_2020] find that in countries with higher public spending on healthcare, individuals are less likely to view income-based access as fair. Similarly, Busemeyer [-@busemeyer_skills_2014] shows that increased public investment in education is associated with lower support for privatized provision. Conversely, Lindh [-@lindh_public_2015] finds that in countries with more market-oriented welfare systems, support for market-based distribution tends to be higher, suggesting that individual attitudes often align with institutional outputs. These findings are consistent with neo-institutionalist and policy feedback theories [@campbell_institutional_2020], which argue that institutions do more than redistribute resources—they also shape the normative categories through which individuals assess who is deserving of support [@immergut_theoretical_1998]. In this view, dominant values and preferences are both shaped by and embedded in institutional configurations [@busemeyer_skills_2014], reinforcing the notion that institutions are not neutral structures, but active producers of the moral frameworks that legitimize or challenge inequality.

## Social mobility

The study of social mobility, its drivers and consequences has long been central to sociology. Classic theorists explored not only movement across social hierarchies but also its broader implications for class conflict, norm stability, and institutional change [@breen_effects_2024]. Sorokin [-@sorokin_social_1927] introduced the concept formally, defining mobility as the shift of individuals, values, or objects between positions within a stratification system, and distinguishing between horizontal and vertical forms. Later work differentiated intergenerational from intragenerational mobility, as well as absolute mobility—driven by structural change—from relative mobility, which captures the extent to which origins constrain destinations [@eyles_social_2022]. In Latin America, research shows high absolute but low relative mobility [@bucca_merit_2016]: although educational expansion and economic modernization have enabled some upward movement, status reproduction remains strong, especially among elites [@torche_intergenerational_2014; @lopez-roldan_comparative_2021]. This reflects deep structural inequality, segmented education systems, stratified labor markets, and legacies of dependent development that restrict access to mobility channels and reinforce the intergenerational transmission of advantage.

Beyond mapping mobility patterns, growing research has examined its subjective and attitudinal effects. Mobility effects—defined as outcomes resulting from movement between origin and destination classes [@breen_effects_2024]—have long attracted theoretical interest. Sorokin’s (1959) dissociative hypothesis posits that mobility, whether upward or downward, may produce psychological strain due to conflicting norms between class contexts, leading to lower life satisfaction. Similarly, Lenski [-@lenski_status_1954] argued that status inconsistency—mismatches among education, income, and occupation—can undermine well-being. These ideas underpin extensive empirical work linking intergenerational mobility to outcomes such as life satisfaction, mental health, and stress [@prag_subjective_2021; @gugushvili_heterogeneous_2024; @hadjar_does_2015].

Research on intergenerational social mobility has increasingly examined its effects on attitudes toward economic inequality. A key area of inquiry focuses on how upward and downward mobility influence support for redistribution, though findings remain mixed. Alesina et al. [-@alesina_intergenerational_2018] show that individuals with pessimistic expectations about their mobility—particularly those anticipating downward movement—are more likely to support generous redistributive policies. Similarly, Ares [-@ares_changing_2020] finds that upwardly mobile individuals tend to be less supportive of state-led redistribution compared to those who have experienced downward mobility. Comparative studies by Schmidt [-@schmidt_experience_2011] and Gugushvili [-@gugushvili_subjective_2017] likewise report that subjective upward mobility is associated with weaker preferences for redistribution, while downward mobility strengthens redistributive support. However, recent causal evidence from Breen and Ermisch [-@breen_effects_2024] suggests the opposite: upward mobility may increase redistributive preferences, while downward mobility may reduce them.

Beyond redistribution, other studies have explored the impact of mobility on broader beliefs about inequality. Gugushvili [-@gugushvili_intergenerational_2016] finds that upwardly mobile individuals are more likely to adopt individualistic attributions of poverty and to legitimize income inequality, particularly in post-socialist societies. Similarly, Bucca [-@bucca_merit_2016] shows that subjective upward mobility reinforces individualistic explanations of wealth in seven Latin American countries. At the macro level, Shariff et al. [-@shariff_income_2016] demonstrate that higher levels of national economic mobility correlate with greater tolerance of inequality. In contrast, Day and Fiske [-@day_movin_2017] find that low perceived mobility undermines belief in meritocracy and a just world, thereby weakening system justification. Taken together, this body of research suggests that mobility shapes attitudes toward inequality and justice through multiple, and sometimes contradictory, mechanisms.

The literature on the effects of social mobility has proposed various mechanisms to explain how and why changes in social position may influence individual outcomes [@helgason_class_2025]. One of the most prominent is the self-interest mechanism, which posits that individuals who experience upward or downward mobility undergo a shift in their material interests, thereby altering their perceptions and preferences [@helgason_longterm_2023; @ares_changing_2020; @langsaether_explaining_2022]. Closely related to this logic is the Prospect of Upward Mobility (POUM) hypothesis, which suggests that individuals may oppose redistribution not because of their current position, but because they anticipate improving their status in the future [@benabou_social_2001]. A second line of explanation draws on the framework of theories of socialization. Within this tradition, hypotheses such as acculturation, socialization, and status maximization propose that individuals adjust their attitudes based on the norms and values of either their class of origin or their destination, or a combination of both [@jaime-castillo_social_2019]. However, most of these mechanisms focus primarily on the indirect effects of origin and destination positions, rather than on the direct effect of experience of movement itself. In response to this gap, recent research has highlighted the role of self-serving bias in causal attribution processes, suggesting that individuals tend to explain their mobility trajectories in ways that justify their current position, which in turn shapes their beliefs and preferences [@schmidt_experience_2011; @gugushvili_intergenerational_2016c; @molina_its_2019].

The self-serving bias mechanism builds on a value-oriented perspective, emphasizing that individuals’ experiences of social mobility shape their causal attributions, which in turn influence their beliefs about justice and distributive preferences [@gugushvili_trends_2014]. Causal attribution refers to the process through which individuals generate explanations for their own behavior and outcomes, as well as those of others [@gugushvili_intergenerational_2016]. In this view, people’s interpretations of economic inequality depend on whether they believe such disparities reflect unequal individual contributions. Individuals who adopt an internal attribution framework tend to see success or failure as rooted in personal characteristics such as effort, talent, or merit. In contrast, those who regard inequality as unjust are more likely to adopt an external attribution model, viewing outcomes as the result of structural barriers beyond individual control [@kluegel_beliefs_1981]. This mechanism, often described as intrapersonal causal attribution, reflects how people explain their own socioeconomic positions—typically attributing their successes to internal qualities while blaming failures on external circumstances [@miller_selfserving_1975]. Over time, individuals may revise their beliefs and attitudes: while early views are shaped by their social origin, these are later adjusted in light of personal experiences of mobility and the perceived role of ascribed versus achieved factors in determining socioeconomic outcomes [@gugushvili_intergenerational_2016].

Empirical research has provided support for the self-serving bias mechanism in shaping redistributive preferences and attitudes toward the legitimacy of inequality. For instance, Schmidt [-@schmidt_experience_2011] finds that individuals who experience upward mobility are more likely to interpret their success as the result of personal effort or merit, and consequently perceive redistribution as less necessary. Conversely, individuals who experience downward mobility tend to attribute their decline to external circumstances—such as structural inequality or unemployment—and show stronger support for redistribution. In a comparative analysis across different welfare domains, Gugushvili [-@gugushvili_subjective_2017] demonstrates that upward mobility is associated with lower support for government spending on housing and pensions, while individuals who experience downward mobility express lower support for healthcare and education spending, but favor increased investment in housing and pensions, reflecting the material nature of these domains. Moreover, Gugushvili [-@gugushvili_intergenerational_2016c] finds that upward mobility is linked to greater justification of income inequality, suggesting that improved social standing reinforces an attributional view in which success is seen as the result of individual characteristics—thus legitimizing inequality as a fair outcome.

Consequently, considering this theoretical and empirical background, the first hypothesis of this research is that:

$H1$: Experiencing upward (downward) social mobility is positively (negatively) associated with greater support for market justice in healthcare, pensions, and education.

## Meritocracy

Meritocracy constitutes a central ideological framework for legitimizing different types of social inequality, for instance through market justice beliefs. Rooted in the belief that rewards and positions should be allocated based on individual effort and talent, meritocracy operates as a normative ideal and a descriptive belief about how society functions. As initially conceptualized by Michael Young [-@young_rise_1958], the term was meant to critique a system in which merit-based stratification becomes a new form of inequality. However, over time, meritocracy has been widely supported in many societies as a fair and desirable principle of distribution, particularly within liberal democracies and market-oriented economies [@mijs_paradox_2019; @sandel_tyranny_2020]. From a sociological perspective, the belief in meritocracy is more than a cognitive assessment; it reflects a moral lens through which individuals interpret inequality. People who believe that success results from hard work and talent are more likely to view social and economic disparities as legitimate [@castillo_meritocracia_2019; @batruch_belief_2023]. Conversely, if they see outcomes as driven by luck, social origin, or systemic barriers, inequality is more likely to be perceived as unjust. This distinction becomes crucial in societies with persistent structural inequality, where public narratives often emphasize personal responsibility and merit while overlooking entrenched disadvantages.

I adopt a multidimensional perspective on meritocracy, distinguishing between two key dimensions: effort-based and talent-based perceptions. This distinction is essential, as it captures different pathways through which individuals justify inequality [@young_rise_1958]. Effort-based meritocracy emphasizes hard work and perseverance as the basis for social rewards, aligning closely with cultural narratives of personal responsibility. A talent-based meritocracy, by contrast, emphasizes intelligence and innate abilities, which are often perceived as less malleable and more unequally distributed. Both dimensions have been shown to correlate with acceptance of inequality, but they may carry distinct implications for how inequality is justified in specific domains [@castillo_multidimensional_2023]. The relevance of this distinction is supported by recent studies, which show that individuals respond differently to these dimensions. For instance, perceptions that effort is rewarded in society are more strongly associated with positive evaluations of fairness and acceptance of unequal outcomes [@batruch_belief_2023; @wiederkehr_belief_2015; @wilson_role_2003]. This may be because effort is seen as a controllable and morally virtuous trait, whereas talent is often perceived as a natural advantage. Consequently, effort-based meritocracy is likely more potent in legitimizing inequality, particularly in neoliberal contexts.

These dimensions of meritocracy reflect how respondents perceive society's distributive logic, regardless of whether they endorse meritocratic principles. This distinction aligns with recent findings indicating that individuals distinguish between how merit is perceived in society and how it should ideally operate, which in turn shapes their preferences for redistribution and justice [@tejero-peregrina_perceived_2025]. Meritocratic beliefs serve as symbolic justifications for unequal outcomes, particularly when access is stratified by income or social opportunity. Prior studies in Chile have shown that individuals who perceive higher levels of meritocracy tend to express stronger support for unequal distributions that reflect market outcomes in healthcare, education and pensions [@castillo_perceptions_2025; @castillo_socialization_2024].

In addition to influencing individual attitudes toward inequality, meritocratic beliefs can contribute to social division and the stigmatization of disadvantaged groups. Recent research has demonstrated that exposure to meritocratic narratives can reinforce the belief that poverty results from individual failings rather than systemic conditions, reducing support for redistributive measures and increasing the stigmatization of the poor [@hoyt_mindsets_2023]. This reinforces negative stereotypes and reduces empathy toward individuals from lower socioeconomic backgrounds. Moreover, Busemeyer et al. [-@busemeyer_positive_2021] argues that meritocratic narratives can serve as feedback mechanisms that shape public opinion and well-being by framing individuals' understanding of welfare outcomes as deserved or undeserved within existing institutional structures. This psychological mechanism highlights the normative power of meritocracy in stabilizing unequal systems by shaping political attitudes and personal perceptions of success and failure.

Importantly, recent research has explored how meritocratic beliefs interact with experiences of intergenerational mobility to shape distributive attitudes. The belief that one’s success is earned can lead upwardly mobile individuals to internalize meritocratic narratives and justify existing inequalities, reinforcing support for market justice [@gugushvili_intergenerational_2016c; @molina_its_2019]. Conversely, downwardly mobile individuals who maintain strong meritocratic beliefs may interpret their status as a personal failure, reducing their support for redistribution [@day_movin_2017]. At the macro level, Shariff et al. [-@shariff_income_2016] show that higher perceived mobility increases tolerance for inequality, suggesting that meritocracy and mobility are mutually reinforcing.

Taken together, this literature supports the idea that meritocratic beliefs can moderate the relationship between mobility and market justice preferences. Individuals who experience mobility—especially upward—may draw on meritocratic narratives to legitimize both their own status and broader inequalities, thereby strengthening their support for market-based distribution. Accordingly, the second hypothesis of this study is:

$H2$: The positive (negative) relationship between upward (downward) social mobility and support for market justice in healthcare, pensions, and education is moderated by meritocratic beliefs; specifically, this association is stronger (weaker) among individuals with higher perceptions of meritocracy.

## The Chilean context

Chile offers a compelling case for examining individual preferences for market justice in the access to core social services. Since the early 1990s, the country experienced a long period of sustained economic growth, with per capita GDP increasing by nearly 4% annually until the mid-2010s [@llorca-jana_historia_2021]. This growth brought notable improvements in living standards, including a strong decline in income poverty—from 38.6% in 1990 to 6.5% in 2022 [@mideplan_evolucion_2017]. However, this success story has been tempered by growing concerns over the unequal distribution of its benefits. In recent years, economic growth has slowed—exacerbated by the global impact of the COVID-19 pandemic—and structural inequalities have become more visible [@barozet_clases_2021]. Despite its economic achievements, Chile remains one of the most unequal countries in Latin America and the OCDE. The poorest 50% of the population captures just 10% of total income and holds negative net wealth, while the wealthiest 1% receives nearly 27% of all income and controls almost half (49.6%) of the country’s wealth [@chancel_world_2022], a figure that has seen little change since the 1990s [@flores_top_2020]. These stark disparities extend beyond income, reflecting deep inequalities in access to essential social services [@pnud_desiguales_2017].

A significant share of Chile’s inequality stems from neoliberal reforms that privatized and commodified key domains of social reproduction [@arrizabalo_milagro_1995; @ferre_welfare_2023]. Introduced during the military dictatorship (1973–1989) and expanded under democratic governments, these reforms embedded market logic into public services through concessions, subsidies, demand-side vouchers, and pro-private regulatory frameworks [@madariaga_three_2020; @boccardo_30_2020]. This "crowded-out" welfare model benefits higher-income groups, leaving lower-income individuals to rely on limited public options. Scholars argue that this neoliberal shift repurposed the state to create niches of capitalist accumulation -from water to education-, giving rise to a model of “public-service capitalism” heavily reliant on state funding [@boccardo_30_2020]. In healthcare, around 79% of the population is covered by the public insurer FONASA, while 15% are enrolled in private ISAPREs, which offer faster and higher-quality services to wealthier groups [@mideplan_estadisticas_2024]. The pension system is based on individual capitalization, with mandatory contributions managed by private administrators investing in the financial market, currently involving 11 million contributors through private fund administrators (AFPs), yet 27% of the labor force remains outside the system due to informality [@superintendenciadepensiones_estadisticas_2024]. Education is similarly stratified: only 30.6% of students attend fully public schools, while 54% are in voucher-funded private institutions and 9.3% in fully private schools that largely serve affluent families [@ministeriodeeducacion_resumen_2023]. This arrangement has created a structurally segmented system, where access and quality of services depend on ability to pay, reinforced by state subsidies to private providers [@pnud_desiguales_2017].

Over recent decades, Chile has undergone a significant transformation in its class structure, shaped by sustained economic growth, the expansion of the service sector, the massification of education, and major changes in the role of the state and social policy [@ruiz_chilenos_2014]. These changes contributed to the growth of the middle sectors, a phenomena labelled as “mesocratrization” of society [@espinoza_estratificacion_2013]. However, this expansion did not reduce structural inequalities nor facilitate upward mobility into elite positions [@espinoza_estratificacion_2013; @lopez-roldan_comparative_2021]. While Chile shows high absolute mobility, most flows occur between adjacent classes—particularly from working classes into lower service-class positions—while access to upper classes remains limited, indicating strong elite closure and reproduction [@espinoza_estratificacion_2013; @lopez-roldan_comparative_2021; @torche_unequal_2005; @espinoza_movilidad_2014; @perez-ahumada_clases_2019]. Though often described as “unequal but fluid”, recent evidence points to growing rigidity in relative mobility, especially after the economic slowdown, increased household debt, labor market precarity, and the COVID-19 crisis [@barozet_clases_2021]. However, the majority of the population identifies themselves as members of the middle class [@castillo_todos_2013]. These dynamics have fueled a legitimacy crisis: meritocracy is increasingly questioned, and many perceive that advancement depends more on informal networks than on effort or earned credentials [@barozet_clases_2021]. The combination of constrained mobility, persistent inequality, and concentrated opportunity has eroded expectations of upward mobility and intensified social discontent [@pnud_desiguales_2017].

Despite recurrent social unrest, Chile presents a paradoxical coexistence between strong conflict over inequality and widespread public legitimation of it. The October 2019 “social outburst”, a period of mass protests and severe political repression throughout the country,  crystallized discontent over the privatization and commodification of social services, alongside a crisis of political legitimacy [@somma_no_2021]. A survey during the protests identified pensions, healthcare, and education as the top demands [@nucleodesociologiacontingente_informe_2020], reflecting dissatisfaction with Chile’s stratified welfare regime. Yet, this unrest coexists with strong meritocratic beliefs and high tolerance for income inequality. Castillo [-@castillo_legitimacy_2011] shows that wage gaps between occupations are consistently justified, and the larger the perceived gap, the greater its legitimation, an effect amplified by Chile’s structural inequality. Moreover, perceptions of unfairness due to non-meritocratic factors often reinforce belief in effort and talent as legitimate bases for success [@castillo_multidimensional_2023], which are associated with lower perceived injustice [@castillo_meritocracia_2019]. Mac-Clure et al. [@mac-clure_justicia_2024] find that low and lower-middle status individuals are less likely than upper-status groups to view educational income differences as unfair. Qualitative research further reveals the internalization of the enterprising self [@mau_inequality_2015], understood as a way in which the subjectivization process arises in line with positioning the self in the market, generating its own market value and subjecting it to competition. Canales et al. [-@canalesceron_sujeto_2021] describes how families navigate school choice as investment decisions, while Panes [-@panes_criticas_2020] finds that many workers frame pension contributions as individual investments aimed at maximizing future returns. These representations suggest that market-based moral orientations are deeply embedded in Chilean subjectivities, offering fertile ground to study how support for market justice varies across welfare domains.

Although still limited, recent research on market justice preferences in Chile indicates that support for these beliefs has increased in recent years across domains such as healthcare, pensions, and education, and that both objective and subjective dimensions of inequality shape them. Castillo et al. [-@castillo_perceptions_2025] show that agreement with the notion that it is fair for higher-income individuals to access better services was relatively low in 2016, but rose significantly by 2023—especially in the pension domain. Otero and Mendoza [-@otero_power_2024] provide evidence that social class, measured using the EGP scheme, influences these preferences: individuals in lower or subordinate class positions are less supportive of market-based access to welfare services compared to those in higher or privileged positions. Additionally, greater diversity in social class networks is associated with lower support for market justice. This is particularly relevant in the Chilean context, where one of the main determinants of class-diverse networks is intergenerational social mobility, which exposes individuals to relationships with people from different class backgrounds [@otero_lives_2022]. Regarding subjective factors, Castillo et al. [@castillo_perceptions_2025; @castillo_socialization_2024] find that individuals who endorse meritocratic beliefs—such as the idea that effort and talent are rewarded in Chile—are more likely to support market justice across welfare domains.

# Data, variables and methods

## Data

This study draws on data from the Chilean Longitudinal Social Survey (ELSOC) of the Center for Social Conflict and Cohesion Studies (COES). The survey is an annual panel collected from 2016 to 2023, comprising two independent samples (original and refreshment), featuring permanent and rotating questionnaire modules. It evaluates how individuals think, feel, and behave regarding conflict and social cohesion in Chile. The sampling design is complex, probabilistic, clustered, multistage, and stratified by city size, targeting men and women aged 18--75 who habitually reside in private dwellings in urban areas spanning 40 cities (92 municipalities, 13 regions). This analysis focuses on 2023, the most recent wave, which includes 1,737 respondents in the original sample and 989 in the refreshment sample. Further details on sampling, attrition, and weighting can be found at [https://coes.cl/encuesta-panel/](https://coes.cl/encuesta-panel/), and the dataset is publicly available at [https://dataverse.harvard.edu/dataverse/elsoc](https://dataverse.harvard.edu/dataverse/elsoc).

## Variables

*Outcome variables*

**Market justice preferences**: The outcome variables in this study are market justice preferences. This construct is operationalized through three variables that address the degree of justification regarding whether access to social services in healthcare, pensions, and education should be income conditional. Specifically, the justification of inequality in healthcare is assessed by the question: "Is it fair in Chile that people with higher incomes can access better healthcare than people with lower incomes?" The same question is asked for pensions and education. In all cases, respondents indicate their level of agreement on a five-point Likert scale ranging from 1 ("strongly disagree") to 5 ("strongly agree"). Although these items theoretically reflect the same underlying concept of market justice [@castillo_socialization_2024; @lindh_public_2015], they are analyzed separately because healthcare, pensions, and education may prompt distinct response patterns [@lee_fairness_2023; @busemeyer_skills_2014; @immergut_it_2020].

*Independent variables*

**Social mobility**: Intergenerational social mobility is treated as an exposure indicating whether respondents occupy a different class position from their fathers. Following Breen and Ermisch's [-@breen_effects_2024] framework for estimating causal effects of mobility, the variable is constructed in two stages.

1.  Class assignment. Using the International Socio-Economic Index of Occupational Status (ISEI) derived from three-digit ISCO-08 codes, both paternal (origin) and respondent (destination) occupations are grouped into three strata:

    -   Upper (codes 100-299: managers, directors, professionals)
    -   Middle (codes 300-499: skilled white-collar workers)
    -   Lower (codes 500 and above: manual, semi-skilled, or unskilled workers)

2.  Propensity-score estimation. To model mobility as a treatment---i.e., a transition from origin *j* to destination *k*---I estimate multinomial logit propensity scores using covariates that influence an individual's likelihood of moving between strata: (a) father's educational level, (b) presence of both parents at age 15, (c) nationality, (d) age, (e) sex, and (f) ethnicity.

These propensity scores are subsequently employed to adjust for selection into mobility when assessing its effect on market justice preferences.

**Meritocracy**: Meritocratic perception is operationalized through two components: one addressing effort and another focusing on talent [@young_rise_1958]. The item used to gauge effort is: "In Chile, people are rewarded for their efforts," while the item for talent is: "In Chile, people are rewarded for their intelligence and skills." In both cases, respondents indicate their level of agreement on a five-point Likert scale, ranging from 1 ("strongly disagree") to 5 ("strongly agree").

*Controls*

Sociodemographic and attitudinal variables are included to control for potential composition effects in the population. In terms of sociodemographic characteristics, I incorporate per capita household income quantile, educational level (1=Primary or below, 2=Secondary, 3=Technical, 4=University or above), age (in years), and sex (1=Male, 2=Female), which have been shown to significantly influence market justice preferences [@castillo_socialization_2024; @lindh_public_2015]. Regarding attitudinal variables, I include political identification (1=Left, 2=Center, 3=Right, 4=No identification) and subjective social status (ranging from 1 to 10) because they may confound the relationship between market justice preferences and social mobility and meritocracy.

## Methods

Following Breen and Ermisch's [-@breen_effects_2024] strategy for estimating the causal effect of social mobility on market justice preferences, I employ linear regression models with inverse-probability weights (IPW) for mobility conditional on class of origin. This approach allows estimation of the average causal effect of moving from an origin class *j* to a destination class *k* by comparing individuals who actually reach *k* with their counterfactual outcome had they instead moved to an alternative destination *k′* [@breen_effects_2024]. The resulting estimand is the average treatment effect on the treated (ATT).

All the analyses will be conducted using R software.

# References

::: {#refs}
:::

::::