### 2025-09-01 presentations/coloquio_2025/borrador_mjp-mobility-coloquio.qmd
Up resultados V1

---
title: "Preferencias por la comodificación: el papel de la movilidad social intergeneracional y las creencias en la meritocracia en Chile"
format: clean-revealjs
html-math-method:
  method: mathjax
  url: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
author: 
  - name: Andreas Laffert
    orcid: 0000-0002-9008-2454
    email: alaffertt@estudiante.uc.cl 
    affiliations:
      - Instituto de Sociología, Pontificia Universidad Católica de Chile
date: "July 3, 2025"
logo: "images/logo_isuc.png"
---

## Example slide

### This is a subtitle

Here we have some text that may run over several lines of the slide frame,
depending on how long it is.

- first item 
    - A sub item

Next, we'll brief review some theme-specific components.

- Note that _all_ of the standard Reveal.js
[features](https://quarto.org/docs/presentations/revealjs/)
can be used with this theme, even if we don't highlight them here.

## Additional theme classes

### Some extra things you can do with the clean theme

Special classes for emphasis

- `.alert` class for default emphasis, e.g. [important note]{.alert}.
- `.fg` class for custom colour, e.g. [important note]{.fg style="--col: #e64173"}.
- `.bg` class for custom background, e.g. [important note]{.bg style="--col: #e64173"}.

Cross-references

- `.button` class provides a Beamer-like button, e.g.
[[Summary]{.button}](#sec-summary)

## Want more?

### See our longer demo slides

We've deliberarely kept this template lean, to get you up and running as fast as
possible.

We provide a separate
[demo template](https://github.com/grantmcdermott/quarto-revealjs-clean-demo),
with more examples for integrating code, tables, figures, etc.

  - See the live demo slides
  [here](https://grantmcdermott.com/quarto-revealjs-clean-demo/template.html).

## Summary {#sec-summary}

### A minimalist and elegant presentation theme

The Quarto reveal.js clean theme aims to be a minimalist and elegant presention
theme. Here are some options to get you started:

Add the theme to an existing project.

```{.bash}
quarto install extension grantmcdermott/quarto-revealjs-clean
```

... or, create a new project using this slide deck as a lean template.

```{.bash}
quarto use template grantmcdermott/quarto-revealjs-clean
```

... or, create a new project using the demo slide deck as a full template.

```{.bash}
quarto use template grantmcdermott/quarto-revealjs-clean-demo
```