### 2025-11-02 paper/paper.qmd

---
title: "Preferences for the commodification of pensions in Chile: the rol of intergenerational social mobility"
format:
  html: 
    code-fold: true
    html-math-method: katex
    number-sections: true
    code-link: true
    title-block-banner: true
    theme:
      - pretty.scss
  pdf: 
    pdf-engine: xelatex
    geometry: margin=2cm
    template-partials: 
      - title.tex
    keep-tex: true
    include-in-header:
      text: |
        \usepackage[noblocks]{authblk}
        \renewcommand*{\Authsep}{, }
        \renewcommand*{\Authand}{, }
        \renewcommand*{\Authands}{, }
        \renewcommand\Affilfont{\small}
    number-sections: true
editor: source
bibliography: "../input/bib/tesis-magister.bib"
csl: "../input/bib/apa6.csl"
author:
  - name: Andreas Laffert Tamayo
    orcid: 0000-0002-9008-2454
    email: alaffertt@estudiante.uc.cl 
    affiliations:
      - ref: 1
affiliations: 
  - id: 1
    name: Instituto de Sociología, Pontificia Universidad Católica de Chile
citeproc: true
abstract: | 
 My abstract
  \newline
 **Keywords**: pension commodification, social mobility, causal inference, Chile, public preferences
link-citations: true
linestretch: 1.5
mainfont: Times New Roman
fontsize: 13pt          
colorlinks: true
fig-height: 4
fig-width: 7.5
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo=FALSE, 
                      warning = FALSE,
                      message = FALSE, 
                      cache = TRUE,
                      out.width = '85%',
                      fig.pos= "H")
options(scipen=999)
options(kableExtra.auto_format = FALSE)
options(knitr.kable.NA = '')
options(knitr.graphics.error = FALSE)
Sys.setlocale("LC_ALL", "ES_ES.UTF-8")
table_format = if (knitr::is_html_output()) {
   #conditional instructions for kable
   "html" #if html set "html" in format
 } else if (knitr::is_latex_output()) {
   "latex"#if latex set "latex" in format
 }

library(conflicted)
conflicts_prefer(dplyr::filter)

```

This document was last modified at {{< modified >}}
and it was last rendered at {{< now >}}.

:::: {.justify}

{{< include 01-introduction.qmd >}}

{{< include 02-antecedents.qmd >}}

{{< include 03-methods.qmd >}}

{{< include 04-analysis.qmd >}}

{{< include 05-discussion.qmd >}}

{{< include 06-conclusion.qmd >}}

# References

::: {#refs}
:::
::::

{{< include 07-supplementary-material.qmd >}}