### 2025-11-02 paper/paper.tex

% Options for packages loaded elsewhere
% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  13pt,
]{article}
\usepackage{xcolor}
\usepackage[margin=2cm]{geometry}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{5}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
  \setmainfont[]{Times New Roman}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{setspace}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\setlength{\emergencystretch}{3em} % prevent overfull lines

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage[noblocks]{authblk}
\renewcommand*{\Authsep}{, }
\renewcommand*{\Authand}{, }
\renewcommand*{\Authands}{, }
\renewcommand\Affilfont{\small}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Preferences for the commodification of pensions in Chile: the rol of intergenerational social mobility},
  pdfauthor={Andreas Laffert Tamayo},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Preferences for the commodification of pensions in Chile: the rol
of intergenerational social mobility}

  \author{Andreas Laffert Tamayo}
            \affil{%
                  Instituto de Sociología, Pontificia Universidad
                  Católica de Chile
              }

\date{}

% Colores extra por nombre
\usepackage[dvipsnames]{xcolor}
% Ajustes de hipervínculos
% \hypersetup{
%   colorlinks = true,
%   citecolor  = DarkSlateBlue,     % << citas
%   urlcolor   = DarkSlateBlue  % URLs/DOIs
% }
% 
\begin{document}
\maketitle
\begin{abstract}
My abstract \newline \textbf{Keywords}: pension commodification, social
mobility, causal inference, Chile, public preferences
\end{abstract}

\setstretch{1.5}
This document was last modified at 2025-11-02 05:55:15 and it was last
rendered at 2025-11-02 05:55:15.

\section{Introduction}\label{introduction}

What is the legitimate extent of market inequality in the eyes of the
public? Since the early 1980s, many countries have experienced a
widespread retreat from universal welfare programs and a shift toward
the privatization and commodification of public goods, welfare policies,
and social services (\citeproc{ref-gingrich_making_2011}{Gingrich,
2011}; \citeproc{ref-streeck_how_2016}{Streeck, 2016}). In Latin
America, as elsewhere, neoliberal reforms reshaped welfare-state
institutions by extending market logic into domains of social
reproduction that were traditionally governed by the state
(\citeproc{ref-ferre_welfare_2023}{Ferre, 2023}). This transformation
reduced the role of public provision and increased the presence of
private actors in core social services
(\citeproc{ref-ferre_welfare_2023}{Ferre, 2023}). Echoing Polanyi's
(\citeproc{ref-polanyi_great_1975}{1975}) insight that markets
constitute a distinct moral order, the institutional diffusion of market
rules has fostered a corresponding moral economy: a constellation of
norms and values concerning fair allocation, embedded in institutions
and shaping individual subjectivities
(\citeproc{ref-mau_inequality_2015}{Mau, 2015};
\citeproc{ref-svallforsMoralEconomyClass2006a}{Svallfors, 2006}). Within
this framework, a growing body of research examines the extent to which,
and the mechanisms by which, citizens consider it fair that the
allocation of services like health care, pensions, and education be
governed by market-based criteria---a phenomenon known as \emph{market
justice preferences} (\citeproc{ref-busemeyer_skills_2014}{Busemeyer,
2014}; \citeproc{ref-castillo_perceptions_2025}{Castillo, Laffert, et
al., 2025}; \citeproc{ref-castillo_socialization_2024}{Castillo et al.,
2024}; \citeproc{ref-immergut_it_2020}{Immergut \& Schneider, 2020};
\citeproc{ref-koos_moral_2019}{Koos \& Sachweh, 2019};
\citeproc{ref-lindh_public_2015}{Lindh, 2015};
\citeproc{ref-lindh_bringing_2023}{Lindh \& McCall, 2023}).
Understanding these preferences is crucial, as they contribute to
legitimizing economic inequality by framing it as the fair result of
individual responsibility and limited state intervention
(\citeproc{ref-mau_inequality_2015}{Mau, 2015}).

Existing literature shows that market justice preferences are shaped by
both the economic and institutional context and individuals' positions
within social stratification. Grounded in the notion that economic
institutions influence people's normative attitudes
(\citeproc{ref-immergut_theoretical_1998}{Immergut, 1998}), studies find
that countries with stronger public provision or more expansive welfare
states exhibit lower levels of market justice preferences
(\citeproc{ref-busemeyer_skills_2014}{Busemeyer, 2014};
\citeproc{ref-immergut_it_2020}{Immergut \& Schneider, 2020}), while
more privatized contexts show stronger support for market-based criteria
(\citeproc{ref-castillo_socialization_2024}{Castillo et al., 2024};
\citeproc{ref-lindh_public_2015}{Lindh, 2015}). In such contexts, market
justice preferences tend to rise as individuals ``ascend'' the social
structure, with those in more privileged positions in terms of class,
education, and income holding stronger preferences for market-based
solutions compared to those in more disadvantaged or at-risk positions
(\citeproc{ref-castillo_socialization_2024}{Castillo et al., 2024};
\citeproc{ref-immergut_it_2020}{Immergut \& Schneider, 2020};
\citeproc{ref-lee_fairness_2023}{Lee \& Stacey, 2023};
\citeproc{ref-lindh_public_2015}{Lindh, 2015};
\citeproc{ref-otero_power_2024}{Otero \& Mendoza, 2024};
\citeproc{ref-svallfors_political_2007}{Svallfors, 2007};
\citeproc{ref-vondemknesebeck_are_2016}{Von Dem Knesebeck et al.,
2016}).

Market justice preferences are shaped not only by objective
socioeconomic conditions but also by popular beliefs about inequality.
Among these, meritocracy is a key normative principle underpinning
market-based distributive preferences
(\citeproc{ref-mau_inequality_2015}{Mau, 2015}). Studies show that
individuals with stronger meritocratic beliefs tend to perceive less
inequality and legitimize it by attributing economic differences to
personal achievement (\citeproc{ref-batruch_belief_2023}{Batruch et al.,
2023}; \citeproc{ref-mijs_paradox_2019}{Mijs, 2019};
\citeproc{ref-wilson_role_2003}{Wilson, 2003}). In highly unequal
societies where access to services is largely governed by market logic,
such beliefs play a critical role in normalizing inequality. Recent
evidence from Chile shows that students who believe effort and talent
are rewarded in their country express stronger preferences for
market-based access to healthcare, pensions, and education
(\citeproc{ref-castillo_socialization_2024}{Castillo et al., 2024}).

Although it is clear that one's social position influences market
justice preferences, the question of how upward or downward mobility
within the social structure affects these preferences remains
unanswered. This question is far from trivial, especially in Latin
America, where many have experienced various forms of mobility amid high
economic inequality and deep welfare privatization
(\citeproc{ref-ferre_welfare_2023}{Ferre, 2023};
\citeproc{ref-lopez-roldan_comparative_2021}{López-Roldán \& Fachelli,
2021}; \citeproc{ref-torche_intergenerational_2014}{Torche, 2014}).
Social origins and destinations affect attitudes toward inequality in
distinct ways (\citeproc{ref-day_movin_2017}{Day \& Fiske, 2017};
\citeproc{ref-gugushvili_intergenerational_2016}{Gugushvili, 2016b},
\citeproc{ref-gugushvili_subjective_2017}{2017};
\citeproc{ref-jaime-castillo_social_2019}{Jaime-Castillo \&
Marqués-Perales, 2019}; \citeproc{ref-mijs_belief_2022}{Mijs et al.,
2022}; \citeproc{ref-wen_does_2021}{Wen \& Witteveen, 2021}), while
movement between these positions exposes individuals to different
experiences and mechanisms that shape their views on what is fair
(\citeproc{ref-gugushvili_trends_2014}{Gugushvili, 2014};
\citeproc{ref-mau_inequality_2015}{Mau, 2015}). Building on this
research, examining the effects of social mobility on market justice
preferences can help to illuminate how inequalities in access to social
services are justified among individuals who have experienced, or not,
changes in their social standing, and what are the normative mechanisms
that guide this justification (\citeproc{ref-mau_inequality_2015}{Mau,
2015}).

Beyond their isolated effects, social mobility and meritocratic beliefs
interact in complex ways to shape market justice preferences. Among
others, a key mechanism proposed in the literature to explain how
mobility influences distributive justice preferences is the
psychological process of self-serving bias in causal attribution
(\citeproc{ref-gugushvili_intergenerational_2016c}{Gugushvili, 2016a};
\citeproc{ref-schmidt_experience_2011}{Schmidt, 2011}). This bias
suggests that individuals attribute failures---such as downward
mobility---to external factors, while crediting successes---such as
upward mobility---to their own merit and effort
(\citeproc{ref-miller_selfserving_1975}{Miller \& Ross, 1975}). Those
who experience upward mobility tend to view their social position as
earned, making them more likely to believe that individuals are
responsible for their own success or failure. Research shows that upward
mobility is associated with weaker preferences for redistribution
(\citeproc{ref-alesina_intergenerational_2018}{Alesina et al., 2018};
\citeproc{ref-gugushvili_intergenerational_2016c}{Gugushvili, 2016a};
\citeproc{ref-jaime-castillo_social_2019}{Jaime-Castillo \&
Marqués-Perales, 2019}; \citeproc{ref-schmidt_experience_2011}{Schmidt,
2011}) and stronger legitimacy of income inequality
(\citeproc{ref-shariff_income_2016}{Shariff et al., 2016}). In contrast,
individuals who experience downward mobility tend to blame structural
factors like inequality and are more supportive of redistribution while
rejecting merit-based explanations
(\citeproc{ref-gugushvili_trends_2014}{Gugushvili, 2014}). Taken
together, I argue that meritocratic beliefs may reinforce this
self-serving attribution mechanism by legitimizing one's social status
as the outcome of personal merit, closely tied to attribution bias.

Against this background, this article pursues two main objectives:
first, to analyze the extent to which intergenerational social mobility
influences market justice preferences regarding healthcare, pensions,
and education; and second, to examine how meritocratic beliefs may
moderate this relationship. Building on a theoretical framework that
emphasizes how neoliberal transformations---particularly through the
privatization and commodification of key areas of social
reproduction---have profoundly reshaped processes of subject formation
(\citeproc{ref-mau_inequality_2015}{Mau, 2015}), the central argument is
that upward mobility increases support for market justice preferences,
while downward mobility decreases it. Moreover, meritocratic beliefs are
expected to moderate this relationship by reflecting a self-serving
attribution bias, whereby individuals justify their social position in
terms of personal merit.

This study focuses on Chile, a particularly intriguing case for
examining market justice preferences. Despite significant economic
growth and poverty reduction, Chile has some of the highest levels of
inequality in Latin America and among OECD countries
(\citeproc{ref-chancel_world_2022}{Chancel et al., 2022};
\citeproc{ref-flores_top_2020}{Flores et al., 2020}). This inequality
coexists with short-range upward mobility among lower-class segments
moving into middle strata, though strong barriers remain to reaching
higher positions (\citeproc{ref-espinoza_estratificacion_2013}{Espinoza
et al., 2013}; \citeproc{ref-torche_intergenerational_2014}{Torche,
2014}). What makes Chile especially salient is that much of this
inequality is rooted in deep neoliberal reforms that institutionalized
the privatization and commodification of key social sectors
(\citeproc{ref-madariaga_three_2020}{Madariaga, 2020}). Introduced
during the dictatorship (1973--1989) and expanded in democracy, these
reforms enabled the unprecedented emergence of markets in health,
pensions, and education, with provision segmented by individuals'
ability to pay and supported by public subsidies
(\citeproc{ref-boccardo_30_2020}{Boccardo, 2020}). In parallel---and
despite waves of protest against inequality and commodification from
2006 to 2019 (\citeproc{ref-somma_no_2021}{Somma et al.,
2021})---Chilean subjectivities have been increasingly shaped by
neoliberal discourses and market logics, influencing their attitudes
toward inequality and welfare distribution
(\citeproc{ref-araujo_desafios_2012}{Araujo \& Martuccelli, 2012};
\citeproc{ref-canalesceron_sujeto_2021}{Canales Cerón et al., 2021}).

In this context, the questions that guide this research are as follows:

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\tightlist
\item
  To what extent does intergenerational social mobility influence market
  justice preferences regarding healthcare, pensions, and education in
  Chile?
\item
  How do meritocratic beliefs condition or moderate this relationship in
  the Chilean context?
\end{enumerate}

To address these questions, this study draws on large-scale,
representative survey data collected in 2018 from the urban Chilean
population aged 18 to 75 (n = 2,726). The next section outlines the
theoretical framework linking market justice preferences, social
mobility, and meritocratic beliefs, and proposes a set of hypotheses.
This is followed by a description of the data, variables, and analytical
strategy. The final sections present the empirical findings, offer an
interpretation of the results, and conclude with a discussion of their
implications.

\section{Theoretical and empirical
background}\label{theoretical-and-empirical-background}

\subsection{Market justice
preferences}\label{market-justice-preferences}

Beyond state's capacity to reallocate resources from the advantaged to
the vulnerable, market institutions also play a central role in
distributing socially valuable goods and rewards
(\citeproc{ref-koos_moral_2019}{Koos \& Sachweh, 2019};
\citeproc{ref-lindh_bringing_2023}{Lindh \& McCall, 2023}). As Polanyi
(\citeproc{ref-polanyi_great_1975}{1975}) observed, economic integration
in capitalist societies is primarily organized through market exchange,
governed by a self-regulating price system embedded in institutional
frameworks. These institutions are not mere aggregates of individual
behavior, but social realities endowed with rules, mechanisms, and
normative meanings that shape everyday thinking
(\citeproc{ref-immergut_theoretical_1998}{Immergut, 1998};
\citeproc{ref-koos_moral_2019}{Koos \& Sachweh, 2019}). In this sense,
the economic order is mirrored in a moral economy: collectively shared
norms and beliefs about justice in distribution, embedded and reinforced
through institutions (\citeproc{ref-mau_inequality_2015}{Mau, 2015};
\citeproc{ref-svallfors_political_2007}{Svallfors, 2007}). While most
research from this perspective has focused on welfare institutions,
recent scholarship has brought the market back into focus as a site of
distributive justice beliefs and institutional responses to inequality
(\citeproc{ref-lindh_bringing_2023}{Lindh \& McCall, 2023}). In many
countries, privatization and commodification have expanded market logic
into core areas of social reproduction, such as healthcare, education,
pensions, and social security, deepening inequality in access to these
services (\citeproc{ref-ferre_welfare_2023}{Ferre, 2023};
\citeproc{ref-gingrich_making_2011}{Gingrich, 2011}). Yet public support
for market-based welfare provision has grown, even in traditional
welfare states, particularly among higher-income groups who view private
alternatives as more efficient or higher in quality
(\citeproc{ref-busemeyer_welfare_2020}{Busemeyer \& Iversen, 2020}).
This shift calls for broader inquiry into how institutions like markets
structure access to resources and legitimize inequalities
(\citeproc{ref-mau_inequality_2015}{Mau, 2015};
\citeproc{ref-satz_por_2019}{Satz, 2019}).

The legitimacy of market-based inequalities is closely tied to beliefs
about distributive justice grounded in market principles. Beyond
stratification structures and socioeconomic conditions, the study of
inequality also focuses on individuals' beliefs about the origins of
inequality, the normative frameworks that sustain those beliefs, the
mechanisms that shape them, and their implications for attitudes and
behavior (\citeproc{ref-kluegel_beliefs_1981}{Kluegel \& Smith, 1981}).
In this regard, empirical research on distributive justice focuses on
individuals' conceptions of how goods and rewards \emph{should} be
distributed in society (\citeproc{ref-jasso_distributive_2016}{Jasso et
al., 2016}). This line of inquiry allows for examining the extent to
which economic inequality is perceived as just, or in a certain sense,
legitimate (\citeproc{ref-castillo_legitimacy_2011}{Castillo, 2011}).
Most of the research on this field explores the legitimacy of wage
inequality, specially salary gaps between jobs at opposite ends of the
occupational hierarchy
(\citeproc{ref-castillo_legitimacy_2011}{Castillo, 2011};
\citeproc{ref-jasso_justice_1978}{Jasso, 1978};
\citeproc{ref-jasso_gender_1999}{Jasso \& Wegener, 1999}). Recently,
others examines how individuals justify market-generated inequalities in
access to core social services such as healthcare, education, pensions
or social security. Here, legitimacy stems from the belief that access
to these goods should follow market-based criteria
(\citeproc{ref-castillo_socialization_2024}{Castillo et al., 2024};
\citeproc{ref-lindh_public_2015}{Lindh, 2015}). In such views, these
services are treated as legitimate commodities; goods that can be
traded, priced, and evaluated through market logic
(\citeproc{ref-busemeyer_welfare_2020}{Busemeyer \& Iversen, 2020}).

Market justice preferences refer to normative beliefs that legitimize
the idea that access to core social services---such as healthcare,
education, or pensions---should be determined by market-based criteria.
Following Janmaat's (\citeproc{ref-janmaat_subjective_2013}{2013, p.
359}) distinction, these preferences fall under the category of
``beliefs,'' understood as normative ideas about what inequality should
look like, as opposed to ``perceptions,'' which refer to subjective
evaluations of existing inequality. Market justice preferences reflect
the view that access to these services should depend on individuals'
ability to pay, thus justifying inequalities generated by market
mechanisms (\citeproc{ref-kluegel_legitimation_1999}{Kluegel et al.,
1999}; \citeproc{ref-lindh_public_2015}{Lindh, 2015}). The concept draws
on Lane's (\citeproc{ref-lane_market_1986}{1986}) classic contrast
between market justice and political justice: the former is grounded in
the idea of earned deserts---where rewards reflect effort, productivity,
and skill---while the latter prioritizes need and equality, typically
expressed in welfare state policies. Lane argued that markets and states
differ in purpose (efficiency vs.~need), logic (individual
vs.~collective), and fairness criteria (merit vs.~equality). Market
justice assumes that markets are neutral, self-regulating systems in
which fair procedures yield outcomes proportional to merit. Inequality,
from this perspective, is not only expected but legitimate---as long as
it arises from fair competition. In this way, market justice offers a
moral lens through which individuals can view the commodification and
stratified access to social services as fair and justified
(\citeproc{ref-kluegel_legitimation_1999}{Kluegel et al., 1999};
\citeproc{ref-lindh_public_2015}{Lindh, 2015}).

The empirical study of market justice preferences has employed various
strategies to capture how individuals assess inequalities arising from
market allocation. A common approach evaluates whether people find it
fair that access to essential social services---such as healthcare,
education, or pensions---depends on income. This builds on foundational
work by Kluegel and Smith
(\citeproc{ref-kluegel_legitimation_1999}{1999}), who examined normative
justifications for capitalist systems. Subsequent studies have extended
this logic to a wider set of welfare goods. For example, Immergut and
Schneider (\citeproc{ref-immergut_it_2020}{2020}) and von dem Knesebeck
et al. (\citeproc{ref-vondemknesebeck_are_2016}{2016}) explore whether
respondents believe it is just that higher-income individuals receive
better healthcare. Similarly, Lee and Stacey
(\citeproc{ref-lee_fairness_2023}{2023}) and Castillo, Iturra, et al.
(\citeproc{ref-castillo_changes_2025}{2025}) apply this framework to
education. Complementing these efforts, more recent research has
introduced composite indicators to capture broader orientations toward
market-based allocation. Lindh (\citeproc{ref-lindh_public_2015}{2015}),
for instance, constructs an index averaging support for income-based
access to healthcare and education in comparative perspective, while
Castillo et al. (\citeproc{ref-castillo_socialization_2024}{2024})
propose a single-item measure covering health, education, and pensions
in Chile. These instruments seek to gauge to what extent individuals
view market-generated inequalities as legitimate. Taken together, they
capture two core dimensions of market distribution: the role of economic
resources as a key determinant of outcomes, and the framing of services
as tradable commodities that can be bought and sold according to ability
to pay (\citeproc{ref-lindh_public_2015}{Lindh, 2015}).

Comparative empirical research has identified several individual-level
factors that influence support for market justice. Individuals in more
advantaged socioeconomic positions---those with higher income,
education, and occupational status---are consistently more likely to
endorse market-based distributive principles
(\citeproc{ref-koos_moral_2019}{Koos \& Sachweh, 2019};
\citeproc{ref-lindh_public_2015}{Lindh, 2015};
\citeproc{ref-svallfors_political_2007}{Svallfors, 2007}). For example,
Lindh (\citeproc{ref-lindh_public_2015}{2015}) finds that individuals
from the service class are more likely to support market-based access to
healthcare and education than skilled and unskilled workers across 17
relatively affluent countries. In a comparative analysis, Svallfors
(\citeproc{ref-svallfors_political_2007}{2007}) observes that this
expected class pattern appears clearly only in Sweden, where support for
private education and healthcare varies systematically by class.
Busemeyer (\citeproc{ref-busemeyer_skills_2014}{2014}) similarly shows
that support for private education is stronger among high-income groups,
while Immergut and Schneider (\citeproc{ref-immergut_it_2020}{2020}) and
von dem Knesebeck et al. (\citeproc{ref-vondemknesebeck_are_2016}{2016})
report comparable findings for healthcare, suggesting that wealthier
individuals perceive private provision as a means of maintaining
relative advantage. In Chile, Otero and Mendoza
(\citeproc{ref-otero_power_2024}{2024}) show that individuals with
higher income and university education express stronger support for
market allocation in healthcare, education, and pensions. Beliefs about
inequality and political orientation also matter. Castillo et al.
(\citeproc{ref-castillo_socialization_2024}{2024}) and Castillo,
Laffert, et al. (\citeproc{ref-castillo_perceptions_2025}{2025}) show
that individuals who have strong meritocratic beliefs are more likely to
support market-based distribution in Chile, while Lee and Stacy
(\citeproc{ref-lee_fairness_2023}{2023}) in Australia suggest that these
preferences are also greater as people lean toward economic
conservatism. In this sense, market justice preferences are shaped by
the interplay between structural position and normative reasoning.

However, individual characteristics alone do not fully explain variation
in support for market justice. A growing body of cross-national research
highlights the importance of institutional arrangements in shaping these
preferences. For instance, Immergut and Schneider
(\citeproc{ref-immergut_it_2020}{2020}) find that in countries with
higher public spending on healthcare, individuals are less likely to
view income-based access as fair. Similarly, Busemeyer
(\citeproc{ref-busemeyer_skills_2014}{2014}) shows that increased public
investment in education is associated with lower support for privatized
provision. Conversely, Lindh (\citeproc{ref-lindh_public_2015}{2015})
finds that in countries with more market-oriented welfare systems,
support for market-based distribution tends to be higher, suggesting
that individual attitudes often align with institutional outputs. These
findings are consistent with neo-institutionalist and policy feedback
theories (\citeproc{ref-campbell_institutional_2020}{Campbell, 2020}),
which argue that institutions do more than redistribute resources---they
also shape the normative categories through which individuals assess who
is deserving of support
(\citeproc{ref-immergut_theoretical_1998}{Immergut, 1998}). In this
view, dominant values and preferences are both shaped by and embedded in
institutional configurations
(\citeproc{ref-busemeyer_skills_2014}{Busemeyer, 2014}), reinforcing the
notion that institutions are not neutral structures, but active
producers of the moral frameworks that legitimize or challenge
inequality.

\subsection{Social mobility}\label{social-mobility}

The study of social mobility, its drivers and consequences has long been
central to sociology. Classic theorists explored not only movement
across social hierarchies but also its broader implications for class
conflict, norm stability, and institutional change
(\citeproc{ref-breen_effects_2024}{Breen \& Ermisch, 2024}). Sorokin
(\citeproc{ref-sorokin_social_1927}{1927}) introduced the concept
formally, defining mobility as the shift of individuals, values, or
objects between positions within a stratification system, and
distinguishing between horizontal and vertical forms. Later work
differentiated intergenerational from intragenerational mobility, as
well as absolute mobility---driven by structural change---from relative
mobility, which captures the extent to which origins constrain
destinations (\citeproc{ref-eyles_social_2022}{Eyles et al., 2022}). In
Latin America, research shows high absolute but low relative mobility
(\citeproc{ref-bucca_merit_2016}{Bucca, 2016}): although educational
expansion and economic modernization have enabled some upward movement,
status reproduction remains strong, especially among elites
(\citeproc{ref-lopez-roldan_comparative_2021}{López-Roldán \& Fachelli,
2021}; \citeproc{ref-torche_intergenerational_2014}{Torche, 2014}). This
reflects deep structural inequality, segmented education systems,
stratified labor markets, and legacies of dependent development that
restrict access to mobility channels and reinforce the intergenerational
transmission of advantage.

Beyond mapping mobility patterns, growing research has examined its
subjective and attitudinal effects. Mobility effects---defined as
outcomes resulting from movement between origin and destination classes
(\citeproc{ref-breen_effects_2024}{Breen \& Ermisch, 2024})---have long
attracted theoretical interest. Sorokin's (1959) dissociative hypothesis
posits that mobility, whether upward or downward, may produce
psychological strain due to conflicting norms between class contexts,
leading to lower life satisfaction. Similarly, Lenski
(\citeproc{ref-lenski_status_1954}{1954}) argued that status
inconsistency---mismatches among education, income, and occupation---can
undermine well-being. These ideas underpin extensive empirical work
linking intergenerational mobility to outcomes such as life
satisfaction, mental health, and stress
(\citeproc{ref-gugushvili_heterogeneous_2024}{Gugushvili, 2024};
\citeproc{ref-hadjar_does_2015}{Hadjar \& Samuel, 2015};
\citeproc{ref-prag_subjective_2021}{Präg \& Gugushvili, 2021}).

Research on intergenerational social mobility has increasingly examined
its effects on attitudes toward economic inequality. A key area of
inquiry focuses on how upward and downward mobility influence support
for redistribution, though findings remain mixed. Alesina et al.
(\citeproc{ref-alesina_intergenerational_2018}{2018}) show that
individuals with pessimistic expectations about their
mobility---particularly those anticipating downward movement---are more
likely to support generous redistributive policies. Similarly, Ares
(\citeproc{ref-ares_changing_2020}{2020}) finds that upwardly mobile
individuals tend to be less supportive of state-led redistribution
compared to those who have experienced downward mobility. Comparative
studies by Schmidt (\citeproc{ref-schmidt_experience_2011}{2011}) and
Gugushvili (\citeproc{ref-gugushvili_subjective_2017}{2017}) likewise
report that subjective upward mobility is associated with weaker
preferences for redistribution, while downward mobility strengthens
redistributive support. However, recent causal evidence from Breen and
Ermisch (\citeproc{ref-breen_effects_2024}{2024}) suggests the opposite:
upward mobility may increase redistributive preferences, while downward
mobility may reduce them.

Beyond redistribution, other studies have explored the impact of
mobility on broader beliefs about inequality. Gugushvili
(\citeproc{ref-gugushvili_intergenerational_2016}{2016b}) finds that
upwardly mobile individuals are more likely to adopt individualistic
attributions of poverty and to legitimize income inequality,
particularly in post-socialist societies. Similarly, Bucca
(\citeproc{ref-bucca_merit_2016}{2016}) shows that subjective upward
mobility reinforces individualistic explanations of wealth in seven
Latin American countries. At the macro level, Shariff et al.
(\citeproc{ref-shariff_income_2016}{2016}) demonstrate that higher
levels of national economic mobility correlate with greater tolerance of
inequality. In contrast, Day and Fiske
(\citeproc{ref-day_movin_2017}{2017}) find that low perceived mobility
undermines belief in meritocracy and a just world, thereby weakening
system justification. Taken together, this body of research suggests
that mobility shapes attitudes toward inequality and justice through
multiple, and sometimes contradictory, mechanisms.

The literature on the effects of social mobility has proposed various
mechanisms to explain how and why changes in social position may
influence individual outcomes
(\citeproc{ref-helgason_class_2025}{Helgason \& Rehm, 2025}). One of the
most prominent is the self-interest mechanism, which posits that
individuals who experience upward or downward mobility undergo a shift
in their material interests, thereby altering their perceptions and
preferences (\citeproc{ref-ares_changing_2020}{Ares, 2020};
\citeproc{ref-helgason_longterm_2023}{Helgason \& Rehm, 2023};
\citeproc{ref-langsaether_explaining_2022}{Langsæther et al., 2022}).
Closely related to this logic is the Prospect of Upward Mobility (POUM)
hypothesis, which suggests that individuals may oppose redistribution
not because of their current position, but because they anticipate
improving their status in the future
(\citeproc{ref-benabou_social_2001}{Benabou \& Ok, 2001}). A second line
of explanation draws on the framework of theories of socialization.
Within this tradition, hypotheses such as acculturation, socialization,
and status maximization propose that individuals adjust their attitudes
based on the norms and values of either their class of origin or their
destination, or a combination of both
(\citeproc{ref-jaime-castillo_social_2019}{Jaime-Castillo \&
Marqués-Perales, 2019}). However, most of these mechanisms focus
primarily on the indirect effects of origin and destination positions,
rather than on the direct effect of experience of movement itself. In
response to this gap, recent research has highlighted the role of
self-serving bias in causal attribution processes, suggesting that
individuals tend to explain their mobility trajectories in ways that
justify their current position, which in turn shapes their beliefs and
preferences
(\citeproc{ref-gugushvili_intergenerational_2016c}{Gugushvili, 2016a};
\citeproc{ref-molina_its_2019}{Molina et al., 2019};
\citeproc{ref-schmidt_experience_2011}{Schmidt, 2011}).

The self-serving bias mechanism builds on a value-oriented perspective,
emphasizing that individuals' experiences of social mobility shape their
causal attributions, which in turn influence their beliefs about justice
and distributive preferences
(\citeproc{ref-gugushvili_trends_2014}{Gugushvili, 2014}). Causal
attribution refers to the process through which individuals generate
explanations for their own behavior and outcomes, as well as those of
others (\citeproc{ref-gugushvili_intergenerational_2016}{Gugushvili,
2016b}). In this view, people's interpretations of economic inequality
depend on whether they believe such disparities reflect unequal
individual contributions. Individuals who adopt an internal attribution
framework tend to see success or failure as rooted in personal
characteristics such as effort, talent, or merit. In contrast, those who
regard inequality as unjust are more likely to adopt an external
attribution model, viewing outcomes as the result of structural barriers
beyond individual control (\citeproc{ref-kluegel_beliefs_1981}{Kluegel
\& Smith, 1981}). This mechanism, often described as intrapersonal
causal attribution, reflects how people explain their own socioeconomic
positions---typically attributing their successes to internal qualities
while blaming failures on external circumstances
(\citeproc{ref-miller_selfserving_1975}{Miller \& Ross, 1975}). Over
time, individuals may revise their beliefs and attitudes: while early
views are shaped by their social origin, these are later adjusted in
light of personal experiences of mobility and the perceived role of
ascribed versus achieved factors in determining socioeconomic outcomes
(\citeproc{ref-gugushvili_intergenerational_2016}{Gugushvili, 2016b}).

Empirical research has provided support for the self-serving bias
mechanism in shaping redistributive preferences and attitudes toward the
legitimacy of inequality. For instance, Schmidt
(\citeproc{ref-schmidt_experience_2011}{2011}) finds that individuals
who experience upward mobility are more likely to interpret their
success as the result of personal effort or merit, and consequently
perceive redistribution as less necessary. Conversely, individuals who
experience downward mobility tend to attribute their decline to external
circumstances---such as structural inequality or unemployment---and show
stronger support for redistribution. In a comparative analysis across
different welfare domains, Gugushvili
(\citeproc{ref-gugushvili_subjective_2017}{2017}) demonstrates that
upward mobility is associated with lower support for government spending
on housing and pensions, while individuals who experience downward
mobility express lower support for healthcare and education spending,
but favor increased investment in housing and pensions, reflecting the
material nature of these domains. Moreover, Gugushvili
(\citeproc{ref-gugushvili_intergenerational_2016c}{2016a}) finds that
upward mobility is linked to greater justification of income inequality,
suggesting that improved social standing reinforces an attributional
view in which success is seen as the result of individual
characteristics---thus legitimizing inequality as a fair outcome.

Consequently, considering this theoretical and empirical background, the
first hypothesis of this research is that:

\(H1\): Experiencing upward (downward) social mobility is positively
(negatively) associated with greater support for market justice in
healthcare, pensions, and education.

\subsection{Meritocracy}\label{meritocracy}

Meritocracy constitutes a central ideological framework for legitimizing
different types of social inequality, for instance through market
justice beliefs. Rooted in the belief that rewards and positions should
be allocated based on individual effort and talent, meritocracy operates
as a normative ideal and a descriptive belief about how society
functions. As initially conceptualized by Michael Young
(\citeproc{ref-young_rise_1958}{1958}), the term was meant to critique a
system in which merit-based stratification becomes a new form of
inequality. However, over time, meritocracy has been widely supported in
many societies as a fair and desirable principle of distribution,
particularly within liberal democracies and market-oriented economies
(\citeproc{ref-mijs_paradox_2019}{Mijs, 2019};
\citeproc{ref-sandel_tyranny_2020}{Sandel, 2020}). From a sociological
perspective, the belief in meritocracy is more than a cognitive
assessment; it reflects a moral lens through which individuals interpret
inequality. People who believe that success results from hard work and
talent are more likely to view social and economic disparities as
legitimate (\citeproc{ref-batruch_belief_2023}{Batruch et al., 2023};
\citeproc{ref-castillo_meritocracia_2019}{Castillo et al., 2019}).
Conversely, if they see outcomes as driven by luck, social origin, or
systemic barriers, inequality is more likely to be perceived as unjust.
This distinction becomes crucial in societies with persistent structural
inequality, where public narratives often emphasize personal
responsibility and merit while overlooking entrenched disadvantages.

I adopt a multidimensional perspective on meritocracy, distinguishing
between two key dimensions: effort-based and talent-based perceptions.
This distinction is essential, as it captures different pathways through
which individuals justify inequality
(\citeproc{ref-young_rise_1958}{Young, 1958}). Effort-based meritocracy
emphasizes hard work and perseverance as the basis for social rewards,
aligning closely with cultural narratives of personal responsibility. A
talent-based meritocracy, by contrast, emphasizes intelligence and
innate abilities, which are often perceived as less malleable and more
unequally distributed. Both dimensions have been shown to correlate with
acceptance of inequality, but they may carry distinct implications for
how inequality is justified in specific domains
(\citeproc{ref-castillo_multidimensional_2023}{Castillo et al., 2023}).
The relevance of this distinction is supported by recent studies, which
show that individuals respond differently to these dimensions. For
instance, perceptions that effort is rewarded in society are more
strongly associated with positive evaluations of fairness and acceptance
of unequal outcomes (\citeproc{ref-batruch_belief_2023}{Batruch et al.,
2023}; \citeproc{ref-wiederkehr_belief_2015}{Wiederkehr et al., 2015};
\citeproc{ref-wilson_role_2003}{Wilson, 2003}). This may be because
effort is seen as a controllable and morally virtuous trait, whereas
talent is often perceived as a natural advantage. Consequently,
effort-based meritocracy is likely more potent in legitimizing
inequality, particularly in neoliberal contexts.

These dimensions of meritocracy reflect how respondents perceive
society's distributive logic, regardless of whether they endorse
meritocratic principles. This distinction aligns with recent findings
indicating that individuals distinguish between how merit is perceived
in society and how it should ideally operate, which in turn shapes their
preferences for redistribution and justice
(\citeproc{ref-tejero-peregrina_perceived_2025}{Tejero-Peregrina et al.,
2025}). Meritocratic beliefs serve as symbolic justifications for
unequal outcomes, particularly when access is stratified by income or
social opportunity. Prior studies in Chile have shown that individuals
who perceive higher levels of meritocracy tend to express stronger
support for unequal distributions that reflect market outcomes in
healthcare, education and pensions
(\citeproc{ref-castillo_perceptions_2025}{Castillo, Laffert, et al.,
2025}; \citeproc{ref-castillo_socialization_2024}{Castillo et al.,
2024}).

In addition to influencing individual attitudes toward inequality,
meritocratic beliefs can contribute to social division and the
stigmatization of disadvantaged groups. Recent research has demonstrated
that exposure to meritocratic narratives can reinforce the belief that
poverty results from individual failings rather than systemic
conditions, reducing support for redistributive measures and increasing
the stigmatization of the poor (\citeproc{ref-hoyt_mindsets_2023}{Hoyt
et al., 2023}). This reinforces negative stereotypes and reduces empathy
toward individuals from lower socioeconomic backgrounds. Moreover,
Busemeyer et al. (\citeproc{ref-busemeyer_positive_2021}{2021}) argues
that meritocratic narratives can serve as feedback mechanisms that shape
public opinion and well-being by framing individuals' understanding of
welfare outcomes as deserved or undeserved within existing institutional
structures. This psychological mechanism highlights the normative power
of meritocracy in stabilizing unequal systems by shaping political
attitudes and personal perceptions of success and failure.

Importantly, recent research has explored how meritocratic beliefs
interact with experiences of intergenerational mobility to shape
distributive attitudes. The belief that one's success is earned can lead
upwardly mobile individuals to internalize meritocratic narratives and
justify existing inequalities, reinforcing support for market justice
(\citeproc{ref-gugushvili_intergenerational_2016c}{Gugushvili, 2016a};
\citeproc{ref-molina_its_2019}{Molina et al., 2019}). Conversely,
downwardly mobile individuals who maintain strong meritocratic beliefs
may interpret their status as a personal failure, reducing their support
for redistribution (\citeproc{ref-day_movin_2017}{Day \& Fiske, 2017}).
At the macro level, Shariff et al.
(\citeproc{ref-shariff_income_2016}{2016}) show that higher perceived
mobility increases tolerance for inequality, suggesting that meritocracy
and mobility are mutually reinforcing.

Taken together, this literature supports the idea that meritocratic
beliefs can moderate the relationship between mobility and market
justice preferences. Individuals who experience mobility---especially
upward---may draw on meritocratic narratives to legitimize both their
own status and broader inequalities, thereby strengthening their support
for market-based distribution. Accordingly, the second hypothesis of
this study is:

\(H2\): The positive (negative) relationship between upward (downward)
social mobility and support for market justice in healthcare, pensions,
and education is moderated by meritocratic beliefs; specifically, this
association is stronger (weaker) among individuals with higher
perceptions of meritocracy.

\subsection{The Chilean context}\label{the-chilean-context}

Chile offers a compelling case for examining individual preferences for
market justice in the access to core social services. Since the early
1990s, the country experienced a long period of sustained economic
growth, with per capita GDP increasing by nearly 4\% annually until the
mid-2010s (\citeproc{ref-llorca-jana_historia_2021}{Llorca-Jaña \&
Miller, 2021}). This growth brought notable improvements in living
standards, including a strong decline in income poverty---from 38.6\% in
1990 to 6.5\% in 2022 (\citeproc{ref-mideplan_evolucion_2017}{MIDEPLAN,
2017}). However, this success story has been tempered by growing
concerns over the unequal distribution of its benefits. In recent years,
economic growth has slowed---exacerbated by the global impact of the
COVID-19 pandemic---and structural inequalities have become more visible
(\citeproc{ref-barozet_clases_2021}{Barozet, Contreras, et al., 2021}).
Despite its economic achievements, Chile remains one of the most unequal
countries in Latin America and the OCDE. The poorest 50\% of the
population captures just 10\% of total income and holds negative net
wealth, while the wealthiest 1\% receives nearly 27\% of all income and
controls almost half (49.6\%) of the country's wealth
(\citeproc{ref-chancel_world_2022}{Chancel et al., 2022}), a figure that
has seen little change since the 1990s
(\citeproc{ref-flores_top_2020}{Flores et al., 2020}). These stark
disparities extend beyond income, reflecting deep inequalities in access
to essential social services (\citeproc{ref-pnud_desiguales_2017}{PNUD,
2017}).

A significant share of Chile's inequality stems from neoliberal reforms
that privatized and commodified key domains of social reproduction
(\citeproc{ref-arrizabalo_milagro_1995}{Arrizabalo, 1995};
\citeproc{ref-ferre_welfare_2023}{Ferre, 2023}). Introduced during the
military dictatorship (1973--1989) and expanded under democratic
governments, these reforms embedded market logic into public services
through concessions, subsidies, demand-side vouchers, and pro-private
regulatory frameworks (\citeproc{ref-boccardo_30_2020}{Boccardo, 2020};
\citeproc{ref-madariaga_three_2020}{Madariaga, 2020}). This
``crowded-out'' welfare model benefits higher-income groups, leaving
lower-income individuals to rely on limited public options. Scholars
argue that this neoliberal shift repurposed the state to create niches
of capitalist accumulation -from water to education-, giving rise to a
model of ``public-service capitalism'' heavily reliant on state funding
(\citeproc{ref-boccardo_30_2020}{Boccardo, 2020}). In healthcare, around
79\% of the population is covered by the public insurer FONASA, while
15\% are enrolled in private ISAPREs, which offer faster and
higher-quality services to wealthier groups
(\citeproc{ref-mideplan_estadisticas_2024}{MIDEPLAN, 2024}). The pension
system is based on individual capitalization, with mandatory
contributions managed by private administrators investing in the
financial market, currently involving 11 million contributors through
private fund administrators (AFPs), yet 27\% of the labor force remains
outside the system due to informality
(\citeproc{ref-superintendenciadepensiones_estadisticas_2024}{Superintendencia
de Pensiones, 2024}). Education is similarly stratified: only 30.6\% of
students attend fully public schools, while 54\% are in voucher-funded
private institutions and 9.3\% in fully private schools that largely
serve affluent families
(\citeproc{ref-ministeriodeeducacion_resumen_2023}{Ministerio de
Educación, 2023}). This arrangement has created a structurally segmented
system, where access and quality of services depend on ability to pay,
reinforced by state subsidies to private providers
(\citeproc{ref-pnud_desiguales_2017}{PNUD, 2017}).

Over recent decades, Chile has undergone a significant transformation in
its class structure, shaped by sustained economic growth, the expansion
of the service sector, the massification of education, and major changes
in the role of the state and social policy
(\citeproc{ref-ruiz_chilenos_2014}{Ruiz \& Boccardo, 2014}). These
changes contributed to the growth of the middle sectors, a phenomena
labelled as ``mesocratrization'' of society
(\citeproc{ref-espinoza_estratificacion_2013}{Espinoza et al., 2013}).
However, this expansion did not reduce structural inequalities nor
facilitate upward mobility into elite positions
(\citeproc{ref-espinoza_estratificacion_2013}{Espinoza et al., 2013};
\citeproc{ref-lopez-roldan_comparative_2021}{López-Roldán \& Fachelli,
2021}). While Chile shows high absolute mobility, most flows occur
between adjacent classes---particularly from working classes into lower
service-class positions---while access to upper classes remains limited,
indicating strong elite closure and reproduction
(\citeproc{ref-espinoza_estratificacion_2013}{Espinoza et al., 2013};
\citeproc{ref-espinoza_movilidad_2014}{Espinoza \& Núñez, 2014};
\citeproc{ref-lopez-roldan_comparative_2021}{López-Roldán \& Fachelli,
2021}; \citeproc{ref-perez-ahumada_clases_2019}{Pérez-Ahumada, 2019};
\citeproc{ref-torche_unequal_2005}{Torche, 2005}). Though often
described as ``unequal but fluid'', recent evidence points to growing
rigidity in relative mobility, especially after the economic slowdown,
increased household debt, labor market precarity, and the COVID-19
crisis (\citeproc{ref-barozet_clases_2021}{Barozet, Contreras, et al.,
2021}). However, the majority of the population identifies themselves as
members of the middle class (\citeproc{ref-castillo_todos_2013}{Castillo
et al., 2013}). These dynamics have fueled a legitimacy crisis:
meritocracy is increasingly questioned, and many perceive that
advancement depends more on informal networks than on effort or earned
credentials (\citeproc{ref-barozet_clases_2021}{Barozet, Contreras, et
al., 2021}). The combination of constrained mobility, persistent
inequality, and concentrated opportunity has eroded expectations of
upward mobility and intensified social discontent
(\citeproc{ref-pnud_desiguales_2017}{PNUD, 2017}).

Despite recurrent social unrest, Chile presents a paradoxical
coexistence between strong conflict over inequality and widespread
public legitimation of it. The October 2019 ``social outburst'', a
period of mass protests and severe political repression throughout the
country, crystallized discontent over the privatization and
commodification of social services, alongside a crisis of political
legitimacy (\citeproc{ref-somma_no_2021}{Somma et al., 2021}). A survey
during the protests identified pensions, healthcare, and education as
the top demands
(\citeproc{ref-nucleodesociologiacontingente_informe_2020}{Núcleo de
Sociología Contingente, 2020}), reflecting dissatisfaction with Chile's
stratified welfare regime. Yet, this unrest coexists with strong
meritocratic beliefs and high tolerance for income inequality. Castillo
(\citeproc{ref-castillo_legitimacy_2011}{2011}) shows that wage gaps
between occupations are consistently justified, and the larger the
perceived gap, the greater its legitimation, an effect amplified by
Chile's structural inequality. Moreover, perceptions of unfairness due
to non-meritocratic factors often reinforce belief in effort and talent
as legitimate bases for success
(\citeproc{ref-castillo_multidimensional_2023}{Castillo et al., 2023}),
which are associated with lower perceived injustice
(\citeproc{ref-castillo_meritocracia_2019}{Castillo et al., 2019}).
Mac-Clure et al. (\citeproc{ref-mac-clure_justicia_2024}{Mac-Clure et
al., 2024}) find that low and lower-middle status individuals are less
likely than upper-status groups to view educational income differences
as unfair. Qualitative research further reveals the internalization of
the enterprising self (\citeproc{ref-mau_inequality_2015}{Mau, 2015}),
understood as a way in which the subjectivization process arises in line
with positioning the self in the market, generating its own market value
and subjecting it to competition. Canales et al.
(\citeproc{ref-canalesceron_sujeto_2021}{2021}) describes how families
navigate school choice as investment decisions, while Panes
(\citeproc{ref-panes_criticas_2020}{2020}) finds that many workers frame
pension contributions as individual investments aimed at maximizing
future returns. These representations suggest that market-based moral
orientations are deeply embedded in Chilean subjectivities, offering
fertile ground to study how support for market justice varies across
welfare domains.

Although still limited, recent research on market justice preferences in
Chile indicates that support for these beliefs has increased in recent
years across domains such as healthcare, pensions, and education, and
that both objective and subjective dimensions of inequality shape them.
Castillo et al. (\citeproc{ref-castillo_perceptions_2025}{2025}) show
that agreement with the notion that it is fair for higher-income
individuals to access better services was relatively low in 2016, but
rose significantly by 2023---especially in the pension domain. Otero and
Mendoza (\citeproc{ref-otero_power_2024}{2024}) provide evidence that
social class, measured using the EGP scheme, influences these
preferences: individuals in lower or subordinate class positions are
less supportive of market-based access to welfare services compared to
those in higher or privileged positions. Additionally, greater diversity
in social class networks is associated with lower support for market
justice. This is particularly relevant in the Chilean context, where one
of the main determinants of class-diverse networks is intergenerational
social mobility, which exposes individuals to relationships with people
from different class backgrounds (\citeproc{ref-otero_lives_2022}{Otero
et al., 2022}). Regarding subjective factors, Castillo et al.
(\citeproc{ref-castillo_perceptions_2025}{Castillo, Laffert, et al.,
2025}; \citeproc{ref-castillo_socialization_2024}{Castillo et al.,
2024}) find that individuals who endorse meritocratic beliefs---such as
the idea that effort and talent are rewarded in Chile---are more likely
to support market justice across welfare domains.

\section{Method}\label{method}

Social mobility effects---understood as impacts on an outcome arising
from movements between an origin state and a destination state (e.g.,
social class)---have long been a focus of sociological research
(\citeproc{ref-eyles_social_2022}{Eyles et al., 2022};
\citeproc{ref-langsaether_explaining_2022}{Langsæther et al., 2022}).
Yet, as Breen and Ermisch (\citeproc{ref-breen_effects_2024}{2024, p.
467}) emphasize, most hypotheses are, at their core, individual-level
counterfactual comparisons between the observed outcome under a mobility
trajectory and the outcome the same person would have shown if they had
remained in their origin class (or moved to an alternative destination).
Standard specifications (e.g., SAM, DRM, and mobility-contrast models)
struggle to retrieve these inherently counterfactual quantities due to
identification constraints and their reliance on between-group contrasts
constructed from additive terms and interactions, thereby yielding
primarily associational evidence
(\citeproc{ref-breen_effects_2024}{Breen \& Ermisch, 2024};
\citeproc{ref-song_there_2025}{Song \& Zhou, 2025}). I therefore adopt
Breen and Ermisch's (\citeproc{ref-breen_effects_2024}{2024}) causal
formulation, treating destination class as the treatment, conditioning
on the class of origin, and estimating heterogeneous destination effects
with observational data under explicit identification assumptions. To
align design and target, I follow the MIDA template
(\citeproc{ref-blair_research_2023}{Blair et al., 2023}): set out the
causal model and assumptions (M), define the inquiry and estimand (I),
describe the data and variables (D), and detail the answer strategy (A)
used to identify the causal effect of intergenerational social mobility
on preferences for the commodification of pensions.

\subsection{A Causal Model for Mobility
Effects}\label{a-causal-model-for-mobility-effects}

Breen and Ermisch (\citeproc{ref-breen_effects_2024}{2024}), building on
a critical reassessment of the inferential limits of standard mobility
models, propose a causal framework that reconceptualizes the ``mobility
effect'' as the treatment effect of reaching a destination class, with
impacts heterogeneous by class of origin. The core claim is that typical
mobility hypotheses are fundamentally within-person counterfactual
comparisons rather than between-person contrasts. Defining mobility as
\((M = D - O)\) ---a change from origin \((O)\) to destination \((D)\),
rather than an additive or interactive combination---places the focus
squarely on the Neyman--Rubin problem: for any individual, we observe
the outcome under the realized destination but not under alternative
destinations or under immobility.

Exploiting the temporal ordering of origin, destination and outcome,
Breen and Ermisch (\citeproc{ref-breen_effects_2024}{2024}) frame
mobility as a treatment process that motivates the causal question:
``how would the outcome among people from origin \emph{j} who entered
destination \emph{k} have been different if those people had,
counterfactually, entered destination \emph{k'} instead?'' (p.472). The
corresponding estimand is the conditional causal effect of destination
given origin. It is of particular interest when \emph{k' = j}, i.e.,
immobility. This estimand compares movers' observed outcomes with the
hypothetical outcomes those same individuals would have exhibited had
they remained in their origin class. This formulation provides a
coherent potential-outcomes basis for studying how class transitions
shape individual preferences and attitudes.

Following Breen and Ermisch (\citeproc{ref-breen_effects_2024}{2024}),
the identification of causal mobility effects from observational data
requires the following assumptions:

\textbf{Positivity}: For all \((j,c)\) in the support of \((O,C)\) and
for each relevant destination \((k)\), the probability of receiving
\(D=k\) is strictly between 0 and 1; substantively, every \((O,C)\)-type
has some chance of each comparison destination.

\begin{equation}\phantomsection\label{eq-posi}{
 0<P(D=k\mid O=j, C=c)<1
}\end{equation}

\textbf{Stable Unit Treatment Value Assumption (SUTVA)}: each unit's
potential outcomes \(Y_i(D)\) does not depend on the mechanism used to
assign treatments (destinations) and by the treatments assigned to other
units (also called no interference), assuming a single, well-defined
version of each treatment (consistency).

\begin{equation}\phantomsection\label{eq-sutva}{
Y_i(D_i) = Y_i\big(D_i, D_{-i}\big)\quad \forall, D_{-i}
}\end{equation}

\textbf{Conditional Independence}: Also known as conditional
unconfoundedness or exchangeability, this assumption emphasizes that,
conditional on origin \(O\) and pre-treatment covariates \(C\) (e.g.,
parental education, ethnicity, cohort, early-life factors), assignment
to \(D\) is as good as random. This underlies IPW and regression
adjustment, which seek exchangeability between mobile and immobile
individuals.

\begin{equation}\phantomsection\label{eq-indep}{
Y(D)\ \perp\kern-5pt \perp D\ \mid\ (O,C)
}\end{equation}

Figure~\ref{fig-dag} depicts the causal model guiding identification. In
this DAG, \(C\) represents pre-treatment confounders (e.g., family
resources, household composition, and ascriptive status) affecting both
origin and destination, and \(U\) denotes unobserved factors influencing
\(C\). Class origin \(O\) is determined by \(C\) and, in turn, affects
destination \(D\). The outcome \(Y\) (preferences for commodification)
is caused by \(D\) and can also be affected by \(C\). There is no direct
arrow from \(O\) to \(Y\), reflecting the assumption that the effect of
origin on the outcome operates entirely through background factors
\((C)\) and destination \((D)\). Accordingly, conditioning on \((O,C)\)
blocks the relevant back-door paths from \(D\) to \(Y\), specifically
\(D\leftarrow O\leftarrow C\rightarrow Y\) and, if present,
\(D \leftarrow O \rightarrow Y\), so \((O,C)\) is a minimal sufficient
adjustment set.

\begin{figure}[H]

\centering{

\includegraphics[width=0.85\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-dag-1.pdf}

}

\caption{\label{fig-dag}Causal graph of the effect of intergenerational
social mobility on preferences for commodification. Y = preferences for
pension commodification, my outcome of interest, D = an individual's
class of destination, O = an individual's class of origin, C = different
attributes determined in childhood or earlier that affects an
individual's class of origin and destination. Finally, U = unobserved
factors influencing origin attributes.}

\end{figure}%

Formally, the DAG encodes the following relations:

\[
U \rightarrow C,\quad C \rightarrow O,\quad C \rightarrow D,\quad C \rightarrow Y,\quad O \rightarrow D,\quad D \rightarrow Y
\]

This structure corresponds to a nonparametric structural causal model in
which \((D)\) is the treatment variable, \((Y)\) the outcome, and
\((O, C)\) the minimal sufficient adjustment set. Conditioning on these
variables satisfies the back-door criterion for identifying the causal
effect of mobility on preferences
(\citeproc{ref-breen_effects_2024}{Breen \& Ermisch, 2024}).

Why are the assumptions plausible here? Conditional independence is
targeted by controlling for a plausible sufficient adjustment set
\((O,C)\); entropy balancing/IPW enforces covariate balance within each
origin stratum. Positivity is assessed via propensity scores and
trimming of extreme weights, restricting to common support. SUTVA holds
by treating the destination class as a single exposure measured before
\((Y)\) and assuming no interference. Together, the DAG and assumptions
define conditions for identifying the causal effect of intergenerational
mobility on preferences for commodification.

\subsection{Inquiry and estimand}\label{inquiry-and-estimand}

The causal inquiry guiding this study asks: How does intergenerational
social mobility affect individuals' preferences for the commodification
of pensions in Chile? Following the potential outcomes framework
proposed by Breen and Ermisch (\citeproc{ref-breen_effects_2024}{2024,
p. 473}), the estimand of interest is the average treatment effect on
the treated (ATT), defined as the mean causal effect of moving from an
origin class \((O=j)\) to a destination class \((D=k)\), compared with
the counterfactual outcome that those same individuals would have
exhibited had they instead reached an alternative destination \((k')\):

\begin{equation}\phantomsection\label{eq-breen}{ 
ATT_{j,k,k′} =E[Y(D=k|O=j, D=k)]−E[Y(D=k′|O=j, D=k)]
}\end{equation}

In this study, I focus on the specific case where the counterfactual
destination corresponds to class immobility \((k' = j)\), that is, how
the outcome among people from origin \emph{k} who entered destination
\emph{k} would have bee different if those people had, counterfactually,
remained in origin \emph{j} instead (inmobility). Accordingly, the
estimand becomes:

\begin{equation}\phantomsection\label{eq-att}{ 
ATT_{j,k,j} =E[Y(D=k|O=j, D=k)]−E[Y(D=j|O=j, D=k)]
}\end{equation}

which represents the average causal effect of attaining class
\emph{(k)}---for instance, moving upward or downward---relative to
remaining in class \emph{(j)}, among individuals who actually
experienced such mobility. This estimand captures the within-origin
counterfactual contrast at the heart of mobility effects: how a person's
preference for market-based pensions would differ if, counterfactually,
they had stayed in their origin class rather than moved to their
observed destination.

\subsection{Data and variables}\label{data-and-variables}

\subsubsection{Data}\label{data}

This study draws on data from the Chilean Longitudinal Social Survey
(ELSOC) of the Center for Social Conflict and Cohesion Studies (COES).
This survey is a nationally representative panel study of the urban
adult population in Chile, conducted annually between 2016 and 2023.
Designed to examine individuals' attitudes, emotions, and behaviors
regarding social conflict and cohesion, ELSOC employs a probabilistic,
stratified, clustered, and multistage sampling design covering both
major urban centers and smaller cities. The sampling frame was
proportionally stratified into six categories of urban population size
(e.g., large and small cities), followed by a random selection of
households within 1,067 city blocks. The target population includes men
and women aged 18 to 75 who are habitual residents of private dwellings.

Because respondent occupation is not measured in every wave, I restrict
the analysis to the 2016, 2018, and 2023 waves. Parental occupation was
collected only in 2023 as open-text labels. I coded these texts to the
ISCO-08 two-digit scheme using the National Institute of Statistics of
Chile's automated coding API and retained cases with high-confidence
matches (\textgreater95\%). Treating parental occupation as a
time-invariant origin attribute, I carried these codes back to 2016 and
2018. After listwise deletion and restricting to key variables
(respondent occupation and the outcome measure), the final analytic
sample comprises \emph{N} = 3,435 observations nested within \emph{N} =
1,787 individuals (2016: 914; 2018: 1,377; 2023: 1,144). Consistent with
the study design, estimation proceeds within trajectory-specific
subsamples (e.g., low→low; low→middle), retaining movers and matched
non-movers from the same origin class. Then, I apply weights to
construct the corresponding counterfactual control distributions for
each trajectory. As Breen and Ermisch
(\citeproc{ref-breen_effects_2024}{2024}) did, I use the three waves in
the estimations that follow, allowing for correlation between the
observations for each individual in calculating the standard errors
(i.e.~clustering on the personal identifier).

\subsubsection{Outcome variable}\label{outcome-variable}

The outcome variable measures preferences regarding the commodification
of pensions, operationalized with a single item addressing the degree of
public support for market-based access to old-age pension benefits.
Respondents were asked: ``Is it fair in Chile that people with higher
incomes have better pensions than people with lower incomes?'' and were
asked to indicate their level of agreement on a five-point Likert scale
ranging from 1 (``strongly disagree'') to 5 (``strongly agree''). For
estimation and interpretability, I dichotomized the item so that values
4-5 indicate agreement with market-based access to old-age pensions
(coded as 1), while values 1-3 indicate disagreement (coded as 0). This
binary coding treats neutrality as substantially closer to non-support
than to support and produces a clear contrast between respondents who
support market-based differentiation in pensions and those who do not. I
use this item for two main reasons: on the one hand, to be able to make
comparisons with existing literature
(\citeproc{ref-castillo_perceptions_2025}{Castillo, Laffert, et al.,
2025}; \citeproc{ref-lindh_public_2015}{Lindh, 2015};
\citeproc{ref-otero_power_2024}{Otero \& Mendoza, 2024}) and, on the
other hand, because it captures two fundamental dimensions of
market-based welfare distribution: the role of economic resources as a
key determinant of outcomes and the framing of services as as tradable
commodities that can be bought and sold based on ability to pay
(\citeproc{ref-lindh_public_2015}{Lindh, 2015}).

Table~\ref{tbl-summary1} shows the descriptive statistics of the
variables used in the study.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.4216}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2647}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2059}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.1078}}@{}}
\caption{Descriptive statistics for all
variables.}\label{tbl-summary1}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Label
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Stats / Values
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Freqs (\% of Valid)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Valid
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Label
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Stats / Values
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Freqs (\% of Valid)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Valid
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Preference for pension commodification &
\begin{minipage}[t]{\linewidth}\raggedright
1. Disagree\\
2. Agree\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
2702 (78.7\%)\\
733 (21.3\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
3435\\
(100.0\%)\strut
\end{minipage} \\
Father stratum & \begin{minipage}[t]{\linewidth}\raggedright
1. Low\\
2. Middle\\
3. High\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
1122 (32.7\%)\\
1104 (32.1\%)\\
1209 (35.2\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
3435\\
(100.0\%)\strut
\end{minipage} \\
Offspring stratum & \begin{minipage}[t]{\linewidth}\raggedright
1. Low\\
2. Middle\\
3. High\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
1186 (34.5\%)\\
1110 (32.3\%)\\
1139 (33.2\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
3435\\
(100.0\%)\strut
\end{minipage} \\
Parental education & \begin{minipage}[t]{\linewidth}\raggedright
Mean (sd) : 4 (2.2)\\
min \textless{} med \textless{} max:\\
1 \textless{} 4 \textless{} 10\\
IQR (CV) : 3 (0.6)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
1 : 222 ( 6.5\%)\\
2 : 905 (26.3\%)\\
3 : 576 (16.8\%)\\
4 : 321 ( 9.3\%)\\
5 : 826 (24.0\%)\\
6 : 30 ( 0.9\%)\\
7 : 214 ( 6.2\%)\\
8 : 75 ( 2.2\%)\\
9 : 245 ( 7.1\%)\\
10 : 21 ( 0.6\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
3435\\
(100.0\%)\strut
\end{minipage} \\
Co-residence with both parents at age 15 &
\begin{minipage}[t]{\linewidth}\raggedright
1. No co-residence\\
2. Co-residence\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
996 (29.0\%)\\
2439 (71.0\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
3435\\
(100.0\%)\strut
\end{minipage} \\
Nacionality & \begin{minipage}[t]{\linewidth}\raggedright
1. Non-Chilean\\
2. Chilean\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
62 ( 1.8\%)\\
3373 (98.2\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
3435\\
(100.0\%)\strut
\end{minipage} \\
Sex & \begin{minipage}[t]{\linewidth}\raggedright
1. Male\\
2. Female\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
1507 (43.9\%)\\
1928 (56.1\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
3435\\
(100.0\%)\strut
\end{minipage} \\
Age (in years) & \begin{minipage}[t]{\linewidth}\raggedright
Mean (sd) : 42.6 (12.5)\\
min \textless{} med \textless{} max:\\
18 \textless{} 43 \textless{} 75\\
IQR (CV) : 21 (0.3)\strut
\end{minipage} & 58 distinct values &
\begin{minipage}[t]{\linewidth}\raggedright
3435\\
(100.0\%)\strut
\end{minipage} \\
Indigenous ethnicity & \begin{minipage}[t]{\linewidth}\raggedright
1. Non-indigenous\\
2. Indigenous\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
3035 (88.4\%)\\
400 (11.6\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
3435\\
(100.0\%)\strut
\end{minipage} \\
Meritocracy perception & \begin{minipage}[t]{\linewidth}\raggedright
1. Low\\
2. High\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
2741 (79.8\%)\\
694 (20.2\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
3435\\
(100.0\%)\strut
\end{minipage} \\
Wave & \begin{minipage}[t]{\linewidth}\raggedright
1. 2016\\
2. 2018\\
3. 2023\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
914 (26.6\%)\\
1377 (40.1\%)\\
1144 (33.3\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
3435\\
(100.0\%)\strut
\end{minipage} \\
\end{longtable}

\subsubsection{Treatment}\label{treatment}

\paragraph{Intergenerational occupational
mobility}\label{intergenerational-occupational-mobility}

I treat intergenerational occupational mobility as an exposure
indicating whether respondents occupy a different socioeconomic status
than their fathers, closely following Breen and Ermisch's
(\citeproc{ref-breen_effects_2024}{2024}) causal framework. Class
assignment proceeds in two steps. First, I derive status for both origin
(father) and destination (respondent) from two-digit ISCO-08 occupations
using the International Socio-Economic Index of Occupational Status
(ISEI). ISEI is a hierarchical, unidimensional scale that locates
occupations by incumbents' average education and earnings and was
explicitly designed to harmonize occupational stratification across
other class schemes and socioeconomic status measures
(\citeproc{ref-ganzeboom_internationally_1996}{Ganzeboom \& Treiman,
1996}), thereby supporting cross-study comparability
(\citeproc{ref-barozet_measurement_2021a}{Barozet, Boado, et al.,
2021}). Second, I group both origin and destination ISEI scores into
terciles (low, middle, high), yielding a three-category schema for
origin and destination stratums (see Table~\ref{tbl-matrix}).
Substantively, I use ISEI because occupation-based measures capture
structural locations that organize life chances and attitudes more
directly than income alone: they bundle access to the means of
production, qualification credentials, and the authority/power inherent
in labor relations; central dimensions of stratification and inequality
(\citeproc{ref-barone_rise_2022}{Barone et al., 2022};
\citeproc{ref-wrightUnderstandingClass2015}{Wright, 2015}). In short,
using ISEI aligns the measurement of origin and destination on a common
vertical status continuum that is theoretically meaningful for social
mobility research and empirically feasible given the available
information of ELSOC's data.

\begin{table}

\caption{\label{tbl-matrix}Occupational mobility by occupational
groups.}

\centering{

\centering
\begin{tabular}{cccccc}
\toprule
Father↓ & Offspring→ & Low & Middle & High & Total\\
\midrule
Low &  & 44.7\%   (501) & 33.6\%   (377) & 21.7\%   (244) & 100.0\% (1,122)\\
Middle &  & 35.1\%   (388) & 34.9\%   (385) & 30.0\%   (331) & 100.0\% (1,104)\\
High &  & 24.6\%   (297) & 28.8\%   (348) & 46.7\%   (564) & 100.0\% (1,209)\\
Total &  & 34.5\% (1,186) & 32.3\% (1,110) & 33.2\% (1,139) & 100.0\% (3,435)\\
\bottomrule
\end{tabular}

}

\end{table}%

\paragraph{Pre-treatment control variables for selection into
treatment}\label{pre-treatment-control-variables-for-selection-into-treatment}

Within each origin stratum \(O=j\), I estimate average treatment effects
on the treated (ATT) by comparing movers to otherwise similar non-movers
from the same origin. To adjust for selection into mobility, I reweight
only the control group (immobile) using entropy balancing so that the
distribution of pre-treatment origin covariates \(C\) matches that of
the treated on selected moments.

The selection of the adjustment set \(C\) follows two principles. First,
it mirrors the logic in Breen and Ermisch
(\citeproc{ref-breen_effects_2024}{2024})---who condition on attributes
determined in childhood or earlier (e.g., parental resources, household
structure, ascriptive traits)---and, second, it is constrained to
origin-side characteristics available in ELSOC. Guided by theory and
Chilean evidence on relative mobility determinants (mainly
\citeproc{ref-brunori_inequality_2025}{Brunori et al., 2025};
\citeproc{ref-celhay_intergenerational_2010}{Celhay et al., 2010};
\citeproc{ref-espinoza_movilidad_2014}{Espinoza \& Núñez, 2014};
\citeproc{ref-torche_unequal_2005}{Torche, 2005}), I include six
covariates covering family resources, household composition, and
ascriptive status: (a) parental education (highest of father/mother), 10
categories coded ordinally from no schooling to postgraduate studies;
(b) co-residence with both parents at age 15 (0 = no, 1 = yes); (c)
nationality (0 = non-Chilean, 1 = Chilean); (d) age in years; (e) sex (0
= male, 1 = female); and (f) indigenous ethnicity (0 = no, 1 = yes).
These variables are fixed before the destination class, plausibly affect
selection into mobility, and constitute the maximal origin set in ELSOC,
thereby making the conditional independence assumption more credible
within each \(O=j\).

After balancing, I construct stabilized IPW-ATT, setting treated weights
to 1 and assigning ebal-derived weights to controls, thus aligning the
counterfactual distribution of non-movers with the covariate profile of
movers in each origin stratum. These weights are then used in the
outcome stage to estimate the causal effect of mobility on preferences
for the commodification of pensions.

\subsubsection{Effect heterogeneity}\label{effect-heterogeneity}

To test heterogeneity in mobility effects, I examine two moderators:
meritocratic perceptions and time. Meritocracy is captured with two
items (\citeproc{ref-young_rise_1958}{Young, 1958}), one referring to
effort (``In Chile, people are rewarded for their efforts'') and one to
talent (``In Chile, people are rewarded for their intelligence and
skills''), each answered on a five-point Likert scale (1 = ``strongly
disagree'' to 5 = ``strongly agree''). I average the two items into a
single index and dichotomize it into low meritocracy (≤ 3) and high
meritocracy (≥ 4). Time is measured by survey wave (2016, 2018, 2023)
and entered as a categorical (dummy) variable. Both variables are used
as moderators by interacting them with the mobility treatment to assess
conditional average treatment effects---i.e., whether the effect of
mobility varies across levels of meritocratic beliefs or across survey
waves.

\subsubsection{Controls}\label{controls}

All models include the same pre-treatment covariates \(C\) used in the
IPW construction, not to re-balance groups (entropy balancing already
does so), but to (i) block backdoor paths from \(C\) to the outcome
\(Y\) as implied by the DAG, and (ii) achieve double robustness:
estimates remain consistent if either the weighting model or the outcome
model is correctly specified. Concretely, I adjust for (a) father's
educational level, (b) co-residence with both parents at age 15, (c)
nationality, (d) age, (e) sex, and (f) ethnicity, along with wave fixed
effects to absorb secular trends.

\subsection{Analytical strategy}\label{analytical-strategy}

Because the dependent variable---the preference for pension
commodification---is binary \((Y_{it}\in{0,1})\), I estimate weighted
linear probability models (WLS) using stabilized inverse probability
weights (IPW--ATT) obtained via entropy balancing. Although OLS
estimator is often questioned with binary outcomes, it's unbiased. It
consistently estimates the conditional expectation \(E(Y|X)\) under
standard exogeneity \(E[\varepsilon_i|X_i]=0\), with heteroskedasticity
addressed by robust or clustered errors
(\citeproc{ref-wooldridge_introductory_2009}{Wooldridge, 2009},
ch.~7.5). WLS further improves efficiency under heteroskedasticity or
unequal sampling probabilities
(\citeproc{ref-wooldridge_introductory_2009}{Wooldridge, 2009},
ch.~8.4). Following Gelman, Hill, and Vehtari
(\citeproc{ref-gelman_regression_2020}{2020, pp. 270--272}),
IPW-weighted regression transparently targets the ATT by recreating the
counterfactual distribution. Linear models are preferred here because
coefficients are average probability differences; nonlinear models
(logit/probit) complicate weighting and interpretation
(\citeproc{ref-gelman_regression_2020}{Gelman et al., 2020}, ch.~13).

Formally, the model is:

\begin{equation}\phantomsection\label{eq-final}{
Y_{it}=\alpha+\beta T_i+X_i'\gamma+\lambda_t+\varepsilon_{it},
}\end{equation}

where \(Y_{it}\) indicates whether individual \(i\) in wave \(t\)
supports more market-based pension access; \(\alpha\) is the baseline
probability for the reference categories; \(\beta T_i\) captures the
intergenerational mobility contrast: \(T_i=1\) when the observed
destination is \(D_i=k\) (mobile to \(k\)) and \(T_i=0\) when \(D_i=j\)
(immobile), so that \(\beta\) is the ATT within origin
\(O=j (\widehat\beta=\widehat{ATT}{j,k\mid j})\), i.e., the
percentage-point change in the probability of preferring pension
commodification from reaching \(k\) rather than remaining in \(j\). The
term \(X_i'\gamma\) includes the pre-treatment covariates \((C)\) used
to construct the entropy-balancing weights; \(\lambda_t\) are wave fixed
effects; and \(\varepsilon_{it}\) is idiosyncratic error. The analytic
sample is restricted to individuals sharing the same origin \((O=j)\).
Estimation uses pooled WLS with stabilized IPW--ATT from entropy
balancing and CR2 standard errors clustered by individual to address
heteroskedasticity and within-person dependence.

The specification is doubly robust: entropy-balancing weights are
combined with covariate adjustment so consistency holds if either the
weighting model or the outcome model is correctly specified, while also
improving efficiency
(\citeproc{ref-wooldridge_introductory_2009}{Wooldridge, 2009}). I
report both IPW-only and doubly robust specifications to document
robustness. In addition, I assess causal heterogeneity by interacting
the mobility treatment with (i) meritocratic perceptions and (ii) survey
wave, reporting simple slopes (origin-specific ATTs) by moderator
levels. To evaluate sensitivity to unmeasured confounding, I compute
E-values for point estimates and confidence limits
(\citeproc{ref-ding_sensitivity_2016}{Ding \& VanderWeele, 2016}). To
address multiplicity across the six mobility trajectories, I also report
Holm-adjusted p-values alongside 95\% confidence intervals.

All the analyses were conducted using R software and the
\emph{lm\_robust} package.

\section{Results}\label{results}

\subsection{Descriptive statistics}\label{descriptive-statistics}

Lorem ipsoThe specification is doubly robust: entropy-balancing weights
are combined with covariate adjustment so consistency holds ifeither the
weighting model or the outcome model is correctly specified, while also
improving efficiency (Wooldridge, 2009).I report both IPW-only and
doubly robust specifications to document robustnessThe specification is
doubly robust: entropy-balancing weights are combined with covariate
adjustment so consistency holds ifeither the weighting model or the
outcome model is correctly specified, while also improving efficiency
(Wooldridge, 2009).I report both IPW-only and doubly robust
specifications to document robustness

\begin{figure}[H]

\centering{

\includegraphics[width=0.85\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-alluvial-1.pdf}

}

\caption{\label{fig-alluvial}Change in preferences for pension
commodification over time (2016, 2018, and 2023).}

\end{figure}%

The specification is doubly robust: entropy-balancing weights are
combined with covariate adjustment so consistency holds ifeither the
weighting model or the outcome model is correctly specified, while also
improving efficiency (Wooldridge, 2009).I report both IPW-only and
doubly robust specifications to document robustness

\begin{figure}[H]

\centering{

\includegraphics[width=0.9\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-mean-1.pdf}

}

\caption{\label{fig-mean}Percentage agreement on preferences for pension
commodification according to mobility trajectory and time (2016, 2018,
and 2023).}

\end{figure}%

\subsection{Models}\label{models}

The specification is doubly robust: entropy-balancing weights are
combined with covariate adjustment so consistency holds ifeither the
weighting model or the outcome model is correctly specified, while also
improving efficiency (Wooldridge, 2009).I report both IPW-only and
doubly robust specifications to document robustness

\begin{figure}[H]

\centering{

\includegraphics[width=0.9\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-reg-1.pdf}

}

\caption{\label{fig-reg}Effects of intergenerational occupational
mobility on preferences for pension commodification. Coefficient plot of
origin-specific ATT estimates comparing mobility versus immobility.
Estimates from pooled WLS with doubly robust adjustment, wave fixed
effects, and standard errors clustered by individual; bars show 95\%
confidence intervals.}

\end{figure}%

The specification is doubly robust: entropy-balancing weights are
combined with covariate adjustment so consistency holds ifeither the
weighting model or the outcome model is correctly specified, while also
improving efficiency (Wooldridge, 2009).I report both IPW-only and
doubly robust specifications to document robustnessThe specification is
doubly robust: entropy-balancing weights are combined with covariate
adjustment so consistency holds ifeither the weighting model or the
outcome model is correctly specified, while also improving efficiency
(Wooldridge, 2009).I report both IPW-only and doubly robust
specifications to document robustness

\subsection{Effect heterogeneity}\label{effect-heterogeneity-1}

The specification is doubly robust: entropy-balancing weights are
combined with covariate adjustment so consistency holds ifeither the
weighting model or the outcome model is correctly specified, while also
improving efficiency (Wooldridge, 2009).I report both IPW-only and
doubly robust specifications to document robustness

The specification is doubly robust: entropy-balancing weights are
combined with covariate adjustment so consistency holds ifeither the
weighting model or the outcome model is correctly specified, while also
improving efficiency (Wooldridge, 2009).I report both IPW-only and
doubly robust specifications to document robustness

\begin{figure}[H]

\centering{

\includegraphics[width=0.9\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-interact-1.pdf}

}

\caption{\label{fig-interact}Effects of intergenerational occupational
mobility on preferences for pension commodification, by meritocratic
perception. Coefficient plot of origin-specific ATT estimates comparing
mobility versus immobility, estimated from interaction models (T ×
Merit). Points show simple slopes of the mobility effect at low merit
and high merit; bars are 95\% confidence intervals. Estimates from
pooled WLS with doubly robust adjustment, wave fixed effects, and
standard errors clustered by individual.}

\end{figure}%

\begin{table}

\caption{\label{tbl-inttime}Effects of intergenerational occupational
mobility on preferences for pension commodification, by survey wave}

\centering{

\begin{center}
\scalebox{0.8}{
\begin{tabular}{l c c c c c c}
\hline
 & Low-Middle & Low-High & Middle-High & Middle-Low & High-Middle & High-Low \\
\hline
Intercept                              & $0.18$           & $-0.04$          & $0.28$           & $0.20$           & $0.18$            & $0.15$            \\
                                       & $ [-0.36; 0.73]$ & $ [-0.36; 0.29]$ & $ [-0.01; 0.57]$ & $ [-0.11; 0.52]$ & $ [-0.15;  0.50]$ & $ [-0.27;  0.58]$ \\
Mobility treatment                     & $-0.05$          & $-0.03$          & $0.13^{*}$       & $0.06$           & $-0.19^{*}$       & $-0.19^{*}$       \\
                                       & $ [-0.16; 0.06]$ & $ [-0.15; 0.10]$ & $ [ 0.05; 0.22]$ & $ [-0.03; 0.15]$ & $ [-0.33; -0.05]$ & $ [-0.32; -0.05]$ \\
Wave (Ref.= 2016)                      &                  &                  &                  &                  &                   &                   \\
                                       &                  &                  &                  &                  &                   &                   \\
\quad Wave 2018                        & $0.01$           & $0.01$           & $0.10^{*}$       & $0.09^{*}$       & $-0.04$           & $-0.02$           \\
                                       & $ [-0.10; 0.12]$ & $ [-0.10; 0.12]$ & $ [ 0.03; 0.18]$ & $ [ 0.01; 0.17]$ & $ [-0.15;  0.07]$ & $ [-0.15;  0.10]$ \\
\quad Wave 2023                        & $0.03$           & $0.01$           & $0.19^{*}$       & $0.18^{*}$       & $0.05$            & $0.04$            \\
                                       & $ [-0.09; 0.14]$ & $ [-0.12; 0.13]$ & $ [ 0.10; 0.28]$ & $ [ 0.08; 0.27]$ & $ [-0.09;  0.18]$ & $ [-0.11;  0.19]$ \\
Mobility treatment x Wave (Ref.= 2016) &                  &                  &                  &                  &                   &                   \\
                                       &                  &                  &                  &                  &                   &                   \\
\quad Mobility treatment x Wave 2018   & $0.03$           & $-0.06$          & $-0.04$          & $-0.09$          & $0.11$            & $0.06$            \\
                                       & $ [-0.11; 0.17]$ & $ [-0.21; 0.09]$ & $ [-0.17; 0.09]$ & $ [-0.20; 0.03]$ & $ [-0.04;  0.26]$ & $ [-0.09;  0.22]$ \\
\quad Mobility treatment x Wave 2023   & $0.07$           & $0.10$           & $-0.09$          & $-0.07$          & $0.19^{*}$        & $0.18$            \\
                                       & $ [-0.09; 0.23]$ & $ [-0.08; 0.28]$ & $ [-0.23; 0.06]$ & $ [-0.22; 0.08]$ & $ [ 0.01;  0.38]$ & $ [-0.01;  0.37]$ \\
\hline
Controls                               & Yes              & Yes              & Yes              & Yes              & Yes               & Yes               \\
R$^2$                                  & $0.04$           & $0.05$           & $0.07$           & $0.03$           & $0.06$            & $0.06$            \\
Adj. R$^2$                             & $0.03$           & $0.04$           & $0.06$           & $0.02$           & $0.05$            & $0.05$            \\
Num. obs.                              & $878$            & $745$            & $716$            & $773$            & $912$             & $861$             \\
RMSE                                   & $0.40$           & $0.37$           & $0.39$           & $0.39$           & $0.44$            & $0.42$            \\
N Clusters                             & $497$            & $449$            & $443$            & $470$            & $536$             & $525$             \\
\hline
\multicolumn{7}{l}{\scriptsize{Note: Cells contain regression coefficients with confidence intervals in parentheses. $^*$ Null hypothesis value outside the confidence interval..}}
\end{tabular}
}
\label{table:coefficients}
\end{center}

}

\end{table}%

\subsection{Sensitivity analysis}\label{sensitivity-analysis}

The specification is doubly robust: entropy-balancing weights are
combined with covariate adjustment so consistency holds ifeither the
weighting model or the outcome model is correctly specified, while also
improving efficiency (Wooldridge, 2009).I report both IPW-only and
doubly robust specifications to document robustness

\section{Discussion}\label{discussion}

\section{Conclusion}\label{conclusion}

\section{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-alesina_intergenerational_2018}
Alesina, A., Stantcheva, S., \& Teso, E. (2018). Intergenerational
{Mobility} and {Preferences} for {Redistribution}. \emph{American
Economic Review}, \emph{108}(2), 521--554.
\url{https://doi.org/10.1257/aer.20162015}

\bibitem[\citeproctext]{ref-araujo_desafios_2012}
Araujo, K., \& Martuccelli, D. (2012). \emph{Desaf{í}os comunes: Retrato
de la sociedad chilena y sus individuos} (1a. ed). Santiago: LOM
Ediciones.

\bibitem[\citeproctext]{ref-ares_changing_2020}
Ares, M. (2020). Changing classes, changing preferences: How social
class mobility affects economic preferences. \emph{West European
Politics}, \emph{43}(6), 1211--1237.
\url{https://doi.org/10.1080/01402382.2019.1644575}

\bibitem[\citeproctext]{ref-arrizabalo_milagro_1995}
Arrizabalo, X. (1995). \emph{{Milagro o quimera: la econom{í}a chilena
durante la dictadura}}. Libros de la Catarata.

\bibitem[\citeproctext]{ref-barone_rise_2022}
Barone, C., Hertel, F. R., \& Smallenbroek, O. (2022). The rise of
income and the demise of class and social status? {A} systematic review
of measures of socio-economic position in stratification research.
\emph{Research in Social Stratification and Mobility}, \emph{78},
100678. \url{https://doi.org/10.1016/j.rssm.2022.100678}

\bibitem[\citeproctext]{ref-barozet_measurement_2021a}
Barozet, E., Boado, M., \& Marqués-Perales, I. (2021). The {Measurement}
of {Social Stratification}: {Comparative Perspectives Between Europe}
and {Latin America}. In P. López-Roldán \& S. Fachelli (Eds.),
\emph{Towards a {Comparative Analysis} of {Social Inequalities} between
{Europe} and {Latin America}} (pp. 171--202). Cham: Springer
International Publishing.
\url{https://doi.org/10.1007/978-3-030-48442-2_6}

\bibitem[\citeproctext]{ref-barozet_clases_2021}
Barozet, E., Contreras, D., Espinoza, V., Gayo, M., \& Méndez, M. L.
(2021). \emph{{Clases medias en tiempos de crisis: vulnerabilidad
persistente, desaf{í}os para la cohesi{ó}n y un nuevo pacto social en
Chile}}. Santiago: Comisi{ó}n Econ{ó}mica para Am{é}rica Latina y el
Caribe (CEPAL).

\bibitem[\citeproctext]{ref-batruch_belief_2023}
Batruch, A., Jetten, J., Van De Werfhorst, H., Darnon, C., \& Butera, F.
(2023). Belief in {School Meritocracy} and the {Legitimization} of
{Social} and {Income Inequality}. \emph{Social Psychological and
Personality Science}, \emph{14}(5), 621--635.
\url{https://doi.org/10.1177/19485506221111017}

\bibitem[\citeproctext]{ref-benabou_social_2001}
Benabou, R., \& Ok, E. A. (2001). Social {Mobility} and the {Demand} for
{Redistribution}: {The Poum Hypothesis}. \emph{The Quarterly Journal of
Economics}, \emph{116}(2), 447--487.
\url{https://doi.org/10.1162/00335530151144078}

\bibitem[\citeproctext]{ref-blair_research_2023}
Blair, G., Coppock, A., \& Humphreys, M. (2023). \emph{Research design
in the social sciences: Declaration, diagnosis, and redesign}. Princeton
(N.J.) Oxford: Princeton University press.

\bibitem[\citeproctext]{ref-boccardo_30_2020}
Boccardo, G. (2020). \emph{30 a{ñ}os de privatizaciones en {Chile}: Lo
que la pandemia revel{ó}} (Nodo XXI). Santiago.

\bibitem[\citeproctext]{ref-breen_effects_2024}
Breen, R., \& Ermisch, J. (2024). The {Effects} of {Social Mobility}.
\emph{Sociological Science}, \emph{11}, 467--488.
\url{https://doi.org/10.15195/v11.a17}

\bibitem[\citeproctext]{ref-brunori_inequality_2025}
Brunori, P., Ferreira, F. H. G., \& Neidhöfer, G. (2025). Inequality of
opportunity and intergenerational persistence in {Latin America}.
\emph{Oxford Open Economics}, \emph{4}(Supplement\_1), i167--i199.
\url{https://doi.org/10.1093/ooec/odae021}

\bibitem[\citeproctext]{ref-bucca_merit_2016}
Bucca, M. (2016). Merit and blame in unequal societies: {Explaining
Latin Americans}' beliefs about wealth and poverty. \emph{Research in
Social Stratification and Mobility}, \emph{44}, 98--112.
\url{https://doi.org/10.1016/j.rssm.2016.02.005}

\bibitem[\citeproctext]{ref-busemeyer_skills_2014}
Busemeyer, M. (2014). \emph{Skills and {Inequality}: {Partisan Politics}
and the {Political Economy} of {Education Reforms} in {Western Welfare
States}}. Cambridge University Press.

\bibitem[\citeproctext]{ref-busemeyer_positive_2021}
Busemeyer, M., Abrassart, A., \& Nezi, R. (2021). Beyond {Positive} and
{Negative}: {New Perspectives} on {Feedback Effects} in {Public Opinion}
on the {Welfare State}. \emph{British Journal of Political Science},
\emph{51}(1), 137--162. \url{https://doi.org/10.1017/S0007123418000534}

\bibitem[\citeproctext]{ref-busemeyer_welfare_2020}
Busemeyer, M., \& Iversen, T. (2020). The {Welfare State} with {Private
Alternatives}: {The Transformation} of {Popular Support} for {Social
Insurance}. \emph{The Journal of Politics}, \emph{82}(2), 671--686.
\url{https://doi.org/10.1086/706980}

\bibitem[\citeproctext]{ref-campbell_institutional_2020}
Campbell, J. L. (2020). \emph{Institutional {Change} and
{Globalization}}. Princeton University Press.
\url{https://doi.org/10.2307/j.ctv131bw68}

\bibitem[\citeproctext]{ref-canalesceron_sujeto_2021}
Canales Cerón, M., Orellana Calderón, V. S., \& Guajardo Mañán, F.
(2021). Sujeto y cotidiano en la era neoliberal: El caso de la
educaci{ó}n chilena. \emph{Revista Mexicana de Ciencias Pol{í}ticas y
Sociales}, \emph{67}(244).
\url{https://doi.org/10.22201/fcpys.2448492xe.2022.244.70386}

\bibitem[\citeproctext]{ref-castillo_legitimacy_2011}
Castillo, J. C. (2011). Legitimacy of {Inequality} in a {Highly Unequal
Context}: {Evidence} from the {Chilean Case}. \emph{Social Justice
Research}, \emph{24}(4), 314--340.
\url{https://doi.org/10.1007/s11211-011-0144-5}

\bibitem[\citeproctext]{ref-castillo_changes_2025}
Castillo, J. C., Iturra, J., \& Carrasco, K. (2025). Changes in the
{Justification} of {Educational Inequalities}: {The Role} of
{Perceptions} of {Inequality} and {Meritocracy During} the {COVID
Pandemic}. \emph{Social Justice Research}.
\url{https://doi.org/10.1007/s11211-025-00458-0}

\bibitem[\citeproctext]{ref-castillo_multidimensional_2023}
Castillo, J. C., Iturra, J., Maldonado, L., Atria, J., \& Meneses, F.
(2023). A {Multidimensional Approach} for {Measuring Meritocratic
Beliefs}: {Advantages}, {Limitations} and {Alternatives} to the {ISSP
Social Inequality Survey}. \emph{International Journal of Sociology},
1--25. \url{https://doi.org/10.1080/00207659.2023.2274712}

\bibitem[\citeproctext]{ref-castillo_perceptions_2025}
Castillo, J. C., Laffert, A., Carrasco, K., \& Iturra, J. (2025).
Perceptions of {Inequality} and {Meritocracy}: {Their Interplay} in
{Shaping Preferences} for {Market Justice} in {Chile} (2016-2023).
\emph{Under Review at Frontiers in Sociology}.

\bibitem[\citeproctext]{ref-castillo_todos_2013}
Castillo, J. C., Miranda, D., \& Cabib, I. M. (2013). Todos somos de
clase media: {Sobre} el estatus social subjetivo en {Chile}. \emph{Latin
American Research Review}, \emph{48}(1), 155--173.
\url{https://doi.org/10.1353/lar.2013.0006}

\bibitem[\citeproctext]{ref-castillo_socialization_2024}
Castillo, J. C., Salgado, M., Carrasco, K., \& Laffert, A. (2024). The
{Socialization} of {Meritocracy} and {Market Justice Preferences} at
{School}. \emph{Societies}, \emph{14}(11), 214.
\url{https://doi.org/10.3390/soc14110214}

\bibitem[\citeproctext]{ref-castillo_meritocracia_2019}
Castillo, J. C., Torres, A., Atria, J., \& Maldonado, L. (2019).
Meritocracia y desigualdad econ{ó}mica: {Percepciones}, preferencias e
implicancias. \emph{Revista Internacional de Sociolog{í}a},
\emph{77}(1), 117. \url{https://doi.org/10.3989/ris.2019.77.1.17.114}

\bibitem[\citeproctext]{ref-celhay_intergenerational_2010}
Celhay, P., Sanhueza, C., \& Zubizarreta, J. R. (2010).
{INTERGENERATIONAL MOBILITY OF INCOME AND SCHOOLING}: {CHILE} 1996-2006.
\emph{Revista de an{á}lisis Econ{ó}mico}, \emph{25}(2), 43--63.
\url{https://doi.org/10.4067/S0718-88702010000200003}

\bibitem[\citeproctext]{ref-chancel_world_2022}
Chancel, L., Piketty, T., Saez, E., \& Zucman, G. (2022). World
inequality report 2022.
https://bibliotecadigital.ccb.org.co/handle/11520/27510.

\bibitem[\citeproctext]{ref-day_movin_2017}
Day, M. V., \& Fiske, S. T. (2017). Movin' on {Up}? {How Perceptions} of
{Social Mobility Affect Our Willingness} to {Defend} the {System}.
\emph{Social Psychological and Personality Science}, \emph{8}(3),
267--274. \url{https://doi.org/10.1177/1948550616678454}

\bibitem[\citeproctext]{ref-ding_sensitivity_2016}
Ding, P., \& VanderWeele, T. J. (2016). Sensitivity {Analysis Without
Assumptions}: \emph{Epidemiology}, \emph{27}(3), 368--377.
\url{https://doi.org/10.1097/EDE.0000000000000457}

\bibitem[\citeproctext]{ref-espinoza_estratificacion_2013}
Espinoza, V., Barozet, E., \& Méndez, M. L. (2013). {Estratificaci{ó}n y
movilidad social bajo un modelo neoliberal: El caso de Chile}.
\emph{Lavboratorio}, (25).

\bibitem[\citeproctext]{ref-espinoza_movilidad_2014}
Espinoza, V., \& Núñez, J. (2014). Movilidad ocupacional en {Chile}
2001-2009. {\textquestiondown}{Desigualdad} de ingresos con igualdad de
oportunidades? \emph{Revista Internacional de Sociolog{í}a},
\emph{72}(1), 57--82. \url{https://doi.org/10.3989/ris.2011.11.08}

\bibitem[\citeproctext]{ref-eyles_social_2022}
Eyles, A., Major, L. E., \& Machin, S. (2022). \emph{Social {Mobility} -
{Past}, {Present} and {Future}: {The State} of {Play} in {Social
Mobility}, on the 25th {Anniversary} of the {Sutton Trust}}. Sutton
Trust.

\bibitem[\citeproctext]{ref-ferre_welfare_2023}
Ferre, J. C. (2023). Welfare regimes in twenty-first-century {Latin
America}. \emph{Journal of International and Comparative Social Policy},
\emph{39}(2), 101--127. \url{https://doi.org/10.1017/ics.2023.16}

\bibitem[\citeproctext]{ref-flores_top_2020}
Flores, I., Sanhueza, C., Atria, J., \& Mayer, R. (2020). Top {Incomes}
in {Chile}: {A Historical Perspective} on {Income Inequality},
1964--2017. \emph{Review of Income and Wealth}, \emph{66}(4), 850--874.
\url{https://doi.org/10.1111/roiw.12441}

\bibitem[\citeproctext]{ref-ganzeboom_internationally_1996}
Ganzeboom, H. B. G., \& Treiman, D. J. (1996). Internationally
{Comparable Measures} of {Occupational Status} for the 1988
{International Standard Classification} of {Occupations}. \emph{Social
Science Research}, \emph{25}(3), 201--239.
\url{https://doi.org/10.1006/ssre.1996.0010}

\bibitem[\citeproctext]{ref-gelman_regression_2020}
Gelman, A., Hill, J., \& Vehtari, A. (2020). \emph{Regression and {Other
Stories}} (1st ed.). Cambridge University Press.
\url{https://doi.org/10.1017/9781139161879}

\bibitem[\citeproctext]{ref-gingrich_making_2011}
Gingrich, J. R. (2011). \emph{Making {Markets} in the {Welfare State}:
{The Politics} of {Varying Market Reforms}} (1st ed.). Cambridge
University Press. \url{https://doi.org/10.1017/CBO9780511791529}

\bibitem[\citeproctext]{ref-gugushvili_trends_2014}
Gugushvili, A. (2014). Trends, {Covariates} and {Consequences} of
{Intergenerational Social Mobility} in {Post- Socialist Societies}.

\bibitem[\citeproctext]{ref-gugushvili_intergenerational_2016c}
Gugushvili, A. (2016a). Intergenerational objective and subjective
mobility and attitudes towards income differences: Evidence from
transition societies. \emph{Journal of International and Comparative
Social Policy}, \emph{32}(3), 199--219.
\url{https://doi.org/10.1080/21699763.2016.1206482}

\bibitem[\citeproctext]{ref-gugushvili_intergenerational_2016}
Gugushvili, A. (2016b). Intergenerational {Social Mobility} and {Popular
Explanations} of {Poverty}: {A Comparative Perspective}. \emph{Social
Justice Research}, \emph{29}(4), 402--428.
\url{https://doi.org/10.1007/s11211-016-0275-9}

\bibitem[\citeproctext]{ref-gugushvili_subjective_2017}
Gugushvili, A. (2017). Subjective {Intergenerational Mobility} and
{Support} for {Welfare State Programmes}.

\bibitem[\citeproctext]{ref-gugushvili_heterogeneous_2024}
Gugushvili, A. (2024). The heterogeneous well-being effects of
intergenerational mobility perceptions. \emph{Journal of Health
Psychology}, \emph{29}(2), 99--112.
\url{https://doi.org/10.1177/13591053231187345}

\bibitem[\citeproctext]{ref-hadjar_does_2015}
Hadjar, A., \& Samuel, R. (2015). Does upward social mobility increase
life satisfaction? {A} longitudinal analysis using {British} and {Swiss}
panel data. \emph{Research in Social Stratification and Mobility},
\emph{39}, 48--58. \url{https://doi.org/10.1016/j.rssm.2014.12.002}

\bibitem[\citeproctext]{ref-helgason_longterm_2023}
Helgason, A. F., \& Rehm, P. (2023). Long-term income trajectories and
the evolution of political attitudes. \emph{European Journal of
Political Research}, \emph{62}(1), 264--284.
\url{https://doi.org/10.1111/1475-6765.12506}

\bibitem[\citeproctext]{ref-helgason_class_2025}
Helgason, A. F., \& Rehm, P. (2025). Class experiences and the long-term
evolution of economic values. \emph{Social Forces}, \emph{103}(3),
1125--1143. \url{https://doi.org/10.1093/sf/soae135}

\bibitem[\citeproctext]{ref-hoyt_mindsets_2023}
Hoyt, C. L., Burnette, J. L., Billingsley, J., Becker, W., \& Babij, A.
D. (2023). Mindsets of poverty: {Implications} for redistributive policy
support. \emph{Analyses of Social Issues and Public Policy},
\emph{23}(3), 668--693. \url{https://doi.org/10.1111/asap.12367}

\bibitem[\citeproctext]{ref-immergut_theoretical_1998}
Immergut, E. M. (1998). The {Theoretical Core} of the {New
Institutionalism}. \emph{Politics \& Society}, \emph{26}(1), 5--34.
\url{https://doi.org/10.1177/0032329298026001002}

\bibitem[\citeproctext]{ref-immergut_it_2020}
Immergut, E. M., \& Schneider, S. M. (2020). Is it unfair for the
affluent to be able to purchase {``better''} healthcare? {Existential}
standards and institutional norms in healthcare attitudes across 28
countries. \emph{Social Science \& Medicine}, \emph{267}, 113146.
\url{https://doi.org/10.1016/j.socscimed.2020.113146}

\bibitem[\citeproctext]{ref-jaime-castillo_social_2019}
Jaime-Castillo, A. M., \& Marqués-Perales, I. (2019). Social mobility
and demand for redistribution in {Europe}: A comparative analysis.
\emph{The British Journal of Sociology}, \emph{70}(1), 138--165.
\url{https://doi.org/10.1111/1468-4446.12363}

\bibitem[\citeproctext]{ref-janmaat_subjective_2013}
Janmaat, J. G. (2013). Subjective {Inequality}: A {Review} of
{International Comparative Studies} on {People}'s {Views} about
{Inequality}. \emph{European Journal of Sociology}, \emph{54}(3),
357--389. \url{https://doi.org/10.1017/S0003975613000209}

\bibitem[\citeproctext]{ref-jasso_justice_1978}
Jasso, G. (1978). On the {Justice} of {Earnings}: {A New Specification}
of the {Justice Evaluation Function}. \emph{American Journal of
Sociology}, \emph{83}(6), 1398--1419.
\url{https://doi.org/10.1086/226706}

\bibitem[\citeproctext]{ref-jasso_distributive_2016}
Jasso, G., Törnblom, K. Y., \& Sabbagh, C. (2016). Distributive
{Justice}. In C. Sabbagh \& M. Schmitt (Eds.), \emph{Handbook of {Social
Justice Theory} and {Research}} (pp. 201--218). New York, NY: Springer.
\url{https://doi.org/10.1007/978-1-4939-3216-0_11}

\bibitem[\citeproctext]{ref-jasso_gender_1999}
Jasso, G., \& Wegener, B. (1999). Gender and {Country Differences} in
the {Sense} of {Justice}: {Justice Evaluation}, {Gender Earnings Gap},
and {Earnings Functions} in {Thirteen Countries}. \emph{International
Journal of Comparative Sociology}, \emph{40}(1), 94--115.
\url{https://doi.org/10.1177/002071529904000106}

\bibitem[\citeproctext]{ref-kluegel_legitimation_1999}
Kluegel, J. R., Mason, D. S., \& Wegener, B. (1999). The {Legitimation}
of {Capitalism} in the {Postcommunist Transition}: {Public Opinion}
about {Market Justice}, 1991-1996. \emph{European Sociological Review},
\emph{15}(3), 251--283. Retrieved from
\url{https://www.jstor.org/stable/522731}

\bibitem[\citeproctext]{ref-kluegel_beliefs_1981}
Kluegel, J. R., \& Smith, E. R. (1981). Beliefs {About Stratification}.
\emph{Annual Review of Sociology}, 29--56.

\bibitem[\citeproctext]{ref-koos_moral_2019}
Koos, S., \& Sachweh, P. (2019). The moral economies of market
societies: Popular attitudes towards market competition, redistribution
and reciprocity in comparative perspective. \emph{Socio-Economic
Review}, \emph{17}(4), 793--821.
\url{https://doi.org/10.1093/ser/mwx045}

\bibitem[\citeproctext]{ref-lane_market_1986}
Lane, R. E. (1986). Market {Justice}, {Political Justice}.
\emph{American Political Science Review}, \emph{80}(2), 383--402.
\url{https://doi.org/10.2307/1958264}

\bibitem[\citeproctext]{ref-langsaether_explaining_2022}
Langsæther, P. E., Evans, G., \& O'Grady, T. (2022). Explaining the
{Relationship Between Class Position} and {Political Preferences}: {A
Long-Term Panel Analysis} of {Intra-Generational Class Mobility}.
\emph{British Journal of Political Science}, \emph{52}(2), 958--967.
\url{https://doi.org/10.1017/S0007123420000599}

\bibitem[\citeproctext]{ref-lee_fairness_2023}
Lee, J.-S., \& Stacey, M. (2023). Fairness perceptions of educational
inequality: The effects of self-interest and neoliberal orientations.
\emph{The Australian Educational Researcher}.
\url{https://doi.org/10.1007/s13384-023-00636-6}

\bibitem[\citeproctext]{ref-lenski_status_1954}
Lenski, G. E. (1954). Status {Crystallization}: {A Non-Vertical
Dimension} of {Social Status}. \emph{American Sociological Review},
\emph{19}(4), 405. \url{https://doi.org/10.2307/2087459}

\bibitem[\citeproctext]{ref-lindh_public_2015}
Lindh, A. (2015). Public {Opinion} against {Markets}? {Attitudes}
towards {Market Distribution} of {Social Services} -- {A Comparison} of
17 {Countries}. \emph{Social Policy \& Administration}, \emph{49}(7),
887--910. \url{https://doi.org/10.1111/spol.12105}

\bibitem[\citeproctext]{ref-lindh_bringing_2023}
Lindh, A., \& McCall, L. (2023). Bringing the market in: An expanded
framework for understanding popular responses to economic inequality.
\emph{Socio-Economic Review}, \emph{21}(2), 1035--1055.
\url{https://doi.org/10.1093/ser/mwac018}

\bibitem[\citeproctext]{ref-llorca-jana_historia_2021}
Llorca-Jaña, M., \& Miller, R. M. D. (2021). \emph{{Historia econ{ó}mica
de Chile desde la independencia}}. Santiago de Chile: RIL editores.

\bibitem[\citeproctext]{ref-lopez-roldan_comparative_2021}
López-Roldán, P., \& Fachelli, S. (Eds.). (2021). \emph{Towards a
{Comparative Analysis} of {Social Inequalities} between {Europe} and
{Latin America}}. Cham: Springer International Publishing.
\url{https://doi.org/10.1007/978-3-030-48442-2}

\bibitem[\citeproctext]{ref-mac-clure_justicia_2024}
Mac-Clure, O., Barozet, E., \& Franetovic, G. (2024). {Justicia
distributiva y posici{ó}n social subjetiva: {\textquestiondown}la
meritocracia justifica la desigualdad de ingresos?} \emph{Convergencia
Revista de Ciencias Sociales}, \emph{31}, 1.
\url{https://doi.org/10.29101/crcs.v31i0.22258}

\bibitem[\citeproctext]{ref-madariaga_three_2020}
Madariaga, A. (2020). The three pillars of neoliberalism: {Chile}'s
economic policy trajectory in comparative perspective.
\emph{Contemporary Politics}, \emph{26}(3), 308--329.
\url{https://doi.org/10.1080/13569775.2020.1735021}

\bibitem[\citeproctext]{ref-mau_inequality_2015}
Mau, S. (2015). \emph{Inequality, {Marketization} and the {Majority
Class}: {Why Did} the {European Middle Classes Accept Neo-Liberalism}?}
Houndmills: Palgrave Macmillan.

\bibitem[\citeproctext]{ref-mideplan_evolucion_2017}
MIDEPLAN. (2017). \emph{Evoluci{ó}n de la pobreza 1990-2017}. Ministerio
de Desarrollo Social y Familia - PNUD.

\bibitem[\citeproctext]{ref-mideplan_estadisticas_2024}
MIDEPLAN. (2024). \emph{Estad{í}sticas {Encuesta Casen} 2022, secci{ó}n
{Salud}}. Observatorio social.

\bibitem[\citeproctext]{ref-mijs_paradox_2019}
Mijs, J. (2019). The paradox of inequality: Income inequality and belief
in meritocracy go hand in hand. \emph{Socio-Economic Review},
\emph{19}(1), 7--35. \url{https://doi.org/10.1093/ser/mwy051}

\bibitem[\citeproctext]{ref-mijs_belief_2022}
Mijs, J., Daenekindt, S., de Koster, W., \& van der Waal, J. (2022).
Belief in {Meritocracy Reexamined}: {Scrutinizing} the {Role} of
{Subjective Social Mobility}. \emph{Social Psychology Quarterly},
\emph{85}(2), 131--141. \url{https://doi.org/10.1177/01902725211063818}

\bibitem[\citeproctext]{ref-miller_selfserving_1975}
Miller, D. T., \& Ross, M. (1975). Self-serving biases in the
attribution of causality: {Fact} or fiction? \emph{Psychological
Bulletin}, \emph{82}(2), 213--225.
\url{https://doi.org/10.1037/h0076486}

\bibitem[\citeproctext]{ref-ministeriodeeducacion_resumen_2023}
Ministerio de Educación. (2023). \emph{Resumen de matr{í}cula por
establecimiento educacional}. Datos abiertos Ministerio de Educaci{ó}n.

\bibitem[\citeproctext]{ref-molina_its_2019}
Molina, M. D., Bucca, M., \& Macy, M. W. (2019). It's not just how the
game is played, it's whether you win or lose. \emph{SCIENCE ADVANCES}.

\bibitem[\citeproctext]{ref-nucleodesociologiacontingente_informe_2020}
Núcleo de Sociología Contingente, {[}NUDESOC{]}. (2020). \emph{Informe
de resultados oficial {Encuesta Zona Cero}}. Santiago de Chile.

\bibitem[\citeproctext]{ref-otero_power_2024}
Otero, G., \& Mendoza, M. (2024). The {Power} of {Diversity}: {Class},
{Networks} and {Attitudes Towards Inequality}. \emph{Sociology},
\emph{58}(4), 851--876. \url{https://doi.org/10.1177/00380385231217625}

\bibitem[\citeproctext]{ref-otero_lives_2022}
Otero, G., Volker, B., Rözer, J., \& Mollenhorst, G. (2022). The lives
of others: {Class} divisions, network segregation, and attachment to
society in {Chile}. \emph{The British Journal of Sociology},
\emph{73}(4), 754--785. \url{https://doi.org/10.1111/1468-4446.12966}

\bibitem[\citeproctext]{ref-panes_criticas_2020}
Panes, D. (2020). \emph{Cr{í}ticas y experiencias obreras en torno al
sistema de {AFP}'s en {Chile} (1981-2020)} (PhD thesis). Universidad de
Chile, Santiago.

\bibitem[\citeproctext]{ref-perez-ahumada_clases_2019}
Pérez-Ahumada, P. (2019). {Clases sociales, sectores econ{ó}micos y
cambios en la estructura social chilena entre 1992 y 2013}.
\emph{Revista de la CEPAL}, \emph{2018}(126), 171--192.
\url{https://doi.org/10.18356/0ce0faed-es}

\bibitem[\citeproctext]{ref-pnud_desiguales_2017}
PNUD (Ed.). (2017). \emph{{Desiguales: or{í}genes, cambios y desaf{í}os
de la brecha social en Chile}}. Santiago, Chile: PNUD : Uqbar Editores.

\bibitem[\citeproctext]{ref-polanyi_great_1975}
Polanyi, K. (1975). \emph{The great transformation} (Repr). New York,
NY: Octagon Books.

\bibitem[\citeproctext]{ref-prag_subjective_2021}
Präg, P., \& Gugushvili, A. (2021). Subjective social mobility and
health in {Germany}. \emph{European Societies}, \emph{23}(4), 464--486.
\url{https://doi.org/10.1080/14616696.2021.1887916}

\bibitem[\citeproctext]{ref-ruiz_chilenos_2014}
Ruiz, C., \& Boccardo, G. (2014). \emph{Los chilenos bajo el
neoliberalismo: Clases y conflicto social} (Primera edici{ó}n). Santiago
de Chile: Ediciones y Publicaciones El Buen Aire.

\bibitem[\citeproctext]{ref-sandel_tyranny_2020}
Sandel, M. J. (2020). \emph{The tyranny of merit: {What}'s become of the
common good?} (First edition). New York: {Farrar, Straus and Giroux}.

\bibitem[\citeproctext]{ref-satz_por_2019}
Satz, D. (2019). \emph{{Por Qu{é} Algunas Cosas No Deber{í}an Estar en
Venta: Los l{í}mites Morales Del Mercado}}. Ciudad Aut{ó}noma de Buenos
Aires: Siglo XXI Editores.

\bibitem[\citeproctext]{ref-schmidt_experience_2011}
Schmidt, A. W. (2011). The experience of social mobility and the
formation of attitudes toward redistribution. In.

\bibitem[\citeproctext]{ref-shariff_income_2016}
Shariff, A. F., Wiwad, D., \& Aknin, L. B. (2016). Income {Mobility
Breeds Tolerance} for {Income Inequality}: {Cross-National} and
{Experimental Evidence}. \emph{Perspectives on Psychological Science},
\emph{11}(3), 373--380. \url{https://doi.org/10.1177/1745691616635596}

\bibitem[\citeproctext]{ref-somma_no_2021}
Somma, N. M., Bargsted, M., Disi Pavlic, R., \& Medel, R. M. (2021). No
water in the oasis: The {Chilean Spring} of 2019--2020. \emph{Social
Movement Studies}, \emph{20}(4), 495--502.
\url{https://doi.org/10.1080/14742837.2020.1727737}

\bibitem[\citeproctext]{ref-song_there_2025}
Song, X., \& Zhou, X. (2025). Is {There} a {Mobility Effect}? {On
Methodological Issues} in the {Mobility Contrast Model}.
\emph{Sociological Methods \& Research}, \emph{54}(4), 1576--1593.
\url{https://doi.org/10.1177/00491241251347983}

\bibitem[\citeproctext]{ref-sorokin_social_1927}
Sorokin, P. A. (1927). \emph{Social {Mobility}}. Harper \& Brothers.

\bibitem[\citeproctext]{ref-streeck_how_2016}
Streeck, W. (2016). \emph{How will capitalism end? Essays on a failing
system}. London: Verso.

\bibitem[\citeproctext]{ref-superintendenciadepensiones_estadisticas_2024}
Superintendencia de Pensiones. (2024). \emph{Estad{í}sticas de
{Afiliados} en {AFP} en {Sistema} de {Pensiones}. {Abril} 2024}.
Superintendencia de Pensiones.

\bibitem[\citeproctext]{ref-svallforsMoralEconomyClass2006a}
Svallfors, S. (2006). \emph{The {Moral Economy} of {Class}: {Class} and
{Attitudes} in {Comparative Perspective}}. Stanford University Press.

\bibitem[\citeproctext]{ref-svallfors_political_2007}
Svallfors, S. (Ed.). (2007). \emph{The {Political Sociology} of the
{Welfare State}: {Institutions}, {Social Cleavages}, and {Orientations}}
(1st ed.). Stanford University Press.
\url{https://doi.org/10.2307/j.ctvr0qv0q}

\bibitem[\citeproctext]{ref-tejero-peregrina_perceived_2025}
Tejero-Peregrina, L., Willis, G., Sánchez-Rodríguez, Á., \&
Rodríguez-Bailón, R. (2025). From {Perceived Economic Inequality} to
{Support} for {Redistribution}: {The Role} of {Meritocracy Perception}.
\emph{International Review of Social Psychology}, \emph{38}(1), 4.
\url{https://doi.org/10.5334/irsp.1013}

\bibitem[\citeproctext]{ref-torche_unequal_2005}
Torche, F. (2005). Unequal {But Fluid}: {Social Mobility} in {Chile} in
{Comparative Perspective}. \emph{American Sociological Review},
\emph{70}(3), 422--450. \url{https://doi.org/10.1177/000312240507000304}

\bibitem[\citeproctext]{ref-torche_intergenerational_2014}
Torche, F. (2014). Intergenerational {Mobility} and {Inequality}: {The
Latin American Case}. \emph{Annual Review of Sociology}, \emph{40}(1),
619--642. \url{https://doi.org/10.1146/annurev-soc-071811-145521}

\bibitem[\citeproctext]{ref-vondemknesebeck_are_2016}
Von Dem Knesebeck, O., Vonneilich, N., \& Kim, T. J. (2016). Are health
care inequalities unfair? {A} study on public attitudes in 23 countries.
\emph{International Journal for Equity in Health}, \emph{15}(1), 61.
\url{https://doi.org/10.1186/s12939-016-0350-8}

\bibitem[\citeproctext]{ref-wen_does_2021}
Wen, F., \& Witteveen, D. (2021). Does perceived social mobility shape
attitudes toward government and family educational investment?
\emph{Social Science Research}, \emph{98}, 102579.
\url{https://doi.org/10.1016/j.ssresearch.2021.102579}

\bibitem[\citeproctext]{ref-wiederkehr_belief_2015}
Wiederkehr, V., Bonnot, V., Krauth-Gruber, S., \& Darnon, C. (2015).
Belief in school meritocracy as a system-justifying tool for low status
students. \emph{Frontiers in Psychology}, \emph{6}.

\bibitem[\citeproctext]{ref-wilson_role_2003}
Wilson, C. (2003). The {Role} of a {Merit Principle} in {Distributive
Justice}. \emph{The Journal of Ethics}, \emph{7}(3), 277--314.
\url{https://doi.org/10.1023/A:1024667228488}

\bibitem[\citeproctext]{ref-wooldridge_introductory_2009}
Wooldridge, J. M. (2009). \emph{{Introductory econometrics: a modern
approach}} (4th ed). Mason, OH: South Western, Cengage Learning.

\bibitem[\citeproctext]{ref-wrightUnderstandingClass2015}
Wright, E. O. (2015). \emph{Understanding {Class}} (Verso).

\bibitem[\citeproctext]{ref-young_rise_1958}
Young, M. (1958). \emph{The rise of the meritocracy}. New Brunswick,
N.J., U.S.A: Transaction Publishers.

\end{CSLReferences}

\section{Supplementary material}\label{supplementary-material}

\end{document}