### 2025-11-17 paper/paper.tex

% Options for packages loaded elsewhere
% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  13pt,
]{article}
\usepackage{xcolor}
\usepackage[margin=2cm]{geometry}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{5}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
  \setmainfont[]{Times New Roman}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{setspace}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\setlength{\emergencystretch}{3em} % prevent overfull lines

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage[noblocks]{authblk}
\renewcommand*{\Authsep}{, }
\renewcommand*{\Authand}{, }
\renewcommand*{\Authands}{, }
\renewcommand\Affilfont{\small}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Preferences for the commodification of pensions in Chile: the rol of intergenerational social mobility},
  pdfauthor={Andreas Laffert Tamayo},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Preferences for the commodification of pensions in Chile: the rol
of intergenerational social mobility}

  \author{Andreas Laffert Tamayo}
            \affil{%
                  Instituto de Sociología, Pontificia Universidad
                  Católica de Chile
              }

\date{}

% Colores extra por nombre
\usepackage[dvipsnames]{xcolor}
% Ajustes de hipervínculos
% \hypersetup{
%   colorlinks = true,
%   citecolor  = DarkSlateBlue,     % << citas
%   urlcolor   = DarkSlateBlue  % URLs/DOIs
% }
% 
\begin{document}
\maketitle
\begin{abstract}
My abstract \newline \textbf{Keywords}: Pension commodification ·
Intergenerational mobility · Causal inference · Meritocracy · Chile
\end{abstract}

\setstretch{1.5}
This document was last modified at 2025-11-17 19:08:04 and it was last
rendered at 2025-11-17 19:08:04.

\section{Introduction}\label{introduction}

What is the legitimate extent of market inequality in the eyes of the
public? Since the early 1980s, many countries have experienced a
widespread retreat from universal welfare programs and a shift toward
the privatization and commodification of public goods, welfare policies,
and social services (\citeproc{ref-gingrich_making_2011}{Gingrich,
2011}; \citeproc{ref-streeck_how_2016}{Streeck, 2016}). In Latin
America, as elsewhere, neoliberal reforms reshaped welfare-state
institutions by extending market logic into domains of social
reproduction that were traditionally governed by the state
(\citeproc{ref-ferre_welfare_2023}{Ferre, 2023}). This transformation
reduced the role of public provision and increased the presence of
private actors in core social services
(\citeproc{ref-ferre_welfare_2023}{Ferre, 2023}). Echoing Polanyi's
(\citeproc{ref-polanyi_great_1975}{1975}) insight that markets
constitute a distinct moral order, the institutional diffusion of market
rules has fostered a corresponding moral economy: a constellation of
norms and values concerning fair allocation, embedded in institutions
and shaping individual subjectivities
(\citeproc{ref-mau_inequality_2015}{Mau, 2015};
\citeproc{ref-svallforsMoralEconomyClass2006a}{Svallfors, 2006}). Within
this framework, a growing body of research examines the extent to which,
and the mechanisms by which, citizens consider it fair that the
allocation of services like health care, pensions, and education be
governed by market-based criteria---a phenomenon known as \emph{market
justice preferences} (\citeproc{ref-busemeyer_skills_2014}{Busemeyer,
2014}; \citeproc{ref-castillo_perceptions_2025}{Castillo, Laffert, et
al., 2025}; \citeproc{ref-castillo_socialization_2024}{Castillo et al.,
2024}; \citeproc{ref-immergut_it_2020}{Immergut \& Schneider, 2020};
\citeproc{ref-koos_moral_2019}{Koos \& Sachweh, 2019};
\citeproc{ref-lindh_public_2015}{Lindh, 2015};
\citeproc{ref-lindh_bringing_2023}{Lindh \& McCall, 2023}).
Understanding these preferences is crucial, as they contribute to
legitimizing economic inequality by framing it as the fair result of
individual responsibility and limited state intervention
(\citeproc{ref-mau_inequality_2015}{Mau, 2015}).

Existing literature shows that market justice preferences are shaped by
both the economic and institutional context and individuals' positions
within social stratification. Grounded in the notion that economic
institutions influence people's normative attitudes
(\citeproc{ref-immergut_theoretical_1998}{Immergut, 1998}), studies find
that countries with stronger public provision or more expansive welfare
states exhibit lower levels of market justice preferences
(\citeproc{ref-busemeyer_skills_2014}{Busemeyer, 2014};
\citeproc{ref-immergut_it_2020}{Immergut \& Schneider, 2020}), while
more privatized contexts show stronger support for market-based criteria
(\citeproc{ref-castillo_socialization_2024}{Castillo et al., 2024};
\citeproc{ref-lindh_public_2015}{Lindh, 2015}). In such contexts, market
justice preferences tend to rise as individuals ``ascend'' the social
structure, with those in more privileged positions in terms of class,
education, and income holding stronger preferences for market-based
solutions compared to those in more disadvantaged or at-risk positions
(\citeproc{ref-castillo_socialization_2024}{Castillo et al., 2024};
\citeproc{ref-immergut_it_2020}{Immergut \& Schneider, 2020};
\citeproc{ref-lee_fairness_2023}{Lee \& Stacey, 2023};
\citeproc{ref-lindh_public_2015}{Lindh, 2015};
\citeproc{ref-otero_power_2024}{Otero \& Mendoza, 2024};
\citeproc{ref-svallfors_political_2007}{Svallfors, 2007};
\citeproc{ref-vondemknesebeck_are_2016}{Von Dem Knesebeck et al.,
2016}).

Market justice preferences are shaped not only by objective
socioeconomic conditions but also by popular beliefs about inequality.
Among these, meritocracy is a key normative principle underpinning
market-based distributive preferences
(\citeproc{ref-mau_inequality_2015}{Mau, 2015}). Studies show that
individuals with stronger meritocratic beliefs tend to perceive less
inequality and legitimize it by attributing economic differences to
personal achievement (\citeproc{ref-batruch_belief_2023}{Batruch et al.,
2023}; \citeproc{ref-mijs_paradox_2019}{Mijs, 2019};
\citeproc{ref-wilson_role_2003}{Wilson, 2003}). In highly unequal
societies where access to services is largely governed by market logic,
such beliefs play a critical role in normalizing inequality. Recent
evidence from Chile shows that students who believe effort and talent
are rewarded in their country express stronger preferences for
market-based access to healthcare, pensions, and education
(\citeproc{ref-castillo_socialization_2024}{Castillo et al., 2024}).

Although it is clear that one's social position influences market
justice preferences, the question of how upward or downward mobility
within the social structure affects these preferences remains
unanswered. This question is far from trivial, especially in Latin
America, where many have experienced various forms of mobility amid high
economic inequality and deep welfare privatization
(\citeproc{ref-ferre_welfare_2023}{Ferre, 2023};
\citeproc{ref-lopez-roldan_comparative_2021}{López-Roldán \& Fachelli,
2021}; \citeproc{ref-torche_intergenerational_2014}{Torche, 2014}).
Social origins and destinations affect attitudes toward inequality in
distinct ways (\citeproc{ref-day_movin_2017}{Day \& Fiske, 2017};
\citeproc{ref-gugushvili_intergenerational_2016}{Gugushvili, 2016b},
\citeproc{ref-gugushvili_subjective_2017}{2017};
\citeproc{ref-jaime-castillo_social_2019}{Jaime-Castillo \&
Marqués-Perales, 2019}; \citeproc{ref-mijs_belief_2022}{Mijs et al.,
2022}; \citeproc{ref-wen_does_2021}{Wen \& Witteveen, 2021}), while
movement between these positions exposes individuals to different
experiences and mechanisms that shape their views on what is fair
(\citeproc{ref-gugushvili_trends_2014}{Gugushvili, 2014};
\citeproc{ref-mau_inequality_2015}{Mau, 2015}). Building on this
research, examining the effects of social mobility on market justice
preferences can help to illuminate how inequalities in access to social
services are justified among individuals who have experienced, or not,
changes in their social standing, and what are the normative mechanisms
that guide this justification (\citeproc{ref-mau_inequality_2015}{Mau,
2015}).

Beyond their isolated effects, social mobility and meritocratic beliefs
interact in complex ways to shape market justice preferences. Among
others, a key mechanism proposed in the literature to explain how
mobility influences distributive justice preferences is the
psychological process of self-serving bias in causal attribution
(\citeproc{ref-gugushvili_intergenerational_2016c}{Gugushvili, 2016a};
\citeproc{ref-schmidt_experience_2011}{Schmidt, 2011}). This bias
suggests that individuals attribute failures---such as downward
mobility---to external factors, while crediting successes---such as
upward mobility---to their own merit and effort
(\citeproc{ref-miller_selfserving_1975}{Miller \& Ross, 1975}). Those
who experience upward mobility tend to view their social position as
earned, making them more likely to believe that individuals are
responsible for their own success or failure. Research shows that upward
mobility is associated with weaker preferences for redistribution
(\citeproc{ref-alesina_intergenerational_2018}{Alesina et al., 2018};
\citeproc{ref-gugushvili_intergenerational_2016c}{Gugushvili, 2016a};
\citeproc{ref-jaime-castillo_social_2019}{Jaime-Castillo \&
Marqués-Perales, 2019}; \citeproc{ref-schmidt_experience_2011}{Schmidt,
2011}) and stronger legitimacy of income inequality
(\citeproc{ref-shariff_income_2016}{Shariff et al., 2016}). In contrast,
individuals who experience downward mobility tend to blame structural
factors like inequality and are more supportive of redistribution while
rejecting merit-based explanations
(\citeproc{ref-gugushvili_trends_2014}{Gugushvili, 2014}). Taken
together, I argue that meritocratic beliefs may reinforce this
self-serving attribution mechanism by legitimizing one's social status
as the outcome of personal merit, closely tied to attribution bias.

Against this background, this article pursues two main objectives:
first, to analyze the extent to which intergenerational social mobility
influences market justice preferences regarding healthcare, pensions,
and education; and second, to examine how meritocratic beliefs may
moderate this relationship. Building on a theoretical framework that
emphasizes how neoliberal transformations---particularly through the
privatization and commodification of key areas of social
reproduction---have profoundly reshaped processes of subject formation
(\citeproc{ref-mau_inequality_2015}{Mau, 2015}), the central argument is
that upward mobility increases support for market justice preferences,
while downward mobility decreases it. Moreover, meritocratic beliefs are
expected to moderate this relationship by reflecting a self-serving
attribution bias, whereby individuals justify their social position in
terms of personal merit.

This study focuses on Chile, a particularly intriguing case for
examining market justice preferences. Despite significant economic
growth and poverty reduction, Chile has some of the highest levels of
inequality in Latin America and among OECD countries
(\citeproc{ref-chancel_world_2022}{Chancel et al., 2022};
\citeproc{ref-flores_top_2020}{Flores et al., 2020}). This inequality
coexists with short-range upward mobility among lower-class segments
moving into middle strata, though strong barriers remain to reaching
higher positions (\citeproc{ref-espinoza_estratificacion_2013}{Espinoza
et al., 2013}; \citeproc{ref-torche_intergenerational_2014}{Torche,
2014}). What makes Chile especially salient is that much of this
inequality is rooted in deep neoliberal reforms that institutionalized
the privatization and commodification of key social sectors
(\citeproc{ref-madariaga_three_2020}{Madariaga, 2020}). Introduced
during the dictatorship (1973--1989) and expanded in democracy, these
reforms enabled the unprecedented emergence of markets in health,
pensions, and education, with provision segmented by individuals'
ability to pay and supported by public subsidies
(\citeproc{ref-boccardo_30_2020}{Boccardo, 2020}). In parallel---and
despite waves of protest against inequality and commodification from
2006 to 2019 (\citeproc{ref-somma_no_2021}{Somma et al.,
2021})---Chilean subjectivities have been increasingly shaped by
neoliberal discourses and market logics, influencing their attitudes
toward inequality and welfare distribution
(\citeproc{ref-araujo_desafios_2012}{Araujo \& Martuccelli, 2012};
\citeproc{ref-canalesceron_sujeto_2021}{Canales Cerón et al., 2021}).

In this context, the questions that guide this research are as follows:

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\tightlist
\item
  To what extent does intergenerational social mobility influence market
  justice preferences regarding healthcare, pensions, and education in
  Chile?
\item
  How do meritocratic beliefs condition or moderate this relationship in
  the Chilean context?
\end{enumerate}

To address these questions, this study draws on large-scale,
representative survey data collected in 2018 from the urban Chilean
population aged 18 to 75 (n = 2,726). The next section outlines the
theoretical framework linking market justice preferences, social
mobility, and meritocratic beliefs, and proposes a set of hypotheses.
This is followed by a description of the data, variables, and analytical
strategy. The final sections present the empirical findings, offer an
interpretation of the results, and conclude with a discussion of their
implications.

\section{Theoretical and empirical
background}\label{theoretical-and-empirical-background}

\subsection{Market justice
preferences}\label{market-justice-preferences}

Beyond state's capacity to reallocate resources from the advantaged to
the vulnerable, market institutions also play a central role in
distributing socially valuable goods and rewards
(\citeproc{ref-koos_moral_2019}{Koos \& Sachweh, 2019};
\citeproc{ref-lindh_bringing_2023}{Lindh \& McCall, 2023}). As Polanyi
(\citeproc{ref-polanyi_great_1975}{1975}) observed, economic integration
in capitalist societies is primarily organized through market exchange,
governed by a self-regulating price system embedded in institutional
frameworks. These institutions are not mere aggregates of individual
behavior, but social realities endowed with rules, mechanisms, and
normative meanings that shape everyday thinking
(\citeproc{ref-immergut_theoretical_1998}{Immergut, 1998};
\citeproc{ref-koos_moral_2019}{Koos \& Sachweh, 2019}). In this sense,
the economic order is mirrored in a moral economy: collectively shared
norms and beliefs about justice in distribution, embedded and reinforced
through institutions (\citeproc{ref-mau_inequality_2015}{Mau, 2015};
\citeproc{ref-svallfors_political_2007}{Svallfors, 2007}). While most
research from this perspective has focused on welfare institutions,
recent scholarship has brought the market back into focus as a site of
distributive justice beliefs and institutional responses to inequality
(\citeproc{ref-lindh_bringing_2023}{Lindh \& McCall, 2023}). In many
countries, privatization and commodification have expanded market logic
into core areas of social reproduction, such as healthcare, education,
pensions, and social security, deepening inequality in access to these
services (\citeproc{ref-ferre_welfare_2023}{Ferre, 2023};
\citeproc{ref-gingrich_making_2011}{Gingrich, 2011}). Yet public support
for market-based welfare provision has grown, even in traditional
welfare states, particularly among higher-income groups who view private
alternatives as more efficient or higher in quality
(\citeproc{ref-busemeyer_welfare_2020}{Busemeyer \& Iversen, 2020}).
This shift calls for broader inquiry into how institutions like markets
structure access to resources and legitimize inequalities
(\citeproc{ref-mau_inequality_2015}{Mau, 2015};
\citeproc{ref-satz_por_2019}{Satz, 2019}).

The legitimacy of market-based inequalities is closely tied to beliefs
about distributive justice grounded in market principles. Beyond
stratification structures and socioeconomic conditions, the study of
inequality also focuses on individuals' beliefs about the origins of
inequality, the normative frameworks that sustain those beliefs, the
mechanisms that shape them, and their implications for attitudes and
behavior (\citeproc{ref-kluegel_beliefs_1981}{Kluegel \& Smith, 1981}).
In this regard, empirical research on distributive justice focuses on
individuals' conceptions of how goods and rewards \emph{should} be
distributed in society (\citeproc{ref-jasso_distributive_2016}{Jasso et
al., 2016}). This line of inquiry allows for examining the extent to
which economic inequality is perceived as just, or in a certain sense,
legitimate (\citeproc{ref-castillo_legitimacy_2011}{Castillo, 2011}).
Most of the research on this field explores the legitimacy of wage
inequality, specially salary gaps between jobs at opposite ends of the
occupational hierarchy
(\citeproc{ref-castillo_legitimacy_2011}{Castillo, 2011};
\citeproc{ref-jasso_justice_1978}{Jasso, 1978};
\citeproc{ref-jasso_gender_1999}{Jasso \& Wegener, 1999}). Recently,
others examines how individuals justify market-generated inequalities in
access to core social services such as healthcare, education, pensions
or social security. Here, legitimacy stems from the belief that access
to these goods should follow market-based criteria
(\citeproc{ref-castillo_socialization_2024}{Castillo et al., 2024};
\citeproc{ref-lindh_public_2015}{Lindh, 2015}). In such views, these
services are treated as legitimate commodities; goods that can be
traded, priced, and evaluated through market logic
(\citeproc{ref-busemeyer_welfare_2020}{Busemeyer \& Iversen, 2020}).

Market justice preferences refer to normative beliefs that legitimize
the idea that access to core social services---such as healthcare,
education, or pensions---should be determined by market-based criteria.
Following Janmaat's (\citeproc{ref-janmaat_subjective_2013}{2013, p.
359}) distinction, these preferences fall under the category of
``beliefs,'' understood as normative ideas about what inequality should
look like, as opposed to ``perceptions,'' which refer to subjective
evaluations of existing inequality. Market justice preferences reflect
the view that access to these services should depend on individuals'
ability to pay, thus justifying inequalities generated by market
mechanisms (\citeproc{ref-kluegel_legitimation_1999}{Kluegel et al.,
1999}; \citeproc{ref-lindh_public_2015}{Lindh, 2015}). The concept draws
on Lane's (\citeproc{ref-lane_market_1986}{1986}) classic contrast
between market justice and political justice: the former is grounded in
the idea of earned deserts---where rewards reflect effort, productivity,
and skill---while the latter prioritizes need and equality, typically
expressed in welfare state policies. Lane argued that markets and states
differ in purpose (efficiency vs.~need), logic (individual
vs.~collective), and fairness criteria (merit vs.~equality). Market
justice assumes that markets are neutral, self-regulating systems in
which fair procedures yield outcomes proportional to merit. Inequality,
from this perspective, is not only expected but legitimate---as long as
it arises from fair competition. In this way, market justice offers a
moral lens through which individuals can view the commodification and
stratified access to social services as fair and justified
(\citeproc{ref-kluegel_legitimation_1999}{Kluegel et al., 1999};
\citeproc{ref-lindh_public_2015}{Lindh, 2015}).

The empirical study of market justice preferences has employed various
strategies to capture how individuals assess inequalities arising from
market allocation. A common approach evaluates whether people find it
fair that access to essential social services---such as healthcare,
education, or pensions---depends on income. This builds on foundational
work by Kluegel and Smith
(\citeproc{ref-kluegel_legitimation_1999}{1999}), who examined normative
justifications for capitalist systems. Subsequent studies have extended
this logic to a wider set of welfare goods. For example, Immergut and
Schneider (\citeproc{ref-immergut_it_2020}{2020}) and von dem Knesebeck
et al. (\citeproc{ref-vondemknesebeck_are_2016}{2016}) explore whether
respondents believe it is just that higher-income individuals receive
better healthcare. Similarly, Lee and Stacey
(\citeproc{ref-lee_fairness_2023}{2023}) and Castillo, Iturra, et al.
(\citeproc{ref-castillo_changes_2025}{2025}) apply this framework to
education. Complementing these efforts, more recent research has
introduced composite indicators to capture broader orientations toward
market-based allocation. Lindh (\citeproc{ref-lindh_public_2015}{2015}),
for instance, constructs an index averaging support for income-based
access to healthcare and education in comparative perspective, while
Castillo et al. (\citeproc{ref-castillo_socialization_2024}{2024})
propose a single-item measure covering health, education, and pensions
in Chile. These instruments seek to gauge to what extent individuals
view market-generated inequalities as legitimate. Taken together, they
capture two core dimensions of market distribution: the role of economic
resources as a key determinant of outcomes, and the framing of services
as tradable commodities that can be bought and sold according to ability
to pay (\citeproc{ref-lindh_public_2015}{Lindh, 2015}).

Comparative empirical research has identified several individual-level
factors that influence support for market justice. Individuals in more
advantaged socioeconomic positions---those with higher income,
education, and occupational status---are consistently more likely to
endorse market-based distributive principles
(\citeproc{ref-koos_moral_2019}{Koos \& Sachweh, 2019};
\citeproc{ref-lindh_public_2015}{Lindh, 2015};
\citeproc{ref-svallfors_political_2007}{Svallfors, 2007}). For example,
Lindh (\citeproc{ref-lindh_public_2015}{2015}) finds that individuals
from the service class are more likely to support market-based access to
healthcare and education than skilled and unskilled workers across 17
relatively affluent countries. In a comparative analysis, Svallfors
(\citeproc{ref-svallfors_political_2007}{2007}) observes that this
expected class pattern appears clearly only in Sweden, where support for
private education and healthcare varies systematically by class.
Busemeyer (\citeproc{ref-busemeyer_skills_2014}{2014}) similarly shows
that support for private education is stronger among high-income groups,
while Immergut and Schneider (\citeproc{ref-immergut_it_2020}{2020}) and
von dem Knesebeck et al. (\citeproc{ref-vondemknesebeck_are_2016}{2016})
report comparable findings for healthcare, suggesting that wealthier
individuals perceive private provision as a means of maintaining
relative advantage. In Chile, Otero and Mendoza
(\citeproc{ref-otero_power_2024}{2024}) show that individuals with
higher income and university education express stronger support for
market allocation in healthcare, education, and pensions. Beliefs about
inequality and political orientation also matter. Castillo et al.
(\citeproc{ref-castillo_socialization_2024}{2024}) and Castillo,
Laffert, et al. (\citeproc{ref-castillo_perceptions_2025}{2025}) show
that individuals who have strong meritocratic beliefs are more likely to
support market-based distribution in Chile, while Lee and Stacy
(\citeproc{ref-lee_fairness_2023}{2023}) in Australia suggest that these
preferences are also greater as people lean toward economic
conservatism. In this sense, market justice preferences are shaped by
the interplay between structural position and normative reasoning.

However, individual characteristics alone do not fully explain variation
in support for market justice. A growing body of cross-national research
highlights the importance of institutional arrangements in shaping these
preferences. For instance, Immergut and Schneider
(\citeproc{ref-immergut_it_2020}{2020}) find that in countries with
higher public spending on healthcare, individuals are less likely to
view income-based access as fair. Similarly, Busemeyer
(\citeproc{ref-busemeyer_skills_2014}{2014}) shows that increased public
investment in education is associated with lower support for privatized
provision. Conversely, Lindh (\citeproc{ref-lindh_public_2015}{2015})
finds that in countries with more market-oriented welfare systems,
support for market-based distribution tends to be higher, suggesting
that individual attitudes often align with institutional outputs. These
findings are consistent with neo-institutionalist and policy feedback
theories (\citeproc{ref-campbell_institutional_2020}{Campbell, 2020}),
which argue that institutions do more than redistribute resources---they
also shape the normative categories through which individuals assess who
is deserving of support
(\citeproc{ref-immergut_theoretical_1998}{Immergut, 1998}). In this
view, dominant values and preferences are both shaped by and embedded in
institutional configurations
(\citeproc{ref-busemeyer_skills_2014}{Busemeyer, 2014}), reinforcing the
notion that institutions are not neutral structures, but active
producers of the moral frameworks that legitimize or challenge
inequality.

\subsection{Social mobility}\label{social-mobility}

The study of social mobility, its drivers and consequences has long been
central to sociology. Classic theorists explored not only movement
across social hierarchies but also its broader implications for class
conflict, norm stability, and institutional change
(\citeproc{ref-breen_effects_2024}{Breen \& Ermisch, 2024}). Sorokin
(\citeproc{ref-sorokin_social_1927}{1927}) introduced the concept
formally, defining mobility as the shift of individuals, values, or
objects between positions within a stratification system, and
distinguishing between horizontal and vertical forms. Later work
differentiated intergenerational from intragenerational mobility, as
well as absolute mobility---driven by structural change---from relative
mobility, which captures the extent to which origins constrain
destinations (\citeproc{ref-eyles_social_2022}{Eyles et al., 2022}). In
Latin America, research shows high absolute but low relative mobility
(\citeproc{ref-bucca_merit_2016}{Bucca, 2016}): although educational
expansion and economic modernization have enabled some upward movement,
status reproduction remains strong, especially among elites
(\citeproc{ref-lopez-roldan_comparative_2021}{López-Roldán \& Fachelli,
2021}; \citeproc{ref-torche_intergenerational_2014}{Torche, 2014}). This
reflects deep structural inequality, segmented education systems,
stratified labor markets, and legacies of dependent development that
restrict access to mobility channels and reinforce the intergenerational
transmission of advantage.

Beyond mapping mobility patterns, growing research has examined its
subjective and attitudinal effects. Mobility effects---defined as
outcomes resulting from movement between origin and destination classes
(\citeproc{ref-breen_effects_2024}{Breen \& Ermisch, 2024})---have long
attracted theoretical interest. Sorokin's (1959) dissociative hypothesis
posits that mobility, whether upward or downward, may produce
psychological strain due to conflicting norms between class contexts,
leading to lower life satisfaction. Similarly, Lenski
(\citeproc{ref-lenski_status_1954}{1954}) argued that status
inconsistency---mismatches among education, income, and occupation---can
undermine well-being. These ideas underpin extensive empirical work
linking intergenerational mobility to outcomes such as life
satisfaction, mental health, and stress
(\citeproc{ref-gugushvili_heterogeneous_2024}{Gugushvili, 2024};
\citeproc{ref-hadjar_does_2015}{Hadjar \& Samuel, 2015};
\citeproc{ref-prag_subjective_2021}{Präg \& Gugushvili, 2021}).

Research on intergenerational social mobility has increasingly examined
its effects on attitudes toward economic inequality. A key area of
inquiry focuses on how upward and downward mobility influence support
for redistribution, though findings remain mixed. Alesina et al.
(\citeproc{ref-alesina_intergenerational_2018}{2018}) show that
individuals with pessimistic expectations about their
mobility---particularly those anticipating downward movement---are more
likely to support generous redistributive policies. Similarly, Ares
(\citeproc{ref-ares_changing_2020}{2020}) finds that upwardly mobile
individuals tend to be less supportive of state-led redistribution
compared to those who have experienced downward mobility. Comparative
studies by Schmidt (\citeproc{ref-schmidt_experience_2011}{2011}) and
Gugushvili (\citeproc{ref-gugushvili_subjective_2017}{2017}) likewise
report that subjective upward mobility is associated with weaker
preferences for redistribution, while downward mobility strengthens
redistributive support. However, recent causal evidence from Breen and
Ermisch (\citeproc{ref-breen_effects_2024}{2024}) suggests the opposite:
upward mobility may increase redistributive preferences, while downward
mobility may reduce them.

Beyond redistribution, other studies have explored the impact of
mobility on broader beliefs about inequality. Gugushvili
(\citeproc{ref-gugushvili_intergenerational_2016}{2016b}) finds that
upwardly mobile individuals are more likely to adopt individualistic
attributions of poverty and to legitimize income inequality,
particularly in post-socialist societies. Similarly, Bucca
(\citeproc{ref-bucca_merit_2016}{2016}) shows that subjective upward
mobility reinforces individualistic explanations of wealth in seven
Latin American countries. At the macro level, Shariff et al.
(\citeproc{ref-shariff_income_2016}{2016}) demonstrate that higher
levels of national economic mobility correlate with greater tolerance of
inequality. In contrast, Day and Fiske
(\citeproc{ref-day_movin_2017}{2017}) find that low perceived mobility
undermines belief in meritocracy and a just world, thereby weakening
system justification. Taken together, this body of research suggests
that mobility shapes attitudes toward inequality and justice through
multiple, and sometimes contradictory, mechanisms.

The literature on the effects of social mobility has proposed various
mechanisms to explain how and why changes in social position may
influence individual outcomes
(\citeproc{ref-helgason_class_2025}{Helgason \& Rehm, 2025}). One of the
most prominent is the self-interest mechanism, which posits that
individuals who experience upward or downward mobility undergo a shift
in their material interests, thereby altering their perceptions and
preferences (\citeproc{ref-ares_changing_2020}{Ares, 2020};
\citeproc{ref-helgason_longterm_2023}{Helgason \& Rehm, 2023};
\citeproc{ref-langsaether_explaining_2022}{Langsæther et al., 2022}).
Closely related to this logic is the Prospect of Upward Mobility (POUM)
hypothesis, which suggests that individuals may oppose redistribution
not because of their current position, but because they anticipate
improving their status in the future
(\citeproc{ref-benabou_social_2001}{Benabou \& Ok, 2001}). A second line
of explanation draws on the framework of theories of socialization.
Within this tradition, hypotheses such as acculturation, socialization,
and status maximization propose that individuals adjust their attitudes
based on the norms and values of either their class of origin or their
destination, or a combination of both
(\citeproc{ref-jaime-castillo_social_2019}{Jaime-Castillo \&
Marqués-Perales, 2019}). However, most of these mechanisms focus
primarily on the indirect effects of origin and destination positions,
rather than on the direct effect of experience of movement itself. In
response to this gap, recent research has highlighted the role of
self-serving bias in causal attribution processes, suggesting that
individuals tend to explain their mobility trajectories in ways that
justify their current position, which in turn shapes their beliefs and
preferences
(\citeproc{ref-gugushvili_intergenerational_2016c}{Gugushvili, 2016a};
\citeproc{ref-molina_its_2019}{Molina et al., 2019};
\citeproc{ref-schmidt_experience_2011}{Schmidt, 2011}).

The self-serving bias mechanism builds on a value-oriented perspective,
emphasizing that individuals' experiences of social mobility shape their
causal attributions, which in turn influence their beliefs about justice
and distributive preferences
(\citeproc{ref-gugushvili_trends_2014}{Gugushvili, 2014}). Causal
attribution refers to the process through which individuals generate
explanations for their own behavior and outcomes, as well as those of
others (\citeproc{ref-gugushvili_intergenerational_2016}{Gugushvili,
2016b}). In this view, people's interpretations of economic inequality
depend on whether they believe such disparities reflect unequal
individual contributions. Individuals who adopt an internal attribution
framework tend to see success or failure as rooted in personal
characteristics such as effort, talent, or merit. In contrast, those who
regard inequality as unjust are more likely to adopt an external
attribution model, viewing outcomes as the result of structural barriers
beyond individual control (\citeproc{ref-kluegel_beliefs_1981}{Kluegel
\& Smith, 1981}). This mechanism, often described as intrapersonal
causal attribution, reflects how people explain their own socioeconomic
positions---typically attributing their successes to internal qualities
while blaming failures on external circumstances
(\citeproc{ref-miller_selfserving_1975}{Miller \& Ross, 1975}). Over
time, individuals may revise their beliefs and attitudes: while early
views are shaped by their social origin, these are later adjusted in
light of personal experiences of mobility and the perceived role of
ascribed versus achieved factors in determining socioeconomic outcomes
(\citeproc{ref-gugushvili_intergenerational_2016}{Gugushvili, 2016b}).

Empirical research has provided support for the self-serving bias
mechanism in shaping redistributive preferences and attitudes toward the
legitimacy of inequality. For instance, Schmidt
(\citeproc{ref-schmidt_experience_2011}{2011}) finds that individuals
who experience upward mobility are more likely to interpret their
success as the result of personal effort or merit, and consequently
perceive redistribution as less necessary. Conversely, individuals who
experience downward mobility tend to attribute their decline to external
circumstances---such as structural inequality or unemployment---and show
stronger support for redistribution. In a comparative analysis across
different welfare domains, Gugushvili
(\citeproc{ref-gugushvili_subjective_2017}{2017}) demonstrates that
upward mobility is associated with lower support for government spending
on housing and pensions, while individuals who experience downward
mobility express lower support for healthcare and education spending,
but favor increased investment in housing and pensions, reflecting the
material nature of these domains. Moreover, Gugushvili
(\citeproc{ref-gugushvili_intergenerational_2016c}{2016a}) finds that
upward mobility is linked to greater justification of income inequality,
suggesting that improved social standing reinforces an attributional
view in which success is seen as the result of individual
characteristics---thus legitimizing inequality as a fair outcome.

Consequently, considering this theoretical and empirical background, the
first hypothesis of this research is that:

\(H1\): Experiencing upward (downward) social mobility is positively
(negatively) associated with greater support for market justice in
healthcare, pensions, and education.

\subsection{Meritocracy}\label{meritocracy}

Meritocracy constitutes a central ideological framework for legitimizing
different types of social inequality, for instance through market
justice beliefs. Rooted in the belief that rewards and positions should
be allocated based on individual effort and talent, meritocracy operates
as a normative ideal and a descriptive belief about how society
functions. As initially conceptualized by Michael Young
(\citeproc{ref-young_rise_1958}{1958}), the term was meant to critique a
system in which merit-based stratification becomes a new form of
inequality. However, over time, meritocracy has been widely supported in
many societies as a fair and desirable principle of distribution,
particularly within liberal democracies and market-oriented economies
(\citeproc{ref-mijs_paradox_2019}{Mijs, 2019};
\citeproc{ref-sandel_tyranny_2020}{Sandel, 2020}). From a sociological
perspective, the belief in meritocracy is more than a cognitive
assessment; it reflects a moral lens through which individuals interpret
inequality. People who believe that success results from hard work and
talent are more likely to view social and economic disparities as
legitimate (\citeproc{ref-batruch_belief_2023}{Batruch et al., 2023};
\citeproc{ref-castillo_meritocracia_2019}{Castillo et al., 2019}).
Conversely, if they see outcomes as driven by luck, social origin, or
systemic barriers, inequality is more likely to be perceived as unjust.
This distinction becomes crucial in societies with persistent structural
inequality, where public narratives often emphasize personal
responsibility and merit while overlooking entrenched disadvantages.

I adopt a multidimensional perspective on meritocracy, distinguishing
between two key dimensions: effort-based and talent-based perceptions.
This distinction is essential, as it captures different pathways through
which individuals justify inequality
(\citeproc{ref-young_rise_1958}{Young, 1958}). Effort-based meritocracy
emphasizes hard work and perseverance as the basis for social rewards,
aligning closely with cultural narratives of personal responsibility. A
talent-based meritocracy, by contrast, emphasizes intelligence and
innate abilities, which are often perceived as less malleable and more
unequally distributed. Both dimensions have been shown to correlate with
acceptance of inequality, but they may carry distinct implications for
how inequality is justified in specific domains
(\citeproc{ref-castillo_multidimensional_2023}{Castillo et al., 2023}).
The relevance of this distinction is supported by recent studies, which
show that individuals respond differently to these dimensions. For
instance, perceptions that effort is rewarded in society are more
strongly associated with positive evaluations of fairness and acceptance
of unequal outcomes (\citeproc{ref-batruch_belief_2023}{Batruch et al.,
2023}; \citeproc{ref-wiederkehr_belief_2015}{Wiederkehr et al., 2015};
\citeproc{ref-wilson_role_2003}{Wilson, 2003}). This may be because
effort is seen as a controllable and morally virtuous trait, whereas
talent is often perceived as a natural advantage. Consequently,
effort-based meritocracy is likely more potent in legitimizing
inequality, particularly in neoliberal contexts.

These dimensions of meritocracy reflect how respondents perceive
society's distributive logic, regardless of whether they endorse
meritocratic principles. This distinction aligns with recent findings
indicating that individuals distinguish between how merit is perceived
in society and how it should ideally operate, which in turn shapes their
preferences for redistribution and justice
(\citeproc{ref-tejero-peregrina_perceived_2025}{Tejero-Peregrina et al.,
2025}). Meritocratic beliefs serve as symbolic justifications for
unequal outcomes, particularly when access is stratified by income or
social opportunity. Prior studies in Chile have shown that individuals
who perceive higher levels of meritocracy tend to express stronger
support for unequal distributions that reflect market outcomes in
healthcare, education and pensions
(\citeproc{ref-castillo_perceptions_2025}{Castillo, Laffert, et al.,
2025}; \citeproc{ref-castillo_socialization_2024}{Castillo et al.,
2024}).

In addition to influencing individual attitudes toward inequality,
meritocratic beliefs can contribute to social division and the
stigmatization of disadvantaged groups. Recent research has demonstrated
that exposure to meritocratic narratives can reinforce the belief that
poverty results from individual failings rather than systemic
conditions, reducing support for redistributive measures and increasing
the stigmatization of the poor (\citeproc{ref-hoyt_mindsets_2023}{Hoyt
et al., 2023}). This reinforces negative stereotypes and reduces empathy
toward individuals from lower socioeconomic backgrounds. Moreover,
Busemeyer et al. (\citeproc{ref-busemeyer_positive_2021}{2021}) argues
that meritocratic narratives can serve as feedback mechanisms that shape
public opinion and well-being by framing individuals' understanding of
welfare outcomes as deserved or undeserved within existing institutional
structures. This psychological mechanism highlights the normative power
of meritocracy in stabilizing unequal systems by shaping political
attitudes and personal perceptions of success and failure.

Importantly, recent research has explored how meritocratic beliefs
interact with experiences of intergenerational mobility to shape
distributive attitudes. The belief that one's success is earned can lead
upwardly mobile individuals to internalize meritocratic narratives and
justify existing inequalities, reinforcing support for market justice
(\citeproc{ref-gugushvili_intergenerational_2016c}{Gugushvili, 2016a};
\citeproc{ref-molina_its_2019}{Molina et al., 2019}). Conversely,
downwardly mobile individuals who maintain strong meritocratic beliefs
may interpret their status as a personal failure, reducing their support
for redistribution (\citeproc{ref-day_movin_2017}{Day \& Fiske, 2017}).
At the macro level, Shariff et al.
(\citeproc{ref-shariff_income_2016}{2016}) show that higher perceived
mobility increases tolerance for inequality, suggesting that meritocracy
and mobility are mutually reinforcing.

Taken together, this literature supports the idea that meritocratic
beliefs can moderate the relationship between mobility and market
justice preferences. Individuals who experience mobility---especially
upward---may draw on meritocratic narratives to legitimize both their
own status and broader inequalities, thereby strengthening their support
for market-based distribution. Accordingly, the second hypothesis of
this study is:

\(H2\): The positive (negative) relationship between upward (downward)
social mobility and support for market justice in healthcare, pensions,
and education is moderated by meritocratic beliefs; specifically, this
association is stronger (weaker) among individuals with higher
perceptions of meritocracy.

\subsection{The Chilean context}\label{the-chilean-context}

Chile offers a compelling case for examining individual preferences for
market justice in the access to core social services. Since the early
1990s, the country experienced a long period of sustained economic
growth, with per capita GDP increasing by nearly 4\% annually until the
mid-2010s (\citeproc{ref-llorca-jana_historia_2021}{Llorca-Jaña \&
Miller, 2021}). This growth brought notable improvements in living
standards, including a strong decline in income poverty---from 38.6\% in
1990 to 6.5\% in 2022 (\citeproc{ref-mideplan_evolucion_2017}{MIDEPLAN,
2017}). However, this success story has been tempered by growing
concerns over the unequal distribution of its benefits. In recent years,
economic growth has slowed---exacerbated by the global impact of the
COVID-19 pandemic---and structural inequalities have become more visible
(\citeproc{ref-barozet_clases_2021}{Barozet et al., 2021}). Despite its
economic achievements, Chile remains one of the most unequal countries
in Latin America and the OCDE. The poorest 50\% of the population
captures just 10\% of total income and holds negative net wealth, while
the wealthiest 1\% receives nearly 27\% of all income and controls
almost half (49.6\%) of the country's wealth
(\citeproc{ref-chancel_world_2022}{Chancel et al., 2022}), a figure that
has seen little change since the 1990s
(\citeproc{ref-flores_top_2020}{Flores et al., 2020}). These stark
disparities extend beyond income, reflecting deep inequalities in access
to essential social services (\citeproc{ref-pnud_desiguales_2017}{PNUD,
2017}).

A significant share of Chile's inequality stems from neoliberal reforms
that privatized and commodified key domains of social reproduction
(\citeproc{ref-arrizabalo_milagro_1995}{Arrizabalo, 1995};
\citeproc{ref-ferre_welfare_2023}{Ferre, 2023}). Introduced during the
military dictatorship (1973--1989) and expanded under democratic
governments, these reforms embedded market logic into public services
through concessions, subsidies, demand-side vouchers, and pro-private
regulatory frameworks (\citeproc{ref-boccardo_30_2020}{Boccardo, 2020};
\citeproc{ref-madariaga_three_2020}{Madariaga, 2020}). This
``crowded-out'' welfare model benefits higher-income groups, leaving
lower-income individuals to rely on limited public options. Scholars
argue that this neoliberal shift repurposed the state to create niches
of capitalist accumulation -from water to education-, giving rise to a
model of ``public-service capitalism'' heavily reliant on state funding
(\citeproc{ref-boccardo_30_2020}{Boccardo, 2020}). In healthcare, around
79\% of the population is covered by the public insurer FONASA, while
15\% are enrolled in private ISAPREs, which offer faster and
higher-quality services to wealthier groups
(\citeproc{ref-mideplan_estadisticas_2024}{MIDEPLAN, 2024}). The pension
system is based on individual capitalization, with mandatory
contributions managed by private administrators investing in the
financial market, currently involving 11 million contributors through
private fund administrators (AFPs), yet 27\% of the labor force remains
outside the system due to informality
(\citeproc{ref-superintendenciadepensiones_estadisticas_2024}{Superintendencia
de Pensiones, 2024}). Education is similarly stratified: only 30.6\% of
students attend fully public schools, while 54\% are in voucher-funded
private institutions and 9.3\% in fully private schools that largely
serve affluent families
(\citeproc{ref-ministeriodeeducacion_resumen_2023}{Ministerio de
Educación, 2023}). This arrangement has created a structurally segmented
system, where access and quality of services depend on ability to pay,
reinforced by state subsidies to private providers
(\citeproc{ref-pnud_desiguales_2017}{PNUD, 2017}).

Over recent decades, Chile has undergone a significant transformation in
its class structure, shaped by sustained economic growth, the expansion
of the service sector, the massification of education, and major changes
in the role of the state and social policy
(\citeproc{ref-ruiz_chilenos_2014}{Ruiz \& Boccardo, 2014}). These
changes contributed to the growth of the middle sectors, a phenomena
labelled as ``mesocratrization'' of society
(\citeproc{ref-espinoza_estratificacion_2013}{Espinoza et al., 2013}).
However, this expansion did not reduce structural inequalities nor
facilitate upward mobility into elite positions
(\citeproc{ref-espinoza_estratificacion_2013}{Espinoza et al., 2013};
\citeproc{ref-lopez-roldan_comparative_2021}{López-Roldán \& Fachelli,
2021}). While Chile shows high absolute mobility, most flows occur
between adjacent classes---particularly from working classes into lower
service-class positions---while access to upper classes remains limited,
indicating strong elite closure and reproduction
(\citeproc{ref-espinoza_estratificacion_2013}{Espinoza et al., 2013};
\citeproc{ref-espinoza_movilidad_2014}{Espinoza \& Núñez, 2014};
\citeproc{ref-lopez-roldan_comparative_2021}{López-Roldán \& Fachelli,
2021}; \citeproc{ref-perez-ahumada_clases_2019}{Pérez-Ahumada, 2019};
\citeproc{ref-torche_unequal_2005}{Torche, 2005}). Though often
described as ``unequal but fluid'', recent evidence points to growing
rigidity in relative mobility, especially after the economic slowdown,
increased household debt, labor market precarity, and the COVID-19
crisis (\citeproc{ref-barozet_clases_2021}{Barozet et al., 2021}).
However, the majority of the population identifies themselves as members
of the middle class (\citeproc{ref-castillo_todos_2013}{Castillo et al.,
2013}). These dynamics have fueled a legitimacy crisis: meritocracy is
increasingly questioned, and many perceive that advancement depends more
on informal networks than on effort or earned credentials
(\citeproc{ref-barozet_clases_2021}{Barozet et al., 2021}). The
combination of constrained mobility, persistent inequality, and
concentrated opportunity has eroded expectations of upward mobility and
intensified social discontent (\citeproc{ref-pnud_desiguales_2017}{PNUD,
2017}).

Despite recurrent social unrest, Chile presents a paradoxical
coexistence between strong conflict over inequality and widespread
public legitimation of it. The October 2019 ``social outburst'', a
period of mass protests and severe political repression throughout the
country, crystallized discontent over the privatization and
commodification of social services, alongside a crisis of political
legitimacy (\citeproc{ref-somma_no_2021}{Somma et al., 2021}). A survey
during the protests identified pensions, healthcare, and education as
the top demands
(\citeproc{ref-nucleodesociologiacontingente_informe_2020}{Núcleo de
Sociología Contingente, 2020}), reflecting dissatisfaction with Chile's
stratified welfare regime. Yet, this unrest coexists with strong
meritocratic beliefs and high tolerance for income inequality. Castillo
(\citeproc{ref-castillo_legitimacy_2011}{2011}) shows that wage gaps
between occupations are consistently justified, and the larger the
perceived gap, the greater its legitimation, an effect amplified by
Chile's structural inequality. Moreover, perceptions of unfairness due
to non-meritocratic factors often reinforce belief in effort and talent
as legitimate bases for success
(\citeproc{ref-castillo_multidimensional_2023}{Castillo et al., 2023}),
which are associated with lower perceived injustice
(\citeproc{ref-castillo_meritocracia_2019}{Castillo et al., 2019}).
Mac-Clure et al. (\citeproc{ref-mac-clure_justicia_2024}{Mac-Clure et
al., 2024}) find that low and lower-middle status individuals are less
likely than upper-status groups to view educational income differences
as unfair. Qualitative research further reveals the internalization of
the enterprising self (\citeproc{ref-mau_inequality_2015}{Mau, 2015}),
understood as a way in which the subjectivization process arises in line
with positioning the self in the market, generating its own market value
and subjecting it to competition. Canales et al.
(\citeproc{ref-canalesceron_sujeto_2021}{2021}) describes how families
navigate school choice as investment decisions, while Panes
(\citeproc{ref-panes_criticas_2020}{2020}) finds that many workers frame
pension contributions as individual investments aimed at maximizing
future returns. These representations suggest that market-based moral
orientations are deeply embedded in Chilean subjectivities, offering
fertile ground to study how support for market justice varies across
welfare domains.

Although still limited, recent research on market justice preferences in
Chile indicates that support for these beliefs has increased in recent
years across domains such as healthcare, pensions, and education, and
that both objective and subjective dimensions of inequality shape them.
Castillo et al. (\citeproc{ref-castillo_perceptions_2025}{2025}) show
that agreement with the notion that it is fair for higher-income
individuals to access better services was relatively low in 2016, but
rose significantly by 2023---especially in the pension domain. Otero and
Mendoza (\citeproc{ref-otero_power_2024}{2024}) provide evidence that
social class, measured using the EGP scheme, influences these
preferences: individuals in lower or subordinate class positions are
less supportive of market-based access to welfare services compared to
those in higher or privileged positions. Additionally, greater diversity
in social class networks is associated with lower support for market
justice. This is particularly relevant in the Chilean context, where one
of the main determinants of class-diverse networks is intergenerational
social mobility, which exposes individuals to relationships with people
from different class backgrounds (\citeproc{ref-otero_lives_2022}{Otero
et al., 2022}). Regarding subjective factors, Castillo et al.
(\citeproc{ref-castillo_perceptions_2025}{Castillo, Laffert, et al.,
2025}; \citeproc{ref-castillo_socialization_2024}{Castillo et al.,
2024}) find that individuals who endorse meritocratic beliefs---such as
the idea that effort and talent are rewarded in Chile---are more likely
to support market justice across welfare domains.

\section{Method}\label{method}

Social mobility effects---understood as impacts on an outcome arising
from movements between an origin state and a destination state (e.g.,
social class)---have long been a focus of sociological research
(\citeproc{ref-eyles_social_2022}{Eyles et al., 2022};
\citeproc{ref-langsaether_explaining_2022}{Langsæther et al., 2022}).
Yet, as Breen and Ermisch (\citeproc{ref-breen_effects_2024}{2024, p.
467}) emphasize, most of the mobility hypotheses are, at their core,
individual-level counterfactual comparisons between the observed outcome
under a mobility trajectory and the outcome the same person would have
shown if they had remained in their origin class (or moved to an
alternative destination). Standard specifications (e.g., SAM, DRM, and
mobility-contrast models) struggle to retrieve these inherently
counterfactual quantities due to identification constraints and their
reliance on between-group contrasts constructed from additive terms and
interactions, thereby yielding primarily associational evidence
(\citeproc{ref-breen_effects_2024}{Breen \& Ermisch, 2024};
\citeproc{ref-song_there_2025}{Song \& Zhou, 2025}). Following the
causal framework of Breen and Ermisch
(\citeproc{ref-breen_effects_2024}{2024}), I conceive of the destination
class as a treatment, conditioning on the origin class and estimating
the heterogeneous effects of the destination with observational data
under explicit identification assumptions. To align design and target, I
follow the MIDA template (\citeproc{ref-blair_research_2023}{Blair et
al., 2023}): set out the causal model and assumptions (M), define the
inquiry and estimand (I), describe the data and variables (D), and
detail the answer strategy (A) used to identify the causal effect of
intergenerational occupational mobility on preferences for the
commodification of pensions.

\subsection{A Causal Model for Mobility
Effects}\label{a-causal-model-for-mobility-effects}

Breen and Ermisch (\citeproc{ref-breen_effects_2024}{2024}), building on
a critical reassessment of the inferential limits of standard mobility
models, propose a causal framework that reconceptualizes the ``mobility
effect'' as the treatment effect of reaching a destination class, with
impacts heterogeneous by class of origin. The core claim is that typical
mobility hypotheses are fundamentally within-person counterfactual
comparisons rather than between-person contrasts. Defining mobility as
\(M = D - O\); a change from origin \((O)\) to destination \((D)\),
rather than an additive or interactive combination, places the focus
squarely on the Neyman--Rubin problem: for any individual, we observe
the result in the social position they currently occupy, but not in the
situation of other alternative destinations or immobility.

Exploiting the temporal ordering of origin, destination and outcome,
Breen and Ermisch (\citeproc{ref-breen_effects_2024}{2024}) frame
mobility as a treatment process that motivates the causal question:
``how would the outcome among people from origin \emph{j} who entered
destination \emph{k} have been different if those people had,
counterfactually, entered destination \emph{k'} instead?'' (p.472). The
corresponding estimand is the conditional causal effect of destination
given origin. It is of particular interest when \emph{k' = j}, i.e.,
immobility. This estimand compares movers' observed outcomes with the
hypothetical outcomes those same individuals would have exhibited had
they remained in their origin class. This formulation provides a
coherent potential-outcomes basis for studying how social mobility shape
individual preferences and attitudes.

Following Breen and Ermisch (\citeproc{ref-breen_effects_2024}{2024}),
the identification of causal mobility effects from observational data
requires the following assumptions:

\textbf{Positivity}: For all \((j,c)\) in the support of \((O,C)\) and
for each relevant destination \((k)\), the probability of receiving
\(D=k\) is strictly between 0 and 1; substantively, each type \((O,C)\)
has a nonzero probability of entering each comparison destination.

\begin{equation}\phantomsection\label{eq-posi}{
 0<P(D=k\mid O=j, C=c)<1
}\end{equation}

\textbf{Stable Unit Treatment Value Assumption (SUTVA)}: each unit's
potential outcomes \(Y_i(D)\) does not depend on the mechanism used to
assign treatments (destinations) and by the treatments assigned to other
units (also called no interference), assuming a single, well-defined
version of each treatment (consistency).

\begin{equation}\phantomsection\label{eq-sutva}{
Y_i(D_i) = Y_i\big(D_i, D_{-i}\big)\quad \forall, D_{-i}
}\end{equation}

\textbf{Conditional Independence}: Also known as conditional
unconfoundedness or exchangeability, this assumption emphasizes that,
conditional on origin \(O\) and pre-treatment covariates \(C\) (e.g.,
parental education, ethnicity, cohort, early-life factors), assignment
to \(D\) is as good as random. This underlies IPW and regression
adjustment, which seek exchangeability between mobile (treated) and
immobile individuals (controls).

\begin{equation}\phantomsection\label{eq-indep}{
Y(D)\ \perp\kern-5pt \perp D\ \mid\ (O,C)
}\end{equation}

Establishing the causal relationship between objective social mobility
and subjective outcomes, such as preferences, poses several challenges
for causal inference. Below, I use directed acyclic graphs (DAGs) to
illustrate some of these challenges and evaluate possible strategies for
identifying the effects of social mobility. Figure~\ref{fig-dag} depicts
the causal model guiding identification. In this model, a respondent's
preference \(Y\) is directly affected by her class destination \(D\),
and indirectly influenced by a set of pre-treatment attributes \(C\)
rooted in childhood---family resources, household composition, parental
education, and other ascriptive characteristics. Class origin \(O\) is
itself shaped by these background factors \(C\) and, in turn, affects
destination \(D\). I allow for unobserved determinants \(U\) that
influence background factors \(C\), but I assume that any such
unobserved variation operates only through \(C\). Thus, there are no
direct paths \(U \rightarrow D\) or \(U \rightarrow Y\) beyond those
mediated by \(C\).

Under this structure, conditioning on \((O, C)\) blocks all relevant
backdoor paths from \(D\) to \(Y\)---notably
\(D \leftarrow O \leftarrow C \rightarrow Y\). This renders \((O, C)\) a
minimal sufficient adjustment set for identifying the causal effect of
mobility on preferences, in line with the framework proposed by Breen \&
Ermisch (\citeproc{ref-breen_effects_2024}{2024}). Identification
therefore relies on the assumption that background factors \(C\)
adequately summarize pre-treatment characteristics that jointly shape
both mobility and attitudes.

\begin{figure}[H]

\centering{

\includegraphics[width=0.85\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-dag-1.pdf}

}

\caption{\label{fig-dag}Causal graph of the effect of intergenerational
social mobility on preferences for commodification. Y = preferences for
pension commodification, D = an individual's class of destination, O =
an individual's class of origin, C = different attributes determined in
childhood or earlier that affects an individual's class of origin and
destination. Finally, U = unobserved factors influencing origin
attributes.}

\end{figure}%

Why are the assumptions plausible here? Conditional independence is
targeted by controlling for a plausible sufficient adjustment set
\((O,C)\) and by using inverse probability weights obtained via entropy
balancing (\citeproc{ref-hainmueller_entropy_2012}{Hainmueller, 2012}),
which enforce covariate balance within each origin stratum and mitigate
selection on observables. Positivity is assessed by inspecting the
distribution of the entropy-balancing weights and trimming observations
with extreme weights, thus restricting inference to regions of common
support. SUTVA holds by treating the destination class as a single
exposure measured before \((Y)\) and assuming no interference. Together,
the DAG and assumptions define conditions for identifying the causal
effect of intergenerational social mobility on preferences for
commodification.

\subsection{Inquiry and estimand}\label{inquiry-and-estimand}

The causal inquiry guiding this study asks: How does intergenerational
social mobility affect individuals' preferences for the commodification
of pensions in Chile? Substantively, I am interested in the effect of
mobility relative to immobility within a given class of origin.

Following the potential outcomes framework proposed by Breen and Ermisch
(\citeproc{ref-breen_effects_2024}{2024, p. 473}), the estimand of
interest is an average treatment effect on the treated (ATT) defined
within each origin class. I focus on the specific case in which the
counterfactual destination corresponds to class immobility \((k' = j)\).
In other words, I ask how the outcome among people from origin \emph{j}
who entered destination \emph{k} would have bee different if those
people had, counterfactually, remained in origin \emph{j} instead
(inmobility).

Formally, for individuals with origin \((O=j)\) who attain destination
\((D=k)\), the estimand is:

\$\$

ATT\_\{j,k,j\} = E\big[Y(D = k) \mid O = j, D = k \big] -
E\big[Y(D = j) \mid O = j, D = k \big]

\$\$ \{\#eq-att\}

which represents the mean causal effect of moving from origin class
\((j)\) to destination class \((k)\), compared to the counterfactual
outcome those same individuals would have exhibited had they instead
remained in \((j)\). Thus, this ATT captures the counterfactual contrast
within origin that lies at the heart of mobility effects: how
preferences for market-based pensions would differ if, for the same
upwardly or downwardly mobile individuals, their observed destination
class were replaced by immobility in their origin class.

\subsection{Data and variables}\label{data-and-variables}

\subsubsection{Data}\label{data}

This study draws on data from the Chilean Longitudinal Social Survey
(ELSOC) of the Center for Social Conflict and Cohesion Studies (COES).
This survey is a nationally representative panel study of the urban
adult population in Chile, conducted annually between 2016 and 2023.
Designed to examine individuals' attitudes, emotions, and behaviors
regarding social conflict and cohesion, ELSOC employs a probabilistic,
stratified, clustered, and multistage sampling design covering both
major urban centers and smaller cities. The sampling frame was
proportionally stratified into six categories of urban population size
(e.g., large and small cities), followed by a random selection of
households within 1,067 city blocks. The target population includes men
and women aged 18 to 75 who are habitual residents of private dwellings.

Because respondent occupation\footnote{Parental occupation was collected
  only in 2023 as open-text labels. I coded these texts to the ISCO-08
  two-digit scheme using the National Institute of Statistics of Chile's
  automated coding API and retained cases with high-confidence matches
  (\textgreater95\%). Treating parental occupation as a time-invariant
  origin attribute, I carried these codes back to 2016 and 2018.} is not
measured in every wave, I restrict the analysis to the 2016, 2018, and
2023 waves. After listwise deletion and restricting to key variables
(respondent occupation and the outcome measure), the final analytic
sample comprises \emph{N} = 3,435 observations nested within \emph{N} =
1,787 individuals (2016: 914; 2018: 1,377; 2023: 1,144). Consistent with
the study design, estimation proceeds within trajectory-specific
subsamples (e.g., low→low; low→middle), retaining movers and matched
non-movers from the same origin class. Then, I apply weights to
construct the corresponding counterfactual control sample for each
trajectory. Following Breen and Ermisch
(\citeproc{ref-breen_effects_2024}{2024}), I use the three waves in the
estimations that follow, allowing for correlation between the
observations for each individual in calculating the standard errors
(i.e.~clustering on the personal identifier).

\subsubsection{Outcome variable}\label{outcome-variable}

The outcome variable measures preferences regarding the commodification
of pensions, operationalized with a single item addressing how strongly
individuals justify conditioning access to old-age pension benefits
based on individual income. Respondents were asked: ``Is it fair in
Chile that people with higher incomes have better pensions than people
with lower incomes?'' and answered on a five-point Likert scale ranging
from 1 (``strongly disagree'') to 5 (``strongly agree''). For estimation
and interpretability, I construct a binary indicator in which values
4--5 indicate agreement with market-based access to old-age pensions
(coded 1), while values 1--3 indicate non-agreement (coded 0).
Substantively, the dependent variable is meant to capture
\emph{endorsement} of pension commodification: only responses in the
upper tail of the scale represent an active, explicit justification of
market-based differentiation, whereas neutrality corresponds to an
absence of such endorsement and is therefore treated as closer to
non-support than to support.\footnote{A potential concern is that the
  neutral category may reflect genuine ambivalence rather than
  disagreement. To address this, I conduct robustness checks using
  alternative codings: (i) excluding neutral respondents from the
  analysis; (ii) combining the neutral and agreement categories (3--5
  vs.~1--2); and (iii) estimating models with the original 5-point scale
  using linear specifications. The main mobility effects reported below
  are substantively unchanged across these specifications (see
  \hyperref[anexo]{Supplementary Material}).} This binary coding thus
yields a clear contrast between respondents who \emph{support}
market-based differentiation in pensions and those who do not. I use
this item for two main reasons: first, to enable comparisons with
existing work on market-based justice in social policy
(\citeproc{ref-castillo_perceptions_2025}{Castillo, Laffert, et al.,
2025}; \citeproc{ref-lindh_public_2015}{Lindh, 2015};
\citeproc{ref-otero_power_2024}{Otero \& Mendoza, 2024}); and second,
because it taps two core dimensions of market-based welfare
distribution---(i) the centrality of economic resources as a criterion
for allocating outcomes and (ii) the framing of pensions as tradable
commodities that can be bought and sold according to ability to pay
(\citeproc{ref-lindh_public_2015}{Lindh, 2015}).

\subsubsection{Treatment}\label{treatment}

\paragraph{Intergenerational occupational
mobility}\label{intergenerational-occupational-mobility}

I treat intergenerational occupational mobility as an exposure
indicating whether respondents occupy a different occupational status
than their fathers, closely following Breen and Ermisch's
(\citeproc{ref-breen_effects_2024}{2024}) causal framework. Occupational
assignment proceeds in two steps. First, I derive occupational status
for both origin (father) and destination (respondent) from two-digit
ISCO-08 codes using the International Socio-Economic Index of
Occupational Status (ISEI). Second, I group these ISEI scores into
terciles (low, middle, high), yielding a three-category schema for both
origin and destination strata (see Table~\ref{tbl-matrix}).
Substantively, I use ISEI because occupational status is a core
indicator of stratification and a reliable mobility measure: it locates
jobs on a hierarchical continuum defined by incumbents' typical
education level and earnings
(\citeproc{ref-hauser_intergenerational_2010}{Hauser, 2010};
\citeproc{ref-salgado_uplifting_2025}{Salgado et al., 2025}), and thus
approximates long-run socio-economic position more closely than
volatile, single-year income, which is also prone to recall and
reporting error for both parental and own resources
(\citeproc{ref-barone_rise_2022}{Barone et al., 2022}). This
occupational focus is consistent with a long tradition that interprets
the occupational structure as the backbone of the stratification system
and a key determinant of life chances
(\citeproc{ref-wrightUnderstandingClass2015}{Wright, 2015}). Moreover,
ISEI was explicitly designed to harmonize occupational stratification
across class schemes and SES measures
(\citeproc{ref-ganzeboom_internationally_1996}{Ganzeboom \& Treiman,
1996}), supporting cross-study comparability. In short, using ISEI
aligns the measurement of origin and destination on a common vertical
status continuum that is theoretically meaningful for social mobility
research and empirically feasible given the available information of
ELSOC's data.

\begin{table}

\caption{\label{tbl-matrix}Occupational mobility by occupational
groups.}

\centering{

\centering
\begin{tabular}{cccccc}
\toprule
Father↓ & Offspring→ & Low & Middle & High & Total\\
\midrule
Low &  & 44.7\%   (501) & 33.6\%   (377) & 21.7\%   (244) & 100.0\% (1,122)\\
Middle &  & 35.1\%   (388) & 34.9\%   (385) & 30.0\%   (331) & 100.0\% (1,104)\\
High &  & 24.6\%   (297) & 28.8\%   (348) & 46.7\%   (564) & 100.0\% (1,209)\\
Total &  & 34.5\% (1,186) & 32.3\% (1,110) & 33.2\% (1,139) & 100.0\% (3,435)\\
\bottomrule
\end{tabular}

}

\end{table}%

\paragraph{Pre-treatment control variables for selection into
treatment}\label{pre-treatment-control-variables-for-selection-into-treatment}

Within each origin stratum (\(O = j\)), I estimate average treatment
effects on the treated (ATT) by comparing movers to otherwise similar
non-movers from the same origin. To reduce bias from non-random
selection into mobility based on observed characteristics, I preprocess
the data using entropy balancing, reweighting only the control group
(the immobile) so that the distribution of pre-treatment origin
covariates \(C\) among non-movers matches that of movers on selected
moments. Entropy balancing implements a maximum-entropy reweighting
scheme that chooses unit weights for non-movers to satisfy a set of
balance constraints (e.g., equality of means and, where relevant, higher
moments), while keeping the new weights as close as possible to the
original weights (\citeproc{ref-hainmueller_entropy_2012}{Hainmueller,
2012}). Substantively, this procedure addresses selection on
observables: it reweights immobile respondents so that, within each
origin class, they resemble mobile respondents in their observed
background characteristics, making any remaining differences in the
outcome more plausibly attributable to mobility rather than to
pre-existing measured advantages.

The selection of the adjustment set \(C\) follows two principles. First,
it mirrors the logic in Breen and Ermisch
(\citeproc{ref-breen_effects_2024}{2024}), who condition on attributes
determined in childhood or earlier (e.g., parental resources, household
structure, ascriptive traits). Second, it is constrained to origin-side
characteristics available in ELSOC. Guided by theory and Chilean
evidence on the determinants of relative mobility (mainly
\citeproc{ref-brunori_inequality_2025}{Brunori et al., 2025};
\citeproc{ref-espinoza_movilidad_2014}{Espinoza \& Núñez, 2014};
\citeproc{ref-salgado_uplifting_2025}{Salgado et al., 2025};
\citeproc{ref-torche_unequal_2005}{Torche, 2005}), I include six
covariates covering family resources, household composition, and
ascriptive status: (a) parental education (highest of father/mother), in
10 ordinal categories from no schooling to postgraduate studies; (b)
co-residence with both parents at age 15 (0 = no, 1 = yes); (c)
nationality (0 = non-Chilean, 1 = Chilean); (d) age in years; (e) sex (0
= male, 1 = female); and (f) indigenous ethnicity (0 = no, 1 = yes).
This set of covariates is fixed prior to the destination class and
constitutes the maximum set on the origin side in ELSOC, thus allowing
for the plausible capture of both family and contextual influences that
may affect career trajectories, making the conditional independence
hypothesis more credible within each \(O = j\). Descriptive statistics
for all variables are presented in Supplementary Material Table A1.

After achieving balance, I construct stabilized IPW-ATT weights by
setting treated (mobile) cases to weight 1 and assigning the
entropy-balancing weights to controls, rescaled within each origin
stratum so that the mean weight remains close to one. This aligns the
counterfactual distribution of non-movers with the covariate profile of
movers in each origin class (see \hyperref[anexo]{Supplementary
Material} for balance diagnostics). The resulting weights are then used
in the outcome stage to estimate the causal effect of intergenerational
mobility on preferences for the commodification of pensions.

\subsubsection{Effect heterogeneity}\label{effect-heterogeneity}

To test heterogeneity in mobility effects, I examine two moderators:
meritocratic perceptions and time. Meritocracy is captured with two
items (\citeproc{ref-young_rise_1958}{Young, 1958}), one referring to
effort (``In Chile, people are rewarded for their efforts'') and one to
talent (``In Chile, people are rewarded for their intelligence and
skills''), each answered on a five-point Likert scale (1 = ``strongly
disagree'' to 5 = ``strongly agree''). I average the two items into a
single index and dichotomize it into low meritocracy (≤ 3) and high
meritocracy (≥ 4). Time is measured by survey wave (2016, 2018, 2023)
and entered as a categorical (dummy) variable. Both variables are used
as moderators by interacting them with the mobility treatment to assess
conditional average treatment effects---i.e., whether the effect of
mobility varies across levels of meritocratic perceptions or across
survey waves.

\subsubsection{Controls}\label{controls}

All models include the same pre-treatment covariates \(C\) used in the
IPW construction, not to re-balance groups (entropy balancing already
does so), but to (i) block backdoor paths from \(C\) to the outcome
\(Y\) as implied by the DAG, and (ii) achieve double robustness:
estimates remain consistent if either the weighting model or the outcome
model is correctly specified. Concretely, I adjust for (a) father's
educational level, (b) co-residence with both parents at age 15, (c)
nationality, (d) age, (e) sex, and (f) ethnicity, along with wave fixed
effects to absorb secular trends.

\subsection{Analytical strategy}\label{analytical-strategy}

Because the dependent variable---the preference for pension
commodification---is binary \((Y_{it}\in{0,1})\), I estimate weighted
linear probability models (WLS) using stabilized inverse probability
weights (IPW) obtained via entropy balancing. Although OLS is often
questioned with binary outcomes, it consistently estimates the
conditional mean \(E(Y|X)\) under standard exogeneity
\(E[\varepsilon_i|X_i]=0\), and heteroskedasticity can be handled with
robust or clustered standard errors
(\citeproc{ref-wooldridge_introductory_2009}{Wooldridge, 2009},
ch.~7.5). WLS with IPW further improves efficiency and implements the
ATT estimand by recreating the counterfactual distribution of non-movers
(\citeproc{ref-gelman_regression_2020}{Gelman et al., 2020, pp.
270--272}). Linear models are preferred here because coefficients are
directly interpretable as average differences in predicted
probabilities, whereas non-linear models complicate both weighting and
interpretation (\citeproc{ref-gelman_regression_2020}{Gelman et al.,
2020}, ch.~13).

Formally, the model is:

\begin{equation}\phantomsection\label{eq-final}{
Y_{it}=\alpha+\beta T_i+X_i'\gamma+\lambda_t+\varepsilon_{it}
}\end{equation}

where \(Y_{it}\) indicates whether individual \(i\) in wave \(t\)
supports more market-based pension access; \(\alpha\) is the baseline
probability for the reference categories; \(\beta T_i\) captures the
intergenerational mobility contrast: \(T_i=1\) when the observed
destination is \(D_i=k\) (mobile to \(k\)) and \(T_i=0\) when \(D_i=j\)
(immobile), so that \(\beta\) is the ATT within origin
\(O=j (\widehat\beta=\widehat{ATT}{j,k\mid j})\), i.e., the
percentage-point change in the probability of preferring pension
commodification from reaching \(k\) rather than remaining in \(j\). The
term \(X_i'\gamma\) includes the pre-treatment covariates \((C)\) used
to construct the entropy-balancing weights; \(\lambda_t\) are wave fixed
effects; and \(\varepsilon_{it}\) is idiosyncratic error. The analytic
sample is restricted to individuals sharing the same origin \((O=j)\).
Therefore, estimation uses pooled WLS with stabilized IPW--ATT from
entropy balancing and CR2 standard errors clustered by individual to
address heteroskedasticity and within-person dependence.

The specification is doubly robust: entropy-balancing weights are
combined with covariate adjustment so consistency holds if either the
weighting model or the outcome model is correctly specified, while also
improving efficiency
(\citeproc{ref-wooldridge_introductory_2009}{Wooldridge, 2009}). These
results are virtually identical to baseline models without covariates
(see \hyperref[anexo]{Supplementary Material} for complete models), with
no substantive changes in the mobility coefficients. Heterogeneous
effects by meritocratic perceptions and survey wave are examined through
interaction models described in the previous section.

To evaluate the robustness of the findings, I implement two sets of
sensitivity analyses. First, I conduct standard robustness checks by
re-estimating the models under alternative codings of the outcome: (i)
excluding neutral responses from the analysis, (ii) combining neutral
responses with agreement (3--5 vs.~1--2), and (iii) using the original
5-point item in a linear specification. Second, I assess sensitivity to
unmeasured confounding using the omitted-variable--bias framework of
Cinelli and Hazlett (\citeproc{ref-cinelli_making_2020}{2020}), which
quantifies how strong an unobserved confounder would need to be to
explain away the estimated ATT.

\section{Results}\label{results}

\subsection{Descriptive statistics}\label{descriptive-statistics}

Figure~\ref{fig-alluvial} shows the annual frequencies of preferences
for pension commodification in 2016, 2018, and 2023. Each year presents
stacked percentages frequencies for the level of agreement and
disagreement. Overall, a large majority rejects market-based access to
old-age pension benefits, but this opposition has eased over time:
83.5\% in 2016, 81.2\% in 2018, and 71.8\% in 2023. The 2016--2018
change is modest, whereas 2023 registers a 9.4 point drop in
disagreement relative to 2018, mirrored by rising agreement: 16.5\%
(2016), 18.8\% (2018), 28.2\% (2023). Substantively, while most
respondents continue to oppose the idea that higher-income individuals
should obtain better pensions via the market, a non-trivial and growing
portion endorses this statement, with the sharpest expansion
concentrated in the latest wave (+9.4 points from 2018 to 2023).

\begin{figure}[H]

\centering{

\includegraphics[width=0.85\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-alluvial-1.pdf}

}

\caption{\label{fig-alluvial}Change in preferences for pension
commodification over time (2016, 2018, and 2023).}

\end{figure}%

Regarding the relationship between preferences for the commodification
of pensions and intergenerational occupational mobility,
Figure~\ref{fig-mean} shows the percentage of agreement with
market-based access to pension benefits according to mobility
trajectories and survey waves. Averaging across waves, agreement is
highest among the High-High immobile (27.8\%), followed by the
Middle-High upwardly mobile (24.6\%) and the High-Middle downwardly
mobile (24.0\%). The lowest levels are found among ascending Low-High
(17.3\%), and immobile Low-Low (17.4\%) groups. The specific profiles of
each wave accentuate these gradients in most trajectories, especially in
2023: agreement reaches 31.9\% for High-High, 29.4\% for upward
Middle-High, and peaks at 36.8\% for downward High-Middle (the highest
of all trajectories for that year). Descriptively, immobility at the top
is associated with greater support for commodification; among those who
move, support is greatest for the upward Middle-High trajectory and the
downward High-Middle trajectory, patterns that intensify in 2023. These
shifts anticipate the heterogeneity documented below.

\begin{figure}[H]

\centering{

\includegraphics[width=0.9\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-mean-1.pdf}

}

\caption{\label{fig-mean}Percentage agreement on preferences for pension
commodification according to mobility trajectory and time (2016, 2018,
and 2023).}

\end{figure}%

\subsection{Mobility effects models}\label{mobility-effects-models}

Figure~\ref{fig-reg} reports the double-robust estimates of the effect
of intergenerational occupational mobility on preferences for pension
commodification, obtained from weighted models that combine stabilized
IPW with covariate adjustment and fixed wave effects. Results reveal a
clear directional asymmetry between upward and downward mobility
trajectories. Among upwardly mobile respondents, only the Middle→High
trajectory exhibits a significant positive effect on support for pension
commodification (\(\beta\) = 0.09, 95\% CI {[}0.02, 0.16{]}).
Interpreted as an origin-specific ATT, this estimate compares---for
individuals who actually moved from the middle to the high socioeconomic
status---their observed endorsement with the counterfactual endorsement
they would have expressed had they remained immobile in the middle
socioeconomic status. Holding all other predictors constant, the point
estimate implies a 9 percentage point higher probability of endorsing
market-based access to pension benefits for Middle→High movers relative
to their own immobility counterfactual, statistically significant at the
95\% confidence level. In contrast, two downward trajectories:
High→Middle (\(\beta\) = −0.07, CI {[}−0.15, −0.00{]}) and High→Low
(\(\beta\) = −0.10, CI {[}−0.18, −0.02{]}), show significant negative
effects, suggesting that individuals who experience downward mobility
from higher-status origins express weaker preferences for market-based
pension benefits. The remaining pathways (Low→Middle, Low→High,
Middle→Low) display no significant differences relative to their
non-mobile counterparts. These patterns indicate that upward movement
within the upper segment (Middle→High) reinforces pro-commodification
attitudes, whereas downward movement from privileged origins reduces
them.

\begin{figure}[H]

\centering{

\includegraphics[width=0.9\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-reg-1.pdf}

}

\caption{\label{fig-reg}Effects of intergenerational occupational
mobility on preferences for pension commodification. Coefficient plot of
origin-specific ATT estimates comparing mobility versus immobility.
Estimates from pooled WLS with doubly robust adjustment, wave fixed
effects, and standard errors clustered by individual; bars show 95\%
confidence intervals.}

\end{figure}%

\subsection{Effect heterogeneity}\label{effect-heterogeneity-1}

To probe potential mechanisms underlying the mobility effects, I
estimate origin-specific ATT models that interact the mobility treatment
with high meritocratic perceptions (ref. = low), using stabilized IPW
weights, covariate adjustment, and wave fixed effects. I also estimate
an analogous specification interacting mobility with survey waves (ref.
= 2016) to assess temporal heterogeneity in the treatment effect (see
\hyperref[anexo]{Supplementary Material} for complete models).

\begin{figure}[H]

\centering{

\includegraphics[width=0.9\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-interact-1.pdf}

}

\caption{\label{fig-interact}Effects of intergenerational occupational
mobility on preferences for pension commodification, by meritocratic
perception. Coefficient plot of origin-specific ATT estimates comparing
mobility versus immobility, estimated from interaction models (T ×
Merit). Points show simple slopes of the mobility effect at low merit
and high merit; bars are 95\% confidence intervals. Estimates from
pooled WLS with doubly robust adjustment, wave fixed effects, and
standard errors clustered by individual.}

\end{figure}%

Figure~\ref{fig-interact} shows that neither the interaction terms nor
the mobility coefficients reach conventional significance across
trajectories, indicating no detectable moderation of the causal mobility
effect by meritocratic perceptions at the 95\% level. Directionally, in
the left panel, the interaction is positive for Middle→High (\(\beta\) =
0.15, 95\% CI {[}−0.03, 0.33{]}) and Low→High (\(\beta\) = 0.05,
{[}−0.14, 0.24{]}), suggesting, if anything, somewhat larger
pro-commodification effects among high-merit respondents when upwardly
mobile, whereas it is negative for High→Low (\(\beta\) = −0.14,
{[}−0.38, 0.10{]}) and High→Middle (\(\beta\) = −0.03, {[}−0.23,
0.17{]}), showing weaker pro-commodification effects under downward
mobility at high merit. Focusing on the treatment effect at low
meritocracy (the reference category) at the right panel, the mobility
coefficient captures the origin-specific ATT comparing movers to their
immobility counterfactual among low-merit respondents. These effects are
small and imprecise across trajectories: Low→Middle (\(\beta\) = −0.02,
{[}−0.09, 0.04{]}) and Low→High (\(\beta\) = −0.03, {[}−0.10, 0.05{]})
are near zero. Middle→High is positive but not significant (\(\beta\) =
0.07, {[}−0.01, 0.14{]}), Middle→Low is null (\(\beta\) = -0.00,
{[}−0.07, 0.06{]}), and downward from high shows negative, borderline
estimates: High→Middle (\(\beta\) = −0.07, {[}−0.15, 0.00{]}) and
High→Low (\(\beta\) = −0.08, {[}−0.15, 0.00{]}). Substantively, I do not
find evidence that meritocratic perceptions moderate the causal effect
of mobility, although the signs align with the idea that upward mobility
at high merit tilts attitudes toward commodification, and downward
mobility at high merit tilts them away.

\begin{table}

\caption{\label{tbl-inttime}Effects of intergenerational occupational
mobility on preferences for pension commodification, by survey wave}

\centering{

\begin{center}
\scalebox{0.8}{
\begin{tabular}{l c c c c c c}
\hline
 & Low-Middle & Low-High & Middle-High & Middle-Low & High-Middle & High-Low \\
\hline
Intercept                              & $0.18$           & $-0.04$          & $0.28$           & $0.20$           & $0.18$            & $0.15$            \\
                                       & $ [-0.36; 0.73]$ & $ [-0.36; 0.29]$ & $ [-0.01; 0.57]$ & $ [-0.11; 0.52]$ & $ [-0.15;  0.50]$ & $ [-0.27;  0.58]$ \\
Mobility treatment                     & $-0.05$          & $-0.03$          & $0.13^{*}$       & $0.06$           & $-0.19^{*}$       & $-0.19^{*}$       \\
                                       & $ [-0.16; 0.06]$ & $ [-0.15; 0.10]$ & $ [ 0.05; 0.22]$ & $ [-0.03; 0.15]$ & $ [-0.33; -0.05]$ & $ [-0.32; -0.05]$ \\
Wave (Ref.= 2016)                      &                  &                  &                  &                  &                   &                   \\
                                       &                  &                  &                  &                  &                   &                   \\
\quad Wave 2018                        & $0.01$           & $0.01$           & $0.10^{*}$       & $0.09^{*}$       & $-0.04$           & $-0.02$           \\
                                       & $ [-0.10; 0.12]$ & $ [-0.10; 0.12]$ & $ [ 0.03; 0.18]$ & $ [ 0.01; 0.17]$ & $ [-0.15;  0.07]$ & $ [-0.15;  0.10]$ \\
\quad Wave 2023                        & $0.03$           & $0.01$           & $0.19^{*}$       & $0.18^{*}$       & $0.05$            & $0.04$            \\
                                       & $ [-0.09; 0.14]$ & $ [-0.12; 0.13]$ & $ [ 0.10; 0.28]$ & $ [ 0.08; 0.27]$ & $ [-0.09;  0.18]$ & $ [-0.11;  0.19]$ \\
Mobility treatment x Wave (Ref.= 2016) &                  &                  &                  &                  &                   &                   \\
                                       &                  &                  &                  &                  &                   &                   \\
\quad Mobility treatment x Wave 2018   & $0.03$           & $-0.06$          & $-0.04$          & $-0.09$          & $0.11$            & $0.06$            \\
                                       & $ [-0.11; 0.17]$ & $ [-0.21; 0.09]$ & $ [-0.17; 0.09]$ & $ [-0.20; 0.03]$ & $ [-0.04;  0.26]$ & $ [-0.09;  0.22]$ \\
\quad Mobility treatment x Wave 2023   & $0.07$           & $0.10$           & $-0.09$          & $-0.07$          & $0.19^{*}$        & $0.18$            \\
                                       & $ [-0.09; 0.23]$ & $ [-0.08; 0.28]$ & $ [-0.23; 0.06]$ & $ [-0.22; 0.08]$ & $ [ 0.01;  0.38]$ & $ [-0.01;  0.37]$ \\
\hline
Controls                               & Yes              & Yes              & Yes              & Yes              & Yes               & Yes               \\
R$^2$                                  & $0.04$           & $0.05$           & $0.07$           & $0.03$           & $0.06$            & $0.06$            \\
Adj. R$^2$                             & $0.03$           & $0.04$           & $0.06$           & $0.02$           & $0.05$            & $0.05$            \\
Num. obs.                              & $878$            & $745$            & $716$            & $773$            & $912$             & $861$             \\
RMSE                                   & $0.40$           & $0.37$           & $0.39$           & $0.39$           & $0.44$            & $0.42$            \\
N Clusters                             & $497$            & $449$            & $443$            & $470$            & $536$             & $525$             \\
\hline
\multicolumn{7}{l}{\scriptsize{Note: Cells contain regression coefficients with confidence intervals in parentheses. $^*$ Null hypothesis value outside the confidence interval..}}
\end{tabular}
}
\label{table:coefficients}
\end{center}

}

\end{table}%

Table~\ref{tbl-inttime} shows the causal heterogeneity of the mobility
effect across the different waves of the survey. Regarding temporal
heterogeneity, the only statistically significant interaction term is
observed for the High→Middle trajectory in 2023
(\(\beta_{\text{int}} = 0.19\)), 95\% CI {[}0.01, 0.38{]}). However,
when this interaction is added to the baseline effect in 2016 (-0.19),
the marginal effect of High→Middle mobility in 2023 is approximately
zero, and its confidence interval includes the null. Thus, rather than
indicating a reversal toward a positive effect, the evidence suggests
that the negative impact of downward mobility from high backgrounds is
confined to 2016 and attenuates over time. No comparable temporal
changes are observed in the remaining trajectories. Overall, the results
provide little support for strong time-conditional heterogeneity in
mobility effects, beyond the fading of the initially negative
High→Middle effect.

\subsection{Robustness check and sensitivity
analysis}\label{robustness-check-and-sensitivity-analysis}

As a robustness check, I re-estimate the models under three alternative
codings of the outcome: (i) recoding the neutral category as agreement
(3--5 vs.~1--2), (ii) excluding neutral responses from the analysis, and
(iii) using the original 5-point item in a linear specification. Across
these alternative models (see \hyperref[anexo]{Supplementary Material}
for complete tables), the substantive pattern of the main effects is
preserved: the Middle→High trajectory remains positively associated with
support for pension commodification, and the High→Low trajectory remains
negatively associated with it, with both coefficients generally
increasing in magnitude. By contrast, the negative High→Middle effect
becomes statistically indistinguishable from zero in the robustness
specifications, indicating that this particular estimate is more
sensitive to how the outcome is coded than the other two trajectories.

To assess the extent to which the estimated mobility effects may be
driven by unmeasured confounding, I implement the sensitivity analysis
proposed by Cinelli and Hazlett
(\citeproc{ref-cinelli_making_2020}{2020}), which quantifies how strong
an unobserved confounder would need to be---in terms of its joint
explanatory power for both treatment and outcome---to attenuate or
explain away the ATT, using gender as a benchmark covariate (the
predictor with the largest partial association with the outcome aside
from treatment). For the Middle→High trajectory
(\(\beta \approx 0.09\)), partial \(R^2 \approx 0.013\), an unobserved
confounder would need to explain about 10--11\% of the residual variance
of both treatment and outcome to reduce the effect to zero and roughly
4\% to render it statistically insignificant, substantially more than
gender does. For the High→Low effect (\(\beta \approx -0.10\)), partial
\(R^2 \approx 0.012\), the required strength is of similar magnitude
(around 10--11\% to explain away the effect and about 4\% to remove
significance), again exceeding the explanatory power of the main
observed covariates. By contrast, the High→Middle effect
(\(\beta \approx -0.07\)), partial \(R^2 \approx 0.007\) is somewhat
less robust: a confounder explaining around 8\% of residual variance in
both treatment and outcome could nullify the estimate, and one
explaining about 2\% could make it non-significant. Overall, the
sensitivity analysis suggests that the positive Middle→High and negative
High→Low effects are moderately robust to unmeasured confounding,
whereas the High→Middle effect is more vulnerable to relatively modest
omitted variables (see \hyperref[anexo]{Supplementary Material} for
complete tables).

\section{Discussion}\label{discussion}

\section{Conclusion}\label{conclusion}

\section{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-alesina_intergenerational_2018}
Alesina, A., Stantcheva, S., \& Teso, E. (2018). Intergenerational
{Mobility} and {Preferences} for {Redistribution}. \emph{American
Economic Review}, \emph{108}(2), 521--554.
\url{https://doi.org/10.1257/aer.20162015}

\bibitem[\citeproctext]{ref-araujo_desafios_2012}
Araujo, K., \& Martuccelli, D. (2012). \emph{Desafíos comunes: Retrato
de la sociedad chilena y sus individuos} (1a. ed). Santiago: LOM
Ediciones.

\bibitem[\citeproctext]{ref-ares_changing_2020}
Ares, M. (2020). Changing classes, changing preferences: How social
class mobility affects economic preferences. \emph{West European
Politics}, \emph{43}(6), 1211--1237.
\url{https://doi.org/10.1080/01402382.2019.1644575}

\bibitem[\citeproctext]{ref-arrizabalo_milagro_1995}
Arrizabalo, X. (1995). \emph{{Milagro o quimera: la economía chilena
durante la dictadura}}. Libros de la Catarata.

\bibitem[\citeproctext]{ref-barone_rise_2022}
Barone, C., Hertel, F. R., \& Smallenbroek, O. (2022). The rise of
income and the demise of class and social status? {A} systematic review
of measures of socio-economic position in stratification research.
\emph{Research in Social Stratification and Mobility}, \emph{78},
100678. \url{https://doi.org/10.1016/j.rssm.2022.100678}

\bibitem[\citeproctext]{ref-barozet_clases_2021}
Barozet, E., Contreras, D., Espinoza, V., Gayo, M., \& Méndez, M. L.
(2021). \emph{{Clases medias en tiempos de crisis: vulnerabilidad
persistente, desafíos para la cohesión y un nuevo pacto social en
Chile}}. Santiago: Comisión Económica para América Latina y el Caribe
(CEPAL).

\bibitem[\citeproctext]{ref-batruch_belief_2023}
Batruch, A., Jetten, J., Van De Werfhorst, H., Darnon, C., \& Butera, F.
(2023). Belief in {School Meritocracy} and the {Legitimization} of
{Social} and {Income Inequality}. \emph{Social Psychological and
Personality Science}, \emph{14}(5), 621--635.
\url{https://doi.org/10.1177/19485506221111017}

\bibitem[\citeproctext]{ref-benabou_social_2001}
Benabou, R., \& Ok, E. A. (2001). Social {Mobility} and the {Demand} for
{Redistribution}: {The Poum Hypothesis}. \emph{The Quarterly Journal of
Economics}, \emph{116}(2), 447--487.
\url{https://doi.org/10.1162/00335530151144078}

\bibitem[\citeproctext]{ref-blair_research_2023}
Blair, G., Coppock, A., \& Humphreys, M. (2023). \emph{Research design
in the social sciences: Declaration, diagnosis, and redesign}. Princeton
(N.J.) Oxford: Princeton University press.

\bibitem[\citeproctext]{ref-boccardo_30_2020}
Boccardo, G. (2020). \emph{30 años de privatizaciones en {Chile}: Lo que
la pandemia reveló} (Nodo XXI). Santiago.

\bibitem[\citeproctext]{ref-breen_effects_2024}
Breen, R., \& Ermisch, J. (2024). The {Effects} of {Social Mobility}.
\emph{Sociological Science}, \emph{11}, 467--488.
\url{https://doi.org/10.15195/v11.a17}

\bibitem[\citeproctext]{ref-brunori_inequality_2025}
Brunori, P., Ferreira, F. H. G., \& Neidhöfer, G. (2025). Inequality of
opportunity and intergenerational persistence in {Latin America}.
\emph{Oxford Open Economics}, \emph{4}(Supplement\_1), i167--i199.
\url{https://doi.org/10.1093/ooec/odae021}

\bibitem[\citeproctext]{ref-bucca_merit_2016}
Bucca, M. (2016). Merit and blame in unequal societies: {Explaining
Latin Americans}' beliefs about wealth and poverty. \emph{Research in
Social Stratification and Mobility}, \emph{44}, 98--112.
\url{https://doi.org/10.1016/j.rssm.2016.02.005}

\bibitem[\citeproctext]{ref-busemeyer_skills_2014}
Busemeyer, M. (2014). \emph{Skills and {Inequality}: {Partisan Politics}
and the {Political Economy} of {Education Reforms} in {Western Welfare
States}}. Cambridge University Press.

\bibitem[\citeproctext]{ref-busemeyer_positive_2021}
Busemeyer, M., Abrassart, A., \& Nezi, R. (2021). Beyond {Positive} and
{Negative}: {New Perspectives} on {Feedback Effects} in {Public Opinion}
on the {Welfare State}. \emph{British Journal of Political Science},
\emph{51}(1), 137--162. \url{https://doi.org/10.1017/S0007123418000534}

\bibitem[\citeproctext]{ref-busemeyer_welfare_2020}
Busemeyer, M., \& Iversen, T. (2020). The {Welfare State} with {Private
Alternatives}: {The Transformation} of {Popular Support} for {Social
Insurance}. \emph{The Journal of Politics}, \emph{82}(2), 671--686.
\url{https://doi.org/10.1086/706980}

\bibitem[\citeproctext]{ref-campbell_institutional_2020}
Campbell, J. L. (2020). \emph{Institutional {Change} and
{Globalization}}. Princeton University Press.
\url{https://doi.org/10.2307/j.ctv131bw68}

\bibitem[\citeproctext]{ref-canalesceron_sujeto_2021}
Canales Cerón, M., Orellana Calderón, V. S., \& Guajardo Mañán, F.
(2021). Sujeto y cotidiano en la era neoliberal: El caso de la educación
chilena. \emph{Revista Mexicana de Ciencias Políticas y Sociales},
\emph{67}(244).
\url{https://doi.org/10.22201/fcpys.2448492xe.2022.244.70386}

\bibitem[\citeproctext]{ref-castillo_legitimacy_2011}
Castillo, J. C. (2011). Legitimacy of {Inequality} in a {Highly Unequal
Context}: {Evidence} from the {Chilean Case}. \emph{Social Justice
Research}, \emph{24}(4), 314--340.
\url{https://doi.org/10.1007/s11211-011-0144-5}

\bibitem[\citeproctext]{ref-castillo_changes_2025}
Castillo, J. C., Iturra, J., \& Carrasco, K. (2025). Changes in the
{Justification} of {Educational Inequalities}: {The Role} of
{Perceptions} of {Inequality} and {Meritocracy During} the {COVID
Pandemic}. \emph{Social Justice Research}.
\url{https://doi.org/10.1007/s11211-025-00458-0}

\bibitem[\citeproctext]{ref-castillo_multidimensional_2023}
Castillo, J. C., Iturra, J., Maldonado, L., Atria, J., \& Meneses, F.
(2023). A {Multidimensional Approach} for {Measuring Meritocratic
Beliefs}: {Advantages}, {Limitations} and {Alternatives} to the {ISSP
Social Inequality Survey}. \emph{International Journal of Sociology},
1--25. \url{https://doi.org/10.1080/00207659.2023.2274712}

\bibitem[\citeproctext]{ref-castillo_perceptions_2025}
Castillo, J. C., Laffert, A., Carrasco, K., \& Iturra, J. (2025).
Perceptions of {Inequality} and {Meritocracy}: {Their Interplay} in
{Shaping Preferences} for {Market Justice} in {Chile} (2016-2023).
\emph{Under Review at Frontiers in Sociology}.

\bibitem[\citeproctext]{ref-castillo_todos_2013}
Castillo, J. C., Miranda, D., \& Cabib, I. M. (2013). Todos somos de
clase media: {Sobre} el estatus social subjetivo en {Chile}. \emph{Latin
American Research Review}, \emph{48}(1), 155--173.
\url{https://doi.org/10.1353/lar.2013.0006}

\bibitem[\citeproctext]{ref-castillo_socialization_2024}
Castillo, J. C., Salgado, M., Carrasco, K., \& Laffert, A. (2024). The
{Socialization} of {Meritocracy} and {Market Justice Preferences} at
{School}. \emph{Societies}, \emph{14}(11), 214.
\url{https://doi.org/10.3390/soc14110214}

\bibitem[\citeproctext]{ref-castillo_meritocracia_2019}
Castillo, J. C., Torres, A., Atria, J., \& Maldonado, L. (2019).
Meritocracia y desigualdad económica: {Percepciones}, preferencias e
implicancias. \emph{Revista Internacional de Sociología}, \emph{77}(1),
117. \url{https://doi.org/10.3989/ris.2019.77.1.17.114}

\bibitem[\citeproctext]{ref-chancel_world_2022}
Chancel, L., Piketty, T., Saez, E., \& Zucman, G. (2022). World
inequality report 2022.
https://bibliotecadigital.ccb.org.co/handle/11520/27510.

\bibitem[\citeproctext]{ref-cinelli_making_2020}
Cinelli, C., \& Hazlett, C. (2020). Making {Sense} of {Sensitivity}:
{Extending Omitted Variable Bias}. \emph{Journal of the Royal
Statistical Society Series B: Statistical Methodology}, \emph{82}(1),
39--67. \url{https://doi.org/10.1111/rssb.12348}

\bibitem[\citeproctext]{ref-day_movin_2017}
Day, M. V., \& Fiske, S. T. (2017). Movin' on {Up}? {How Perceptions} of
{Social Mobility Affect Our Willingness} to {Defend} the {System}.
\emph{Social Psychological and Personality Science}, \emph{8}(3),
267--274. \url{https://doi.org/10.1177/1948550616678454}

\bibitem[\citeproctext]{ref-espinoza_estratificacion_2013}
Espinoza, V., Barozet, E., \& Méndez, M. L. (2013). {Estratificación y
movilidad social bajo un modelo neoliberal: El caso de Chile}.
\emph{Lavboratorio}, (25).

\bibitem[\citeproctext]{ref-espinoza_movilidad_2014}
Espinoza, V., \& Núñez, J. (2014). Movilidad ocupacional en {Chile}
2001-2009. \textquestiondown{{Desigualdad}} de ingresos con igualdad de
oportunidades? \emph{Revista Internacional de Sociología}, \emph{72}(1),
57--82. \url{https://doi.org/10.3989/ris.2011.11.08}

\bibitem[\citeproctext]{ref-eyles_social_2022}
Eyles, A., Major, L. E., \& Machin, S. (2022). \emph{Social {Mobility} -
{Past}, {Present} and {Future}: {The State} of {Play} in {Social
Mobility}, on the 25th {Anniversary} of the {Sutton Trust}}. Sutton
Trust.

\bibitem[\citeproctext]{ref-ferre_welfare_2023}
Ferre, J. C. (2023). Welfare regimes in twenty-first-century {Latin
America}. \emph{Journal of International and Comparative Social Policy},
\emph{39}(2), 101--127. \url{https://doi.org/10.1017/ics.2023.16}

\bibitem[\citeproctext]{ref-flores_top_2020}
Flores, I., Sanhueza, C., Atria, J., \& Mayer, R. (2020). Top {Incomes}
in {Chile}: {A Historical Perspective} on {Income Inequality},
1964--2017. \emph{Review of Income and Wealth}, \emph{66}(4), 850--874.
\url{https://doi.org/10.1111/roiw.12441}

\bibitem[\citeproctext]{ref-ganzeboom_internationally_1996}
Ganzeboom, H. B. G., \& Treiman, D. J. (1996). Internationally
{Comparable Measures} of {Occupational Status} for the 1988
{International Standard Classification} of {Occupations}. \emph{Social
Science Research}, \emph{25}(3), 201--239.
\url{https://doi.org/10.1006/ssre.1996.0010}

\bibitem[\citeproctext]{ref-gelman_regression_2020}
Gelman, A., Hill, J., \& Vehtari, A. (2020). \emph{Regression and {Other
Stories}} (1st ed.). Cambridge University Press.
\url{https://doi.org/10.1017/9781139161879}

\bibitem[\citeproctext]{ref-gingrich_making_2011}
Gingrich, J. R. (2011). \emph{Making {Markets} in the {Welfare State}:
{The Politics} of {Varying Market Reforms}} (1st ed.). Cambridge
University Press. \url{https://doi.org/10.1017/CBO9780511791529}

\bibitem[\citeproctext]{ref-gugushvili_trends_2014}
Gugushvili, A. (2014). Trends, {Covariates} and {Consequences} of
{Intergenerational Social Mobility} in {Post- Socialist Societies}.

\bibitem[\citeproctext]{ref-gugushvili_intergenerational_2016c}
Gugushvili, A. (2016a). Intergenerational objective and subjective
mobility and attitudes towards income differences: Evidence from
transition societies. \emph{Journal of International and Comparative
Social Policy}, \emph{32}(3), 199--219.
\url{https://doi.org/10.1080/21699763.2016.1206482}

\bibitem[\citeproctext]{ref-gugushvili_intergenerational_2016}
Gugushvili, A. (2016b). Intergenerational {Social Mobility} and {Popular
Explanations} of {Poverty}: {A Comparative Perspective}. \emph{Social
Justice Research}, \emph{29}(4), 402--428.
\url{https://doi.org/10.1007/s11211-016-0275-9}

\bibitem[\citeproctext]{ref-gugushvili_subjective_2017}
Gugushvili, A. (2017). Subjective {Intergenerational Mobility} and
{Support} for {Welfare State Programmes}.

\bibitem[\citeproctext]{ref-gugushvili_heterogeneous_2024}
Gugushvili, A. (2024). The heterogeneous well-being effects of
intergenerational mobility perceptions. \emph{Journal of Health
Psychology}, \emph{29}(2), 99--112.
\url{https://doi.org/10.1177/13591053231187345}

\bibitem[\citeproctext]{ref-hadjar_does_2015}
Hadjar, A., \& Samuel, R. (2015). Does upward social mobility increase
life satisfaction? {A} longitudinal analysis using {British} and {Swiss}
panel data. \emph{Research in Social Stratification and Mobility},
\emph{39}, 48--58. \url{https://doi.org/10.1016/j.rssm.2014.12.002}

\bibitem[\citeproctext]{ref-hainmueller_entropy_2012}
Hainmueller, J. (2012). Entropy {Balancing} for {Causal Effects}: {A
Multivariate Reweighting Method} to {Produce Balanced Samples} in
{Observational Studies}. \emph{Political Analysis}, \emph{20}(1),
25--46. \url{https://doi.org/10.1093/pan/mpr025}

\bibitem[\citeproctext]{ref-hauser_intergenerational_2010}
Hauser, R. M. (2010). Intergenerational {Economic Mobility} in the
{United States} ---{Measures}, {Differentials}, and {Trends}.

\bibitem[\citeproctext]{ref-helgason_longterm_2023}
Helgason, A. F., \& Rehm, P. (2023). Long-term income trajectories and
the evolution of political attitudes. \emph{European Journal of
Political Research}, \emph{62}(1), 264--284.
\url{https://doi.org/10.1111/1475-6765.12506}

\bibitem[\citeproctext]{ref-helgason_class_2025}
Helgason, A. F., \& Rehm, P. (2025). Class experiences and the long-term
evolution of economic values. \emph{Social Forces}, \emph{103}(3),
1125--1143. \url{https://doi.org/10.1093/sf/soae135}

\bibitem[\citeproctext]{ref-hoyt_mindsets_2023}
Hoyt, C. L., Burnette, J. L., Billingsley, J., Becker, W., \& Babij, A.
D. (2023). Mindsets of poverty: {Implications} for redistributive policy
support. \emph{Analyses of Social Issues and Public Policy},
\emph{23}(3), 668--693. \url{https://doi.org/10.1111/asap.12367}

\bibitem[\citeproctext]{ref-immergut_theoretical_1998}
Immergut, E. M. (1998). The {Theoretical Core} of the {New
Institutionalism}. \emph{Politics \& Society}, \emph{26}(1), 5--34.
\url{https://doi.org/10.1177/0032329298026001002}

\bibitem[\citeproctext]{ref-immergut_it_2020}
Immergut, E. M., \& Schneider, S. M. (2020). Is it unfair for the
affluent to be able to purchase {``better''} healthcare? {Existential}
standards and institutional norms in healthcare attitudes across 28
countries. \emph{Social Science \& Medicine}, \emph{267}, 113146.
\url{https://doi.org/10.1016/j.socscimed.2020.113146}

\bibitem[\citeproctext]{ref-jaime-castillo_social_2019}
Jaime-Castillo, A. M., \& Marqués-Perales, I. (2019). Social mobility
and demand for redistribution in {Europe}: A comparative analysis.
\emph{The British Journal of Sociology}, \emph{70}(1), 138--165.
\url{https://doi.org/10.1111/1468-4446.12363}

\bibitem[\citeproctext]{ref-janmaat_subjective_2013}
Janmaat, J. G. (2013). Subjective {Inequality}: A {Review} of
{International Comparative Studies} on {People}'s {Views} about
{Inequality}. \emph{European Journal of Sociology}, \emph{54}(3),
357--389. \url{https://doi.org/10.1017/S0003975613000209}

\bibitem[\citeproctext]{ref-jasso_justice_1978}
Jasso, G. (1978). On the {Justice} of {Earnings}: {A New Specification}
of the {Justice Evaluation Function}. \emph{American Journal of
Sociology}, \emph{83}(6), 1398--1419.
\url{https://doi.org/10.1086/226706}

\bibitem[\citeproctext]{ref-jasso_distributive_2016}
Jasso, G., Törnblom, K. Y., \& Sabbagh, C. (2016). Distributive
{Justice}. In C. Sabbagh \& M. Schmitt (Eds.), \emph{Handbook of {Social
Justice Theory} and {Research}} (pp. 201--218). New York, NY: Springer.
\url{https://doi.org/10.1007/978-1-4939-3216-0_11}

\bibitem[\citeproctext]{ref-jasso_gender_1999}
Jasso, G., \& Wegener, B. (1999). Gender and {Country Differences} in
the {Sense} of {Justice}: {Justice Evaluation}, {Gender Earnings Gap},
and {Earnings Functions} in {Thirteen Countries}. \emph{International
Journal of Comparative Sociology}, \emph{40}(1), 94--115.
\url{https://doi.org/10.1177/002071529904000106}

\bibitem[\citeproctext]{ref-kluegel_legitimation_1999}
Kluegel, J. R., Mason, D. S., \& Wegener, B. (1999). The {Legitimation}
of {Capitalism} in the {Postcommunist Transition}: {Public Opinion}
about {Market Justice}, 1991-1996. \emph{European Sociological Review},
\emph{15}(3), 251--283. Retrieved from
\url{https://www.jstor.org/stable/522731}

\bibitem[\citeproctext]{ref-kluegel_beliefs_1981}
Kluegel, J. R., \& Smith, E. R. (1981). Beliefs {About Stratification}.
\emph{Annual Review of Sociology}, 29--56.

\bibitem[\citeproctext]{ref-koos_moral_2019}
Koos, S., \& Sachweh, P. (2019). The moral economies of market
societies: Popular attitudes towards market competition, redistribution
and reciprocity in comparative perspective. \emph{Socio-Economic
Review}, \emph{17}(4), 793--821.
\url{https://doi.org/10.1093/ser/mwx045}

\bibitem[\citeproctext]{ref-lane_market_1986}
Lane, R. E. (1986). Market {Justice}, {Political Justice}.
\emph{American Political Science Review}, \emph{80}(2), 383--402.
\url{https://doi.org/10.2307/1958264}

\bibitem[\citeproctext]{ref-langsaether_explaining_2022}
Langsæther, P. E., Evans, G., \& O'Grady, T. (2022). Explaining the
{Relationship Between Class Position} and {Political Preferences}: {A
Long-Term Panel Analysis} of {Intra-Generational Class Mobility}.
\emph{British Journal of Political Science}, \emph{52}(2), 958--967.
\url{https://doi.org/10.1017/S0007123420000599}

\bibitem[\citeproctext]{ref-lee_fairness_2023}
Lee, J.-S., \& Stacey, M. (2023). Fairness perceptions of educational
inequality: The effects of self-interest and neoliberal orientations.
\emph{The Australian Educational Researcher}.
\url{https://doi.org/10.1007/s13384-023-00636-6}

\bibitem[\citeproctext]{ref-lenski_status_1954}
Lenski, G. E. (1954). Status {Crystallization}: {A Non-Vertical
Dimension} of {Social Status}. \emph{American Sociological Review},
\emph{19}(4), 405. \url{https://doi.org/10.2307/2087459}

\bibitem[\citeproctext]{ref-lindh_public_2015}
Lindh, A. (2015). Public {Opinion} against {Markets}? {Attitudes}
towards {Market Distribution} of {Social Services} -- {A Comparison} of
17 {Countries}. \emph{Social Policy \& Administration}, \emph{49}(7),
887--910. \url{https://doi.org/10.1111/spol.12105}

\bibitem[\citeproctext]{ref-lindh_bringing_2023}
Lindh, A., \& McCall, L. (2023). Bringing the market in: An expanded
framework for understanding popular responses to economic inequality.
\emph{Socio-Economic Review}, \emph{21}(2), 1035--1055.
\url{https://doi.org/10.1093/ser/mwac018}

\bibitem[\citeproctext]{ref-llorca-jana_historia_2021}
Llorca-Jaña, M., \& Miller, R. M. D. (2021). \emph{{Historia económica
de Chile desde la independencia}}. Santiago de Chile: RIL editores.

\bibitem[\citeproctext]{ref-lopez-roldan_comparative_2021}
López-Roldán, P., \& Fachelli, S. (Eds.). (2021). \emph{Towards a
{Comparative Analysis} of {Social Inequalities} between {Europe} and
{Latin America}}. Cham: Springer International Publishing.
\url{https://doi.org/10.1007/978-3-030-48442-2}

\bibitem[\citeproctext]{ref-mac-clure_justicia_2024}
Mac-Clure, O., Barozet, E., \& Franetovic, G. (2024). {Justicia
distributiva y posición social subjetiva: \textquestiondown la
meritocracia justifica la desigualdad de ingresos?} \emph{Convergencia
Revista de Ciencias Sociales}, \emph{31}, 1.
\url{https://doi.org/10.29101/crcs.v31i0.22258}

\bibitem[\citeproctext]{ref-madariaga_three_2020}
Madariaga, A. (2020). The three pillars of neoliberalism: {Chile}'s
economic policy trajectory in comparative perspective.
\emph{Contemporary Politics}, \emph{26}(3), 308--329.
\url{https://doi.org/10.1080/13569775.2020.1735021}

\bibitem[\citeproctext]{ref-mau_inequality_2015}
Mau, S. (2015). \emph{Inequality, {Marketization} and the {Majority
Class}: {Why Did} the {European Middle Classes Accept Neo-Liberalism}?}
Houndmills: Palgrave Macmillan.

\bibitem[\citeproctext]{ref-mideplan_evolucion_2017}
MIDEPLAN. (2017). \emph{Evolución de la pobreza 1990-2017}. Ministerio
de Desarrollo Social y Familia - PNUD.

\bibitem[\citeproctext]{ref-mideplan_estadisticas_2024}
MIDEPLAN. (2024). \emph{Estadísticas {Encuesta Casen} 2022, sección
{Salud}}. Observatorio social.

\bibitem[\citeproctext]{ref-mijs_paradox_2019}
Mijs, J. (2019). The paradox of inequality: Income inequality and belief
in meritocracy go hand in hand. \emph{Socio-Economic Review},
\emph{19}(1), 7--35. \url{https://doi.org/10.1093/ser/mwy051}

\bibitem[\citeproctext]{ref-mijs_belief_2022}
Mijs, J., Daenekindt, S., de Koster, W., \& van der Waal, J. (2022).
Belief in {Meritocracy Reexamined}: {Scrutinizing} the {Role} of
{Subjective Social Mobility}. \emph{Social Psychology Quarterly},
\emph{85}(2), 131--141. \url{https://doi.org/10.1177/01902725211063818}

\bibitem[\citeproctext]{ref-miller_selfserving_1975}
Miller, D. T., \& Ross, M. (1975). Self-serving biases in the
attribution of causality: {Fact} or fiction? \emph{Psychological
Bulletin}, \emph{82}(2), 213--225.
\url{https://doi.org/10.1037/h0076486}

\bibitem[\citeproctext]{ref-ministeriodeeducacion_resumen_2023}
Ministerio de Educación. (2023). \emph{Resumen de matrícula por
establecimiento educacional}. Datos abiertos Ministerio de Educación.

\bibitem[\citeproctext]{ref-molina_its_2019}
Molina, M. D., Bucca, M., \& Macy, M. W. (2019). It's not just how the
game is played, it's whether you win or lose. \emph{SCIENCE ADVANCES}.

\bibitem[\citeproctext]{ref-nucleodesociologiacontingente_informe_2020}
Núcleo de Sociología Contingente, {[}NUDESOC{]}. (2020). \emph{Informe
de resultados oficial {Encuesta Zona Cero}}. Santiago de Chile.

\bibitem[\citeproctext]{ref-otero_power_2024}
Otero, G., \& Mendoza, M. (2024). The {Power} of {Diversity}: {Class},
{Networks} and {Attitudes Towards Inequality}. \emph{Sociology},
\emph{58}(4), 851--876. \url{https://doi.org/10.1177/00380385231217625}

\bibitem[\citeproctext]{ref-otero_lives_2022}
Otero, G., Volker, B., Rözer, J., \& Mollenhorst, G. (2022). The lives
of others: {Class} divisions, network segregation, and attachment to
society in {Chile}. \emph{The British Journal of Sociology},
\emph{73}(4), 754--785. \url{https://doi.org/10.1111/1468-4446.12966}

\bibitem[\citeproctext]{ref-panes_criticas_2020}
Panes, D. (2020). \emph{Críticas y experiencias obreras en torno al
sistema de {AFP}'s en {Chile} (1981-2020)} (PhD thesis). Universidad de
Chile, Santiago.

\bibitem[\citeproctext]{ref-perez-ahumada_clases_2019}
Pérez-Ahumada, P. (2019). {Clases sociales, sectores económicos y
cambios en la estructura social chilena entre 1992 y 2013}.
\emph{Revista de la CEPAL}, \emph{2018}(126), 171--192.
\url{https://doi.org/10.18356/0ce0faed-es}

\bibitem[\citeproctext]{ref-pnud_desiguales_2017}
PNUD (Ed.). (2017). \emph{{Desiguales: orígenes, cambios y desafíos de
la brecha social en Chile}}. Santiago, Chile: PNUD : Uqbar Editores.

\bibitem[\citeproctext]{ref-polanyi_great_1975}
Polanyi, K. (1975). \emph{The great transformation} (Repr). New York,
NY: Octagon Books.

\bibitem[\citeproctext]{ref-prag_subjective_2021}
Präg, P., \& Gugushvili, A. (2021). Subjective social mobility and
health in {Germany}. \emph{European Societies}, \emph{23}(4), 464--486.
\url{https://doi.org/10.1080/14616696.2021.1887916}

\bibitem[\citeproctext]{ref-ruiz_chilenos_2014}
Ruiz, C., \& Boccardo, G. (2014). \emph{Los chilenos bajo el
neoliberalismo: Clases y conflicto social} (Primera edición). Santiago
de Chile: Ediciones y Publicaciones El Buen Aire.

\bibitem[\citeproctext]{ref-salgado_uplifting_2025}
Salgado, M., Díaz, M., Gamarra, C., \& Núñez, J. (2025). Uplifting
mechanism or equalizing force? {The} expansion of higher education and
its role in intergenerational mobility in {Chile}. \emph{Higher
Education}. \url{https://doi.org/10.1007/s10734-025-01560-7}

\bibitem[\citeproctext]{ref-sandel_tyranny_2020}
Sandel, M. J. (2020). \emph{The tyranny of merit: {What}'s become of the
common good?} (First edition). New York: {Farrar, Straus and Giroux}.

\bibitem[\citeproctext]{ref-satz_por_2019}
Satz, D. (2019). \emph{{Por Qué Algunas Cosas No Deberían Estar en
Venta: Los límites Morales Del Mercado}}. Ciudad Autónoma de Buenos
Aires: Siglo XXI Editores.

\bibitem[\citeproctext]{ref-schmidt_experience_2011}
Schmidt, A. W. (2011). The experience of social mobility and the
formation of attitudes toward redistribution. In.

\bibitem[\citeproctext]{ref-shariff_income_2016}
Shariff, A. F., Wiwad, D., \& Aknin, L. B. (2016). Income {Mobility
Breeds Tolerance} for {Income Inequality}: {Cross-National} and
{Experimental Evidence}. \emph{Perspectives on Psychological Science},
\emph{11}(3), 373--380. \url{https://doi.org/10.1177/1745691616635596}

\bibitem[\citeproctext]{ref-somma_no_2021}
Somma, N. M., Bargsted, M., Disi Pavlic, R., \& Medel, R. M. (2021). No
water in the oasis: The {Chilean Spring} of 2019--2020. \emph{Social
Movement Studies}, \emph{20}(4), 495--502.
\url{https://doi.org/10.1080/14742837.2020.1727737}

\bibitem[\citeproctext]{ref-song_there_2025}
Song, X., \& Zhou, X. (2025). Is {There} a {Mobility Effect}? {On
Methodological Issues} in the {Mobility Contrast Model}.
\emph{Sociological Methods \& Research}, \emph{54}(4), 1576--1593.
\url{https://doi.org/10.1177/00491241251347983}

\bibitem[\citeproctext]{ref-sorokin_social_1927}
Sorokin, P. A. (1927). \emph{Social {Mobility}}. Harper \& Brothers.

\bibitem[\citeproctext]{ref-streeck_how_2016}
Streeck, W. (2016). \emph{How will capitalism end? Essays on a failing
system}. London: Verso.

\bibitem[\citeproctext]{ref-superintendenciadepensiones_estadisticas_2024}
Superintendencia de Pensiones. (2024). \emph{Estadísticas de {Afiliados}
en {AFP} en {Sistema} de {Pensiones}. {Abril} 2024}. Superintendencia de
Pensiones.

\bibitem[\citeproctext]{ref-svallforsMoralEconomyClass2006a}
Svallfors, S. (2006). \emph{The {Moral Economy} of {Class}: {Class} and
{Attitudes} in {Comparative Perspective}}. Stanford University Press.

\bibitem[\citeproctext]{ref-svallfors_political_2007}
Svallfors, S. (Ed.). (2007). \emph{The {Political Sociology} of the
{Welfare State}: {Institutions}, {Social Cleavages}, and {Orientations}}
(1st ed.). Stanford University Press.
\url{https://doi.org/10.2307/j.ctvr0qv0q}

\bibitem[\citeproctext]{ref-tejero-peregrina_perceived_2025}
Tejero-Peregrina, L., Willis, G., Sánchez-Rodríguez, Á., \&
Rodríguez-Bailón, R. (2025). From {Perceived Economic Inequality} to
{Support} for {Redistribution}: {The Role} of {Meritocracy Perception}.
\emph{International Review of Social Psychology}, \emph{38}(1), 4.
\url{https://doi.org/10.5334/irsp.1013}

\bibitem[\citeproctext]{ref-torche_unequal_2005}
Torche, F. (2005). Unequal {But Fluid}: {Social Mobility} in {Chile} in
{Comparative Perspective}. \emph{American Sociological Review},
\emph{70}(3), 422--450. \url{https://doi.org/10.1177/000312240507000304}

\bibitem[\citeproctext]{ref-torche_intergenerational_2014}
Torche, F. (2014). Intergenerational {Mobility} and {Inequality}: {The
Latin American Case}. \emph{Annual Review of Sociology}, \emph{40}(1),
619--642. \url{https://doi.org/10.1146/annurev-soc-071811-145521}

\bibitem[\citeproctext]{ref-vondemknesebeck_are_2016}
Von Dem Knesebeck, O., Vonneilich, N., \& Kim, T. J. (2016). Are health
care inequalities unfair? {A} study on public attitudes in 23 countries.
\emph{International Journal for Equity in Health}, \emph{15}(1), 61.
\url{https://doi.org/10.1186/s12939-016-0350-8}

\bibitem[\citeproctext]{ref-wen_does_2021}
Wen, F., \& Witteveen, D. (2021). Does perceived social mobility shape
attitudes toward government and family educational investment?
\emph{Social Science Research}, \emph{98}, 102579.
\url{https://doi.org/10.1016/j.ssresearch.2021.102579}

\bibitem[\citeproctext]{ref-wiederkehr_belief_2015}
Wiederkehr, V., Bonnot, V., Krauth-Gruber, S., \& Darnon, C. (2015).
Belief in school meritocracy as a system-justifying tool for low status
students. \emph{Frontiers in Psychology}, \emph{6}.

\bibitem[\citeproctext]{ref-wilson_role_2003}
Wilson, C. (2003). The {Role} of a {Merit Principle} in {Distributive
Justice}. \emph{The Journal of Ethics}, \emph{7}(3), 277--314.
\url{https://doi.org/10.1023/A:1024667228488}

\bibitem[\citeproctext]{ref-wooldridge_introductory_2009}
Wooldridge, J. M. (2009). \emph{{Introductory econometrics: a modern
approach}} (4th ed). Mason, OH: South Western, Cengage Learning.

\bibitem[\citeproctext]{ref-wrightUnderstandingClass2015}
Wright, E. O. (2015). \emph{Understanding {Class}} (Verso).

\bibitem[\citeproctext]{ref-young_rise_1958}
Young, M. (1958). \emph{The rise of the meritocracy}. New Brunswick,
N.J., U.S.A: Transaction Publishers.

\end{CSLReferences}

\section{Supplementary material}\label{anexo}

This section presents the supplementary material for this study.

\subsection{Descriptive statistics}\label{descriptive-statistics-1}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.4216}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2647}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2059}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.1078}}@{}}
\caption{Descriptive statistics for all
variables.}\label{tbl-summary1}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Label
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Stats / Values
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Freqs (\% of Valid)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Valid
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Label
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Stats / Values
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Freqs (\% of Valid)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Valid
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Preference for pension commodification &
\begin{minipage}[t]{\linewidth}\raggedright
1. Disagree\\
2. Agree\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
2702 (78.7\%)\\
733 (21.3\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
3435\\
(100.0\%)\strut
\end{minipage} \\
Father stratum & \begin{minipage}[t]{\linewidth}\raggedright
1. Low\\
2. Middle\\
3. High\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
1122 (32.7\%)\\
1104 (32.1\%)\\
1209 (35.2\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
3435\\
(100.0\%)\strut
\end{minipage} \\
Offspring stratum & \begin{minipage}[t]{\linewidth}\raggedright
1. Low\\
2. Middle\\
3. High\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
1186 (34.5\%)\\
1110 (32.3\%)\\
1139 (33.2\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
3435\\
(100.0\%)\strut
\end{minipage} \\
Parental education & \begin{minipage}[t]{\linewidth}\raggedright
Mean (sd) : 4 (2.2)\\
min \textless{} med \textless{} max:\\
1 \textless{} 4 \textless{} 10\\
IQR (CV) : 3 (0.6)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
1 : 222 ( 6.5\%)\\
2 : 905 (26.3\%)\\
3 : 576 (16.8\%)\\
4 : 321 ( 9.3\%)\\
5 : 826 (24.0\%)\\
6 : 30 ( 0.9\%)\\
7 : 214 ( 6.2\%)\\
8 : 75 ( 2.2\%)\\
9 : 245 ( 7.1\%)\\
10 : 21 ( 0.6\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
3435\\
(100.0\%)\strut
\end{minipage} \\
Co-residence with both parents at age 15 &
\begin{minipage}[t]{\linewidth}\raggedright
1. No co-residence\\
2. Co-residence\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
996 (29.0\%)\\
2439 (71.0\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
3435\\
(100.0\%)\strut
\end{minipage} \\
Nacionality & \begin{minipage}[t]{\linewidth}\raggedright
1. Non-Chilean\\
2. Chilean\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
62 ( 1.8\%)\\
3373 (98.2\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
3435\\
(100.0\%)\strut
\end{minipage} \\
Sex & \begin{minipage}[t]{\linewidth}\raggedright
1. Male\\
2. Female\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
1507 (43.9\%)\\
1928 (56.1\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
3435\\
(100.0\%)\strut
\end{minipage} \\
Age (in years) & \begin{minipage}[t]{\linewidth}\raggedright
Mean (sd) : 42.6 (12.5)\\
min \textless{} med \textless{} max:\\
18 \textless{} 43 \textless{} 75\\
IQR (CV) : 21 (0.3)\strut
\end{minipage} & 58 distinct values &
\begin{minipage}[t]{\linewidth}\raggedright
3435\\
(100.0\%)\strut
\end{minipage} \\
Indigenous ethnicity & \begin{minipage}[t]{\linewidth}\raggedright
1. Non-indigenous\\
2. Indigenous\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
3035 (88.4\%)\\
400 (11.6\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
3435\\
(100.0\%)\strut
\end{minipage} \\
Meritocracy perception & \begin{minipage}[t]{\linewidth}\raggedright
1. Low\\
2. High\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
2741 (79.8\%)\\
694 (20.2\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
3435\\
(100.0\%)\strut
\end{minipage} \\
Wave & \begin{minipage}[t]{\linewidth}\raggedright
1. 2016\\
2. 2018\\
3. 2023\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
914 (26.6\%)\\
1377 (40.1\%)\\
1144 (33.3\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
3435\\
(100.0\%)\strut
\end{minipage} \\
\end{longtable}

\subsection{Balance evaluation}\label{balance-evaluation}

\begin{figure}[H]

\centering{

\includegraphics[width=0.9\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-balance-1.pdf}

}

\caption{\label{fig-balance}Balance SMD --- Mobility treatment (ATT)}

\end{figure}%

\subsection{Mobility effects models}\label{mobility-effects-models-1}

\begin{table}

\caption{\label{tbl-complete1}Effects of intergenerational occupational
mobility on preferences for pension commodification, with covariates and
wave fixed effects.}

\centering{

\begin{center}
\scalebox{0.7}{
\begin{tabular}{l c c c c c c}
\hline
 & Low-Middle & Low-High & Middle-High & Middle-Low & High-Middle & High-Low \\
\hline
Intercept                                               & $0.18$            & $-0.02$           & $0.30^{*}$        & $0.23$            & $0.13$            & $0.14$            \\
                                                        & $ [-0.36;  0.71]$ & $ [-0.35;  0.31]$ & $ [ 0.01;  0.58]$ & $ [-0.08;  0.54]$ & $ [-0.19;  0.46]$ & $ [-0.28;  0.56]$ \\
Mobility treatment                                      & $-0.02$           & $-0.02$           & $0.09^{*}$        & $0.00$            & $-0.07^{*}$       & $-0.10^{*}$       \\
                                                        & $ [-0.08;  0.05]$ & $ [-0.09;  0.05]$ & $ [ 0.02;  0.16]$ & $ [-0.06;  0.06]$ & $ [-0.15; -0.00]$ & $ [-0.18; -0.02]$ \\
Age (in years)                                          & $0.00$            & $0.00$            & $-0.00$           & $-0.00$           & $0.00$            & $0.00$            \\
                                                        & $ [-0.00;  0.00]$ & $ [-0.00;  0.00]$ & $ [-0.00;  0.00]$ & $ [-0.00;  0.00]$ & $ [-0.00;  0.01]$ & $ [-0.00;  0.01]$ \\
Female (Ref. = Male)                                    & $-0.14^{*}$       & $-0.14^{*}$       & $-0.13^{*}$       & $-0.07^{*}$       & $-0.11^{*}$       & $-0.12^{*}$       \\
                                                        & $ [-0.20; -0.08]$ & $ [-0.22; -0.06]$ & $ [-0.20; -0.06]$ & $ [-0.14; -0.01]$ & $ [-0.19; -0.03]$ & $ [-0.21; -0.03]$ \\
Chilean nacionality (Ref. = Non-Chilean)                & $-0.03$           & $0.20$            & $0.08$            & $0.04$            & $-0.00$           & $0.06$            \\
                                                        & $ [-0.63;  0.57]$ & $ [-0.05;  0.46]$ & $ [-0.19;  0.36]$ & $ [-0.26;  0.34]$ & $ [-0.27;  0.27]$ & $ [-0.30;  0.42]$ \\
Indigenous ethnicity (Ref. = Non-indigenous)            & $0.03$            & $-0.05$           & $0.05$            & $-0.01$           & $0.07$            & $0.01$            \\
                                                        & $ [-0.06;  0.12]$ & $ [-0.15;  0.05]$ & $ [-0.08;  0.18]$ & $ [-0.10;  0.08]$ & $ [-0.08;  0.22]$ & $ [-0.13;  0.14]$ \\
Co-residence with both parents (Ref. = No co-residence) & $-0.00$           & $0.04$            & $-0.06$           & $-0.01$           & $0.10^{*}$        & $0.10^{*}$        \\
                                                        & $ [-0.08;  0.07]$ & $ [-0.04;  0.12]$ & $ [-0.14;  0.01]$ & $ [-0.09;  0.06]$ & $ [ 0.01;  0.19]$ & $ [ 0.01;  0.19]$ \\
Parental education                                      & $0.02$            & $0.01$            & $-0.02^{*}$       & $-0.01$           & $-0.00$           & $-0.01$           \\
                                                        & $ [-0.00;  0.05]$ & $ [-0.01;  0.04]$ & $ [-0.03; -0.01]$ & $ [-0.02;  0.01]$ & $ [-0.02;  0.02]$ & $ [-0.03;  0.02]$ \\
Wave (Ref.= 2016)                                       &                   &                   &                   &                   &                   &                   \\
                                                        &                   &                   &                   &                   &                   &                   \\
\quad Wave 2018                                         & $0.02$            & $-0.01$           & $0.08^{*}$        & $0.04$            & $-0.00$           & $-0.00$           \\
                                                        & $ [-0.05;  0.09]$ & $ [-0.09;  0.07]$ & $ [ 0.02;  0.14]$ & $ [-0.02;  0.10]$ & $ [-0.08;  0.08]$ & $ [-0.09;  0.09]$ \\
\quad Wave 2023                                         & $0.06$            & $0.04$            & $0.15^{*}$        & $0.14^{*}$        & $0.12^{*}$        & $0.10$            \\
                                                        & $ [-0.02;  0.13]$ & $ [-0.06;  0.13]$ & $ [ 0.08;  0.22]$ & $ [ 0.07;  0.21]$ & $ [ 0.02;  0.21]$ & $ [-0.00;  0.21]$ \\
\hline
R$^2$                                                   & $0.04$            & $0.04$            & $0.07$            & $0.03$            & $0.05$            & $0.06$            \\
Adj. R$^2$                                              & $0.03$            & $0.03$            & $0.06$            & $0.02$            & $0.04$            & $0.05$            \\
Num. obs.                                               & $878$             & $745$             & $716$             & $773$             & $912$             & $861$             \\
RMSE                                                    & $0.40$            & $0.37$            & $0.39$            & $0.39$            & $0.44$            & $0.42$            \\
N Clusters                                              & $497$             & $449$             & $443$             & $470$             & $536$             & $525$             \\
\hline
\multicolumn{7}{l}{\scriptsize{Note: Cells contain regression coefficients with confidence intervals in parentheses. $^*$ Null hypothesis value outside the confidence interval..}}
\end{tabular}
}
\label{table:coefficients}
\end{center}

}

\end{table}%

\begin{table}

\caption{\label{tbl-complete2}Effects of intergenerational occupational
mobility on preferences for pension commodification.}

\centering{

\begin{center}
\scalebox{0.8}{
\begin{tabular}{l c c c c c c}
\hline
 & Low-Middle & Low-High & Middle-High & Middle-Low & High-Middle & High-Low \\
\hline
Intercept          & $0.21^{*}$       & $0.18^{*}$       & $0.17^{*}$      & $0.19^{*}$       & $0.32^{*}$       & $0.28^{*}$        \\
                   & $ [ 0.16; 0.26]$ & $ [ 0.13; 0.23]$ & $ [0.12; 0.21]$ & $ [ 0.15; 0.24]$ & $ [ 0.26; 0.38]$ & $ [ 0.21;  0.35]$ \\
Mobility treatment & $-0.01$          & $-0.02$          & $0.08^{*}$      & $-0.00$          & $-0.07$          & $-0.09^{*}$       \\
                   & $ [-0.08; 0.05]$ & $ [-0.08; 0.05]$ & $ [0.01; 0.15]$ & $ [-0.06; 0.06]$ & $ [-0.15; 0.00]$ & $ [-0.17; -0.02]$ \\
\hline
R$^2$              & $0.00$           & $0.00$           & $0.01$          & $0.00$           & $0.01$           & $0.01$            \\
Adj. R$^2$         & $-0.00$          & $-0.00$          & $0.01$          & $-0.00$          & $0.01$           & $0.01$            \\
Num. obs.          & $878$            & $745$            & $716$           & $773$            & $912$            & $861$             \\
RMSE               & $0.40$           & $0.38$           & $0.40$          & $0.39$           & $0.45$           & $0.43$            \\
N Clusters         & $497$            & $449$            & $443$           & $470$            & $536$            & $525$             \\
\hline
\multicolumn{7}{l}{\scriptsize{Note: Cells contain regression coefficients with confidence intervals in parentheses. $^*$ Null hypothesis value outside the confidence interval..}}
\end{tabular}
}
\label{table:coefficients}
\end{center}

}

\end{table}%

\begin{table}

\caption{\label{tbl-complete3}Interactions effects between
intergenerational occupational mobility and perceived meritocracy on
preferences for pension commodification, with covariates and wave fixed
effects.}

\centering{

\begin{center}
\scalebox{0.7}{
\begin{tabular}{l c c c c c c}
\hline
 & Low-Middle & Low-High & Middle-High & Middle-Low & High-Middle & High-Low \\
\hline
Intercept                                                    & $0.15$           & $-0.04$          & $0.30^{*}$       & $0.21$           & $0.08$           & $0.05$           \\
                                                             & $ [-0.36; 0.67]$ & $ [-0.40; 0.31]$ & $ [ 0.04; 0.56]$ & $ [-0.09; 0.51]$ & $ [-0.25; 0.40]$ & $ [-0.30; 0.41]$ \\
Mobility treatment                                           & $-0.02$          & $-0.03$          & $0.07$           & $-0.00$          & $-0.07$          & $-0.08$          \\
                                                             & $ [-0.09; 0.04]$ & $ [-0.10; 0.05]$ & $ [-0.01; 0.14]$ & $ [-0.07; 0.06]$ & $ [-0.15; 0.00]$ & $ [-0.15; 0.00]$ \\
High meritocracy perception (Ref.= Low)                      & $0.05$           & $0.06$           & $-0.01$          & $0.03$           & $0.17^{*}$       & $0.26^{*}$       \\
                                                             & $ [-0.06; 0.16]$ & $ [-0.07; 0.19]$ & $ [-0.09; 0.08]$ & $ [-0.07; 0.13]$ & $ [ 0.00; 0.33]$ & $ [ 0.04; 0.48]$ \\
Mobility treatment x High meritocracy perception (Ref.= Low) & $0.04$           & $0.05$           & $0.15$           & $0.04$           & $-0.03$          & $-0.14$          \\
                                                             & $ [-0.12; 0.20]$ & $ [-0.14; 0.24]$ & $ [-0.03; 0.33]$ & $ [-0.11; 0.19]$ & $ [-0.23; 0.17]$ & $ [-0.38; 0.10]$ \\
\hline
Controls                                                     & Yes              & Yes              & Yes              & Yes              & Yes              & Yes              \\
R$^2$                                                        & $0.04$           & $0.05$           & $0.08$           & $0.04$           & $0.07$           & $0.10$           \\
Adj. R$^2$                                                   & $0.03$           & $0.04$           & $0.07$           & $0.02$           & $0.06$           & $0.08$           \\
Num. obs.                                                    & $878$            & $745$            & $716$            & $773$            & $912$            & $861$            \\
RMSE                                                         & $0.40$           & $0.37$           & $0.39$           & $0.39$           & $0.44$           & $0.41$           \\
N Clusters                                                   & $497$            & $449$            & $443$            & $470$            & $536$            & $525$            \\
\hline
\multicolumn{7}{l}{\scriptsize{Note: Cells contain regression coefficients with confidence intervals in parentheses. $^*$ Null hypothesis value outside the confidence interval..}}
\end{tabular}
}
\label{table:coefficients}
\end{center}

}

\end{table}%

\newpage{}

\subsection{Robustness check and sensitivity
analysis}\label{robustness-check-and-sensitivity-analysis-1}

\begin{table}

\caption{\label{tbl-robus1}Effects of intergenerational occupational
mobility on preferences for pension commodification coded as 3-5
vs.~1-2, with covariates and wave fixed effects.}

\centering{

\begin{center}
\scalebox{0.7}{
\begin{tabular}{l c c c c c c}
\hline
 & Low-Middle & Low-High & Middle-High & Middle-Low & High-Middle & High-Low \\
\hline
Intercept                                               & $0.16$            & $0.01$            & $0.59^{*}$        & $0.57^{*}$        & $0.48^{*}$        & $0.29$            \\
                                                        & $ [-0.38;  0.70]$ & $ [-0.37;  0.40]$ & $ [ 0.17;  1.01]$ & $ [ 0.23;  0.92]$ & $ [ 0.04;  0.92]$ & $ [-0.32;  0.90]$ \\
Mobility treatment                                      & $0.01$            & $0.05$            & $0.09^{*}$        & $-0.02$           & $-0.08$           & $-0.13^{*}$       \\
                                                        & $ [-0.07;  0.08]$ & $ [-0.02;  0.13]$ & $ [ 0.01;  0.16]$ & $ [-0.09;  0.05]$ & $ [-0.15;  0.00]$ & $ [-0.21; -0.04]$ \\
Age (in years)                                          & $0.00$            & $-0.00$           & $-0.00^{*}$       & $-0.00$           & $0.00$            & $0.00$            \\
                                                        & $ [-0.00;  0.00]$ & $ [-0.01;  0.00]$ & $ [-0.01; -0.00]$ & $ [-0.00;  0.00]$ & $ [-0.00;  0.00]$ & $ [-0.00;  0.01]$ \\
Female (Ref. = Male)                                    & $-0.15^{*}$       & $-0.16^{*}$       & $-0.17^{*}$       & $-0.09^{*}$       & $-0.11^{*}$       & $-0.10^{*}$       \\
                                                        & $ [-0.22; -0.08]$ & $ [-0.24; -0.07]$ & $ [-0.25; -0.09]$ & $ [-0.16; -0.02]$ & $ [-0.19; -0.03]$ & $ [-0.20; -0.00]$ \\
Chilean nacionality (Ref. = Non-Chilean)                & $0.03$            & $0.28$            & $-0.10$           & $-0.24$           & $-0.21$           & $-0.02$           \\
                                                        & $ [-0.57;  0.63]$ & $ [-0.01;  0.57]$ & $ [-0.54;  0.33]$ & $ [-0.56;  0.08]$ & $ [-0.71;  0.29]$ & $ [-0.79;  0.76]$ \\
Indigenous ethnicity (Ref. = Non-indigenous)            & $-0.03$           & $-0.10$           & $0.02$            & $-0.02$           & $0.02$            & $-0.07$           \\
                                                        & $ [-0.12;  0.07]$ & $ [-0.22;  0.01]$ & $ [-0.11;  0.15]$ & $ [-0.12;  0.08]$ & $ [-0.13;  0.16]$ & $ [-0.21;  0.07]$ \\
Co-residence with both parents (Ref. = No co-residence) & $0.01$            & $0.06$            & $0.01$            & $0.00$            & $0.15^{*}$        & $0.12^{*}$        \\
                                                        & $ [-0.08;  0.09]$ & $ [-0.03;  0.14]$ & $ [-0.07;  0.10]$ & $ [-0.08;  0.08]$ & $ [ 0.06;  0.24]$ & $ [ 0.02;  0.22]$ \\
Parental education                                      & $0.02$            & $0.02$            & $-0.02$           & $0.00$            & $0.00$            & $0.00$            \\
                                                        & $ [-0.01;  0.05]$ & $ [-0.01;  0.05]$ & $ [-0.03;  0.00]$ & $ [-0.02;  0.02]$ & $ [-0.01;  0.02]$ & $ [-0.02;  0.02]$ \\
Wave (Ref.= 2016)                                       &                   &                   &                   &                   &                   &                   \\
                                                        &                   &                   &                   &                   &                   &                   \\
\quad Wave 2018                                         & $0.05$            & $0.00$            & $0.07$            & $0.03$            & $0.02$            & $-0.00$           \\
                                                        & $ [-0.02;  0.12]$ & $ [-0.08;  0.08]$ & $ [-0.01;  0.15]$ & $ [-0.04;  0.11]$ & $ [-0.06;  0.10]$ & $ [-0.09;  0.09]$ \\
\quad Wave 2023                                         & $0.13^{*}$        & $0.11^{*}$        & $0.23^{*}$        & $0.20^{*}$        & $0.15^{*}$        & $0.15^{*}$        \\
                                                        & $ [ 0.04;  0.22]$ & $ [ 0.00;  0.21]$ & $ [ 0.14;  0.32]$ & $ [ 0.12;  0.28]$ & $ [ 0.05;  0.25]$ & $ [ 0.05;  0.26]$ \\
\hline
R$^2$                                                   & $0.05$            & $0.07$            & $0.09$            & $0.05$            & $0.06$            & $0.07$            \\
Adj. R$^2$                                              & $0.04$            & $0.05$            & $0.08$            & $0.04$            & $0.05$            & $0.06$            \\
Num. obs.                                               & $878$             & $745$             & $716$             & $773$             & $912$             & $861$             \\
RMSE                                                    & $0.44$            & $0.43$            & $0.45$            & $0.44$            & $0.48$            & $0.45$            \\
N Clusters                                              & $497$             & $449$             & $443$             & $470$             & $536$             & $525$             \\
\hline
\multicolumn{7}{l}{\scriptsize{Note: Cells contain regression coefficients with confidence intervals in parentheses. $^*$ Null hypothesis value outside the confidence interval..}}
\end{tabular}
}
\label{table:coefficients}
\end{center}

}

\end{table}%

\begin{table}

\caption{\label{tbl-robus2}Effects of intergenerational occupational
mobility on preferences for pension commodification excluding
intermediate category, with covariates and wave fixed effects.}

\centering{

\begin{center}
\scalebox{0.7}{
\begin{tabular}{l c c c c c c}
\hline
 & Low-Middle & Low-High & Middle-High & Middle-Low & High-Middle & High-Low \\
\hline
Intercept                                               & $1.84^{*}$        & $1.76^{*}$        & $2.39^{*}$        & $2.40^{*}$        & $2.14^{*}$        & $2.35^{*}$        \\
                                                        & $ [ 1.08;  2.60]$ & $ [ 1.02;  2.49]$ & $ [ 1.80;  2.97]$ & $ [ 1.88;  2.92]$ & $ [ 1.45;  2.82]$ & $ [ 1.53;  3.18]$ \\
Mobility treatment                                      & $-0.03$           & $-0.03$           & $0.24^{*}$        & $-0.01$           & $-0.12$           & $-0.22^{*}$       \\
                                                        & $ [-0.16;  0.09]$ & $ [-0.16;  0.11]$ & $ [ 0.10;  0.38]$ & $ [-0.13;  0.11]$ & $ [-0.26;  0.01]$ & $ [-0.37; -0.07]$ \\
Age (in years)                                          & $0.00$            & $-0.00$           & $-0.00$           & $-0.00$           & $0.00$            & $-0.00$           \\
                                                        & $ [-0.01;  0.01]$ & $ [-0.01;  0.01]$ & $ [-0.01;  0.00]$ & $ [-0.01;  0.00]$ & $ [-0.00;  0.01]$ & $ [-0.01;  0.01]$ \\
Female (Ref. = Male)                                    & $-0.24^{*}$       & $-0.23^{*}$       & $-0.27^{*}$       & $-0.16^{*}$       & $-0.16^{*}$       & $-0.22^{*}$       \\
                                                        & $ [-0.35; -0.12]$ & $ [-0.39; -0.07]$ & $ [-0.42; -0.12]$ & $ [-0.29; -0.03]$ & $ [-0.30; -0.02]$ & $ [-0.39; -0.06]$ \\
Chilean nacionality (Ref. = Non-Chilean)                & $0.07$            & $0.28$            & $-0.16$           & $-0.34$           & $-0.20$           & $-0.07$           \\
                                                        & $ [-0.73;  0.87]$ & $ [-0.33;  0.89]$ & $ [-0.69;  0.37]$ & $ [-0.76;  0.07]$ & $ [-0.82;  0.43]$ & $ [-0.75;  0.61]$ \\
Indigenous ethnicity (Ref. = Non-indigenous)            & $0.03$            & $-0.08$           & $-0.02$           & $0.04$            & $0.05$            & $-0.21$           \\
                                                        & $ [-0.14;  0.20]$ & $ [-0.28;  0.12]$ & $ [-0.28;  0.25]$ & $ [-0.15;  0.23]$ & $ [-0.20;  0.31]$ & $ [-0.49;  0.07]$ \\
Co-residence with both parents (Ref. = No co-residence) & $0.00$            & $0.05$            & $-0.09$           & $-0.01$           & $0.15$            & $0.15$            \\
                                                        & $ [-0.14;  0.15]$ & $ [-0.10;  0.21]$ & $ [-0.23;  0.06]$ & $ [-0.16;  0.13]$ & $ [-0.01;  0.32]$ & $ [-0.03;  0.33]$ \\
Parental education                                      & $0.01$            & $0.02$            & $-0.04^{*}$       & $-0.01$           & $0.00$            & $-0.01$           \\
                                                        & $ [-0.04;  0.07]$ & $ [-0.04;  0.07]$ & $ [-0.08; -0.01]$ & $ [-0.05;  0.02]$ & $ [-0.03;  0.04]$ & $ [-0.05;  0.03]$ \\
Wave (Ref.= 2016)                                       &                   &                   &                   &                   &                   &                   \\
                                                        &                   &                   &                   &                   &                   &                   \\
\quad Wave 2018                                         & $0.02$            & $-0.13$           & $0.10$            & $-0.01$           & $-0.08$           & $-0.18$           \\
                                                        & $ [-0.11;  0.16]$ & $ [-0.28;  0.03]$ & $ [-0.05;  0.26]$ & $ [-0.14;  0.13]$ & $ [-0.24;  0.07]$ & $ [-0.37;  0.01]$ \\
\quad Wave 2023                                         & $0.21^{*}$        & $0.03$            & $0.39^{*}$        & $0.37^{*}$        & $0.30^{*}$        & $0.16$            \\
                                                        & $ [ 0.06;  0.36]$ & $ [-0.15;  0.21]$ & $ [ 0.23;  0.54]$ & $ [ 0.23;  0.51]$ & $ [ 0.13;  0.46]$ & $ [-0.04;  0.36]$ \\
\hline
R$^2$                                                   & $0.04$            & $0.04$            & $0.10$            & $0.06$            & $0.06$            & $0.08$            \\
Adj. R$^2$                                              & $0.03$            & $0.02$            & $0.09$            & $0.05$            & $0.05$            & $0.07$            \\
Num. obs.                                               & $812$             & $677$             & $629$             & $693$             & $808$             & $770$             \\
RMSE                                                    & $0.77$            & $0.73$            & $0.75$            & $0.75$            & $0.82$            & $0.79$            \\
N Clusters                                              & $480$             & $431$             & $415$             & $445$             & $499$             & $494$             \\
\hline
\multicolumn{7}{l}{\scriptsize{Note: Cells contain regression coefficients with confidence intervals in parentheses. $^*$ Null hypothesis value outside the confidence interval..}}
\end{tabular}
}
\label{table:coefficients}
\end{center}

}

\end{table}%

\begin{table}

\caption{\label{tbl-robus3}Effects of intergenerational occupational
mobility on preferences for pension commodification using the original
5-point likert scale, with covariates and wave fixed effects.}

\centering{

\begin{center}
\scalebox{0.7}{
\begin{tabular}{l c c c c c c}
\hline
 & Low-Middle & Low-High & Middle-High & Middle-Low & High-Middle & High-Low \\
\hline
Intercept                                               & $1.99^{*}$        & $1.76^{*}$        & $2.91^{*}$        & $2.86^{*}$        & $2.53^{*}$        & $2.59^{*}$        \\
                                                        & $ [ 0.81;  3.17]$ & $ [ 0.82;  2.70]$ & $ [ 2.19;  3.64]$ & $ [ 2.24;  3.47]$ & $ [ 1.59;  3.47]$ & $ [ 1.34;  3.83]$ \\
Mobility treatment                                      & $-0.03$           & $0.03$            & $0.30^{*}$        & $-0.02$           & $-0.19$           & $-0.33^{*}$       \\
                                                        & $ [-0.21;  0.15]$ & $ [-0.16;  0.22]$ & $ [ 0.11;  0.48]$ & $ [-0.18;  0.14]$ & $ [-0.38;  0.00]$ & $ [-0.54; -0.12]$ \\
Age (in years)                                          & $0.00$            & $-0.00$           & $-0.00$           & $-0.00$           & $0.00$            & $0.00$            \\
                                                        & $ [-0.01;  0.01]$ & $ [-0.01;  0.01]$ & $ [-0.01;  0.00]$ & $ [-0.01;  0.00]$ & $ [-0.01;  0.01]$ & $ [-0.01;  0.01]$ \\
Female (Ref. = Male)                                    & $-0.37^{*}$       & $-0.37^{*}$       & $-0.40^{*}$       & $-0.23^{*}$       & $-0.26^{*}$       & $-0.32^{*}$       \\
                                                        & $ [-0.54; -0.20]$ & $ [-0.58; -0.15]$ & $ [-0.60; -0.21]$ & $ [-0.41; -0.06]$ & $ [-0.46; -0.06]$ & $ [-0.55; -0.08]$ \\
Chilean nacionality (Ref. = Non-Chilean)                & $0.10$            & $0.56$            & $-0.25$           & $-0.49^{*}$       & $-0.32$           & $-0.05$           \\
                                                        & $ [-1.17;  1.38]$ & $ [-0.08;  1.20]$ & $ [-0.89;  0.39]$ & $ [-0.93; -0.05]$ & $ [-1.23;  0.59]$ & $ [-1.24;  1.15]$ \\
Indigenous ethnicity (Ref. = Non-indigenous)            & $-0.00$           & $-0.18$           & $-0.00$           & $0.02$            & $0.08$            & $-0.26$           \\
                                                        & $ [-0.24;  0.24]$ & $ [-0.46;  0.10]$ & $ [-0.35;  0.35]$ & $ [-0.24;  0.28]$ & $ [-0.29;  0.44]$ & $ [-0.66;  0.13]$ \\
Co-residence with both parents (Ref. = No co-residence) & $0.01$            & $0.11$            & $-0.06$           & $-0.01$           & $0.28^{*}$        & $0.26^{*}$        \\
                                                        & $ [-0.20;  0.22]$ & $ [-0.11;  0.33]$ & $ [-0.27;  0.15]$ & $ [-0.21;  0.19]$ & $ [ 0.05;  0.50]$ & $ [ 0.01;  0.50]$ \\
Parental education                                      & $0.04$            & $0.03$            & $-0.06^{*}$       & $-0.01$           & $0.01$            & $-0.01$           \\
                                                        & $ [-0.04;  0.11]$ & $ [-0.04;  0.11]$ & $ [-0.10; -0.01]$ & $ [-0.05;  0.04]$ & $ [-0.04;  0.05]$ & $ [-0.07;  0.05]$ \\
Wave (Ref.= 2016)                                       &                   &                   &                   &                   &                   &                   \\
                                                        &                   &                   &                   &                   &                   &                   \\
\quad Wave 2018                                         & $0.08$            & $-0.12$           & $0.16$            & $0.03$            & $-0.05$           & $-0.17$           \\
                                                        & $ [-0.11;  0.26]$ & $ [-0.32;  0.09]$ & $ [-0.04;  0.36]$ & $ [-0.15;  0.21]$ & $ [-0.26;  0.16]$ & $ [-0.42;  0.09]$ \\
\quad Wave 2023                                         & $0.33^{*}$        & $0.14$            & $0.55^{*}$        & $0.53^{*}$        & $0.42^{*}$        & $0.30^{*}$        \\
                                                        & $ [ 0.11;  0.54]$ & $ [-0.11;  0.39]$ & $ [ 0.35;  0.76]$ & $ [ 0.35;  0.72]$ & $ [ 0.18;  0.65]$ & $ [ 0.02;  0.57]$ \\
\hline
R$^2$                                                   & $0.04$            & $0.05$            & $0.10$            & $0.06$            & $0.06$            & $0.08$            \\
Adj. R$^2$                                              & $0.03$            & $0.04$            & $0.09$            & $0.05$            & $0.05$            & $0.07$            \\
Num. obs.                                               & $878$             & $745$             & $716$             & $773$             & $912$             & $861$             \\
RMSE                                                    & $1.10$            & $1.04$            & $1.06$            & $1.07$            & $1.17$            & $1.12$            \\
N Clusters                                              & $497$             & $449$             & $443$             & $470$             & $536$             & $525$             \\
\hline
\multicolumn{7}{l}{\scriptsize{Note: Cells contain regression coefficients with confidence intervals in parentheses. $^*$ Null hypothesis value outside the confidence interval..}}
\end{tabular}
}
\label{table:coefficients}
\end{center}

}

\end{table}%

\begin{table}

\caption{\label{tbl-evalues1}E-values for ATT estimates for Middle-High
trajectory (Cinelli \& Hazlett, 2020 approximation)}

\centering{

\centering
\begin{tabular}{lrrrrrr}
\multicolumn{7}{c}{Outcome: \textit{y}} \\
\hline \hline 
Treatment: & Est. & S.E. & t-value & $R^2_{Y \sim D |{\bf X}}$ & $RV_{q = 1}$ & $RV_{q = 1, \alpha = 0.05}$  \\ 
\hline 
\textit{t} & 0.088 & 0.029 & 2.995 & 1.3\% & 10.7\% & 3.8\% \\ 
\hline 
df = 706 & & \multicolumn{5}{r}{ \small \textit{Bound (1x sexo)}: $R^2_{Y\sim Z| {\bf X}, D}$ = 2.7\%, $R^2_{D\sim Z| {\bf X} }$ = 3.6\%} \\
\end{tabular}

}

\end{table}%

\begin{table}

\caption{\label{tbl-evalues2}E-values for ATT estimates for High-Middle
trajectory (Cinelli \& Hazlett, 2020 approximation)}

\centering{

\centering
\begin{tabular}{lrrrrrr}
\multicolumn{7}{c}{Outcome: \textit{y}} \\
\hline \hline 
Treatment: & Est. & S.E. & t-value & $R^2_{Y \sim D |{\bf X}}$ & $RV_{q = 1}$ & $RV_{q = 1, \alpha = 0.05}$  \\ 
\hline 
\textit{t} & -0.074 & 0.03 & -2.447 & 0.7\% & 7.8\% & 1.6\% \\ 
\hline 
df = 902 & & \multicolumn{5}{r}{ \small \textit{Bound (1x sexo)}: $R^2_{Y\sim Z| {\bf X}, D}$ = 1.5\%, $R^2_{D\sim Z| {\bf X} }$ = 3.9\%} \\
\end{tabular}

}

\end{table}%

\begin{table}

\caption{\label{tbl-evalues3}E-values for ATT estimates for High-Low
trajectory (Cinelli \& Hazlett, 2020 approximation)}

\centering{

\centering
\begin{tabular}{lrrrrrr}
\multicolumn{7}{c}{Outcome: \textit{y}} \\
\hline \hline 
Treatment: & Est. & S.E. & t-value & $R^2_{Y \sim D |{\bf X}}$ & $RV_{q = 1}$ & $RV_{q = 1, \alpha = 0.05}$  \\ 
\hline 
\textit{t} & -0.098 & 0.03 & -3.245 & 1.2\% & 10.5\% & 4.3\% \\ 
\hline 
df = 851 & & \multicolumn{5}{r}{ \small \textit{Bound (1x sexo)}: $R^2_{Y\sim Z| {\bf X}, D}$ = 1.7\%, $R^2_{D\sim Z| {\bf X} }$ = 0.8\%} \\
\end{tabular}

}

\end{table}%

\end{document}