### 2025-11-21 paper/01-introduction.qmd
up antes de borrar tiempo

# Introduction

What is the legitimate scope of market inequality in the eyes of the public? Since the early 1980s, many countries have experienced a rollback of universal welfare programs and a shift toward the privatisation and commodification of public goods and social services [@gingrich_making_2011; @streeck_how_2016]. In Latin America, these reforms extended market logic into domains of social reproduction long governed by the state, shrinking public provision and expanding private actors in core welfare sectors [@ferre_welfare_2023]. From a moral-economy perspective, this diffusion of market rules has reconfigured the normative order: through policy feedback, welfare institutions increasingly embed market-based criteria of “fair” allocation and shape how citizens understand deservingness and the balance between state and market [@svallforsMoralEconomyClass2006a; @koos_moral_2019]. Against this backdrop, growing scholarship examines *market justice preferences*—the extent to which citizens regard market-based criteria as a fair basis for access to essential services such as healthcare, education and pensions [@castillo_perceptions_2025; @svallfors_political_2007; @lindh_public_2015]. These orientations matter because they help legitimise economic inequality by portraying stratified outcomes as a fair result of individual responsibility and limited state intervention.

This article examines public support for the commodification of pension welfare—specifically, whether people consider it fair that access to better pensions depends on income or contributions. Pension privatisation lies at the core of Latin America’s trajectory of commodification [@huber_political_2000], making pension systems a strategic domain for analysing how market principles become normatively embedded [@mau_inequality_2015]. As many countries expanded mandatory or voluntary individual-capitalisation schemes—most radically in Chile—pensions became organised around contribution-based entitlements and market performance [@arenas_sistemas_2019; @verbic_political_2019]. This shift reshapes not only how pension institutions allocate old-age risk, heightening exposure to poverty in retirement and tying benefit levels to lifetime labour-market trajectories [@oecd_pensions_2023; @madero-cabib_private_2019], but also reshape the justice principles citizens associate with old-age security: people have to judge whether higher benefits should follow market-derived criteria or solidaristic principles [@arrizabalomontoro_debate_2019; @borzutzky_pension_2012]. Although we know something about what people consider a “just” monetary pension [@castillo_deserving_2019] and about support for public versus private provision [@busemeyer_welfare_2020; @jaime-castillo_public_2013], we still lack systematic evidence on how citizens assess the fairness of market-dependent access to pensions.

Research shows that support for market-based access to welfare services is strongly stratified: individuals with higher income, education or class position are more likely to endorse market-based criteria than disadvantaged groups [@busemeyer_skills_2014; @immergut_it_2020;  @lindh_public_2015;  @svallfors_political_2007]. Evidence for pensions points in the same direction: higher income and greater financial gains in funded schemes correlate with stronger support for market-oriented pension policies [@kerner_pension_2020]. Yet we know very little about how movement within the class structure shapes preferences for pension commodification. This gap is especially salient in Latin America, where mobility has unfolded under high inequality and deeply privatised welfare regimes [@lopez-roldan_comparative_2021]. Social origins and destinations anchor material interests and justice principles [@alesina_intergenerational_2018; @langsaether_explaining_2022; @gugushvili_subjective_2017], while movement between these positions exposes individuals to new risks, resources and normative environments that reshape fairness judgements [@gugushvili_trends_2014; @helgason_longterm_2023]. As @mau_inequality_2015 argues, social mobility can foster market-justice orientations, helping explain how inequalities in access to welfare services become normatively sustained.

The literature identifies several mechanisms through which mobility may shape distributive-justice preferences—material self-interest [@benabou_social_2001], socialisation and acculturation [@jaime-castillo_social_2019], learning through long-term exposure [@helgason_longterm_2023], and cognitive attributions such as self-serving bias or privilege awareness [@gugushvili_intergenerational_2016c; @molina_its_2019]. While distinct, these accounts converge in highlighting two broad channels relevant for market-based welfare: shifts in material stakes and shifts in normative frames [@mau_inequality_2015]. In this paper, I do not treat these mechanisms as mediators but use them to guide hypotheses about heterogeneity: if mobility alters preferences through material interests and normative updating, its effects should be most potent among those who interpret social outcomes through merit-based evaluations of effort [@castillo_perceptions_2025; @mijs_belief_2022].

Against this background, this article pursues two aims: first, to assess whether intergenerational mobility causally affects support for the commodification of pensions; and second, to examine whether this effect varies by meritocratic beliefs. I argue that mobility works through two main channels. Upward mobility increases earnings, expected returns in a funded system and perceived economic security, making contribution-based access more appealing. At the same time, integration into higher-class networks exposes people to more individualistic norms and effort-based understandings of fairness. Downward mobility, in turn, heightens economic insecurity, dependence on collective protection and attachment to solidaristic principles, lowering support for market allocation. Meritocratic beliefs are conceptualised not as mediators but as moderators. Among those who see mobility as a fair reward for effort, both material and normative effects should be amplified, producing higher support for pension commodification. The paper contributes to research on market justice by focusing on a specific welfare domain and by providing causal evidence on how mobility shapes distributive preferences.

This study focuses on Chile, a particularly instructive case for examining preferences for pension commodification. Despite sustained economic growth, Chile combines high and persistent inequality with short-range upward mobility and substantial barriers to higher-class positions [@flores_top_2020; @salgado_uplifting_2025; @torche_intergenerational_2014]. It was the first country worldwide to fully replace a public pay-as-you-go system with a mandatory, privately administered (through pension fund administrators, AFPs) defined-contribution scheme, while the state has progressively expanded a segmented solidarity pillar to compensate for insufficient benefits [@madariaga_three_2020; @boccardo_30_2020; @solimano_rise_2021]. In parallel—and despite waves of protest from 2006 to 2019 targeting the AFP system [@somma_no_2021]—Chilean subjectivities have been increasingly shaped by neoliberal discourses and market logics, influencing attitudes toward pension distribution and inequality [@araujo_desafios_2012; @galvez_con_2023]. The combination of profound inequality, market-driven old-age security and contested legitimacy makes Chile a critical case for analysing how mobility and meritocratic beliefs structure support for pension commodification.

To address these questions, the study draws on large-scale, representative survey data for the urban Chilean population collected in 2016, 2018 and 2023. It adopts a causal framework of mobility effects inspired by @breen_effects_2024. The following section outlines the theoretical framework linking preferences for market-based welfare, social mobility and expected heterogeneity by meritocratic beliefs and time, and derives hypotheses. A subsequent section presents the causal identification strategy, the data, the variables, and the analytical approach. The empirical sections report the findings and interpret their implications. The article concludes by discussing what these results reveal about the politics of market-based pension welfare and social mobility.