### 2025-11-30 paper/01-introduction.qmd
Up antecedentes

# Introduction

What is the legitimate scope of market inequality in the eyes of the public? Since the early 1980s, many countries have experienced a rollback of universal welfare programs and a shift toward the privatisation and commodification of public goods and social services [@gingrich_making_2011; @streeck_how_2016]. In Latin America, these reforms extended market logic into domains of social reproduction long governed by the state, shrinking public provision and expanding private actors in core welfare sectors [@ferre_welfare_2023]. From a moral-economy perspective, this diffusion of market rules has reconfigured the normative order: through policy feedback, welfare institutions embed market-based criteria of “fair” allocation and shape how citizens understand deservingness and the balance between state and market [@svallforsMoralEconomyClass2006a; @koos_moral_2019]. Against this backdrop, growing scholarship examines *market justice preferences*—the extent to which citizens regard market-based criteria as a fair basis for the allocation of essential services such as healthcare, education and pensions [@castillo_perceptions_2025; @svallfors_political_2007; @lindh_public_2015]. These orientations matter because they legitimize unequal outcomes as the product of individual responsibility and promote the conception of basic social services as commodities.

This article examines public support for the commodification of pension welfare—specifically, whether people consider it fair that access to better pensions depends on income or contributions. Pension privatisation lies at the core of Latin America’s trajectory of commodification [@huber_political_2000]. As many countries expanded mandatory or voluntary individual-capitalisation schemes—most radically in Chile—pensions became organised around contribution-based entitlements and market performance [@arenas_sistemas_2019; @verbic_political_2019]. This shift reshapes not only how pension institutions allocate old-age risk and tie benefit levels to lifetime labour-market trajectories [@oecd_pensions_2023; @madero-cabib_private_2019], but also transforms the justice principles citizens associate with old-age security: they must judge whether higher benefits should follow market-derived criteria or solidaristic principles [@mau_inequality_2015; @borzutzky_pension_2012]. Although we know something about what people consider a “just” monetary pension [@castillo_deserving_2019] and about support for public versus private provision [@busemeyer_welfare_2020; @jaime-castillo_public_2013], we still lack systematic evidence on how citizens assess the fairness of market-dependent access to pensions.

Research shows that support for market-based access to welfare services is strongly stratified: people with higher socioeconomic status are more likely to support market-based criteria than disadvantaged groups [@busemeyer_skills_2014; @immergut_it_2020; @lindh_public_2015; @svallfors_political_2007]. Evidence for pensions points in the same direction: higher income and greater financial gains in funded schemes correlate with stronger support for market-oriented pension policies [@busemeyer_welfare_2020; @kerner_pension_2020]. Yet we know very little about how movement within the class structure shapes preferences for pension commodification. This gap is especially salient in Latin America, where mobility has unfolded under high inequality and deeply privatised welfare regimes [@lopez-roldan_comparative_2021]. Social origins and destinations anchor material interests and justice principles [@alesina_intergenerational_2018; @langsaether_explaining_2022; @gugushvili_subjective_2017], while movement between these positions exposes individuals to new risks, resources and normative environments that reshape fairness judgements [@gugushvili_trends_2014; @helgason_longterm_2023]. Social mobility can thus foster market-justice orientations, helping explain how inequalities in access to welfare services become normatively sustained [@mau_inequality_2015].

Existing accounts identify several mechanisms through which intergenerational mobility may shape distributive justice preferences but converge on two channels especially relevant for market-based welfare: changes in material interests and shifts in normative frames [@ares_changing_2020; @jaime-castillo_social_2019; @helgason_longterm_2023; @gugushvili_intergenerational_2016c]. Mobility alters individuals’ material stakes by changing their occupational status, income, and risk exposure. At the same time, it reshapes their fairness judgments. Empirically, upward mobility is often associated with stronger meritocratic interpretations of success and fairness, whereas downward mobility is more closely linked to structural explanations of inequality without necessarily eroding meritocratic convictions [@mijs_belief_2022; @bucca_merit_2016; @deng_its_2025]. Regarding market-based preferences, recent research shows that meritocratic beliefs – the conviction that effort and ability are rewarded – are strongly associated with support for market allocation [@castillo_perceptions_2025]. In such contexts, the upwardly mobile are especially likely to view markets as just distributors, since their own trajectories appear to confirm that markets appropriately reward effort; those who experience downward mobility may still interpret outcomes through a meritocratic lens, often reading losses as personal failure rather than as evidence against market fairness. Thus, mobility and meritocracy evolve on partly separate tracks yet intersect systematically [@mau_inequality_2015].

Against this backdrop, the central aim of this article is to assess whether intergenerational occupational mobility causally affects support for the commodification of pensions. A second, more exploratory aim is to examine whether this effect varies by meritocratic beliefs, using this heterogeneity as an indication of possible underlying mechanisms. I argue that social mobility operates through two main pathways [@mau_inequality_2015]. Upward mobility increases earnings, expected returns in a funded system, and perceived economic security, making contribution-based access more appealing. Simultaneously, upward moves reinforce individualistic, effort-based understandings of success and, thus, the view that markets are just assignors. Downward mobility tends to heighten economic insecurity, dependence on public protection, and attachment to external attributions of poverty, reducing support for market allocation. Meritocratic beliefs are treated not as mediators but as moderators of these effects: among strong meritocrats, upward mobility should amplify pro-commodification attitudes, whereas downward mobility should attenuate – or even reverse – its expected de-commodifying impact, as meritocratic interpretations of effort lead individuals to view market outcomes as fair regardless of direction. This inquiry contributes to research on market justice by identifying the causal effects of mobility and using variation in meritocratic beliefs to assess whether these effects follow a meritocratic pathway.

This study focuses on Chile, an instructive case for examining preferences for pension commodification. Despite sustained economic growth, Chile combines high and persistent inequality with short-range upward mobility and substantial barriers to higher-class positions [@flores_top_2020; @salgado_uplifting_2025; @torche_intergenerational_2014]. It was the first country worldwide in 1981 to fully replace a public pay-as-you-go system with a mandatory, privately administered defined-contribution scheme, while the state has progressively expanded a segmented solidarity pillar to compensate for insufficient benefits [@madariaga_three_2020; @boccardo_30_2020; @solimano_rise_2021]. In parallel—and despite waves of protest from 2006 to 2019 against the private pension system [@somma_no_2021]—Chilean subjectivities have been increasingly shaped by neoliberal discourses and market logics, influencing attitudes toward pension distribution and inequality [@araujo_desafios_2012; @galvez_con_2023]. The combination of profound inequality, market-driven old-age security and contested legitimacy makes Chile a critical case for analysing how mobility and meritocratic beliefs structure support for pension commodification.

In this context, the questions that guide this research are as follows:

(1) How does intergenerational social mobility affect individuals’ preferences for the commodification of pensions in Chile?
(2) How do meritocratic beliefs condition or moderate this relationship?

To address these questions, the study draws on large-scale, representative survey data for the urban Chilean population collected in 2016, 2018 and 2023. It adopts a causal framework of mobility effects inspired by @breen_effects_2024. The next section outlines the theoretical framework linking preferences for market-based welfare, social mobility and expected heterogeneity by meritocratic beliefs and time, and derives hypotheses. A subsequent section presents the causal identification strategy, the data and the analytical approach. The empirical sections report the findings, and the article concludes by discussing what these results reveal about the politics of market-based pension welfare and social mobility.