### 2025-12-01 paper/02-antecedents.qmd

---
bibliography: "../input/bib/tesis-magister.bib"
csl: "../input/bib/apa6.csl"
---

# Theoretical and empirical background

## Preferences for pensions commodification

Beyond the state’s capacity to redistribute resources, market institutions are central arenas for allocating socially valuable goods and risks [@lindh_bringing_2023]. Markets are not mere aggregates of individual choices but social institutions endowed with rules and normative meanings, built through “moral projects, saturated with normativity” that permeate everyday thinking around fairness, effort and deservingness [@fourcade_moral_2007; @koos_moral_2019]. In this sense, the economic order is mirrored in a moral economy: shared norms and beliefs about what is considered a just distribution, embedded and reinforced through institutions and policy feedback [@svallforsMoralEconomyClass2006a]. In an era of privatisation and commodification, market logic has expanded into core areas of social reproduction—childcare, healthcare, education and pensions—stratifying access and quality [@ferre_welfare_2023; @gingrich_making_2011] and generating feedback effects on public beliefs about welfare and market inequality [@lindh_public_2015]. In this context, support for market-based welfare services provision has grown worldwide, particularly among higher-income groups who view private alternatives as more efficient or higher-quality [@busemeyer_welfare_2020; @lindh_public_2015]. These institutional transformations provide the backdrop for analysing how far citizens accept the commodification of social protection [@satz_por_2019].

The legitimacy of market-based welfare is closely tied to beliefs about distributive justice grounded in market principles. Beyond stratification structures, citizens hold structured beliefs about how resources ought to be allocated, reflecting causal attributions for inequality, normative principles and expectations about deservingness [@kluegel_beliefs_1981]. Research on distributive justice examines conceptions of how goods and rewards should be distributed and when inequality is considered just [@castillo_legitimacy_2011; @jasso_gender_1999], while work on redistributive preferences focuses on support for state-led mechanisms that reduce inequality [@cavaille_fair_2025]. Situated at this intersection, the emerging literature on market justice preferences conceptualises the market itself as a site of redistribution and asks whether individuals regard market criteria as fair bases for allocating wages and access to core welfare services [@lindh_public_2015; @lindh_bringing_2023; @castillo_perceptions_2025]. In this perspective, the expansion of market logic into welfare domains implies that social services are treated as legitimate commodities that can be priced and stratified [@busemeyer_welfare_2020], and empirical work suggests that such institutional changes feed back into attitudes, reinforcing market-conforming understandings of fairness and welfare [@zhu_policy_2015].

Market justice preferences refer to normative beliefs that legitimise allocating goods and services according to market criteria such as income and purchasing power [@kluegel_legitimation_1999; @lindh_public_2015]. Building on Lane’s [-@lane_market_1986] contrast between market and political justice, market justice is a distributive principle that endorses rewarding effort, productivity and skill rather than need and equality as emphasised in welfare-state policies. In this study, I treat market justice less as an abstract doctrine and more concretely as preferences for commodification: support for treating welfare services as commodities, so that better services follow from higher income or ability to pay, and the resulting stratified access is seen as fair [@lindh_public_2015]. Empirically, such orientations are typically measured with survey items asking whether it is fair that access to better services depends on income, a strategy rooted in research on justifications of capitalist inequality [@kluegel_legitimation_1999] and applied to healthcare and education [@lee_fairness_2023; @vondemknesebeck_are_2016; @immergut_it_2020; @castillo_changes_2025]. Moving beyond single domains, @lindh_public_2015 develops a comparative index of income-based access to healthcare and education, while @castillo_perceptions_2025 use the same single-item measure but add the pensions domain in Chile. These constructs gauge how individuals view market-generated inequalities as legitimate and capture two core dimensions of market distribution: the role of income in determining attainment and the framing of services as commodities that can be bought and sold according to ability to pay [@lindh_public_2015].

This study examines support for the commodification of pensions, defined as normative judgements on whether better pensions should depend on income or individual contributions. I treat this as a domain-specific form of market justice: pensions are conceived as commodities whose allocation should follow contribution- and return-based,  and whose outcome is seen as fair [@lane_market_1986; @mau_inequality_2015]. The spread of funded, privately managed schemes built on individual capitalisation accounts has transformed both the distributive and normative structure of old-age security [@huber_political_2000; @arenas_sistemas_2019]. Distributively, these schemes reassign risk from the collective to the individual by tying benefits to formal employment, contribution density, earnings and financial returns, making retirement income heavily dependent on labour-market trajectories and capital-market success [@madero-cabib_private_2019; @oecd_pensions_2023]. Normatively, they recast pensions as the outcome of personal investment decisions rather than collective obligations, promoting ideals of individual responsibility, self-economisation and contribution-based desert [@arza_pension_2008; @borzutzky_pension_2012]. Contributions are increasingly framed as private assets rather than risk pooling, and workers are expected to act as employee–entrepreneurs who choose providers, monitor portfolios and “make their money work” over the life course [@kerner_pension_2020; @mau_inequality_2015; @borzutzky_chiles_2016]. As more people tie their security expectations to private investments and financial markets, their material interests become more deeply rooted in the market, and their imperatives gain stronger normative support [@mau_inequality_2015]. When citizens are asked whether higher pensions should be allocated to those who earn and contribute more, they reflect support for the market allocation of a critical social service and the legitimacy of the resulting risks and inequalities in old age [@busemeyer_welfare_2020; @lindh_public_2015].

Most empirical work on market-based welfare preferences is cross-national and shows that support for income-based access to services is closely tied to countries’ economic and institutional contexts. Where public spending and social investment are higher, support for market allocation of health and education tends to be weaker, whereas in regimes with more prominent private providers and heavily privatized services, it is stronger [@busemeyer_skills_2014; @immergut_it_2020; @lindh_public_2015]. This literature suggests that market justice preferences align with existing institutional designs and outputs, echoing welfare-attitudes and policy-feedback research, which argues that institutions not only redistribute resources but also shape the categories through which people judge deservingness [@campbell_institutional_2020]. Dominant values and preferences are thus embedded in institutional configurations [@busemeyer_skills_2014], which actively produce moral frameworks that legitimize or contest inequality. Longitudinal evidence from Chile shows that agreement that higher-income groups should have access to better welfare services rose markedly between 2016 and 2023, especially for pensions [@castillo_perceptions_2025], plausibly reflecting mixed policy feedback effects of the private pension system [@busemeyer_positive_2021]. Comparative research also shows that market justice preferences are structured by individuals’ socioeconomic positions and broader attitudes to inequality.

## Social mobility

The drivers and consequences of social mobility have long been central to sociology. Social mobility denotes movements between positions in a stratification system. Classical and later work distinguish intragenerational (within the life course) from intergenerational mobility (between parents and children), as well as absolute mobility—driven by structural change—from relative mobility, which captures the extent to which social origins constrain destinations [@eyles_social_2022]. Beyond mapping mobility patterns, a large body of research examines how mobility shapes attitudes and behaviours. In this literature, “mobility effects” refer to outcomes that arise from movement between origin and destination classes [@breen_effects_2024]. A growing number of studies analyse the consequences of intergenerational mobility for attitudes towards economic inequality [e.g. @alesina_intergenerational_2018; @bucca_merit_2016; @day_movin_2017; @gugushvili_intergenerational_2016; @helgason_class_2025]. These inquiries rest on the idea that, if attitudes to inequality are strongly stratified by class—because classes embody distinct material interests and moral economies [@edlundDemocraticClassStruggle2015a; @kulinClassValuesAttitudes2013]—, then people who move between classes should adjust their attitudes towards those typical of their new class location. This study contributes to this line of work by examining how intergenerational mobility is associated with preferences for the commodification of pensions in a highly stratified and commodified welfare regime such as Chile’s.

The literature on the effects of social mobility identifies several mechanisms through which changes in class position may affect individual attitudes [@helgason_longterm_2023]. A first family of accounts emphasises material self-interest and can be divided into myopic and anticipatory theories. Myopic theories assume that individuals rapidly align their attitudes with their current class position, so mobility triggers shifts in material interests, information and perceived risks that update preferences [@helgason_longterm_2023; @ares_changing_2020; @langsaether_explaining_2022]. Anticipatory theories stress that forward-looking agents align their current attitudes with expected future income, as in the Prospect of Upward Mobility (POUM) hypothesis, in which individuals may oppose redistribution because they anticipate moving up [@benabou_social_2001].

A second family foregrounds culture and socialisation. Acculturation accounts posit that people gradually adapt their views to those prevailing in their destination class through exposure to new information, class-homophilous networks, group pressure and the assimilation of principles and normative beliefs [@jaime-castillo_social_2019; @helgason_class_2025]. Socialisation perspectives, instead, argue that core attitudes are formed early in life in the class of origin through families, schools, and critical events, and remain largely stable despite later mobility [@jaime-castillo_social_2019]. Status-maximisation theories propose that individuals align their attitudes with the highest class they have occupied, so upwardly mobile respondents converge towards their destination class, while downwardly mobile respondents retain attitudes from their higher-status origins [@jaime-castillo_social_2019]. Helgason and Rehm [-@helgason_longterm_2023;-@helgason_class_2025] refine these ideas in a learning framework, in which attitudes evolve slowly through cumulative class experiences that combine economic resources, socialisation into class-specific values, information, and networks. Finally, cognitive mechanisms highlight the self-serving bias in causal attribution: individuals interpret their mobility trajectories through internal or external causes in ways that legitimise their current socioeconomic position, thereby shaping their beliefs about effort, deservingness and inequality[@gugushvili_intergenerational_2016; @molina_its_2019; @miller_selfserving_1975]. In what follows, I focus on two mechanisms that are particularly relevant for market-based pensions: material self-interest and acculturation into meritocratic distributive norms.

How intergenerational social mobility shapes preferences for the commodification of pensions? A plausible mechanism runs through changes in material self-interest. Social mobility alters individuals’ incomes, job security and access to economic resources [@helgason_longterm_2023; @mau_inequality_2015], thereby modifying the costs and benefits they face in a private individual pension capitalization system. Material interests can be understood as the pursuit of economic well-being under the constraints and trade-offs implied by one’s class position [@wrightClassCountsComparative1997a]. Upward mobility typically entails higher and more stable earnings, greater capacity to contribute to funded schemes, greater tolerance for investment risk, and the ability to purchase supplementary coverage, whereas downward mobility reduces income, undermines contribution histories, and increases dependence on public or solidarity-based pillars [@madero-cabib_private_2019; @galvez_pensiones_2024]. These diverging pension prospects imply that upwardly and downwardly mobile individuals face systematically different stakes in market-based pension systems, making material self-interest a first channel through which mobility may shape support for pension commodification.

Consistent with this reasoning, existing research indicates that support for welfare commodification and pension marketisation is strongly structured by material advantage. 
Individuals with higher income, education and occupational class are more likely to endorse market-based distributive principles and to view private provision as a way to secure or enhance relative advantage  [@koos_moral_2019; @vondemknesebeck_are_2016; @immergut_it_2020; @busemeyer_skills_2014; @lindh_public_2015; @svallfors_political_2007]. In the pension field, institutional design sharpens these incentives. @busemeyer_welfare_2020 show that in countries where private alternatives to public pensions exist, support for public pension spending among upper-income groups erodes as they exit public schemes into private plans whose contributions and benefits closely track their income and risk. In Latin America, @kerner_pension_2020 finds that higher income and higher perceived financial returns from pension funds are associated with preferences for a stronger market role in pensions, suggesting that favourable returns are interpreted as evidence that pension neoliberalism “works”. Evidence from Chile points in the same direction: higher-income and university-educated respondents express stronger support for market allocation in healthcare, education and pensions [@castillo_socialization_2024; @otero_power_2024], and economic elites—especially those socialised in highly prestigious schools—hold more favourable attitudes towards inequality that can be traced to their material interests against redistribution [@carranza_what_2024]. Systems that link benefits to individual contributions thus allow better-off groups to capitalise on their financial independence and minimise reliance on public support, implying that mobility into or out of these advantaged positions should alter individuals’ stakes in pension commodification.

Research on redistribution and inequality offers additional, indirect support for this self-interest mechanism. Individuals with pessimistic mobility expectations—especially those anticipating downward movement—display stronger support for redistributive policies [@alesina_intergenerational_2018]. Studies tracking realised mobility show that, compared to immobility, upward moves are associated with more economically conservative, less redistributive attitudes, whereas downward moves strengthen demand for state intervention and redistribution [@ares_changing_2020; @langsaether_explaining_2022; @schmidt_experience_2011; @jaime-castillo_social_2019; @rodriguez_intergenerational_2024]. Over the life course, cumulative exposure to higher-class positions—whether by staying or moving up—is linked to more market-liberal economic values, while sustained exposure to lower-class positions—through stability or downward mobility—reinforces pro-redistributive orientations [@helgason_longterm_2023;@helgason_class_2025]. Domain-specific evidence also aligns with changing material stakes: @gugushvili_subjective_2017 finds that upward mobility is associated with lower support for government spending on housing and pensions, whereas downward mobility reduces support for healthcare and education but increases support for housing and pensions, reflecting the material nature of these domains. Beyond redistribution, upward mobility fosters more individualistic attributions of poverty and wealth and greater legitimation of income inequality [@bucca_merit_2016; @gugushvili_intergenerational_2016; @gugushvili_intergenerational_2016c], and societies with higher economic mobility display more tolerance of inequality [@shariff_income_2016]. Taken together, this evidence suggests that social mobility realigns attitudes with changing material stakes and distributive interests. Although the present study cannot directly test the underlying mechanism, it treats myopic self-interest as the central interpretive framework for the expected effects of intergenerational mobility on support for pension commodification.

Consequently, the main hypothesis of this research is:

$H_1$: Compared to immobility, intergenerational upward mobility is associated with stronger, and downward mobility with weaker, support for the commodification of pensions.

While material self-interest is the main lens, mobility also reshapes norms and interpretive frames. Mobile individuals move across social milieus and networks and are exposed to new standards and beliefs about fairness and deservingness [@helgason_longterm_2023]. Acculturation accounts hold that people gradually adopt the attitudes of their destination class [@jaime-castillo_social_2019]. Mobility can thus reshape how individuals interpret success, failure and inequality, with upward moves fostering more individualistic and meritocratic understandings and downward moves encouraging structural accounts of disadvantage [@gugushvili_subjective_2017; @mau_inequality_2015; @mijs_belief_2022] Applied to pensions, those who rise may come to see market-based schemes as fair systems that reward effort and foresight. In contrast, those who fall may view commodified pensions as unjust and  and that the pension system does not reward merit. Therefore, shifts in meritocratic norms can be a second channel through which the mobility effect operates, serving as a complementary acculturation mechanism that may influence support for pension commodification, as examined in the next section.

## Meritocracy

Meritocracy refers to a distributive system in which individual merit—typically defined as effort and talent—is the primary criterion for allocating resources and rewards, rather than social origins or inherited privilege [@young_rise_1958; @sen_merit_2000]. Originally coined by Young in a dystopian critique of a society where power and status are justified by achievement and mobility only driven by “merit”, the term has been re-appropriated as a positive ideal of fairness, especially in liberal and market-oriented societies [@dubet_repensar_2011; @mijs_paradox_2019; @vandewerfhorst_meritocracy_2024]. From a sociological perspective, belief in meritocracy is not just a cognitive judgement; it constitutes a moral lens through which individuals interpret social and economic disparities [@castillo_multidimensional_2023]. The conviction that economic inequalities are justified because they reflect differential merit has been identified as a key mechanism behind the persistence of inequality, since it recasts structural advantages and disadvantages as outcomes of individual performance and encourages “winners” to see their position as earned and “losers” to internalise blame rather than question the underlying structure of opportunity [@mijs_unfulfillable_2016; @sandel_tyranny_2020; @wilson_role_2003]. Understanding how these beliefs in meritocracy are structured is therefore crucial for analysing attitudes toward inequality.

Recent research has emphasised the need to decompose the term “meritocratic beliefs” and to distinguish between meritocratic preferences and meritocratic perceptions [@castillo_multidimensional_2023; @zhu_meritocratic_2025]. Preferences refer to normative ideas about how rewards should be allocated—whether people think effort and ability ought to determine life chances—whereas perceptions capture evaluations of how meritocracy actually operates in society, that is, whether people believe that existing inequalities reflect merit-based processes [@janmaat_subjective_2013]. This distinction is crucial for understanding how people interpret inequality in stratified societies because it clarifies the (mis)alignment between normative support for meritocracy and evaluations about its actual implementation [@castillo_multidimensional_2023; @mijs_paradox_2019; @mijs_how_2022; @newman_economic_2023; @lindner_does_2024]. In this study, I focus on meritocratic perceptions of effort and talent as key dimensions of broader meritocratic beliefs, given their central role in shaping attitudes toward inequality and welfare.

Several empirical studies have examined the social foundations and consequences of meritocratic perceptions. Individuals with higher levels of education, income and occupational prestige are more likely to endorse merit-based explanations for social outcomes [@mijs_paradox_2019; @duru-bellat_who_2012; @garcia-sanchez_perceptions_2018]. These perceptions function as a normative framework that legitimises several unequal outcomes, especially when access to social goods is stratified by income or social background. Individuals who see their society as meritocratic tend to show lower support for redistribution [@hoyt_mindsets_2023; @tejero-peregrina_perceived_2025], greater legitimisation of class inequality [@darnon_where_2018], increased tolerance of income gaps [@batruch_belief_2023], more support for system-justifying ideologies [@wiederkehr_belief_2015], and even lower perceived levels of economic inequality itself [@castillo_meritocracia_2019]. This evidence provides a basis for extending the study of the attitudinal consequences of meritocracy to preferences for market-based welfare.

Preferences for the commodification of social welfare services are anchored in normative frames. Studies show that support for market-based welfare is higher among individuals who endorse meritocratic and liberal principles—that benefits should depend on individual effort and contributions—and among those with economically conservative orientations, whereas egalitarian views reduce support for market allocation [@lee_fairness_2023; @quadagno_has_2012; @jaime-castillo_public_2013]. In the pension field, factorial evidence from Chile indicates that merit-based criteria such as educational attainment and years in the labour force weigh more heavily than need-based attributes in judgments of just pension levels, with respondents willing to accept very low pensions for those perceived as low achievers [@castillo_deserving_2019]. Recently, @castillo_perceptions_2025 show that agreement with the notion that it is fair for higher-income individuals to access better welfare services was relatively low in 2016 but rose significantly by 2023, especially in the pension domain. They also find that individuals who endorse more meritocratic perceptions–specifically the perception that effort is rewarded– are more likely to support market-based access for healthcare, education and pensions. All in all, meritocratic perceptions are a powerful source of inequality legitimation, anchoring who is seen as deserving to be better or worse off, and are strongly linked to how individuals understand unequal life chances and movements within the social hierarchy.

Social mobility and meritocratic perceptions are closely intertwined in the acculturation mechanism, although the direction of this link can vary across trajectories and contexts. A large body of work consistently shows that upwardly mobile individuals tend to interpret their trajectory through a meritocratic lens, reading their success as evidence that effort and ability are rewarded and reinforcing individualistic attributions of poverty and wealth [@bucca_merit_2016; @gugushvili_intergenerational_2016; @shariff_income_2016; @mijs_belief_2022], even in a highly unequal and low mobility setting were the narratives of individual success reinforces [@deng_its_2025]. Studies of economic elites in Chile similarly report strong endorsement of meritocracy, with respondents attributing their ascent mainly to talent and business or leadership skills [@atria_economic_2020]. Evidence on downward or blocked mobility is more ambivalent. Exposure to low-mobility contexts and experiences of downward mobility are associated with weaker beliefs that effort is rewarded, lower just-world and opportunity beliefs, and stronger structural attributions for success and failure [@gugushvili_intergenerational_2016; @davidai_why_2018; @day_movin_2017; @mijs_how_2022]. Yet, under strong meritocratic convictions, status loss can also be internalised as personal failure, sustaining system-justifying attitudes and the legitimation of existing hierarchies as a psychological defence strategy [@day_movin_2017; @mau_inequality_2015; @deng_its_2025].  In sum, upward mobility is generally associated with stronger meritocratic perceptions, whereas downward mobility is more closely linked to structural explanations of inequality, without necessarily eroding meritocratic convictions.

The ways in which social mobility trajectories shape interpretations of meritocracy can determine whether mobility translates into stronger or weaker support for market-based pension welfare. Following Mau’s [-@mau_inequality_2015] account of the “majority class”, upward mobility is expected to strengthen market-based welfare not only because material interests change, but because those who rise come to read their story as one of individual effort and talent rewarded. In this view, markets are seen as fair arenas that properly allocate pensions and other welfare benefits according to merit, so upwardly mobile individuals with strong meritocratic convictions should be especially prone to endorse commodified pensions. In contrast, downward mobility typically fosters de-commodifying demands, yet under strong meritocratic perceptions, status loss may be read as individually deserved, weakening—or even reversing—its usual de-commodifying pattern, as market outcomes are still viewed as broadly fair regardless of direction. Empirical research on redistribution and attitudes toward inequality supports this mechanism. Survey studies suggest that subjective social mobility shapes support for redistribution and pension spending through self-serving bias [@schmidt_experience_2011; @gugushvili_subjective_2017], although recent work indicates that this mechanism is evident only under explicit constraints in experimental designs [@molina_its_2019]. More in line with the acculturation theories, experimental reserach has shown that exposure to upward-mobility information or self-made narratives reduces support for redistribution, increases tolerance of inequality and system justification, and even legitimises exploitative work conditions, whereas exposure to low or downward mobility produces the opposite pattern; crucially, these effects are partly mediated by stronger meritocratic perceptions [@matamoros-lima_social_2025; @shariff_income_2016; @deng_its_2025]. In this framework, meritocratic perceptions condition the attitudinal consequences of mobility by amplifying the pro-market effects of upward mobility and muting the critical potential of downward mobility; thus, I treat meritocratic perceptions as a moderating mechanism linking intergenerational mobility to preferences for pension commodification.

Accordingly, the second set of hypotheses of this study is:

$H_{2a}$: Under high meritocracy, upward mobility should generate a stronger increase in support for market-based pension access than under low meritocracy.

$H_{2b}$: Under low meritocracy, downward mobility should moderately reduce support for market-based pension access, whereas under high meritocracy, this negative effect should be attenuated toward neutral or slightly positive attitudes.

A summary of the hypotheses and their expected directional patterns is presented in @tbl-hipo.

\begin{table}[htbp]
\centering
\caption{Overview of hypotheses with expected directional patterns {#tbl-hipo}}
\begin{tabular}{p{3.2cm} p{3.2cm} p{4cm} p{5cm}}
\toprule
\textbf{Hypothesis} & \textbf{Focus} & \textbf{Expected pattern} & \textbf{Interpretation} \\
\midrule
\textbf{H1: Average mobility effect} &
Effect of intergenerational mobility on preferences for pension commodification &
Upward mobility $\uparrow$ higher support; downward mobility $\downarrow$ lower support relative to immobility &
Mobility shapes preferences through material stakes and normative frames \\
\midrule
\textbf{H2a: Upward $\times$ Meritocracy} &
Conditional effect of upward mobility by meritocracy &
Under high meritocracy: amplified pro-market effect $\uparrow\uparrow$; under low meritocracy: modest increase $\uparrow$ &
Meritocratic lenses strengthen pro-market reactions to upward mobility \\
\midrule
\textbf{H2b: Downward $\times$ Meritocracy} &
Conditional effect of downward mobility by meritocracy &
Under low meritocracy: moderate decrease $\downarrow$; under high meritocracy: negative effect attenuated $\searrow$ toward neutral or slightly positive attitudes &
Meritocratic lenses weaken de-commodifying reactions to downward mobility \\
\bottomrule
\end{tabular}
\end{table}

## The Chilean context

Chile offers a compelling case for examining preferences for market-based pension welfare. Despite sustained economic growth, it remains one of the most unequal countries in Latin America and the OECD. The poorest 50% of the population captures just 10% of total income and holds negative net wealth, while the wealthiest 1% receives nearly 27% of all income and controls almost half of the country’s wealth [@chancel_world_2022], a pattern that has changed little since the 1990s [@flores_top_2020]. Research on intergenerational mobility shows that Chile combines high absolute mobility with mostly short-range moves—mainly from working classes into lower service-class positions—while access to upper and elite classes remains tightly restricted, and that despite the expansion of higher education, relative mobility remains rigid and intergenerational inequalities persist [@espinoza_movilidad_2014; @lopez-roldan_comparative_2021; @salgado_uplifting_2025; @torche_intergenerational_2014]. Crucially, a significant share of Chile’s inequality is rooted in neoliberal reforms that privatised and commodified key social services, stratifying and segmenting access to high-quality education, healthcare and housing between classes [@ferre_welfare_2023; @pnud_desiguales_2017].

Introduced under the military dictatorship (1973–1989) and expanded by democratic governments, neoliberal reforms in Chile embedded market logic into public services through concessions, subsidies, vouchers and pro-private regulation [@madariaga_three_2020; @boccardo_30_2020]. This “crowded-out” welfare model benefits higher-income groups, leaving lower-income individuals to rely on limited public options. Scholars argue that this neoliberal shift repurposed the state to create niches of capitalist accumulation—from water to education—giving rise to a model of “public-service capitalism” heavily reliant on state funding [@boccardo_30_2020]. In this setting, class trajectories unfold under conditions of high inequality and market-based provision in core welfare domains, such as pensions.

Chile was the first country to fully replace a public pay-as-you-go social insurance system with mandatory individual capitalisation: in 1981, the tripartite *"cajas de previsión"*(insurance schemes) were replaced by a fully funded defined-contribution scheme managed by private Pension Fund Administrators (AFP) [@arenas_sistemas_2019; @huber_political_2000]. Framed by its principal architect, José Piñera, as “the mother of all battles”, the reform was a way to roll back the redistributive role of the state, shift “freedom of choice” to individuals and create a “society of worker-capitalists” [@borzutzky_pension_2012; @solimano_rise_2021]. Dependent workers contribute 10% of their wages to individual accounts, plus an administrative fee, while employers contribute nothing; AFPs invest these mandatory savings in domestic and international financial markets, and since the 2000s affiliates choose (or are defaulted into) “multifunds” (A–E) with different risk profiles, directly exposing pensions to market volatility [@hyde_chiles_2015; @barriga_pensiones_2024]. After more than four decades, this architecture has produced an open pension crisis: for most retirees, AFP-financed pensions fall below the minimum wage and are clearly insufficient to avoid old-age poverty, with replacement rates around 30% even among cohorts with long contribution histories [@galvez_pensiones_2024; @madero-cabib_private_2019]. To prevent destitution, the state has been forced to re-enter through tax-funded solidarity pillars —the Nuevo Pilar Solidario (2008) and the Pensión Garantizada Universal (PGU, 2022)— that partially repair these failures [@boccardo_30_2020; @solimano_rise_2021] by topping up low or non-existent contributory pensions for the poorest and lower-middle older adults, so that for many low earners these tax-funded contributions are the main component of their total pension income [@galvez_pensiones_2024]. Meanwhile, the AFP industry has remained highly concentrated [@hyde_chiles_2015], posting very high returns on equity (around US$ 613 million in 2024) and fostering powerful national economic actors [@ruiz_formacion_2020], leaving Chile with a structurally fragile, segmented and regressive pension system that combines poor pensions, heavy reliance on public subsidies and high, sustained profits for private financial intermediaries. Chile’s combination of constrained mobility, persistent inequality and deep privatisation has eroded expectations of upward mobility and fuelled social discontent [@pnud_desiguales_2017], with pensions at the centre of this conflict.

Despite recurrent social unrest, Chile exhibits a paradoxical coexistence of intense conflict over inequality and its widespread legitimation, vividly illustrated by pensions. The October 2019 “social outburst”, a period of mass protests and severe political repression throughout the country, crystallised discontent over the privatisation and commodification of social services, alongside a crisis of political legitimacy [@somma_no_2021]. One of the main conflicts in this period centred on grievances over low pensions and the profits of private pension funds, building on earlier mobilisation by the *"No+AFP"* movement since 2016; a survey during the 2019 protests identified pensions as the top public demand [@nudesoc_informe_2020]. This conflict helped place comprehensive pension reform on the political agenda after 2022. However, contestation coexists with enduring market-conforming orientations. Research documents strong meritocratic beliefs and high tolerance for inequality [@castillo_meritocracia_2019; @mac-clure_justicia_2024], and the COVID-19 emergency withdrawals of pension savings gave rise to the slogan *“Con mi plata NO”*, expressing opposition both to greater pay-as-you-go elements and to continued AFP control [@galvez_con_2023]. This stance reflects the internalisation of the “enterprising self” [@mau_inequality_2015]: families navigate school choice as investment strategies [@canalesceron_sujeto_2021], and workers denounce the pension system as illegitimate in its design and outcomes while simultaneously treating their contributions as personal investments aimed at maximising returns and asserting that they themselves —rather than AFPs, the state or “society”— are the legitimate owners of their pension funds [@panes_criticas_2020]. These ambivalent representations indicate that market-based moral orientations are deeply embedded, making Chile a fertile setting to study how support for pension commodification varies across social groups and meritocratic beliefs.