### 2025-12-03 paper/05-discussion.qmd
Up draft presentation

---
bibliography: "../input/bib/tesis-magister.bib"
csl: "../input/bib/apa6.csl"
---

# Discussion and conclusions

Pension privatisation lies at the heart of Latin America’s trajectory of commodification since the 1980s, reshaping how societies organise old-age security [@ferre_welfare_2023]. By replacing public pay-as-you-go pillars with individual capitalisation schemes—privately managed defined-contribution accounts—pensions became governed by contribution-based entitlements and market performance [@huber_political_2000; @oecd_pensions_2023], with Chile as a compelling example. This shift both restructured who bears old-age risks and redefined the principles deemed legitimate for allocating pension resources [@borzutzky_pension_2012; @madero-cabib_private_2019; @solimano_rise_2021]. From a moral-economy perspective, these agreements generate feedback on public beliefs, fostering opinions that access to core social services should reflect income and individual responsibility, extending the principles of market justice [@mau_inequality_2015]. Existing research shows that socioeconomic position and beliefs about inequality shape support for market-based access to welfare services [@busemeyer_skills_2014; @castillo_perceptions_2025; @immergut_it_2020; @lindh_public_2015; @svallfors_political_2007]. However, we still know little about how citizens assess the fairness of pension commodification in such settings, and how social mobility and meritocratic perceptions shape these judgements. This article addresses this gap by examining whether intergenerational mobility causally affects support for pension commodification in Chile and whether this effect varies by meritocratic perceptions. Drawing on survey data and under explicit identification assumptions, I apply Breen and Ermisch’s [-@breen_effects_2024] causal framework of mobility effects, treating destination class as a treatment, conditioning on origin class, and estimating heterogeneous destination effects.

The findings reveal a tension in Chileans’ judgements of market-based pension welfare. Across 2016, 2018 and 2023, large majorities reject the claim that it is fair for higher-income people to obtain better pensions than those with lower incomes. However, a non-trivial—and growing—minority endorses this principle, with the sharpest increase between 2018 and 2023. Market-based pension allocation thus remains a minority view. However, it is becoming more common, in line with longitudinal evidence of rising support for income-based access to welfare services in Chile [@castillo_perceptions_2025]. This coexistence unfolds within a system of mandatory, privately managed, defined-contribution accounts that tightly link old-age income to earnings, contribution density and financial returns [@madero-cabib_private_2019; @galvez_pensiones_2024], and that frame pensions as personal investments rather than solidarity-based rights, promoting individual responsibility, self-economisation, contribution-based desert and casting workers as employee-entrepreneurs who must navigate financial risks and “make their money work” [@borzutzky_pension_2012]. The Chilean case thus fits evidence that income-based welfare access is more strongly endorsed in liberal, marketised regimes [@lindh_public_2015].This coexistence also echoes recent pension conflict—from the “No+AFP” protests to post-pandemic withdrawals and the “Con mi plata NO” mobilisation [@galvez_con_2023]—and illustrates mixed policy feedback: the same regime generates rights-based critiques and proprietarian defences of pension commodification. Together, these patterns suggest that, despite discontent and a legitimacy crisis, support for pension commodification remains a salient and contested orientation, raising the question of which social trajectories and beliefs underpin it.

Turning to the social mobility hypothesis, the results support a causal role of intergenerational mobility in shaping preferences for pension commodification. Rise from middle to high occupational status is causally related to significantly greater support for pension commodification, whereas the opposite holds for descent from high to middle and from high to low status, relative to immobility in the origin class. These patterns are consistent with a myopic material self-interest mechanism [@helgason_longterm_2023] in Chile’s pension system, where benefits mainly depend on individual contributions: upwardly mobile middle→high individuals gain income, employment stability and contribution capacity, and can better bear investment risk, making a contribution-linked, market-based pension regime more attractive. By contrast, those moving down from high positions lose resources and security, become more exposed to labour-market risks and depend more on public or solidarity-based pillars, which helps explain their lower support for income-based allocation [@busemeyer_welfare_2020; @lindh_public_2015]. The fact that significant effects emerge only for trajectories into and out of the upper occupational status resonates with research on static social class and pensions, which shows that higher socioeconomic groups and those obtaining higher financial returns are more supportive of market-based welfare and private pension arrangements [@busemeyer_welfare_2020; @kerner_pension_2020; @castillo_socialization_2024; @otero_power_2024]. It also aligns with evidence on social mobility and redistribution, where upward mobility is linked to less redistributive attitudes and lower support for public spending on pensions, while downward mobility has the opposite effect [@ares_changing_2020; @gugushvili_subjective_2017; @schmidt_experience_2011; @rodriguez_intergenerational_2024; @alesina_intergenerational_2018]. Still, material self-interest may not be the only channel through which mobility affects these judgements, raising the question of whether meritocratic perceptions condition these effects by providing a normative lens to interpret mobility experiences.

The second set of hypotheses examined whether meritocratic perceptions condition the effects of intergenerational mobility. Building on acculturation accounts and self-serving bias arguments [@jaime-castillo_social_2019; @gugushvili_subjective_2017], I expected mobility to affect preferences for pension commodification not only through material self-interest but also through shifts in normative frames of distributive justice. Prior research shows that upwardly mobile individuals tend to interpret their trajectory through a meritocratic lens, reading their success as evidence that effort and ability are rewarded and reinforcing individualistic attributions of poverty and wealth. In contrast, downward or blocked mobility is more often associated with structural explanations of inequality and more ambivalent meritocratic beliefs, sometimes including the internalisation of status loss as personal failure [@traini_climbing_2025; @gugushvili_intergenerational_2016; @davidai_why_2018; @day_movin_2017; @mijs_how_2022; @deng_its_2025]. Following Mau’s [-@mau_inequality_2015] account of the “majority class”, I therefore expected upward mobility under strong meritocratic convictions to be especially conducive to endorsing commodified pensions, and strong meritocracy to dampen the usual de-commodifying effect of downward mobility. Empirically, however, the interaction models offer little support for this mechanism. Meritocratic perceptions do not systematically moderate the mobility effects identified in the primary analysis, and the few trajectory-specific patterns that tentatively point in the expected direction are small and imprecisely estimated. This result contrasts with survey and experimental studies showing that upward mobility experiences or exposure to self-made narratives reduce support for redistribution and public pension spending partly by reinforcing meritocratic convictions [@schmidt_experience_2011; @gugushvili_subjective_2017; @molina_its_2019; @shariff_income_2016; @matamoros-lima_social_2025; @deng_its_2025], and aligns with recent research that has shown that meritocracy does not mediate the effect of social mobility on redistributive preferences [@rodriguez_intergenerational_2024]. One possibility is that, in this context, intergenerational occupational mobility—as operationalised here—has limited impact on meritocratic perceptions, which appear more weakly tied to mobility trajectories than previous research suggests and more strongly anchored in broader socialisation processes and class-specific cultures.

The contributions of this study can be summarized as follows. Substantively, it offers one of the first systematic analyses of preferences for pension commodification in Chile’s highly unequal and marketised pension regime, addressing Mau’s [-@mau_inequality_2015] question of why core elements of neoliberal welfare arrangements gain public acceptance. By showing that upward and downward intergenerational mobility into and out of upper-class positions causally shape support for market-based pension allocation, the findings highlight social mobility as a key source of legitimation for commodified pensions. Second, the article advances research on market justice and market-based welfare preferences by moving beyond static class locations and showing how intergenerational occupational mobility structures support for pension commodification in ways that partly echo, but also refine, existing evidence on class, income and financial returns. Third, the null results for the moderating role of meritocratic perceptions nuance accounts that tightly link mobility, meritocracy and the embrace of neoliberal welfare: in this context, mobility effects on pension commodification preferences appear to operate independently of meritocratic beliefs. Finally, the study contributes methodologically by applying Breen and Ermisch’s [-@breen_effects_2024] causal framework of mobility effects to survey data, formalising mobility as origin-specific treatment effects and estimating these using linear probability models with inverse probability weighting.

Despite its merits, the study has limitations that should be acknowledged and read as avenues for future improvement. First, the causal estimates hinge on the construction of inverse probability weighting within Breen and Ermisch’s [-@breen_effects_2024] framework. Identification relies on conditional independence given a relatively limited but relevant set of origin-side covariates, together with positivity and SUTVA. Entropy balancing helps to enforce covariate balance and mitigate selection on observables, but unmeasured confounding cannot be ruled out. Sensitivity analyses suggest moderate robustness to omitted variables; therefore, the estimates should be interpreted with caution. Second, while the analysis identifies mobility effects on preferences, it does not explain why or how they arise. The results are analysed through material self-interest and acculturation mechanisms. Still, the study does not implement formal causal mediation analysis, nor can it determine whether meritocratic perceptions function as mediators rather than mere correlates. Third, the findings are context-specific: they are derived from Chile, a highly unequal country with a strongly marketised pension regime, which enhances internal validity but limits external validity. The patterns documented here, therefore, cannot be straightforwardly generalised to countries with different welfare institutions, mobility regimes or levels of inequality, and should be seen as a theoretically informed case study rather than a cross-national pattern.

Future research could extend these findings in several directions. First, there is a need for richer measures of preferences for pension commodification. This study relies on a single item that conflates income-based allocation and fairness judgments; better items that separate justice principles and dimensions of commodification would allow a finer assessment of how people evaluate pensions as income-conditioned commodities. Second, methodological limitations could be addressed by including more detailed origin-side covariates and life-course data, repeated measures of meritocratic perceptions, and causal mediation analyses that unpack mechanisms such as material self-interest, risk exposure, and meritocratic frames, alongside alternative mobility measures based on class schemas, income trajectories, or wealth. Third, comparative designs would clarify how institutional context shapes the attitudinal consequences of intergenerational mobility. Applying similar causal strategies in countries with different welfare regimes, levels of inequality, mobility patterns, and degrees of pension marketisation would show whether the Chilean pattern is context-specific or part of a broader configuration linking mobility, merit and support for market-based welfare [@mau_inequality_2015]. Ultimately, clarifying when and for whom market allocation of pensions is judged fair is crucial to understanding the legitimate reach of welfare commodification in unequal, marketised societies.