### 2025-12-03 paper/paper.qmd

---
title: "Preferences for the commodification of pensions in Chile: the rol of intergenerational social mobility"
date: today
date-format: "MMMM, YYYY"
format:
  html: 
    latex-environment: true
    code-fold: true
    html-math-method: katex
    number-sections: true
    code-link: true
    title-block-banner: true
    theme:
      - pretty.scss
  pdf: 
    pdf-engine: xelatex
    geometry: margin=2cm
    template-partials: 
      - title.tex
    keep-tex: true
    include-in-header:
      text: |
        \usepackage[noblocks]{authblk}
        \renewcommand*{\Authsep}{, }
        \renewcommand*{\Authand}{, }
        \renewcommand*{\Authands}{, }
        \renewcommand\Affilfont{\small}
    number-sections: true
editor: source
bibliography: "../input/bib/tesis-magister.bib"
csl: "../input/bib/apa6.csl"
author:
  - name: 'Andreas Laffert Tamayo\thanks{This M.A. thesis in Sociology is one of my first achievements on the long path toward becoming an academic committed to the study of inequalities and to a public sociology. This journey would not have been possible without my relentless effort and the support of my family, friends and colleagues. Last but certainly not least, I want to acknowledge the motivation behind my research: the everyday processes through which neoliberalism is legitimised by ordinary people. This project was made possible through the financial support of ANID (BECAS/MAGÍSTER NACIONAL 2024) and the intellectual support and academic environment provided by the FONDECYT Regular project Nº 1250518. Any errors or omissions in this work are my sole responsibility. Comments are welcome at: andreas.laffert@uchile.cl}'
    orcid: 0000-0002-9008-2454
    email: alaffertt@estudiante.uc.cl 
    affiliations:
      - ref: 1
affiliations: 
  - id: 1
    name: Instituto de Sociología, Pontificia Universidad Católica de Chile
citeproc: true
abstract: | 
 The privatisation and marketisation of social services have been linked to feedback effects on public attitudes towards welfare and market inequality. From a moral-economy perspective, prior research shows that support for market-based access to core services depends on socio-structural position, inequality beliefs and institutional context. Yet, we know little about how citizens judge the fairness of pension commodification, and in particular how intergenerational mobility shapes these judgements and whether meritocratic perceptions condition this relationship. I address this gap by analysing preferences for market-based pension access in Chile, a highly unequal country with constrained mobility and a private pension system based on mandatory individual capitalisation that depends heavily on contribution histories. Drawing on three waves of the Chilean Longitudinal Social Survey (ELSOC; 3,435 observations on 1,787 individuals) and under explicit identification assumptions, I apply a recent causal framework for mobility effects that treats destination class as a treatment conditional on origin class, and estimate heterogeneous destination effects on support for pension commodification. Results show that upward mobility from middle to upper occupational status increases support for market-based pension access relative to immobility. In contrast, downward mobility from upper to middle or lower status reduces it. Meritocratic perceptions do not significantly moderate these effects, suggesting that mobility and meritocracy operate through largely separate pathways. These findings advance research on welfare marketisation by providing causal evidence on how intergenerational mobility shapes preferences for market-based welfare, and by assessing the potential moderating role of meritocratic perceptions in a highly commodified regime.
  \newline
 **Keywords**: Pension commodification · Intergenerational mobility · Causal inference · Meritocracy · Chile
link-citations: true
linestretch: 1.5
mainfont: Times New Roman
fontsize: 13pt          
colorlinks: true
fig-height: 4
fig-width: 7.5
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo=FALSE, 
                      warning = FALSE,
                      message = FALSE, 
                      cache = TRUE,
                      out.width = '85%',
                      fig.pos= "H")
options(scipen=999)
options(kableExtra.auto_format = FALSE)
options(knitr.kable.NA = '')
options(knitr.graphics.error = FALSE)
Sys.setlocale("LC_ALL", "ES_ES.UTF-8")
table_format = if (knitr::is_html_output()) {
   #conditional instructions for kable
   "html" #if html set "html" in format
 } else if (knitr::is_latex_output()) {
   "latex"#if latex set "latex" in format
 }

library(conflicted)
conflicts_prefer(dplyr::filter)

```

:::: {.justify}

{{< include 01-introduction.qmd >}}

{{< include 02-antecedents.qmd >}}

{{< include 03-methods.qmd >}}

{{< include 04-analysis.qmd >}}

{{< include 05-discussion.qmd >}}

# References

::: {#refs}
:::
::::

::: {pagebreak}
:::

{{< include 06-supplementary-material.qmd >}}