### 2025-12-03 presentations/social_mobility_2025/mjp-mobility-conf-2025.qmd

---
bibliography: "bib/tesis-magister.bib"
csl: "bib/apa6.csl"
lang: es
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
    R.options:

      knitr.graphics.auto_pdf: true
format:
  revealjs:
    #logo: images/edumer.png
    slide-number: true
    theme: "libs/edumer.scss"
    auto-stretch: false
    title-slide-attributes:
      visibility: false
    transition: fade
    transition-speed: slow
# data-background-image: images/cover.jpg
# data-background-size: cover
    auto-play-media: true
    mathjax: "default"
editor_options: 
  chunk_output_type: console
---

```{r setup2, include=FALSE}
 knitr::opts_chunk$set(echo=FALSE, warning = FALSE,message = FALSE, cache = TRUE,out.width = '85%',fig.pos= "H"
                       # , fig.align = 'center'
                       )
 # knitr::opts_knit$set(base.url = "../") #relative path for .html output file
 # knitr::opts_knit$set(root.dir = "../") #relative path for chunks within .rmd files
 options(scipen=999)
 options(kableExtra.auto_format = FALSE)
 options(knitr.kable.NA = '')
 options(knitr.graphics.error = FALSE)
 Sys.setlocale("LC_ALL", "ES_ES.UTF-8")

 table_format = if (knitr::is_html_output()) {
   #conditional instructions for kable
   "html" #if html set "html" in format
 } else if (knitr::is_latex_output()) {
   "latex"#if latex set "latex" in format
 }

 library(conflicted)

conflicts_prefer(dplyr::filter)
```

```{r}
#| label: packages
#| include: false

if (! require("pacman")) install.packages("pacman")

pacman::p_load(here,
               kableExtra)

options(scipen=999)
rm(list = ls())

```

```{r}
#| label: data
#| include: false

source(file = here("processing/prod_analysis_multinom.R"))
library(here)
library(janitor)
library(dagitty)
library(ggdag)
library(tidyverse)

```

::: columns

::: {.column width="20%"}

![](images/logo_isuc.png)

</br>

![](images/jusmer_trans.png)

:::

::: {.column .column-right width="80%"}

# **Preferences for the commodification of pensions in Chile: the rol of intergenerational social mobility**

------------------------------------------------------------------------

**Andreas Laffert^1^**

::: {.blue2 .medium}

**^1^Institute of Sociology, Pontifical Catholic University of Chile**

:::

International Conference: The Future of Social Mobility

3-4 December 2025, Chile

:::
:::

# Context and motivation {.xlarge data-background-color="#523870"}

## Context: marketisation and pensions

::: {.box-inv-4 .sp-after .fragment style="font-size: 110%;"}

Privatization and commodification of public goods and social services [@gingrich_making_2011; @ferre_welfare_2023]

:::

::: {.box-inv-4 .sp-after .fragment style="font-size: 110%;"}

In Latin America, pension privatization and individual capitalization have been pivotal [@huber_political_2000; @verbic_political_2019]

:::

::: {.box-inv-4 .sp-after .fragment style="font-size: 110%;"}

Reallocate old-age risks and reshape public norms about what is fair in welfare [@busemeyer_welfare_2020; @madero-cabib_private_2019; @borzutzky_pension_2012]

:::

::: {.box-inv-5 .sp-after-half .fragment style="font-size: 110%;"}

Preferences for the market-based access to social welfare [@busemeyer_skills_2014; @castillo_perceptions_2025; @lindh_public_2015; @immergut_it_2020; @svallfors_political_2007]

:::

## Chile and the research gap

::: {.box-inv-4 .sp-after .fragment style="font-size: 110%;"}

Chile combines high inequality, deep welfare privatisation and contested private pensions [@flores_top_2020; @somma_no_2021; @madariaga_three_2020]

:::

::: {.box-inv-4 .sp-after .fragment style="font-size: 110%;"}

High absolute mobility coexist with short-range class moves, restricted access to elite positions, and persistent intergenerational inequalities [@lopez-roldan_comparative_2021; @salgado_uplifting_2025; @torche_intergenerational_2014]
:::

::: {.box-inv-4 .sp-after .fragment style="font-size: 110%;"}

We know that support for market-based services is stratified [@lindh_public_2015; @busemeyer_welfare_2020; @kerner_pension_2020; @otero_power_2024] and linked to meritocracy [@castillo_perceptions_2025], but less about how social mobility and meritocratic beliefs shape these preferences in the pension domain.
:::

## Research questions and H1

::: {.incremental .highlight-last style="font-size: 120%;"}

- RQ1: How does intergenerational social mobility affect preferences for pension commodification in Chile?

- Social mobility changes incomes, security and risk exposure, altering material interests in a funded DC system [@busemeyer_welfare_2020; @helgason_longterm_2023; @mau_inequality_2015]

- Social mobility and preferences for redistribution and social spending [@alesina_intergenerational_2018; @rodriguez_intergenerational_2024; @langsaether_explaining_2022; @ares_changing_2020; @gugushvili_subjective_2017]

- $H_1$: Compared to immobility, intergenerational upward mobility is associated with stronger, and downward mobility with weaker, support for the commodification of pensions

:::

## Research questions and H2

::: {.incremental .highlight-last style="font-size: 120%;"}

- RQ2: How do meritocratic beliefs condition the mobility–commodification relationship?

- Social mobility reshapes inequality attitudes by pushing the upwardly mobile toward meritocratic views and the downwardly mobile toward more structural accounts, with strong meritocratic beliefs acting as a key channel of these effects [@mijs_belief_2022; @gugushvili_intergenerational_2016; @bucca_merit_2016; @traini_climbing_2025; @matamoros-lima_social_2025; @deng_its_2025]

- $H_{2}$: Under high meritocracy, upward mobility should be more pro-commodification and downward mobility less de-commodifying than under low meritocracy

:::

# Method {.xlarge data-background-color="#523870"}

(M)odel, (I)nquiry, (D)ata, (A)nswer estrategy [@blair_research_2023]

## Causal model (M): @breen_effects_2024

::::: columns
::: {.column width="50%"}
::: {.fragment}

![](images/breen.png)

:::

:::

::: {.column width="50%" .incremental .highlight-last style="font-size: 110%;"}

- I follow @breen_effects_2024 to estimate causal effects of intergenerational mobility.
- For each origin class $O=j$, I compare movers from $j$ to $k$ with their counterfactual outcome if they had remained in $j$

:::
:::::

## A Causal Model for Mobility Effects

```{r}
#| label: fig-dag
#| out-width: '95%'
#| fig-cap: "Causal graph of the effect of intergenerational social mobility on preferences for commodification. Y = preferences for pension commodification, D = an individual’s class of destination, O = an individual’s class of origin, C = different attributes determined in childhood or earlier that affects an individual’s class of origin and destination. Finally, U = unobserved factors influencing origin attributes."
#| fig-cap-location: bottom
#| echo: false
#| warning: false
#| message: false

dag_mjp <- dagify(
  Y ~ D + C,
  D ~ O + C,
  O ~ C,
  C ~ U,
  coords = time_ordered_coords(
    list(
      "U",
      c("C", "O"),
      "D",
      "Y"
    )
  ),
  exposure = "D",
  outcome = "Y",
  labels = c(
    Y = "Y",
    D = "D",
    O = "O",
    C = "C",
    U = "U"
  )
)

curvatures <- rep(0, 8)
curvatures[3] <- -.25

dag_mjp |>
  tidy_dagitty() |>
  ggplot(aes(x, y, xend = xend, yend = yend)) +
  geom_dag_point(shape = 21, fill = "black", size = 14, stroke = 0.8) +
  geom_dag_edges_arc(curvature = curvatures) +
  geom_dag_text(aes(label = label), color = "white", fontface = "bold", size = 4) +
  theme_dag()

```

## Inquiry and estimand
::: {.incremental .highlight-last style="font-size: 120%;"}

- How does intergenerational social mobility affect individuals’ preferences for the commodification of pensions in Chile?

- @breen_effects_2024: Average causal effect of moving from origin class *j* to destination class *k*, comparing individuals who actually reach *k* with their counterfactual outcome had they remained in their origin *j* (@eq-att).

- Formally,  for individuals with origin $(O=j)$ who attain destination $(D=k)$, the estimand is:

:::

::: {.fragment}
$$ 
ATT_{j,k,j} = E\big[Y(D = k) \mid O = j, D = k \big] - E\big[Y(D = j) \mid O = j, D = k \big]
$${#eq-att}
:::

## Data (D)

::: {.incremental .highlight-last style="font-size: 120%;"}

- [ELSOC](https://coes.cl/elsoc/) (COES): national urban panel 2016–2023; stratified, cluster, multistage probability sampling; target population: regular residents aged 18–75

- I analyze 2016, 2018, and 2023 based on the availability of occupational variables

- Analytical sample: *N*=3,435 observations nested in *N*=1,787 individuals (2016: 914; 2018: 1,377; 2023: 1,144)

- Subsamples by mobility trajectories (e.g., low→low, low→medium), retaining movers and non-movers paired from the same class of origin
:::

## Outcome

```{r}
#| label: fig-desc
#| fig-cap-location: top
#| fig-cap: "Distribution of responses for pension commodification preferences in Chile (2016-2023)"
#| fig-align: center
#| out-width: '130%'

g1 +
  labs(title = "It is fair that people with higher incomes have better pensions than people\nwith lower incomes?",
       subtitle = "Likert: 1 (Strongly disagree) - 5 (Strongly agree)") +
  theme(legend.position = "top",
        plot.title = element_text(face = "bold"),
        text = element_text(size = 16))

```

## Treatment: mobility and class matrix (D)

::: {.incremental .highlight-last style="font-size: 120%;"}

- **Class assignment (O and D):** Using ISCO-08/ISEI, I define origin $O$ and destination $D$ based on tertiles ∈ {low, medium, high}
- For each origin $(O=j)$, the treatment is to reach $(D=k)$ (mobile) vs. immobile $(D=j)$

:::

::: {.fragment}

```{r}
#| label: tbl-matrix
#| tbl-cap: "Occupational mobility by occupational groups"
#| tbl-cap-location: top
#| results: asis
#| echo: false
library(janitor)
db %>%
  tabyl(estrato_orig, estrato_ocupa) %>%
  adorn_totals(where = c("row", "col")) %>%     # agrega totales
  adorn_percentages("row") %>%                  # porcentajes por fila
  adorn_pct_formatting(digits = 1) %>%          # formato %
  adorn_ns() %>% 
  as_tibble() %>% 
  mutate(Offspring = NA) %>% 
  select(estrato_orig, Offspring, everything()) %>% 
  kableExtra::kable(format = "html", 
                    booktabs= T,
                    align = 'c', 
                    col.names = c("Father↓", "Offspring→", "Low", "Middle", "High", "Total"),
                    caption = NULL) %>%
  kableExtra::kable_styling(latex_options = "hold_position", 
                            bootstrap_options = "striped",
                            full_width = F)

```

:::

## Weights and covariates (D → A)

::: {.incremental .highlight-last style="font-size: 120%;"}

- Pre-treatment covariates $C$: (a) parental education (highest of father/mother), in 10 ordinal categories from no schooling to postgraduate studies; (b) co-residence with both parents at age 15 (0 = no, 1 = yes); (c) nationality (0 = non-Chilean, 1 = Chilean); (d) age in years; (e) sex (0 = male, 1 = female); and (f) indigenous ethnicity (0 = no, 1 = yes).
- Within each origin, entropy balancing reweights non-movers so their covariate distribution matches that of movers → stabilised IPW–ATT.
- These weights target the ATT and help adjust for selection into mobility.

:::

## Heterogeneity and controls (D → A)

::: {.incremental .highlight-last style="font-size: 120%;"}

- Meritocracy: two items (effort and intelligence/skills rewarded), averaged and dichotomised into low vs. high meritocracy.

- I estimate models with mobility × meritocracy interactions to study heterogeneous treatment effects.

- The same pre-treatment covariates $C$ enter the outcome model to block backdoor paths and achieve double robustness.

:::

## Analytic strategy (A)

::: {.incremental .highlight-last style="font-size: 120%;"}

- For each origin class, I estimate weighted linear probability models:

:::

::: {.fragment style="font-size: 120%;"}
$$
Y_{it}=\alpha+\beta T_i+X_i'\gamma+\lambda_t+\varepsilon_{it}
$$ {#eq-final}
:::

::: {.incremental style="font-size:120%;"}

- Models use stabilised IPW–ATT weights, wave fixed effects and CR2 cluster-robust SEs by individual.
- $\beta$ is interpreted as $\widehat{ATT}_{j,k,j}$: the causal effect of reaching destination $k$ versus staying in origin $j$

:::

# Results {.xlarge data-background-color="#523870"}

## Mobility effects

::: {.fragment style="font-size: 120%;"}

```{r}
#| label: fig-mp
#| fig-cap-location: top
#| fig-cap: "Mobility effects on preferences for pension comodification"
#| fig-align: center
#| out-width: '180%'
#| fig-asp: 0.6

g3 +
  theme(text = element_text(size = 14))
```

:::

## The rol of meritocracy

::: {.fragment style="font-size: 120%;"}

```{r}
#| label: fig-mp_i
#| fig-cap-location: top
#| fig-cap: "Mobility effects on preferences for pension comodification, by meritocratic perceptions"
#| fig-align: center
#| out-width: '160%'
#| fig-asp: 0.6

g4+
  theme(text = element_text(size = 14))
```

:::

## Robustness check and sensitivity analysis

::: {.incremental .highlight-last style="font-size: 120%;"}

- **Unmeasured confounding** [@cinelli_making_2020]: **Moderate robustness** in Middle→High and High→Low; borderline in High→Middle.

- **Robustness check**: I re-estimate the models under three alternative codings of the outcome: (i) recoding the neutral category as agreement (3–5 vs. 1–2), (ii) excluding neutral responses from the analysis, and (iii) using the original 5-point item in a linear specification. The results are maintained and even improved (only in ii).
:::

# Discussion and conclusions {data-background-color="#523870"}

## Discussion and conclusions

::: {.incremental style="font-size: 120%;"}

1) Most Chileans reject income-based pension allocation, yet support for this market principle has grown in a highly marketised and unequal setting [@lindh_public_2015; @castillo_perceptions_2025].

2) Intergenerational moves into and out of upper-status positions causally shape support for pension commodification, while donward High-Middle y High-Low decrease it [@gugushvili_subjective_2017; @rodriguez_intergenerational_2024; @ares_changing_2020]; meritocratic beliefs correlate with these preferences but seem to operate on a parallel track.

3) Future work: better origin-side measures, richer batteries of market-justice items and comparative applications beyond Chile.

:::

# Thanks!

-   **Github project:** <https://github.com/Andreas-Lafferte/mobility-market-justice>

## References